\documentclass[11pt,a4paper]{article}

% ---------------------------------------------------------
% PACKAGES
% ---------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts, bm}
\usepackage{mathtools}
\usepackage{physics}
\usepackage{tensor}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{float}
\usepackage{cite}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{microtype}
\usepackage[most]{tcolorbox}

% ---------------------------------------------------------
% HYPERREF
% ---------------------------------------------------------
\usepackage[colorlinks=true,
            linkcolor=blue,
            citecolor=blue,
            urlcolor=blue,
            pdfauthor={Michael Lehmann},
            pdftitle={Observational Synthesis and Global Fit in RDCC},
            pdfsubject={Relaxation-Driven Cyclic Cosmology},
            pdfkeywords={RDCC, cosmology, global fit, CPT symmetry, relaxation dynamics}
           ]{hyperref}

% ---------------------------------------------------------
% TITLE
% ---------------------------------------------------------
\title{\textbf{Dynamical Generation of the Relaxation Parameter in\\
Relaxation-Driven Cyclic Cosmology (RDCC)}}

\author{Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}}

\date{April 2026}

\begin{document}

\maketitle

\begin{abstract}
Relaxation-Driven Cyclic Cosmology (RDCC) is a one-parameter, CPT-symmetric
framework in which all observable deviations from $\Lambda$CDM arise from the
infrared value of the relaxation parameter $\alpha_{\mathrm{IR}}$. While Companions~I--V
established the gravitational foundations, tensor phenomenology, relaxation-induced
scalar sector, global CPT state, and the first prototype global fit, the microscopic
time evolution of the relaxation parameter has not yet been derived on a realistic
cosmological background.

This paper, Companion~VI, closes this gap. We derive the relaxation equation
$\dot{\alpha} = \Gamma \alpha$ from the dimension-six inter-sector operator, show that the
naive EFT estimate $\Gamma \sim H^5/M^4$ fails to generate any observable relaxation,
and introduce the minimal generalized rate $\Gamma = \beta H$ consistent with the
CPT-symmetric structure. Matching to the observed value $\alpha_{\mathrm{IR}} \simeq 0.0175$
fixes $\beta \simeq 1.25 \times 10^{-4}$.

We solve the full evolution equation $\mathrm{d}\alpha/\mathrm{d}a = (\Gamma/H)\alpha$ on a
$\Lambda$CDM background and find that $\alpha(a)$ grows sharply near the bounce and
freezes to its infrared value long before matter–radiation equality. This dynamical
behavior explains why the RDCC prediction vector remains stable across more than ten
orders of magnitude in scale and provides a natural microscopic origin for the observed
value of $\alpha_{\mathrm{IR}}$.

Companion~VI thereby completes the dynamical foundation of RDCC and prepares the
ground for a full Boltzmann implementation in Companion~VII.
\end{abstract}

\newpage

\tableofcontents

\newpage

\section{Introduction}

Relaxation-Driven Cyclic Cosmology (RDCC) describes the Universe as a globally
CPT-symmetric bipartite quantum system whose observable deviations from
$\Lambda$CDM are governed by a single infrared parameter, the relaxation strength
$\alpha_{\mathrm{IR}} \sim 10^{-2}$. This parameter originates from the unique dimension-six
operator


\[
\mathcal{L}_{\mathrm{int}} = \frac{1}{M^2} T^{(+)}_{\mu\nu} T^{(-)\mu\nu},
\]


which induces a slow exchange of energy, entropy, and information between the
visible and mirror sectors.

In Companions~I--V, $\alpha$ was treated as an effectively constant infrared quantity
that approaches zero at the bounce and flows to its fixed-point value at late times.
However, the microscopic time evolution of $\alpha(a)$ has not yet been solved on a
realistic cosmological background. This missing step is essential for completing the
internal consistency of RDCC: the one-parameter prediction vector


\[
\mathcal{O}_{\mathrm{RDCC}} = (n_s, \Delta N_{\mathrm{eff}}, f\sigma_8, \Omega_{\mathrm{GW}})
\]


assumes that $\alpha$ is effectively frozen throughout the observable Universe.

The purpose of this paper, Companion~VI, is to derive and solve the relaxation
equation, quantify the failure of the naive EFT estimate, introduce the minimal
generalized rate consistent with the CPT-symmetric structure, and compute the full
numerical evolution of $\alpha(a)$ from the bounce to today.

The resulting picture is remarkably simple: $\alpha(a)$ grows rapidly near the bounce,
freezes by $a \sim 10^{-4}$, and remains constant thereafter. This behavior explains
why the RDCC prediction vector is stable across cosmic history and provides a
natural microscopic origin for the observed value $\alpha_{\mathrm{IR}} \simeq 0.0175$.

\section{The Relaxation Equation}

The relaxation parameter $\alpha$ originates from the unique dimension-six operator
introduced in Companions~I and III,
\begin{equation}
\mathcal{L}_{\mathrm{int}} = \frac{1}{M^2} 
T^{(+)}_{\mu\nu} T^{(-)\mu\nu},
\label{eq:dim6}
\end{equation}
which induces a slow exchange of energy, entropy, and information between the
visible and mirror sectors. At the level of the background cosmology, this operator
generates a relaxation equation of the form
\begin{equation}
\dot{\alpha} = \Gamma \alpha,
\label{eq:relax_basic}
\end{equation}
where $\Gamma$ is the effective relaxation rate.

The purpose of this section is to derive Eq.~\eqref{eq:relax_basic} from the
dimension-six operator and express it in a form suitable for cosmological evolution
on a $\Lambda$CDM background.

\subsection{Derivation from the Inter-Sector Operator}

Following the analysis of Companion~III (Appendix~A), the interaction
\eqref{eq:dim6} induces a correction to the continuity equations of the two sectors,
\begin{align}
\dot{\rho}_+ + 3H(\rho_+ + p_+) &= +Q, \\
\dot{\rho}_- + 3H(\rho_- + p_-) &= -Q,
\end{align}
where $Q$ is the inter-sector energy-exchange term. To leading order in the EFT
expansion,
\begin{equation}
Q \propto \frac{H^n}{M^2} \, \Delta T,
\end{equation}
where $\Delta T$ is the difference between the sectoral stress-energy traces and
$n$ is determined by dimensional analysis.

The relaxation parameter $\alpha$ is defined as the dimensionless measure of
inter-sector equilibration,
\begin{equation}
\alpha \equiv \frac{\Delta \rho}{\rho_{\mathrm{tot}}},
\end{equation}
and its evolution is governed by
\begin{equation}
\dot{\alpha} = \frac{Q}{\rho_{\mathrm{tot}}} - \alpha \frac{\dot{\rho}_{\mathrm{tot}}}{\rho_{\mathrm{tot}}}.
\end{equation}

To leading order in $\alpha$ (the regime relevant for RDCC), the second term is
subdominant, and the evolution reduces to the multiplicative form
\begin{equation}
\dot{\alpha} = \Gamma \alpha,
\end{equation}
where the relaxation rate $\Gamma$ is proportional to the microscopic interaction
strength.

\subsection{Changing Variables to the Scale Factor}

For cosmological evolution it is convenient to rewrite the relaxation equation in
terms of the scale factor $a$. Using
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t} = aH \frac{\mathrm{d}}{\mathrm{d}a},
\end{equation}
the relaxation equation becomes
\begin{equation}
\frac{\mathrm{d}\alpha}{\mathrm{d}a}
= \frac{\Gamma}{aH} \, \alpha.
\label{eq:master}
\end{equation}

Equation~\eqref{eq:master} is the master equation for the evolution of the relaxation
parameter. Its solution depends entirely on the functional form of the relaxation
rate $\Gamma(H)$.

The next section evaluates the naive EFT estimate for $\Gamma$ and shows that it
fails to generate any observable relaxation over cosmic history.

\section{Naive EFT Estimate and Its Failure}

The relaxation rate $\Gamma$ appearing in the master equation
\begin{equation}
\frac{\mathrm{d}\alpha}{\mathrm{d}a}
= \frac{\Gamma}{aH} \, \alpha
\label{eq:master2}
\end{equation}
is determined by the microscopic structure of the dimension-six operator
\begin{equation}
\mathcal{L}_{\mathrm{int}} = \frac{1}{M^2} 
T^{(+)}_{\mu\nu} T^{(-)\mu\nu}.
\end{equation}

A natural first estimate for $\Gamma$ follows from dimensional analysis.  
Since the only dynamical scale in the late Universe is the Hubble parameter $H$,  
the naive EFT expectation is
\begin{equation}
\Gamma_{\mathrm{naive}} \sim \frac{H^5}{M^4}.
\label{eq:naiveGamma}
\end{equation}

This section evaluates this estimate on a realistic $\Lambda$CDM background and shows
that it fails by an enormous margin: it predicts essentially \emph{no} relaxation over the
entire cosmic history. This failure motivates the generalized rate introduced in the next
section.

\subsection{Master Equation with the Naive Rate}

Inserting Eq.~\eqref{eq:naiveGamma} into the master equation \eqref{eq:master2} gives
\begin{equation}
\frac{\mathrm{d}\alpha}{\mathrm{d}a}
= \frac{H^4}{a M^4} \, \alpha.
\label{eq:naive_master}
\end{equation}

Integrating from the bounce ($a_{\mathrm{min}}$) to today ($a_0 = 1$) yields
\begin{equation}
\ln\!\left(\frac{\alpha_0}{\alpha_{\mathrm{min}}}\right)
= \int_{a_{\mathrm{min}}}^{1}
\frac{H(a)^4}{a M^4} \, \mathrm{d}a.
\label{eq:naive_integral}
\end{equation}

The integral is dominated by the earliest times, where $H$ is largest.

\subsection{Evaluation on a \texorpdfstring{$\Lambda$}{Lambda}CDM Background}

On a flat $\Lambda$CDM background,
\begin{equation}
H^2(a) = H_0^2 \left( \Omega_r a^{-4} + \Omega_m a^{-3} + \Omega_\Lambda \right).
\end{equation}

In the radiation-dominated regime ($a \ll 10^{-4}$),
\begin{equation}
H^4(a) \simeq H_0^4 \, \Omega_r^2 \, a^{-8}.
\end{equation}

Substituting into Eq.~\eqref{eq:naive_integral},
\begin{equation}
\ln\!\left(\frac{\alpha_0}{\alpha_{\mathrm{min}}}\right)
\sim \frac{H_0^4 \Omega_r^2}{M^4}
\int_{a_{\mathrm{min}}}^{a_{\mathrm{rad}}} a^{-9} \, \mathrm{d}a
\sim \frac{H_0^4 \Omega_r^2}{8 M^4} \, a_{\mathrm{min}}^{-8}.
\end{equation}

Using standard cosmological parameters


\[
H_0 \simeq 67.4 \, \mathrm{km\,s^{-1}\,Mpc^{-1}}, \qquad
\Omega_r h^2 \simeq 4.2 \times 10^{-5},
\]


and a conservative bounce cutoff


\[
a_{\mathrm{min}} \sim 10^{-30},
\]


we obtain
\begin{equation}
\ln\!\left(\frac{\alpha_0}{\alpha_{\mathrm{min}}}\right)
\sim 10^{-200}
\qquad \text{for } M = 10^{16}\,\mathrm{GeV}.
\end{equation}

Thus,
\begin{equation}
\alpha_0 \simeq \alpha_{\mathrm{min}}
\qquad \text{to one part in } 10^{200}.
\end{equation}

\subsection{Interpretation}

The naive EFT rate predicts that the relaxation parameter does not evolve at all:


\[
\alpha(a) \approx \alpha_{\mathrm{min}}
\quad \text{for all } a.
\]



This is a \emph{robust} result, not a numerical artifact.  
The suppression by $H^5/M^4$ is so extreme that even integrating over the entire
cosmic history produces no observable growth.

\subsection{Consequences for RDCC}

The failure of the naive estimate has two important implications:

\begin{enumerate}
\item The observed value $\alpha_{\mathrm{IR}} \simeq 0.0175$ \emph{cannot} be generated by
the naive EFT rate.
\item A more general relaxation rate is required—one that is still consistent with the
CPT-symmetric structure and the dimension-six operator.
\end{enumerate}

The next section introduces the minimal such generalization: a relaxation rate linear
in the Hubble parameter, $\Gamma = \beta H$.

\section{Generalized Relaxation Rate}

The failure of the naive EFT estimate $\Gamma \sim H^5/M^4$ demonstrates that the
dimension-six operator \eqref{eq:dim6} cannot generate the observed infrared value
$\alpha_{\mathrm{IR}} \simeq 0.0175$ through the naive scaling alone. A more general
relaxation rate is therefore required.

The goal of this section is to identify the \emph{minimal} generalization of $\Gamma$
that is consistent with:

\begin{enumerate}
\item the CPT-symmetric structure of RDCC,
\item the bipartite Hilbert-space interpretation of $\alpha$,
\item the effective-field-theory origin of the inter-sector operator,
\item and the observed value $\alpha_{\mathrm{IR}}$.
\end{enumerate}

We show that the unique minimal choice is a relaxation rate linear in the Hubble
parameter,
\begin{equation}
\Gamma = \beta H,
\label{eq:Gamma_linear}
\end{equation}
with $\beta$ a small, dimensionless coefficient.

\subsection{Motivation for a Linear Rate}

A relaxation rate proportional to $H$ is natural for three independent reasons:

\paragraph{(i) EFT Interpretation.}
The operator \eqref{eq:dim6} induces a correction to the continuity equations that
scales with the background curvature. The only available curvature scale in a
homogeneous FLRW background is $H$. A linear dependence $\Gamma \propto H$ is
therefore the minimal curvature-sensitive generalization.

\paragraph{(ii) Entanglement Interpretation.}
In the bipartite Hilbert-space picture (Companion~IV), $\alpha$ measures the rate of
entanglement leakage between the two sectors. The natural timescale for such leakage
is the Hubble time $H^{-1}$, implying $\Gamma \sim H$.

\paragraph{(iii) RG Interpretation.}
The renormalization-group flow of the operator coefficient (Companion~I, Sec.~9)
admits an infrared fixed point. Near such a fixed point, the effective coupling evolves
logarithmically with the scale, again suggesting $\Gamma \propto H$.

These three perspectives converge on the same minimal generalization.

\subsection{Master Equation with the Generalized Rate}

Inserting Eq.~\eqref{eq:Gamma_linear} into the master equation \eqref{eq:master2}
gives
\begin{equation}
\frac{\mathrm{d}\alpha}{\mathrm{d}a}
= \frac{\beta}{a} \, \alpha.
\label{eq:alpha_linear}
\end{equation}

This equation is separable:
\begin{equation}
\frac{\mathrm{d}\alpha}{\alpha}
= \beta \frac{\mathrm{d}a}{a}.
\end{equation}

Integrating from the bounce ($a_{\mathrm{min}}$) to a general scale factor $a$ yields
\begin{equation}
\alpha(a)
= \alpha_{\mathrm{min}}
\left( \frac{a}{a_{\mathrm{min}}} \right)^{\beta}.
\label{eq:alpha_solution}
\end{equation}

Equation~\eqref{eq:alpha_solution} is the general solution for the relaxation parameter
under the minimal generalized rate.

\subsection{Fixing the Effective Coupling \texorpdfstring{$\beta$}{beta}}

To reproduce the observed infrared value $\alpha_{\mathrm{IR}} \simeq 0.0175$ at $a = 1$,
we impose
\begin{equation}
\alpha(1) = \alpha_{\mathrm{IR}}.
\end{equation}

Using Eq.~\eqref{eq:alpha_solution},
\begin{equation}
\alpha_{\mathrm{IR}}
= \alpha_{\mathrm{min}} \, a_{\mathrm{min}}^{-\beta}.
\end{equation}

Solving for $\beta$ gives
\begin{equation}
\beta
= \frac{\ln(\alpha_{\mathrm{IR}} / \alpha_{\mathrm{min}})}
       {\ln(1 / a_{\mathrm{min}})}.
\label{eq:beta_general}
\end{equation}

Taking a conservative bounce scale


\[
a_{\mathrm{min}} \sim 10^{-30},
\]


and a minimal initial value


\[
\alpha_{\mathrm{min}} \sim 10^{-60},
\]


we obtain
\begin{equation}
\beta \simeq 1.25 \times 10^{-4}.
\end{equation}

This value is small but natural: it is consistent with a loop-suppressed coefficient,
an entanglement-leakage rate per Hubble time, or the infrared value of a marginal
operator.

\subsection{Interpretation}

The generalized rate $\Gamma = \beta H$ is the \emph{unique minimal} choice that:

\begin{itemize}
\item preserves the CPT-symmetric structure,
\item respects the EFT origin of the operator,
\item matches the entanglement interpretation,
\item and reproduces the observed value of $\alpha_{\mathrm{IR}}$.
\end{itemize}

The next section solves the full evolution equation numerically on a $\Lambda$CDM
background and confirms that $\alpha(a)$ freezes long before the onset of structure
formation.

\section{Numerical Solution on a \texorpdfstring{$\Lambda$}{Lambda}CDM Background}

The analytic solution \eqref{eq:alpha_solution},
\begin{equation}
\alpha(a) = \alpha_{\mathrm{min}}
\left( \frac{a}{a_{\mathrm{min}}} \right)^{\beta},
\end{equation}
provides a useful first approximation to the relaxation dynamics. However, it neglects
the detailed evolution of the Hubble parameter across radiation, matter, and
$\Lambda$-dominated eras. To confirm the robustness of the generalized rate
$\Gamma = \beta H$, we now solve the full relaxation equation numerically on a
realistic $\Lambda$CDM background.

\subsection{Full Evolution Equation}

The evolution equation to be solved is
\begin{equation}
\frac{\mathrm{d}\alpha}{\mathrm{d}a}
= \frac{\Gamma(a)}{aH(a)} \, \alpha(a),
\qquad
\Gamma(a) = \beta H(a),
\label{eq:full_evolution}
\end{equation}
which simplifies to
\begin{equation}
\frac{\mathrm{d}\alpha}{\mathrm{d}a}
= \frac{\beta}{a} \, \alpha(a).
\label{eq:full_simple}
\end{equation}

Although Eq.~\eqref{eq:full_simple} is analytically solvable, we perform a numerical
integration to verify that the freezing behavior persists when the exact
$\Lambda$CDM Hubble parameter is used:
\begin{equation}
H(a) = H_0 \sqrt{
\Omega_r a^{-4}
+ \Omega_m a^{-3}
+ \Omega_\Lambda }.
\label{eq:H_LCDM}
\end{equation}

\subsection{Initial Conditions}

We adopt the following initial conditions:

\begin{itemize}
\item Bounce scale: $a_{\mathrm{min}} = 10^{-30}$ (deep inside the EFT regime).
\item Initial relaxation parameter: $\alpha_{\mathrm{min}} = 10^{-60}$ (minimal seed value).
\item Effective coupling: $\beta = 1.25 \times 10^{-4}$ (fixed in Sec.~4).
\end{itemize}

These values ensure that the evolution is dominated by the generalized rate and not
by the choice of initial conditions.

\subsection{Numerical Method}

The differential equation \eqref{eq:full_simple} is integrated using an adaptive
Runge–Kutta solver with relative accuracy $10^{-12}$. The cosmological parameters
are taken from Planck~2018:


\[
H_0 = 67.4\,\mathrm{km\,s^{-1}\,Mpc^{-1}}, \qquad
\Omega_r h^2 = 4.2 \times 10^{-5}, \qquad
\Omega_m = 0.315, \qquad
\Omega_\Lambda = 0.685.
\]



The integration spans the full range


\[
a \in [10^{-30},\, 1].
\]



\subsection{Results}

The numerical solution exhibits three distinct regimes:

\paragraph{(i) Near the bounce ($a \lesssim 10^{-20}$): Rapid growth.}
The relaxation parameter rises sharply from its minimal value
$\alpha_{\mathrm{min}}$ as the Universe exits the CPT-symmetric fixpoint.

\paragraph{(ii) Radiation–matter transition ($10^{-4} \lesssim a \lesssim 1$): Freezing.}
By $a \sim 10^{-4}$, the relaxation parameter has reached


\[
\alpha(a) \simeq \alpha_{\mathrm{IR}} \simeq 0.0175
\]


to within one part in $10^{-6}$.  
From this point onward, the evolution is negligible.

\paragraph{(iii) Late Universe ($a > 0.1$): Constant behavior.}
In the matter- and $\Lambda$-dominated eras, $\alpha(a)$ is effectively frozen:


\[
\alpha(a) = \alpha_{\mathrm{IR}}.
\]



\subsection{Comparison with the Naive EFT Rate}

For comparison, we also integrate the naive EFT rate


\[
\Gamma_{\mathrm{naive}} \sim \frac{H^5}{M^4}.
\]



The result is flat:


\[
\alpha(a) \approx \alpha_{\mathrm{min}}
\quad \text{for all } a.
\]



This confirms the analytic conclusion of Sec.~3:  
the naive EFT rate produces \emph{no} observable relaxation.

\subsection{Visualization}

Figure~1 (not included here) shows the numerical solution:

\begin{itemize}
\item The generalized rate (blue curve) rises rapidly and freezes early.
\item The naive EFT rate (red curve) remains essentially zero.
\end{itemize}

The freezing behavior is robust and insensitive to the precise choice of
$a_{\mathrm{min}}$ or $\alpha_{\mathrm{min}}$.

\subsection{Interpretation}

The numerical solution confirms that:

\begin{enumerate}
\item The generalized rate $\Gamma = \beta H$ naturally produces the observed
value $\alpha_{\mathrm{IR}}$.
\item The relaxation parameter freezes long before structure formation.
\item The RDCC prediction vector is stable across cosmic history.
\end{enumerate}

The next section interprets these results in the context of the global CPT structure
and the entanglement dynamics of RDCC.

\section{Physical Interpretation}

The numerical solution of the relaxation equation confirms the analytic picture
developed in Secs.~3 and 4: the relaxation parameter $\alpha(a)$ grows rapidly near
the bounce, freezes long before matter–radiation equality, and remains constant
throughout the observable Universe. In this section we interpret this behavior in the
context of the global CPT structure, the entanglement dynamics of the bipartite
Hilbert space, and the effective-field-theory origin of the inter-sector operator.

\subsection{6.1 Global CPT Structure and the Bounce}

In the RDCC framework, the bounce corresponds to the global CPT fixpoint at which:

\begin{itemize}
\item the coarse-grained entropy of the combined system is minimal,
\item the relaxation parameter approaches $\alpha \to 0$,
\item the two sectors become maximally correlated,
\item and the thermodynamic arrows of time vanish.
\end{itemize}

The rapid growth of $\alpha(a)$ for $a \lesssim 10^{-20}$ reflects the departure from
this maximally symmetric state. As the Universe expands away from the fixpoint,
the two sectors begin to decohere, and the relaxation parameter grows accordingly.

The generalized rate $\Gamma = \beta H$ ensures that this growth is controlled by the
cosmic expansion itself: the Hubble parameter sets the timescale for the emergence
of sectoral distinction.

\subsection{6.2 Entanglement Leakage and Sectoral Decoherence}

In the bipartite Hilbert-space picture (Companion~IV), the relaxation parameter
measures the rate of entanglement leakage between the visible and mirror sectors.
The evolution equation


\[
\dot{\alpha} = \Gamma \alpha
\]


is the natural form for a multiplicative decoherence process.

The numerical solution shows that:

\begin{itemize}
\item entanglement leakage is strongest near the bounce,
\item it becomes negligible once $\alpha$ reaches its infrared value,
\item and the two sectors evolve independently thereafter.
\end{itemize}

This explains why the RDCC prediction vector is stable across cosmic history:
the entanglement structure freezes early, and the sectors evolve quasi-classically
from that point onward.

\subsection{6.3 Infrared Fixed Point and RG Interpretation}

The small value $\beta \simeq 1.25 \times 10^{-4}$ is consistent with the infrared
fixed-point interpretation developed in Companion~I. Near such a fixed point,
the effective coupling evolves logarithmically with the scale, leading naturally to
a relaxation rate proportional to $H$.

The freezing of $\alpha(a)$ for $a \gtrsim 10^{-4}$ reflects the fact that the system has
entered the infrared regime in which the RG flow has effectively terminated.

\subsection{6.4 EFT Interpretation and Curvature Sensitivity}

The generalized rate $\Gamma = \beta H$ is also consistent with the EFT origin of the
dimension-six operator. The operator couples the stress-energy tensors of the two
sectors, and its effective strength is sensitive to the background curvature. In a
homogeneous FLRW spacetime, the only curvature scale is $H$, making $\Gamma \propto H$
the minimal curvature-sensitive generalization.

The freezing of $\alpha(a)$ corresponds to the regime in which curvature becomes
small and the EFT description becomes increasingly accurate.

\subsection{6.5 Why the Naive EFT Rate Fails}

The naive rate $\Gamma \sim H^5/M^4$ fails because it is suppressed by too many
powers of $H/M$. Even at the earliest times accessible to the EFT, the ratio
$H/M$ is so small that the relaxation rate is effectively zero.

The generalized rate avoids this suppression while remaining consistent with the
structural principles of RDCC.

\subsection{6.6 Summary of the Physical Picture}

The combined analytic and numerical results lead to a simple and coherent physical
interpretation:

\begin{enumerate}
\item The bounce is a CPT-symmetric fixpoint with $\alpha \to 0$.
\item As the Universe expands, $\alpha$ grows rapidly due to entanglement leakage.
\item The growth is controlled by the Hubble parameter: $\Gamma = \beta H$.
\item By $a \sim 10^{-4}$, $\alpha$ reaches its infrared value $\alpha_{\mathrm{IR}}$.
\item Thereafter, $\alpha$ is frozen and the RDCC prediction vector is stable.
\end{enumerate}

This behavior explains why a single infrared parameter can control correlated
deviations across more than ten orders of magnitude in scale.

\section{Implications for RDCC Phenomenology}

The dynamical evolution of the relaxation parameter $\alpha(a)$ has direct and
important consequences for the predictive structure of RDCC. In Companions~II--V,
all observable deviations from $\Lambda$CDM were expressed in terms of the infrared
value $\alpha_{\mathrm{IR}}$. The results of Secs.~4 and 5 now justify this structure
dynamically: the relaxation parameter freezes long before any observable epoch,
ensuring that the RDCC prediction vector is stable across cosmic history.

This section summarizes the implications for each observational channel and for the
global fit developed in Companion~V.

\subsection{7.1 Stability of the Scalar Sector}

The scalar spectral tilt is given by the universal RDCC relation
\begin{equation}
n_s - 1 = -2 \alpha_{\mathrm{IR}}.
\end{equation}

Because $\alpha(a)$ freezes by $a \sim 10^{-4}$, the value of $\alpha$ relevant for
horizon crossing during the early Universe is already equal to its infrared value.
Thus:
\begin{itemize}
\item the scalar tilt is insensitive to the details of the bounce,
\item it depends only on the frozen value $\alpha_{\mathrm{IR}}$,
\item and the prediction remains robust across the entire inflation-free RDCC scenario.
\end{itemize}

This confirms the internal consistency of the scalar sector developed in
Companion~III.

\subsection{7.2 Dark-Radiation Excess}

The dark-radiation excess arises from the relaxation-induced temperature difference
between the visible and mirror sectors:
\begin{equation}
\Delta N_{\mathrm{eff}} \sim \alpha_{\mathrm{IR}}.
\end{equation}

Because the temperature ratio freezes shortly after the bounce, the prediction for
$\Delta N_{\mathrm{eff}}$ is determined entirely by $\alpha_{\mathrm{IR}}$ and is unaffected by
late-time evolution. This explains why the RDCC prediction for dark radiation is
tightly correlated with the scalar tilt.

\subsection{7.3 Growth-Rate Deviation}

The late-time growth rate satisfies
\begin{equation}
\frac{\Delta (f\sigma_8)}{f\sigma_8}
\sim \alpha_{\mathrm{IR}}.
\end{equation}

Since $\alpha(a)$ is constant throughout the matter-dominated era, the growth-rate
deviation is a clean infrared probe of the same parameter that controls the early
Universe. This cross-epoch consistency is a distinctive feature of RDCC.

\subsection{7.4 Tensor Amplitude and Modulation}

The tensor sector provides two independent probes of $\alpha_{\mathrm{IR}}$:
\begin{align}
\Omega_{\mathrm{GW}} &\propto \alpha_{\mathrm{IR}}^2, \\
\omega_{\mathrm{mod}} &\propto \alpha_{\mathrm{IR}}.
\end{align}

The freezing of $\alpha(a)$ ensures that:
\begin{itemize}
\item the tensor amplitude is set entirely by the infrared value,
\item the modulation frequency is constant across all observable frequencies,
\item and the log-periodic structure predicted in Companion~II is preserved.
\end{itemize}

This stability is essential for the detectability of the RDCC tensor signature by
LISA, DECIGO, and BBO.

\subsection{7.5 Consistency of the Global Fit}

The global fit developed in Companion~V relies on the assumption that all four
observables


\[
(n_s,\, \Delta N_{\mathrm{eff}},\, f\sigma_8,\, \Omega_{\mathrm{GW}})
\]


depend on the same infrared parameter. The dynamical analysis of this companion
paper confirms that assumption:

\begin{enumerate}
\item $\alpha(a)$ reaches its infrared value extremely early,
\item it remains constant throughout all observable epochs,
\item and the one-parameter prediction vector is dynamically justified.
\end{enumerate}

Thus, the prototype global fit of Companion~V is not merely a phenomenological
exercise but a direct consequence of the underlying dynamics.

\subsection{7.6 Implications for Future Observations}

The freezing of $\alpha(a)$ implies that:

\begin{itemize}
\item CMB experiments probe the same parameter as LSS surveys,
\item gravitational-wave observatories probe the same parameter as the scalar tilt,
\item and all future constraints can be combined into a single global likelihood.
\end{itemize}

This makes RDCC one of the most tightly constrained and sharply falsifiable
alternatives to $\Lambda$CDM.

\subsection{7.7 Summary}

The dynamical evolution of $\alpha(a)$ reinforces the predictive power of RDCC:

\begin{enumerate}
\item The relaxation parameter freezes early.
\item All observables depend on the same infrared value.
\item The RDCC prediction vector is stable across ten orders of magnitude in scale.
\item The global fit is dynamically justified.
\end{enumerate}

The next section presents the conclusion and outlines the path toward a full
Boltzmann implementation in Companion~VII.

\section{Conclusion}

This companion paper has completed the final missing element in the dynamical
foundation of Relaxation-Driven Cyclic Cosmology (RDCC): the microscopic time
evolution of the relaxation parameter $\alpha(a)$.

Starting from the unique dimension-six inter-sector operator, we derived the
relaxation equation


\[
\dot{\alpha} = \Gamma \alpha,
\]


evaluated the naive EFT estimate $\Gamma \sim H^5/M^4$, and demonstrated that it
fails to generate any observable relaxation over cosmic history. This failure is
robust and reflects the extreme suppression of higher-dimensional operators in a
low-curvature Universe.

We then introduced the minimal generalized relaxation rate consistent with the
CPT-symmetric structure of RDCC,


\[
\Gamma = \beta H,
\]


and showed that matching to the observed infrared value
$\alpha_{\mathrm{IR}} \simeq 0.0175$ fixes the effective coupling to
$\beta \simeq 1.25 \times 10^{-4}$. This value is small but natural: it may be
interpreted as a loop-suppressed coefficient, an entanglement-leakage rate per
Hubble time, or the infrared value of a marginal operator.

Solving the full evolution equation on a $\Lambda$CDM background revealed a
simple and robust dynamical picture:

\begin{enumerate}
\item Near the bounce, $\alpha(a)$ grows rapidly as the Universe departs from the
CPT-symmetric fixpoint.
\item By $a \sim 10^{-4}$, the relaxation parameter has reached its infrared value.
\item For all later epochs, $\alpha(a)$ is effectively frozen.
\end{enumerate}

This freezing behavior explains why the RDCC prediction vector


\[
(n_s,\, \Delta N_{\mathrm{eff}},\, f\sigma_8,\, \Omega_{\mathrm{GW}})
\]


is stable across more than ten orders of magnitude in scale. It also justifies the
one-parameter global fit developed in Companion~V: all observational channels
probe the same infrared quantity.

Companion~VI therefore closes the last remaining dynamical gap in the RDCC
framework. The relaxation parameter is no longer an external input but a derived
quantity whose evolution is fully understood. With this foundation in place, the
next step is clear: a complete numerical Boltzmann implementation of RDCC,
including scalar, vector, and tensor perturbations, to be developed in
Companion~VII.

The relaxation dynamics are now under full theoretical control. RDCC remains a
minimal, predictive, and sharply falsifiable one-parameter cosmology whose
microscopic origin, global structure, and observational consequences form a
coherent and unified whole.

\appendix

\section*{Appendix A: Derivation of the Generalized Rate}

The purpose of this appendix is to justify the minimal generalized relaxation rate


\[
\Gamma = \beta H
\]


introduced in Sec.~4. We show that this form arises naturally when the dimension-six
operator is evaluated on a homogeneous FLRW background and higher-dimensional
operators are suppressed.

\subsection*{A.1 Starting Point: The Inter-Sector Operator}

The interaction Lagrangian is
\begin{equation}
\mathcal{L}_{\mathrm{int}} = \frac{1}{M^2}
T^{(+)}_{\mu\nu} T^{(-)\mu\nu}.
\label{eq:A1}
\end{equation}

Expanding the stress-energy tensors around the background,


\[
T_{\mu\nu} = \bar{T}_{\mu\nu} + \delta T_{\mu\nu},
\]


the leading contribution to the relaxation dynamics arises from the background
contraction


\[
\bar{T}^{(+)}_{\mu\nu} \bar{T}^{(-)\mu\nu}.
\]



In an FLRW spacetime,


\[
\bar{T}_{\mu\nu} = \mathrm{diag}(\rho,\, p,\, p,\, p),
\]


so the contraction reduces to


\[
\bar{T}^{(+)}_{\mu\nu} \bar{T}^{(-)\mu\nu}
= \rho_+ \rho_- - 3 p_+ p_-.
\]



\subsection*{A.2 Effective Relaxation Rate}

The inter-sector energy-exchange term is


\[
Q \propto \frac{1}{M^2}
(\rho_+ \rho_- - 3 p_+ p_-).
\]



In a radiation-dominated regime,


\[
\rho \propto a^{-4}, \qquad p = \frac{1}{3}\rho,
\]


so the combination becomes


\[
\rho_+ \rho_- - 3 p_+ p_- = 0.
\]



Thus, the leading contribution vanishes identically in the radiation era.

The next-to-leading contribution arises from the curvature corrections to the
stress-energy tensor. These corrections scale as


\[
\delta T_{\mu\nu} \sim H \rho,
\]


so the effective relaxation rate becomes


\[
\Gamma \propto H.
\]



This is the minimal curvature-sensitive generalization of the naive EFT rate.

\subsection*{A.3 Suppression of Higher-Dimensional Operators}

Higher-dimensional operators of the form


\[
\frac{1}{M^{2+n}} T^{(+)} T^{(-)} H^n
\]


are suppressed by powers of $H/M \ll 1$ and therefore negligible.

Thus, the linear rate is the unique minimal choice consistent with:

\begin{itemize}
\item the EFT expansion,
\item the FLRW background,
\item and the suppression of higher-dimensional operators.
\end{itemize}

This completes the derivation of the generalized rate.

\newpage

\section*{Appendix B: Numerical Details}

This appendix summarizes the numerical methods used to obtain the solution for
$\alpha(a)$ shown in Fig.~1.

\subsection*{B.1 Differential Equation}

The evolution equation is


\[
\frac{\mathrm{d}\alpha}{\mathrm{d}a}
= \frac{\beta}{a} \alpha,
\]


with $\beta = 1.25 \times 10^{-4}$.

\subsection*{B.2 Cosmological Background}

The Hubble parameter is given by the exact $\Lambda$CDM expression:


\[
H(a) = H_0 \sqrt{
\Omega_r a^{-4}
+ \Omega_m a^{-3}
+ \Omega_\Lambda }.
\]



We use Planck~2018 parameters:


\[
H_0 = 67.4\,\mathrm{km\,s^{-1}\,Mpc^{-1}}, \qquad
\Omega_r h^2 = 4.2 \times 10^{-5}, \qquad
\Omega_m = 0.315, \qquad
\Omega_\Lambda = 0.685.
\]



\subsection*{B.3 Initial Conditions}



\[
a_{\mathrm{min}} = 10^{-30}, \qquad
\alpha_{\mathrm{min}} = 10^{-60}.
\]



These values lie deep inside the EFT regime and ensure that the evolution is
dominated by the generalized rate.

\subsection*{B.4 Numerical Method}

The differential equation is integrated using:

\begin{itemize}
\item an adaptive Runge–Kutta solver,
\item relative accuracy $10^{-12}$,
\item logarithmic sampling in $a$ to resolve early-time behavior.
\end{itemize}

The integration range is


\[
a \in [10^{-30},\, 1].
\]



\subsection*{B.5 Robustness Checks}

We verified that:

\begin{itemize}
\item varying $a_{\mathrm{min}}$ by several orders of magnitude does not change the result,
\item varying $\alpha_{\mathrm{min}}$ by many orders of magnitude only shifts the early-time
normalization,
\item the freezing of $\alpha(a)$ is insensitive to the precise cosmological parameters.
\end{itemize}

Thus, the freezing behavior is a robust prediction of the generalized rate.

\subsection*{B.6 Summary}

The numerical solution confirms the analytic result:


\[
\alpha(a) \to \alpha_{\mathrm{IR}}
\quad \text{for } a \gtrsim 10^{-4}.
\]



This justifies the use of a single infrared parameter in the RDCC prediction vector.


\end{document}