\documentclass[11pt,a4paper]{article}

% ---------------------------------------------------------
% Encoding, Fonts, Language
% ---------------------------------------------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{lmodern}

% ---------------------------------------------------------
% Math Packages
% ---------------------------------------------------------
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{mathtools}
\usepackage{bm}            % bold math symbols
\usepackage{physics}       % \dv, \pdv, \ket, \bra, etc.
\usepackage{tensor}        % tensor indices
\usepackage{siunitx}       % units

% ---------------------------------------------------------
% Figures, Tables, Layout
% ---------------------------------------------------------
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{geometry}
\geometry{margin=1in}

% ---------------------------------------------------------
% Bibliography
% ---------------------------------------------------------

\usepackage[numbers,sort&compress]{natbib}


% ---------------------------------------------------------
% Hyperref (must be loaded LAST except cleveref)
% ---------------------------------------------------------
\usepackage[
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue,
    pdfencoding=auto,
    psdextra
]{hyperref}

% ---------------------------------------------------------
% Cleveref (after hyperref)
% ---------------------------------------------------------
\usepackage{cleveref}

% ---------------------------------------------------------
% Fix for math in section titles (renamed to avoid conflicts)
% ---------------------------------------------------------
\newcommand{\mathsafe}[2]{\texorpdfstring{#1}{#2}}

% ---------------------------------------------------------
% Paragraph spacing
% ---------------------------------------------------------
\setlength{\parskip}{1em}
\setlength{\parindent}{0pt}

% ---------------------------------------------------------
% Title
% ---------------------------------------------------------
\title{\textbf{Boltzmann Hierarchy and Linear Perturbations in\\
Relaxation-Driven Cyclic Cosmology (RDCC)}}

\author{Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}}

\date{May 2026}

\begin{document}

\maketitle


\begin{abstract}
Relaxation-Driven Cyclic Cosmology (RDCC) is a one-parameter, CPT-symmetric
framework in which all observable deviations from $\Lambda$CDM arise from the
infrared relaxation strength $\alpha_{\mathrm{IR}}$. Companions~I--VI established the
gravitational foundations, tensor phenomenology, relaxation dynamics, global CPT
structure, observational synthesis, and the microscopic evolution of $\alpha(a)$.
The resulting framework is dynamically complete and empirically predictive.

The present paper, Companion~VII, develops the full linear perturbation theory of
RDCC and constructs the Boltzmann hierarchy for scalar, vector, and tensor modes
in both the visible and mirror sectors. We derive the modified Einstein equations,
the coupled continuity and Euler equations, the photon and neutrino Boltzmann
hierarchies, and the tensor-mode evolution including the RDCC modulation
mechanism. The relaxation parameter enters the perturbation system through a
modified Poisson equation, an effective gravitational coupling, and a small
inter-sector energy-exchange term.

We show that the freezing of $\alpha(a)$ established in Companion~VI implies that
all perturbative corrections are controlled by the single constant parameter
$\alpha_{\mathrm{IR}}$. This leads to a unified one-parameter modification of the CMB
anisotropies, matter power spectrum, and stochastic gravitational-wave background.
We provide the complete set of equations required for numerical implementation in
Boltzmann codes such as CLASS or CAMB, thereby enabling a full confrontation of
RDCC with precision cosmological data.

Companion~VII completes the predictive architecture of RDCC and establishes the
framework as a fully testable alternative to inflationary $\Lambda$CDM.
\end{abstract}

\newpage

\tableofcontents

\newpage

\section{Introduction}

Relaxation-Driven Cyclic Cosmology (RDCC) is a minimal, CPT-symmetric framework
in which all observable deviations from $\Lambda$CDM arise from a single infrared
parameter, the relaxation strength $\alpha_{\mathrm{IR}} \sim 10^{-2}$. Companions~I--VI
developed the geometric foundations, tensor phenomenology, relaxation dynamics,
global CPT structure, observational synthesis, and the microscopic evolution of
$\alpha(a)$. The resulting picture is simple and robust: the relaxation parameter
grows rapidly near the bounce, freezes by $a \sim 10^{-4}$, and remains constant
throughout all observable epochs.

This freezing behavior implies that the RDCC prediction vector


\[
(n_s,\, \Delta N_{\mathrm{eff}},\, f\sigma_8,\, \Omega_{\mathrm{GW}})
\]


is controlled entirely by the single infrared value $\alpha_{\mathrm{IR}}$. Companion~V
demonstrated that this one-parameter structure is compatible with current CMB,
LSS, and GW data, and Companion~VI provided the dynamical justification for the
freeze-out of $\alpha(a)$.

The purpose of the present paper, Companion~VII, is to develop the full linear
perturbation theory of RDCC and construct the Boltzmann hierarchy for scalar,
vector, and tensor modes. This is the final step required to make RDCC a fully
testable cosmological framework. The Boltzmann hierarchy determines the evolution
of:

\begin{itemize}
\item photon temperature and polarization multipoles,
\item neutrino multipoles,
\item baryon and CDM density and velocity perturbations,
\item mirror-sector analogues of all above quantities,
\item and the tensor-mode spectrum including RDCC modulation.
\end{itemize}

The relaxation parameter enters the perturbation system in three ways:

\begin{enumerate}
\item a modified Poisson equation with an effective gravitational coupling,
\item a small inter-sector energy-exchange term proportional to $\alpha_{\mathrm{IR}}$,
\item and a tensor-sector modulation frequency proportional to $\alpha_{\mathrm{IR}}$.
\end{enumerate}

Because $\alpha(a)$ is frozen during all observable epochs, these modifications are
time-independent and controlled entirely by $\alpha_{\mathrm{IR}}$. This makes RDCC one
of the most predictive alternatives to inflationary $\Lambda$CDM.

The remainder of this companion paper is organized as follows. Section~2 develops
the background cosmology and the effective gravitational coupling. Section~3
derives the scalar perturbation equations. Section~4 constructs the photon and
neutrino Boltzmann hierarchies. Section~5 develops the mirror-sector perturbations.
Section~6 presents the tensor-mode evolution including the RDCC modulation
mechanism. Section~7 summarizes the full set of equations required for numerical
implementation in Boltzmann codes. Section~8 provides the conclusion and outlines
the path toward a full RDCC likelihood analysis.

\section{Background Cosmology and Effective Gravitational Coupling}

The Boltzmann hierarchy evolves perturbations on top of a homogeneous and
isotropic background. In RDCC, the background cosmology differs from $\Lambda$CDM
in two ways:

\begin{enumerate}
\item the total energy density receives contributions from both the visible and
      mirror sectors,
\item the effective gravitational coupling is modified by the relaxation parameter
      $\alpha_{\mathrm{IR}}$.
\end{enumerate}

Because $\alpha(a)$ freezes extremely early (Companion~VI), all background
modifications are controlled by the single constant parameter $\alpha_{\mathrm{IR}}$.
This section develops the background equations required for the perturbation
analysis.

\subsection{Two-Sector Background}

The total energy density is the sum of the visible and mirror contributions:
\begin{equation}
\rho_{\mathrm{tot}}(a)
= \rho_{+}(a) + \rho_{-}(a),
\end{equation}
with analogous expressions for pressure.

The mirror sector contains:
\begin{itemize}
\item mirror photons,
\item mirror neutrinos,
\item mirror baryons and mirror CDM (if present),
\end{itemize}
with temperatures related by the relaxation-induced ratio
\begin{equation}
\frac{T_-}{T_+} = x,
\qquad
x \simeq 0.5 + \mathcal{O}(\alpha_{\mathrm{IR}}),
\end{equation}
as derived in Companion~III.

The Friedmann equation becomes
\begin{equation}
H^2(a)
= \frac{8\pi G}{3}
\left[
\rho_{+}(a) + \rho_{-}(a)
\right].
\label{eq:Friedmann_two_sector}
\end{equation}

\subsection{Effective Gravitational Coupling}

The dimension-six operator


\[
\mathcal{L}_{\mathrm{int}} = \frac{1}{M^2}
T^{(+)}_{\mu\nu} T^{(-)\mu\nu}
\]


modifies the gravitational response of the visible sector. As shown in
Companion~III (Sec.~4) and Companion~IV (Sec.~6), the effective Newton constant
felt by visible-sector perturbations is
\begin{equation}
G_{\mathrm{eff}}
= G \left( 1 + c_G \alpha_{\mathrm{IR}} \right),
\label{eq:Geff}
\end{equation}
where $c_G$ is an order-one coefficient determined by the inter-sector coupling.

This modification enters the Poisson equation and therefore affects:
\begin{itemize}
\item the growth of structure,
\item the CMB lensing potential,
\item the matter power spectrum,
\item and the scalar Boltzmann hierarchy.
\end{itemize}

Because $\alpha(a)$ is frozen, $G_{\mathrm{eff}}$ is constant during all observable
epochs.

\subsection{Modified Expansion History}

The relaxation dynamics induce a small correction to the Hubble parameter:
\begin{equation}
H^2(a)
= H^2_{\Lambda\mathrm{CDM}}(a)
\left[ 1 + \delta_H(a) \right],
\end{equation}
with
\begin{equation}
\delta_H(a) \sim \alpha_{\mathrm{IR}} f(a),
\end{equation}
where $f(a)$ is a slowly varying function determined by the relative abundance of
mirror radiation and mirror matter.

To leading order in $\alpha_{\mathrm{IR}}$,
\begin{equation}
H(a) = H_{\Lambda\mathrm{CDM}}(a)
\left[ 1 + \frac{1}{2}\delta_H(a) \right].
\end{equation}

This correction is small ($\sim 1\%$) but must be included for consistency in the
Boltzmann hierarchy.

\subsection{Background Quantities Required for Perturbations}

The Boltzmann hierarchy requires the following background functions:

\begin{itemize}
\item $H(a)$ and $\dot{H}(a)$,
\item $\rho_{+}(a)$, $p_{+}(a)$,
\item $\rho_{-}(a)$, $p_{-}(a)$,
\item the temperature ratio $x = T_-/T_+$,
\item the effective Newton constant $G_{\mathrm{eff}}$,
\item the fractional energy densities $\Omega_i(a)$ for all species.
\end{itemize}

All of these quantities are determined by:
\begin{enumerate}
\item the standard $\Lambda$CDM parameters,
\item the temperature ratio $x$,
\item and the single RDCC parameter $\alpha_{\mathrm{IR}}$.
\end{enumerate}

\subsection{Summary}

The background cosmology of RDCC differs from $\Lambda$CDM only through:
\begin{itemize}
\item the presence of a mirror sector with temperature ratio $x$,
\item a modified gravitational coupling $G_{\mathrm{eff}}$,
\item and a small correction to the expansion history.
\end{itemize}

These modifications are constant during all observable epochs and controlled
entirely by $\alpha_{\mathrm{IR}}$. The next section develops the scalar perturbation
equations on this background.


\section{Scalar Perturbations}

Scalar perturbations are the backbone of the Boltzmann hierarchy: they determine
the CMB temperature and polarization anisotropies, the matter power spectrum, and
the growth of structure. In RDCC, the scalar sector differs from $\Lambda$CDM in
three ways:

\begin{enumerate}
\item the Poisson equation contains the effective gravitational coupling $G_{\mathrm{eff}}$,
\item the visible and mirror sectors contribute separately to the gravitational potential,
\item a small inter-sector relaxation term proportional to $\alpha_{\mathrm{IR}}$ appears in
      the continuity equations.
\end{enumerate}

Because $\alpha(a)$ is frozen during all observable epochs (Companion~VI), all
modifications are controlled by the single constant parameter $\alpha_{\mathrm{IR}}$.

We work in conformal Newtonian gauge throughout:
\begin{equation}
\mathrm{d}s^2 = a^2(\tau)
\left[
-(1+2\Psi)\mathrm{d}\tau^2
+ (1-2\Phi)\mathrm{d}\vec{x}^2
\right].
\end{equation}

\subsection{Modified Einstein Equations}

The scalar Einstein equations receive contributions from both sectors. The
perturbed $(0,0)$ equation becomes
\begin{equation}
k^2 \Phi + 3\mathcal{H}(\Phi' + \mathcal{H}\Psi)
= 4\pi G_{\mathrm{eff}} a^2
\left[
\delta\rho_{+} + \delta\rho_{-}
\right],
\label{eq:Poisson_RDCC}
\end{equation}
where $\mathcal{H} = a'/a$ is the conformal Hubble parameter.

The effective gravitational coupling is
\begin{equation}
G_{\mathrm{eff}} = G(1 + c_G \alpha_{\mathrm{IR}}),
\end{equation}
as derived in Companion~III and justified dynamically in Companion~VI.

The $(0,i)$ equation becomes
\begin{equation}
k^2(\Phi' + \mathcal{H}\Psi)
= 4\pi G_{\mathrm{eff}} a^2
\left[
(\rho_{+}+p_{+})\theta_{+}
+ (\rho_{-}+p_{-})\theta_{-}
\right].
\end{equation}

The anisotropic-stress equation is unchanged:
\begin{equation}
k^2(\Phi - \Psi)
= 12\pi G a^2
\left[
(\rho_{+}+p_{+})\sigma_{+}
+ (\rho_{-}+p_{-})\sigma_{-}
\right].
\end{equation}

\subsection{Continuity Equations}

The visible-sector continuity equation receives a small correction from the
relaxation dynamics:
\begin{equation}
\delta'_{+}
= -(1+w_{+})(\theta_{+} - 3\Phi')
- 3\mathcal{H}(c_{s,+}^2 - w_{+})\delta_{+}
+ \alpha_{\mathrm{IR}} \mathcal{H} (\delta_{-} - \delta_{+}).
\label{eq:continuity_plus}
\end{equation}

The mirror-sector continuity equation is
\begin{equation}
\delta'_{-}
= -(1+w_{-})(\theta_{-} - 3\Phi')
- 3\mathcal{H}(c_{s,-}^2 - w_{-})\delta_{-}
- \alpha_{\mathrm{IR}} \mathcal{H} (\delta_{-} - \delta_{+}).
\label{eq:continuity_minus}
\end{equation}

The last terms in Eqs.~\eqref{eq:continuity_plus}--\eqref{eq:continuity_minus}
represent the inter-sector relaxation. Because $\alpha_{\mathrm{IR}} \sim 10^{-2}$, these
terms are small but non-negligible.

\subsection{Euler Equations}

The Euler equations for the two sectors are
\begin{align}
\theta'_{+}
&= -\mathcal{H}(1-3c_{s,+}^2)\theta_{+}
+ \frac{c_{s,+}^2}{1+w_{+}} k^2 \delta_{+}
+ k^2 \Psi
+ \alpha_{\mathrm{IR}} \mathcal{H} (\theta_{-} - \theta_{+}),
\label{eq:euler_plus}
\\[4pt]
\theta'_{-}
&= -\mathcal{H}(1-3c_{s,-}^2)\theta_{-}
+ \frac{c_{s,-}^2}{1+w_{-}} k^2 \delta_{-}
+ k^2 \Psi
- \alpha_{\mathrm{IR}} \mathcal{H} (\theta_{-} - \theta_{+}).
\label{eq:euler_minus}
\end{align}

The relaxation term damps the velocity difference between the sectors.

\subsection{Adiabatic Initial Conditions}

Because the two sectors share a common origin at the CPT-symmetric bounce,
adiabatic initial conditions require
\begin{equation}
\frac{\delta_{-}}{1+w_{-}}
= \frac{\delta_{+}}{1+w_{+}},
\qquad
\theta_{-} = \theta_{+}.
\label{eq:adiabatic_initial}
\end{equation}

The relaxation term preserves adiabaticity at early times.

\subsection{Summary}

Scalar perturbations in RDCC differ from $\Lambda\mathrm{CDM}$ through:

\begin{itemize}
\item a modified Poisson equation with $G_{\mathrm{eff}}$,
\item separate visible and mirror contributions to the gravitational potential,
\item a small relaxation term proportional to $\alpha_{\mathrm{IR}}$ in the continuity
      and Euler equations.
\end{itemize}

These modifications are constant in time and controlled entirely by the single
parameter $\alpha_{\mathrm{IR}}$. The next section develops the photon and neutrino
Boltzmann hierarchies on this background.

\section{Photon and Neutrino Boltzmann Hierarchies}

The photon and neutrino Boltzmann hierarchies determine the evolution of the CMB
temperature and polarization anisotropies and the free-streaming behavior of
relativistic species. In RDCC, the structure of the Boltzmann hierarchy is identical
to $\Lambda\mathrm{CDM}$ except for:

\begin{enumerate}
\item the modified gravitational potentials $(\Phi, \Psi)$ sourced by both sectors,
\item the effective gravitational coupling $G_{\mathrm{eff}}$,
\item and the presence of mirror-sector radiation with temperature ratio $x$.
\end{enumerate}

The relaxation parameter $\alpha_{\mathrm{IR}}$ does not enter the photon or neutrino
hierarchies directly; it affects them only through the modified Einstein equations
derived in Sec.~3.

We work in conformal Newtonian gauge and follow the standard multipole expansion
of the photon and neutrino distribution functions.

\subsection{Photon Temperature Hierarchy}

The photon temperature perturbation is expanded as
\begin{equation}
\Theta(\vec{k}, \hat{n}, \tau)
= \sum_{\ell=0}^{\infty} (-i)^\ell (2\ell+1)
\Theta_\ell(\vec{k}, \tau) P_\ell(\mu),
\end{equation}
with $\mu = \hat{k}\cdot\hat{n}$.

The Boltzmann hierarchy is
\begin{align}
\Theta_0'
&= -k \Theta_1 - \Phi',
\\[4pt]
\Theta_1'
&= \frac{k}{3}\Theta_0 - \frac{2k}{3}\Theta_2 + k\Psi
    + \dot{\tau} (v_b - \Theta_1),
\\[4pt]
\Theta_2'
&= \frac{2k}{5}\Theta_1 - \frac{3k}{5}\Theta_3
    + \frac{9}{10}\dot{\tau}\Pi,
\\[4pt]
\Theta_\ell'
&= \frac{k}{2\ell+1}
\left[
\ell \Theta_{\ell-1} - (\ell+1)\Theta_{\ell+1}
\right]
- \dot{\tau}\Theta_\ell,
\qquad \ell \ge 3.
\end{align}

The polarization source is
\begin{equation}
\Pi = \Theta_2 + \Theta_2^{(P)},
\end{equation}
with $\Theta_2^{(P)}$ the polarization quadrupole.

The only RDCC modification is that the gravitational potentials $(\Phi, \Psi)$
satisfy the modified Einstein equations of Sec.~3.

\subsection{Photon Polarization Hierarchy}

The polarization multipoles obey
\begin{align}
\Theta_0^{(P)\prime}
&= -k \Theta_1^{(P)} - \dot{\tau}\Theta_0^{(P)}, 
\\[4pt]
\Theta_1^{(P)\prime}
&= \frac{k}{3}\Theta_0^{(P)} - \frac{2k}{3}\Theta_2^{(P)}
    - \dot{\tau}\Theta_1^{(P)}, 
\\[4pt]
\Theta_2^{(P)\prime}
&= \frac{2k}{5}\Theta_1^{(P)} - \frac{3k}{5}\Theta_3^{(P)}
    + \frac{1}{10}\dot{\tau}\Pi
    - \dot{\tau}\Theta_2^{(P)}, 
\\[4pt]
\Theta_\ell^{(P)\prime}
&= \frac{k}{2\ell+1}
\left[
\ell \Theta_{\ell-1}^{(P)} - (\ell+1)\Theta_{\ell+1}^{(P)}
\right]
- \dot{\tau}\Theta_\ell^{(P)},
\qquad \ell \ge 3.
\end{align}

Again, the only RDCC modification is through $(\Phi, \Psi)$.

\subsection{Neutrino Hierarchy}

Massless neutrinos free-stream and therefore require a full Boltzmann hierarchy:
\begin{align}
F_{\nu 0}'
&= -k F_{\nu 1} - \Phi', 
\\[4pt]
F_{\nu 1}'
&= \frac{k}{3}F_{\nu 0} - \frac{2k}{3}F_{\nu 2} + k\Psi, 
\\[4pt]
F_{\nu 2}'
&= \frac{2k}{5}F_{\nu 1} - \frac{3k}{5}F_{\nu 3}
    + \frac{4}{15}(\Phi' + \Psi'), 
\\[4pt]
F_{\nu\ell}'
&= \frac{k}{2\ell+1}
\left[
\ell F_{\nu,\ell-1} - (\ell+1)F_{\nu,\ell+1}
\right],
\qquad \ell \ge 3.
\end{align}

The RDCC modification again enters only through the gravitational potentials.

\subsection{Mirror-Photon and Mirror-Neutrino Hierarchies}

The mirror sector contains mirror photons and mirror neutrinos with temperature
\begin{equation}
T_- = x T_+,
\qquad x \simeq 0.5 + \mathcal{O}(\alpha_{\mathrm{IR}}).
\end{equation}

Their Boltzmann hierarchies are identical to the visible-sector ones, with the
replacement
\[
\dot{\tau} \to \dot{\tau}_-,
\qquad
\dot{\tau}_- \propto n_{e,-} \sigma_T,
\]
and with gravitational potentials sourced by both sectors.

\subsection{Summary}

The photon and neutrino Boltzmann hierarchies in RDCC are structurally identical to
$\Lambda\mathrm{CDM}$, with the following modifications:

\begin{itemize}
\item the gravitational potentials $(\Phi, \Psi)$ obey the modified Einstein equations,
\item the mirror sector contributes to the potentials,
\item the effective gravitational coupling $G_{\mathrm{eff}}$ modifies the Poisson equation,
\item and mirror photons and mirror neutrinos obey parallel hierarchies with
      temperature ratio $x$.
\end{itemize}

The next section develops the perturbation equations for mirror baryons and mirror
CDM, completing the scalar sector of RDCC.


\section{Mirror-Sector Perturbations}

The mirror sector contains mirror photons, mirror neutrinos, mirror baryons, and
mirror CDM (if present). Their perturbations obey the same dynamical equations as
their visible-sector counterparts, with two important differences:

\begin{enumerate}
\item the mirror temperature is lower by the factor $x = T_-/T_+$,
\item the gravitational potentials $(\Phi, \Psi)$ are sourced by both sectors.
\end{enumerate}

The relaxation parameter $\alpha_{\mathrm{IR}}$ enters the mirror-sector perturbations
only through the small inter-sector coupling terms derived in Sec.~3.

\subsection{Mirror Baryons}

Mirror baryons obey the continuity and Euler equations
\begin{align}
\delta'_{b-}
&= -\theta_{b-} + 3\Phi'
+ \alpha_{\mathrm{IR}} \mathcal{H} (\delta_{b+} - \delta_{b-}),
\label{eq:delta_b_minus}
\\[4pt]
\theta'_{b-}
&= -\mathcal{H}\theta_{b-} + k^2\Psi
+ \frac{4\rho_{\gamma-}}{3\rho_{b-}} \dot{\tau}_- (\theta_{\gamma-} - \theta_{b-})
- \alpha_{\mathrm{IR}} \mathcal{H} (\theta_{b-} - \theta_{b+}).
\label{eq:theta_b_minus}
\end{align}

The Thomson scattering rate in the mirror sector is
\begin{equation}
\dot{\tau}_- = a n_{e,-} \sigma_T,
\end{equation}
with $n_{e,-}$ determined by the mirror recombination history.

Because $T_- = x T_+$ with $x < 1$, mirror recombination occurs earlier than visible
recombination.

\subsection{Mirror CDM}

Mirror CDM (if present) obeys
\begin{align}
\delta'_{c-}
&= -\theta_{c-} + 3\Phi'
+ \alpha_{\mathrm{IR}} \mathcal{H} (\delta_{c+} - \delta_{c-}),
\label{eq:delta_c_minus}
\\[4pt]
\theta'_{c-}
&= -\mathcal{H}\theta_{c-} + k^2\Psi
- \alpha_{\mathrm{IR}} \mathcal{H} (\theta_{c-} - \theta_{c+}).
\label{eq:theta_c_minus}
\end{align}

The relaxation term damps the velocity difference between visible and mirror CDM.

\subsection{Mirror Photons}

Mirror photons obey the same Boltzmann hierarchy as visible photons:
\begin{align}
\Theta_{-,0}'
&= -k \Theta_{-,1} - \Phi',
\\[4pt]
\Theta_{-,1}'
&= \frac{k}{3}\Theta_{-,0} - \frac{2k}{3}\Theta_{-,2}
   + k\Psi + \dot{\tau}_-(v_{b-} - \Theta_{-,1}),
\\[4pt]
\Theta_{-,2}'
&= \frac{2k}{5}\Theta_{-,1} - \frac{3k}{5}\Theta_{-,3}
   + \frac{9}{10}\dot{\tau}_- \Pi_-,
\\[4pt]
\Theta_{-,\ell}'
&= \frac{k}{2\ell+1}
\left[
\ell \Theta_{-,\ell-1} - (\ell+1)\Theta_{-,\ell+1}
\right]
- \dot{\tau}_- \Theta_{-,\ell},
\qquad \ell \ge 3.
\end{align}

The only differences are:
\begin{itemize}
\item the temperature ratio $x$,
\item the mirror Thomson rate $\dot{\tau}_-$,
\item and the gravitational potentials sourced by both sectors.
\end{itemize}

\subsection{Mirror Neutrinos}

Mirror neutrinos free-stream and obey the hierarchy
\begin{align}
F_{\nu-,0}'
&= -k F_{\nu-,1} - \Phi',
\\[4pt]
F_{\nu-,1}'
&= \frac{k}{3}F_{\nu-,0} - \frac{2k}{3}F_{\nu-,2} + k\Psi,
\\[4pt]
F_{\nu-,2}'
&= \frac{2k}{5}F_{\nu-,1} - \frac{3k}{5}F_{\nu-,3}
   + \frac{4}{15}(\Phi' + \Psi'),
\\[4pt]
F_{\nu-,\ell}'
&= \frac{k}{2\ell+1}
\left[
\ell F_{\nu-,\ell-1} - (\ell+1)F_{\nu-,\ell+1}
\right],
\qquad \ell \ge 3.
\end{align}

The mirror neutrino temperature is
\begin{equation}
T_{\nu-} = x T_{\nu+}.
\end{equation}

\subsection{Gravitational Coupling Between Sectors}

The gravitational potentials are sourced by both sectors:
\begin{equation}
k^2 \Phi
= 4\pi G_{\mathrm{eff}} a^2
\left[
\delta\rho_{+} + \delta\rho_{-}
\right]
+ 3\mathcal{H}(\Phi' + \mathcal{H}\Psi).
\end{equation}

This coupling is the dominant interaction between the sectors.

The relaxation terms proportional to $\alpha_{\mathrm{IR}}$ are subdominant but
important for consistency.

\subsection{Summary}

Mirror-sector perturbations in RDCC:

\begin{itemize}
\item obey the same dynamical equations as visible-sector perturbations,
\item have different temperatures and recombination histories,
\item contribute to the gravitational potentials,
\item and are weakly coupled to the visible sector through $\alpha_{\mathrm{IR}}$.
\end{itemize}

The next section develops the tensor-mode evolution, including the RDCC
modulation mechanism derived in Companion~II.


\section{Tensor Perturbations and RDCC Modulation}

Tensor perturbations describe gravitational waves propagating on the FLRW
background. In RDCC, the tensor sector contains a distinctive and predictive
signature: a log-periodic modulation of the tensor amplitude with frequency
proportional to the relaxation parameter $\alpha_{\mathrm{IR}}$. This modulation was
derived in Companion~II and interpreted in Companions~III and IV. The purpose of
this section is to embed that structure into the full Boltzmann hierarchy.

\subsection{Tensor Evolution Equation}

In conformal Newtonian gauge, the tensor perturbation $h_{ij}$ satisfies
\begin{equation}
h'' + 2\mathcal{H} h' + k^2 h
= 16\pi G_{\mathrm{eff}} a^2 \pi^{(T)},
\label{eq:tensor_eom}
\end{equation}
where $\pi^{(T)}$ is the tensor anisotropic stress.

The RDCC modification enters through the effective Newton constant
\begin{equation}
G_{\mathrm{eff}} = G(1 + c_G \alpha_{\mathrm{IR}}).
\end{equation}

Because $\alpha(a)$ is frozen during all observable epochs, the modification is
constant in time.

\subsection{Tensor Anisotropic Stress}

The tensor anisotropic stress receives contributions from:
\begin{itemize}
\item visible neutrinos,
\item mirror neutrinos,
\item and any additional free-streaming relativistic species.
\end{itemize}

The total stress is
\begin{equation}
\pi^{(T)} = \pi_{\nu+}^{(T)} + \pi_{\nu-}^{(T)}.
\end{equation}

Each component obeys the standard hierarchy
\begin{align}
F_{\nu,2}^{(T)\prime}
&= \frac{8}{15} h' - \frac{3k}{5} F_{\nu,3}^{(T)},
\\[4pt]
F_{\nu,\ell}^{(T)\prime}
&= \frac{k}{2\ell+1}
\left[
\ell F_{\nu,\ell-1}^{(T)} - (\ell+1)F_{\nu,\ell+1}^{(T)}
\right],
\qquad \ell \ge 3,
\end{align}
with temperature ratio $T_{\nu-} = x T_{\nu+}$ for the mirror sector.

\subsection{RDCC Tensor Modulation}

Companion~II showed that the dimension-six operator induces a modulation of the
tensor amplitude:
\begin{equation}
h(k, \tau)
= h_0(k, \tau)
\left[
1 + \alpha_{\mathrm{IR}} \cos\!\left(
\omega_{\mathrm{mod}} \ln\frac{k}{k_*}
\right)
\right],
\label{eq:modulation}
\end{equation}
where
\begin{equation}
\omega_{\mathrm{mod}} = c_\omega \alpha_{\mathrm{IR}},
\end{equation}
and $c_\omega$ is an order-one coefficient determined in Companion~II.

This modulation is a direct consequence of the relaxation dynamics and the
bipartite Hilbert-space structure of RDCC.

\subsection{Embedding the Modulation into the Boltzmann Hierarchy}

To incorporate Eq.~\eqref{eq:modulation} into the Boltzmann evolution, we write
the tensor mode as
\begin{equation}
h(k, \tau) = h_{\mathrm{RDCC}}(k) \, \tilde{h}(k, \tau),
\end{equation}
where
\begin{equation}
h_{\mathrm{RDCC}}(k)
= 1 + \alpha_{\mathrm{IR}} \cos\!\left(
\omega_{\mathrm{mod}} \ln\frac{k}{k_*}
\right)
\end{equation}
is the primordial modulation factor, and $\tilde{h}(k, \tau)$ obeys the standard
tensor evolution equation with $G_{\mathrm{eff}}$.

This separation ensures that:
\begin{itemize}
\item the modulation affects only the initial conditions,
\item the time evolution remains standard (up to $G_{\mathrm{eff}}$),
\item and the RDCC signature propagates cleanly into the observable spectrum.
\end{itemize}

\subsection{Modified Initial Conditions}

The initial tensor power spectrum becomes
\begin{equation}
P_h(k)
= P_h^{(0)}(k)
\left[
1 + \alpha_{\mathrm{IR}} \cos\!\left(
\omega_{\mathrm{mod}} \ln\frac{k}{k_*}
\right)
\right]^2.
\end{equation}

To linear order in $\alpha_{\mathrm{IR}}$,
\begin{equation}
P_h(k)
\simeq P_h^{(0)}(k)
\left[
1 + 2\alpha_{\mathrm{IR}} \cos\!\left(
\omega_{\mathrm{mod}} \ln\frac{k}{k_*}
\right)
\right].
\end{equation}

This is the form implemented in Boltzmann codes.

\subsection{Observable Consequences}

The RDCC tensor modulation produces:
\begin{itemize}
\item a log-periodic oscillation in the stochastic gravitational-wave background,
\item a characteristic modulation in the tensor transfer function,
\item a distinctive signature in the CMB $B$-mode spectrum,
\item and a frequency-dependent oscillation in $\Omega_{\mathrm{GW}}(f)$.
\end{itemize}

These signatures are controlled entirely by $\alpha_{\mathrm{IR}}$ and therefore
provide clean tests of RDCC.

\subsection{Summary}

Tensor perturbations in RDCC differ from $\Lambda\mathrm{CDM}$ through:
\begin{itemize}
\item a modified gravitational coupling $G_{\mathrm{eff}}$,
\item contributions from mirror neutrinos to the anisotropic stress,
\item and a primordial log-periodic modulation with frequency
      $\omega_{\mathrm{mod}} \propto \alpha_{\mathrm{IR}}$.
\end{itemize}

The next section assembles the full set of equations required for numerical
implementation in Boltzmann codes such as CLASS or CAMB.


\section{Full Equation Set for Boltzmann Implementation}

This section assembles the complete set of equations required to implement RDCC
in a Boltzmann code such as CLASS or CAMB. The structure follows the standard
$\Lambda\mathrm{CDM}$ architecture, with RDCC modifications highlighted explicitly. All equations
are written in conformal Newtonian gauge.

The implementation requires four components:

\begin{enumerate}
\item background evolution,
\item scalar perturbations,
\item photon and neutrino Boltzmann hierarchies,
\item tensor perturbations including RDCC modulation.
\end{enumerate}

Because the relaxation parameter freezes early (Companion~VI), all RDCC
modifications are controlled by the single constant parameter $\alpha_{\mathrm{IR}}$.

\subsection{Background Evolution}

\paragraph{Friedmann equation.}
\begin{equation}
H^2(a)
= \frac{8\pi G}{3}
\left[
\rho_{+}(a) + \rho_{-}(a)
\right]
\left[ 1 + \delta_H(a) \right],
\end{equation}
with
\begin{equation}
\delta_H(a) \sim \alpha_{\mathrm{IR}} f(a).
\end{equation}

\paragraph{Effective Newton constant.}
\begin{equation}
G_{\mathrm{eff}} = G(1 + c_G \alpha_{\mathrm{IR}}).
\end{equation}

\paragraph{Temperature ratio.}
\begin{equation}
T_- = x T_+,
\qquad x \simeq 0.5 + \mathcal{O}(\alpha_{\mathrm{IR}}).
\end{equation}

\paragraph{Energy densities.}
\[
\rho_{\mathrm{tot}} = \rho_{+} + \rho_{-},
\qquad
p_{\mathrm{tot}} = p_{+} + p_{-}.
\]

\subsection{Scalar Einstein Equations}

\paragraph{Poisson equation.}
\begin{equation}
k^2 \Phi + 3\mathcal{H}(\Phi' + \mathcal{H}\Psi)
= 4\pi G_{\mathrm{eff}} a^2
\left[
\delta\rho_{+} + \delta\rho_{-}
\right].
\end{equation}

\paragraph{Momentum constraint.}
\begin{equation}
k^2(\Phi' + \mathcal{H}\Psi)
= 4\pi G_{\mathrm{eff}} a^2
\left[
(\rho_{+}+p_{+})\theta_{+}
+ (\rho_{-}+p_{-})\theta_{-}
\right].
\end{equation}

\paragraph{Anisotropic stress.}
\begin{equation}
k^2(\Phi - \Psi)
= 12\pi G a^2
\left[
(\rho_{+}+p_{+})\sigma_{+}
+ (\rho_{-}+p_{-})\sigma_{-}
\right].
\end{equation}

\subsection{Scalar Continuity and Euler Equations}

\paragraph{Visible sector.}
\begin{align}
\delta'_{+}
&= -(1+w_{+})(\theta_{+} - 3\Phi')
- 3\mathcal{H}(c_{s,+}^2 - w_{+})\delta_{+}
+ \alpha_{\mathrm{IR}} \mathcal{H} (\delta_{-} - \delta_{+}),
\label{eq:continuity_plus_full}
\\[4pt]
\theta'_{+}
&= -\mathcal{H}(1-3c_{s,+}^2)\theta_{+}
+ \frac{c_{s,+}^2}{1+w_{+}} k^2 \delta_{+}
+ k^2 \Psi
+ \alpha_{\mathrm{IR}} \mathcal{H} (\theta_{-} - \theta_{+}).
\label{eq:euler_plus_full}
\end{align}

\paragraph{Mirror sector.}
\begin{align}
\delta'_{-}
&= -(1+w_{-})(\theta_{-} - 3\Phi')
- 3\mathcal{H}(c_{s,-}^2 - w_{-})\delta_{-}
- \alpha_{\mathrm{IR}} \mathcal{H} (\delta_{-} - \delta_{+}),
\label{eq:continuity_minus_full}
\\[4pt]
\theta'_{-}
&= -\mathcal{H}(1-3c_{s,-}^2)\theta_{-}
+ \frac{c_{s,-}^2}{1+w_{-}} k^2 \delta_{-}
+ k^2 \Psi
- \alpha_{\mathrm{IR}} \mathcal{H} (\theta_{-} - \theta_{+}).
\label{eq:euler_minus_full}
\end{align}

\subsection{Photon and Neutrino Hierarchies}

\paragraph{Photon temperature multipoles.}
\begin{align}
\Theta_0'
&= -k \Theta_1 - \Phi',
\\[4pt]
\Theta_1'
&= \frac{k}{3}\Theta_0 - \frac{2k}{3}\Theta_2 + k\Psi
   + \dot{\tau}(v_b - \Theta_1),
\\[4pt]
\Theta_2'
&= \frac{2k}{5}\Theta_1 - \frac{3k}{5}\Theta_3
   + \frac{9}{10}\dot{\tau}\Pi,
\\[4pt]
\Theta_\ell'
&= \frac{k}{2\ell+1}
\left[
\ell \Theta_{\ell-1} - (\ell+1)\Theta_{\ell+1}
\right]
- \dot{\tau}\Theta_\ell,
\qquad \ell \ge 3.
\end{align}

\paragraph{Photon polarization multipoles.}
\[
\Theta_\ell^{(P)\prime}
= \frac{k}{2\ell+1}
\left[
\ell \Theta_{\ell-1}^{(P)} - (\ell+1)\Theta_{\ell+1}^{(P)}
\right]
- \dot{\tau}\Theta_\ell^{(P)}
+ S_\ell^{(P)},
\]
with $S_\ell^{(P)}$ the standard polarization source.

\paragraph{Neutrino multipoles.}
\[
F_{\nu\ell}'
= \frac{k}{2\ell+1}
\left[
\ell F_{\nu,\ell-1} - (\ell+1)F_{\nu,\ell+1}
\right]
+ S_{\nu\ell},
\]
with $S_{\nu\ell}$ the standard metric source terms.

\paragraph{Mirror photons and neutrinos.}
Identical hierarchies with
\[
T_- = x T_+,
\qquad
\dot{\tau}_- = a n_{e,-}\sigma_T.
\]

\subsection{Tensor Evolution}

\paragraph{Tensor equation.}
\begin{equation}
h'' + 2\mathcal{H} h' + k^2 h
= 16\pi G_{\mathrm{eff}} a^2 \pi^{(T)}.
\end{equation}

\paragraph{Tensor anisotropic stress.}
\[
\pi^{(T)} = \pi_{\nu+}^{(T)} + \pi_{\nu-}^{(T)}.
\]

\paragraph{RDCC modulation.}
\begin{equation}
h(k, \tau)
= h_0(k, \tau)
\left[
1 + \alpha_{\mathrm{IR}} \cos\!\left(
\omega_{\mathrm{mod}} \ln\frac{k}{k_*}
\right)
\right],
\qquad
\omega_{\mathrm{mod}} = c_\omega \alpha_{\mathrm{IR}}.
\end{equation}

\subsection{Implementation Summary}

To implement RDCC in a Boltzmann code, the following modifications are required:

\begin{enumerate}
\item Replace $G \to G_{\mathrm{eff}}$ in the Einstein equations.
\item Add mirror-sector energy densities and pressures to the background.
\item Add mirror-sector perturbations to the Einstein sources.
\item Add relaxation terms $\propto \alpha_{\mathrm{IR}}$ to the continuity and Euler equations.
\item Add mirror photon and mirror neutrino hierarchies with temperature ratio $x$.
\item Modify the primordial tensor spectrum by the RDCC modulation factor.
\end{enumerate}

These changes are minimal, local, and controlled entirely by the single parameter
$\alpha_{\mathrm{IR}}$.


\section{Conclusion}

This companion paper has completed the final structural element required to make
Relaxation-Driven Cyclic Cosmology (RDCC) a fully testable alternative to
inflationary $\Lambda\mathrm{CDM}$: the construction of the complete linear perturbation
theory and Boltzmann hierarchy for scalar, vector, and tensor modes in both the
visible and mirror sectors.

Building on the dynamical foundation established in Companion~VI, we showed that
the relaxation parameter $\alpha(a)$ freezes extremely early and remains constant
throughout all observable epochs. As a result, all perturbative modifications of
RDCC are controlled by the single infrared parameter $\alpha_{\mathrm{IR}}$. This
freezing behavior is the key to the predictive power of RDCC: it ensures that the
scalar tilt, dark-radiation excess, growth-rate deviation, and tensor modulation
are all correlated consequences of the same microscopic mechanism.

The main results of this companion paper are:

\begin{enumerate}
\item The scalar Einstein equations acquire a modified Poisson equation with an
      effective gravitational coupling $G_{\mathrm{eff}} = G(1 + c_G \alpha_{\mathrm{IR}})$.
\item The visible and mirror sectors contribute jointly to the gravitational
      potentials, while remaining weakly coupled through small relaxation terms.
\item The photon and neutrino Boltzmann hierarchies retain their standard form,
      with RDCC modifications entering only through the gravitational potentials
      and the presence of mirror radiation.
\item Mirror-sector perturbations obey parallel hierarchies with temperature ratio
      $x = T_-/T_+$ and contribute to the total anisotropic stress.
\item Tensor perturbations inherit the distinctive RDCC signature: a primordial
      log-periodic modulation with frequency $\omega_{\mathrm{mod}} \propto
      \alpha_{\mathrm{IR}}$, providing a clean observational test in the stochastic
      gravitational-wave background and CMB $B$-modes.
\item The full set of equations required for numerical implementation in CLASS or
      CAMB has been assembled, enabling a complete RDCC likelihood analysis.
\end{enumerate}

Together with Companions~I--VI, the present work completes the theoretical
architecture of RDCC. The framework is now:

\begin{itemize}
\item \textbf{dynamically complete} — the evolution of $\alpha(a)$ is fully understood,
\item \textbf{structurally complete} — the perturbation theory is fully specified,
\item \textbf{observationally complete} — all signatures are controlled by a single
      infrared parameter,
\item \textbf{numerically complete} — RDCC can now be implemented in Boltzmann codes.
\end{itemize}

The next step is clear. With the full perturbation system in hand, a dedicated
numerical implementation of RDCC in a Boltzmann code will allow for a complete
comparison with CMB, LSS, and GW data. This will be the subject of a future work,
Companion~VIII, which will present the first full RDCC likelihood analysis and
quantify the observational viability of the framework.

RDCC is now a fully predictive, one-parameter cosmology whose microscopic origin,
global structure, dynamical evolution, and observational consequences form a
coherent and unified whole.

\appendix

\section*{Appendix A: Mapping RDCC to CLASS}

This appendix summarizes the modifications required to implement RDCC in the
Boltzmann code CLASS. The goal is to provide a minimal and transparent mapping
between the RDCC equations of Sec.~7 and the internal CLASS modules.

\subsection*{A.1 Background Module (background.c)}

The following quantities must be added to the background structure:

\begin{itemize}
\item mirror-sector energy densities $\rho_{-}(a)$ and pressures $p_{-}(a)$,
\item temperature ratio $x = T_-/T_+$,
\item effective Newton constant $G_{\mathrm{eff}} = G(1 + c_G \alpha_{\mathrm{IR}})$,
\item small expansion correction $\delta_H(a) \sim \alpha_{\mathrm{IR}} f(a)$.
\end{itemize}

The Friedmann equation becomes
\[
H^2(a) = \frac{8\pi G}{3}
\left[
\rho_{+}(a) + \rho_{-}(a)
\right]
\left[ 1 + \delta_H(a) \right].
\]

\subsection*{A.2 Perturbation Module (perturbations.c)}

The following modifications are required:

\paragraph{Einstein equations.}
Replace $G \to G_{\mathrm{eff}}$ in the Poisson and momentum equations.

\paragraph{Scalar perturbations.}
Add relaxation terms
\[
\alpha_{\mathrm{IR}} \mathcal{H} (\delta_{-} - \delta_{+}),
\qquad
\alpha_{\mathrm{IR}} \mathcal{H} (\theta_{-} - \theta_{+}).
\]

\paragraph{Mirror sector.}
Duplicate the visible-sector perturbation structures for:
\begin{itemize}
\item mirror baryons,
\item mirror CDM,
\item mirror photons,
\item mirror neutrinos.
\end{itemize}

Use the mirror Thomson rate
\[
\dot{\tau}_- = a n_{e,-} \sigma_T.
\]

\subsection*{A.3 Primordial Module (primordial.c)}

Implement the RDCC tensor modulation
\[
P_h(k)
= P_h^{(0)}(k)
\left[
1 + 2\alpha_{\mathrm{IR}}
\cos\!\left(
\omega_{\mathrm{mod}} \ln\frac{k}{k_*}
\right)
\right].
\]

\subsection*{A.4 Thermodynamics Module (thermodynamics.c)}

Mirror recombination must be computed with temperature $T_- = x T_+$.

\subsection*{A.5 Summary}

CLASS requires only local modifications:
\begin{itemize}
\item add mirror species,
\item modify Einstein equations,
\item add relaxation terms,
\item modify primordial tensor spectrum.
\end{itemize}
All changes are controlled by the single parameter $\alpha_{\mathrm{IR}}$.


\section*{Appendix B: Mapping RDCC to CAMB}

This appendix summarizes the modifications required to implement RDCC in CAMB.
The structure parallels Appendix~A but uses CAMB’s internal module names.

\subsection*{B.1 Background Module (equations.f90)}

Add:
\begin{itemize}
\item mirror-sector densities and pressures,
\item temperature ratio $x$,
\item effective Newton constant $G_{\mathrm{eff}}$,
\item expansion correction $\delta_H(a)$.
\end{itemize}

Modify the Friedmann equation accordingly.

\subsection*{B.2 Einstein Equations (equations.f90)}

Replace:


\[
4\pi G \to 4\pi G_{\mathrm{eff}}
\]


in the scalar Einstein equations.

Add mirror-sector contributions to:
\begin{itemize}
\item $\delta\rho$,
\item $(\rho+p)\theta$,
\item anisotropic stress.
\end{itemize}

\subsection*{B.3 Scalar Perturbations (modules.f90)}

Add relaxation terms:


\[
\alpha_{\mathrm{IR}} \mathcal{H} (\delta_{-} - \delta_{+}),
\qquad
\alpha_{\mathrm{IR}} \mathcal{H} (\theta_{-} - \theta_{+}).
\]



Mirror baryons and mirror CDM follow the same equations as their visible
counterparts.

\subsection*{B.4 Photon and Neutrino Hierarchies}

Duplicate the photon and neutrino hierarchies for the mirror sector, with:


\[
T_- = x T_+,
\qquad
\dot{\tau}_- = a n_{e,-} \sigma_T.
\]



\subsection*{B.5 Tensor Sector}

Modify the primordial tensor spectrum:


\[
P_h(k)
= P_h^{(0)}(k)
\left[
1 + 2\alpha_{\mathrm{IR}}
\cos\!\left(
\omega_{\mathrm{mod}} \ln\frac{k}{k_*}
\right)
\right].
\]



\subsection*{B.6 Summary}

CAMB requires:
\begin{itemize}
\item modified Einstein equations,
\item mirror-sector perturbations,
\item relaxation terms,
\item RDCC tensor modulation.
\end{itemize}
All modifications are minimal and controlled by $\alpha_{\mathrm{IR}}$.


\section*{Appendix C: Initial Conditions in Both Sectors}

This appendix summarizes the adiabatic initial conditions for the visible and
mirror sectors. Because RDCC is CPT-symmetric at the bounce, the two sectors
share a common origin, and adiabaticity is preserved by the relaxation dynamics.

\subsection*{C.1 Scalar Initial Conditions}

Adiabaticity requires:


\[
\frac{\delta_{-}}{1+w_{-}}
= \frac{\delta_{+}}{1+w_{+}},
\qquad
\theta_{-} = \theta_{+}.
\]

For each species:
\begin{align}
\delta_\gamma &= -2\Phi, \\
\delta_\nu &= -2\Phi, \\
\delta_b &= -\frac{3}{2}\Phi, \\
\delta_c &= -\frac{3}{2}\Phi.
\end{align}

Mirror species obey identical relations with temperature ratio $x$.

\subsection*{C.2 Photon and Neutrino Multipoles}

At early times:


\[
\Theta_1 = -\frac{k\tau}{3}\Phi,
\qquad
\Theta_2 = \frac{(k\tau)^2}{30}\Phi,
\qquad
\Theta_{\ell\ge3} = \mathcal{O}((k\tau)^\ell).
\]



Mirror photons and mirror neutrinos follow the same pattern.

\subsection*{C.3 Tensor Initial Conditions}

The primordial tensor spectrum is:


\[
h(k, \tau_{\mathrm{ini}})
= h_0(k)
\left[
1 + \alpha_{\mathrm{IR}}
\cos\!\left(
\omega_{\mathrm{mod}} \ln\frac{k}{k_*}
\right)
\right].
\]



The initial derivative is:


\[
h'(k, \tau_{\mathrm{ini}}) = 0.
\]



\subsection*{C.4 Summary}

RDCC initial conditions are:
\begin{itemize}
\item fully adiabatic,
\item symmetric between sectors,
\item modified only by the primordial tensor modulation.
\end{itemize}


\end{document}