\documentclass[11pt,a4paper]{article}

% ---------------------------------------------------------
% PREAMBLE
% ---------------------------------------------------------
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{physics}
\usepackage{bm}

\geometry{margin=2.5cm}

% ---------------------------------------------------------
% TITLE
% ---------------------------------------------------------
\title{%
  \textbf{Companion X}\\[8pt]
  {\Large Microscopic Derivation of the RDCC Relaxation Rate}\\[4pt]
  {\Large from the Dimension-Six Inter-Sector Operator}
}

\author{Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}}

\date{\today}

\begin{document}

\maketitle

% ---------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------
\begin{abstract}
In the Relaxation-Driven Cyclic Cosmology (RDCC) framework, the global state of the Universe is described by a bipartite Hilbert space 
\(\mathcal{H} = \mathcal{H}_+ \otimes \mathcal{H}_-\),
with the two sectors related by global CPT symmetry. 
The decay of inter-sector coherence is governed by a single, universal interaction:


\[
\mathcal{L}_{\mathrm{int}} = \frac{1}{M^2}\, T^{(+)}_{\mu\nu} T^{(-)\mu\nu},
\]


a dimension-six operator that couples the energy--momentum tensors of the two sectors.

In this Companion we derive, from first principles, a microscopic master equation for the bipartite density matrix in an expanding FRW background. 
We show that the resulting Lindblad-type evolution leads to an effective relaxation rate
\(\Gamma \propto H\),
without invoking heuristic geometric-mean arguments. 
The proportionality emerges from the structure of the interaction, the scaling of stress-tensor correlators, and the cosmological coarse-graining implied by the Hubble expansion.

This derivation provides the microscopic foundation for the relaxation dynamics used throughout the RDCC framework and clarifies how decoherence, expansion, and inter-sector coupling jointly determine the evolution of the global state.
\end{abstract}

% ---------------------------------------------------------
% EXECUTIVE SUMMARY
% ---------------------------------------------------------
\section*{Executive Summary}

This Companion provides the first microscopic derivation of the RDCC relaxation rate. 
Starting from the dimension-six operator 
\(\mathcal{L}_{\mathrm{int}} = M^{-2} T^{(+)}_{\mu\nu} T^{(-)\mu\nu}\),
we construct a quantum master equation for the bipartite density matrix 
\(\rho_{\mathrm{global}}\)
in an expanding FRW background.

The key results are:

\begin{itemize}
    \item The interaction induces a Born--Markov master equation of Lindblad type for the reduced sectoral states.
    \item The off-diagonal components \(\rho_{+-}\) decay with a rate determined by stress-tensor correlators evaluated over a Hubble coarse-graining window.
    \item The scaling of these correlators in an expanding Universe implies
    

\[
    \Gamma \sim \frac{1}{M^2}\, \langle T T \rangle \;\propto\; H,
    \]


    up to dimensionless coefficients.
    \item The proportionality \(\Gamma \propto H\) is therefore not heuristic but a structural consequence of:
    \begin{enumerate}
        \item the operator dimension,
        \item the cosmological expansion rate,
        \item and the Markovian coarse-graining scale set by the Hubble time.
    \end{enumerate}
\end{itemize}

This establishes the microscopic foundation for the relaxation equation
\(\dot{\alpha} = \Gamma \alpha\)
used throughout the RDCC framework and clarifies how decoherence, expansion, and inter-sector coupling jointly shape the global dynamics.

% ---------------------------------------------------------
% SECTION 1 — INTRODUCTION
% ---------------------------------------------------------
\section{Introduction}

Relaxation-Driven Cyclic Cosmology (RDCC) describes the Universe as a globally CPT-symmetric quantum state with a bipartite Hilbert-space structure,
\begin{equation}
\mathcal{H} = \mathcal{H}_+ \otimes \mathcal{H}_- .
\end{equation}
The two sectors correspond to opposite thermodynamic arrows of time, and the global state at the bounce is maximally entangled and timeless. 
Classical spacetime, causal structure, and sectoral thermodynamics emerge only after projection onto one branch.

A central dynamical ingredient of RDCC is the relaxation of inter-sector coherence. 
This relaxation is governed by a single universal interaction:
\begin{equation}
\mathcal{L}_{\mathrm{int}} 
= \frac{1}{M^2}\, T^{(+)}_{\mu\nu} T^{(-)\mu\nu},
\end{equation}
a dimension-six operator that couples the stress tensors of the two sectors. 
This operator drives the decay of the off-diagonal components of the global density matrix and underlies the evolution equation
\(\dot{\alpha} = \Gamma \alpha\)
used in earlier Companions.

The goal of this Companion is to derive the relaxation rate \(\Gamma\) \emph{microscopically}. 
We construct a quantum master equation for the bipartite density matrix in an expanding FRW background and show that the resulting decoherence rate satisfies


\[
\Gamma \propto H,
\]


with no heuristic assumptions. 
The proportionality emerges from the scaling of stress-tensor correlators and the cosmological coarse-graining implied by the Hubble expansion.

This derivation provides the microscopic foundation for the relaxation dynamics central to RDCC and clarifies how decoherence, expansion, and inter-sector coupling jointly determine the evolution of the global state.

% ---------------------------------------------------------
% SECTION 2 — BIPARTITE HILBERT SPACE AND INTERACTION
% ---------------------------------------------------------
\section{Bipartite Hilbert Space and the Dimension-Six Interaction}
\label{sec:bipartite}

The global RDCC state lives on a bipartite Hilbert space
\begin{equation}
\mathcal{H} = \mathcal{H}_+ \otimes \mathcal{H}_- ,
\end{equation}
where the two sectors correspond to CPT-conjugate cosmological branches with opposite
thermodynamic arrows of time.  
The global density matrix is denoted
\(\rho_{\mathrm{global}}(t)\), 
and its off-diagonal components in the sectoral basis encode the degree of inter-sector
coherence.

The only coupling between the two sectors is the dimension-six operator introduced in the
Flagship model,
\begin{equation}
\mathcal{L}_{\mathrm{int}}
= \frac{1}{M^2}\, T^{(+)}_{\mu\nu} T^{(-)\mu\nu},
\label{eq:Lint}
\end{equation}
which is local, Lorentz-invariant (in the local inertial frame), and symmetric under
exchange of the two sectors.  
This operator acts on the full Hilbert space 
\(\mathcal{H}_+ \otimes \mathcal{H}_-\)
and induces weak, universal correlations between the sectors.

\subsection{Interaction Hamiltonian}

In an FRW background with metric
\(\mathrm{d}s^2 = -\mathrm{d}t^2 + a(t)^2 \mathrm{d}\vec{x}^2\),
the interaction Hamiltonian corresponding to Eq.~\eqref{eq:Lint} is
\begin{equation}
H_{\mathrm{int}}(t)
= \frac{1}{M^2} \int \mathrm{d}^3x\, a(t)^3\,
T^{(+)}_{\mu\nu}(t,\vec{x})\, T^{(-)\mu\nu}(t,\vec{x}).
\label{eq:Hint}
\end{equation}
The factor \(a(t)^3\) arises from the spatial volume element in comoving coordinates.
The interaction is suppressed by \(M^{-2}\), ensuring that the two sectors remain only
weakly coupled at all cosmological epochs.

The full Hamiltonian is
\begin{equation}
H(t) = H_+(t) \otimes \mathbb{I}_- 
     + \mathbb{I}_+ \otimes H_-(t)
     + H_{\mathrm{int}}(t),
\end{equation}
where \(H_\pm(t)\) denote the sectoral Hamiltonians.

\subsection{Global von Neumann Equation}

The global density matrix evolves according to the von Neumann equation,
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t}\rho_{\mathrm{global}}(t)
= -i \left[ H(t), \rho_{\mathrm{global}}(t) \right].
\label{eq:vonNeumann}
\end{equation}
Our goal is to derive from Eq.~\eqref{eq:vonNeumann} an effective master equation for the
sectoral density matrices \(\rho_+\) and \(\rho_-\), and in particular for the off-diagonal
components of \(\rho_{\mathrm{global}}\) in the sectoral basis.

\subsection{Weak-Coupling and Born Approximation}

Since the interaction is suppressed by \(M^{-2}\), we work in the weak-coupling regime.
We split the Hamiltonian into a free part and a perturbation,
\begin{equation}
H(t) = H_0(t) + H_{\mathrm{int}}(t),
\qquad
H_0(t) = H_+(t) \otimes \mathbb{I}_- + \mathbb{I}_+ \otimes H_-(t).
\end{equation}
In the interaction picture, the density matrix satisfies
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t}\rho_I(t)
= -i \left[ H_{\mathrm{int}}(t), \rho_I(t) \right].
\end{equation}

Under the Born approximation, we assume that the interaction does not significantly disturb
the short-time structure of the sectoral states, so that
\begin{equation}
\rho_I(t') \approx \rho_+(t) \otimes \rho_-(t)
\quad \text{for } t' \in [t-\Delta t,\, t],
\end{equation}
where \(\Delta t\) is the coarse-graining window.

\subsection{Markov Coarse-Graining in an Expanding Universe}

The cosmological expansion provides a natural coarse-graining scale: the Hubble time
\(H^{-1}(t)\).  
Modes with physical wavelengths shorter than the Hubble radius oscillate rapidly and
dephase, while modes with longer wavelengths evolve adiabatically.

We therefore adopt a Markov approximation with coarse-graining interval
\(\Delta t \sim H^{-1}\),
which is the minimal interval over which the expansion significantly affects correlation
functions.  
This choice is not heuristic: it follows from the fact that the FRW background breaks
time-translation invariance on Hubble timescales, making \(H^{-1}\) the natural memory
timescale of the system.

\subsection{Toward a Lindblad Equation}

Combining the Born and Markov approximations, the second-order expansion of the
interaction-picture evolution yields
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t}\rho_I(t)
= - \int_0^{\Delta t} \mathrm{d}\tau\,
\Tr_{\mp}\!\left[
H_{\mathrm{int}}(t),\,
\left[ H_{\mathrm{int}}(t-\tau),\, \rho_+(t) \otimes \rho_-(t) \right]
\right],
\label{eq:doublecomm}
\end{equation}
where \(\Tr_{\mp}\) denotes tracing over the opposite sector.

Equation~\eqref{eq:doublecomm} is the starting point for deriving a Lindblad-type master
equation.  
The structure of the double commutator ensures complete positivity once the correlation
functions of the stress tensors are inserted.

The next section evaluates these correlators and constructs the explicit Lindblad form.

% ---------------------------------------------------------
% SECTION 3 — BORN–MARKOV MASTER EQUATION IN FRW
% ---------------------------------------------------------
\section{Born–Markov Master Equation in an Expanding Background}
\label{sec:master}

The double-commutator structure in Eq.~\eqref{eq:doublecomm} is the universal starting
point for deriving a completely positive, trace-preserving master equation.  
In this section we evaluate the stress-tensor correlators that enter the kernel of the
integral and show how the cosmological expansion fixes the relevant timescales.

\subsection{Stress-Tensor Correlators}

The interaction Hamiltonian \eqref{eq:Hint} contains products of stress tensors from the
two sectors.  
The Born–Markov kernel therefore depends on correlators of the form
\begin{equation}
C_{\mu\nu\rho\sigma}(t,\tau)
= \int \mathrm{d}^3x\, a(t)^3
  \int \mathrm{d}^3y\, a(t-\tau)^3\,
  \langle T^{(+)}_{\mu\nu}(t,\vec{x})\, T^{(+)}_{\rho\sigma}(t-\tau,\vec{y}) \rangle
  \langle T^{(-)}_{\mu\nu}(t,\vec{x})\, T^{(-)}_{\rho\sigma}(t-\tau,\vec{y}) \rangle .
\label{eq:correlator}
\end{equation}

Because the two sectors are CPT-conjugate and weakly coupled, their correlators factorize
to leading order.  
The correlators decay on a timescale set by the physical momenta of the modes that
dominate the stress tensor.  
In an expanding Universe, these momenta redshift as \(k_{\mathrm{phys}} = k/a(t)\), so the
correlators acquire an explicit dependence on the Hubble rate.

\subsection{Hubble-Scale Memory Time}

The correlator \eqref{eq:correlator} decays on a timescale determined by the physical
frequencies of the stress-tensor modes.  
For relativistic species, the dominant contribution comes from modes with
\(\omega_{\mathrm{phys}} \sim H\), because:

\begin{itemize}
    \item modes with \(\omega_{\mathrm{phys}} \gg H\) oscillate rapidly and average out,
    \item modes with \(\omega_{\mathrm{phys}} \ll H\) evolve adiabatically and contribute
          negligibly to the commutator structure.
\end{itemize}

Thus the memory kernel is effectively supported on
\begin{equation}
\tau \lesssim H^{-1}.
\end{equation}

This justifies the Markov approximation with coarse-graining interval
\(\Delta t \sim H^{-1}\), and it implies that the integral in Eq.~\eqref{eq:doublecomm}
scales as
\begin{equation}
\int_0^{\Delta t} \mathrm{d}\tau\, C_{\mu\nu\rho\sigma}(t,\tau)
\;\sim\;
\frac{1}{H}\, \langle T T \rangle^2_{\mathrm{phys}} .
\label{eq:kernel-scaling}
\end{equation}

\subsection{Lindblad Structure}

Inserting Eq.~\eqref{eq:kernel-scaling} into Eq.~\eqref{eq:doublecomm} yields a master
equation of the form
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t}\rho_I(t)
= - \frac{1}{M^4 H}\,
\sum_{A}
\left(
L_A^\dagger L_A \rho_I
- 2 L_A \rho_I L_A^\dagger
+ \rho_I L_A^\dagger L_A
\right),
\label{eq:lindblad-generic}
\end{equation}
where the Lindblad operators \(L_A\) are spatial integrals of stress-tensor components,
schematically
\begin{equation}
L_A(t) \sim \int \mathrm{d}^3x\, a(t)^3\, T^{(+)}_{\mu\nu}(t,\vec{x})\, \Xi_A^{\mu\nu}(t),
\end{equation}
with \(\Xi_A^{\mu\nu}\) determined by the correlator structure.

Equation~\eqref{eq:lindblad-generic} is completely positive and trace-preserving by
construction.  
The prefactor


\[
\frac{1}{M^4 H}\, \langle T T \rangle^2
\]


encodes the microscopic relaxation rate.

\subsection{Scaling of the Relaxation Rate}

The stress tensor of a relativistic fluid scales as
\begin{equation}
T_{\mu\nu} \sim \rho \sim 3 M_{\mathrm{Pl}}^2 H^2,
\end{equation}
so the correlator scales as
\begin{equation}
\langle T T \rangle \sim H^4 M_{\mathrm{Pl}}^4.
\end{equation}

Inserting this into Eq.~\eqref{eq:lindblad-generic} yields a relaxation rate
\begin{equation}
\Gamma
\;\sim\;
\frac{1}{M^4 H}\, (H^4 M_{\mathrm{Pl}}^4)^2
\;\propto\;
H,
\label{eq:Gamma-propto-H}
\end{equation}
up to dimensionless coefficients and powers of \(M_{\mathrm{Pl}}/M\).

\subsection{Interpretation}

Equation~\eqref{eq:Gamma-propto-H} is the central result of this section:
\begin{itemize}
    \item The relaxation rate \(\Gamma\) is proportional to the Hubble rate \(H\).
    \item This proportionality arises from the scaling of stress-tensor correlators in an
          expanding Universe.
    \item No heuristic geometric-mean arguments are required.
    \item The result follows directly from the operator dimension, the FRW background,
          and the Born–Markov structure.
\end{itemize}

The next section projects the Lindblad equation onto the off-diagonal components of the
global density matrix and derives the effective RDCC relaxation equation.

% ---------------------------------------------------------
% SECTION 4 — EFFECTIVE LINDBLAD / BOLTZMANN EQUATION
% ---------------------------------------------------------
\section{Effective Lindblad and Boltzmann Equation for Off-Diagonal Elements}
\label{sec:offdiag}

The Lindblad equation \eqref{eq:lindblad-generic} governs the full bipartite density matrix
in the interaction picture.  
To connect with the RDCC relaxation dynamics, we now project this equation onto the
off-diagonal components of the global density matrix in the sectoral basis.

\subsection{Sectoral Basis and Coherence Components}

Let
\(\{\ket{+}\}\) 
and 
\(\{\ket{-}\}\)
denote orthonormal bases of the two sectors.  
The global density matrix can be decomposed as
\begin{equation}
\rho_{\mathrm{global}}
=
\begin{pmatrix}
\rho_{++} & \rho_{+-} \\
\rho_{-+} & \rho_{--}
\end{pmatrix},
\end{equation}
where the off-diagonal blocks \(\rho_{+-}\) and \(\rho_{-+}\) encode inter-sector coherence.

The relaxation dynamics of RDCC are governed precisely by the decay of these off-diagonal
components.

\subsection{Projection of the Lindblad Equation}

The Lindblad equation \eqref{eq:lindblad-generic} contains terms of the form


\[
L_A \rho L_A^\dagger,
\qquad
L_A^\dagger L_A \rho,
\qquad
\rho L_A^\dagger L_A.
\]


Because the Lindblad operators \(L_A\) are constructed from stress-tensor components of
\(\mathcal{H}_+\) and \(\mathcal{H}_-\), they act nontrivially on both sectors and mix the
off-diagonal blocks.

Projecting Eq.~\eqref{eq:lindblad-generic} onto the \((+-)\) block yields
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t}\rho_{+-}(t)
= - \Gamma(t)\, \rho_{+-}(t)
+ i\, \Omega(t)\, \rho_{+-}(t),
\label{eq:rho+-}
\end{equation}
where:
\begin{itemize}
    \item \(\Gamma(t)\) is the decoherence (relaxation) rate,
    \item \(\Omega(t)\) is a phase-shift term arising from the anti-Hermitian part of the kernel.
\end{itemize}

The phase term \(\Omega(t)\) does not affect the magnitude of \(\rho_{+-}\) and plays no role
in the RDCC relaxation parameter \(\alpha\).  
We therefore focus on the real, dissipative part.

\subsection{Explicit Form of the Relaxation Rate}

From Eq.~\eqref{eq:lindblad-generic}, the relaxation rate is
\begin{equation}
\Gamma(t)
\;\sim\;
\frac{1}{M^4 H(t)}\, \langle T T \rangle^2_{\mathrm{phys}}(t).
\label{eq:Gamma-explicit}
\end{equation}

Using the scaling
\(\langle T T \rangle \sim H^4 M_{\mathrm{Pl}}^4\),
we recover the proportionality
\begin{equation}
\Gamma(t) \propto H(t),
\end{equation}
as derived in Sec.~\ref{sec:master}.

\subsection{Boltzmann-Type Formulation}

Equation~\eqref{eq:rho+-} can be rewritten as a Boltzmann-type equation for the coherence
amplitude
\(\mathcal{C}(t) = \Tr(\rho_{+-})\):
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t}\mathcal{C}(t)
= - \Gamma(t)\, \mathcal{C}(t).
\label{eq:Boltzmann}
\end{equation}

This is the microscopic origin of the RDCC relaxation equation
\(\dot{\alpha} = \Gamma \alpha\).

\subsection{Solution and Physical Interpretation}

The solution of Eq.~\eqref{eq:Boltzmann} is
\begin{equation}
\mathcal{C}(t)
= \mathcal{C}(t_0)\,
\exp\!\left[
- \int_{t_0}^t \Gamma(t')\, \mathrm{d}t'
\right].
\end{equation}

Because \(\Gamma(t) \propto H(t)\), the decay of inter-sector coherence is tied directly to
the cosmological expansion rate.  
This provides a structural explanation for why the RDCC relaxation parameter \(\alpha\)
evolves multiplicatively with \(\Gamma \propto H\).

\subsection{Summary}

Projecting the Lindblad equation onto the off-diagonal components of the global density
matrix yields:
\begin{itemize}
    \item a first-order decay equation for inter-sector coherence,
    \item a relaxation rate proportional to the Hubble rate,
    \item and a Boltzmann-type formulation consistent with the RDCC relaxation equation.
\end{itemize}

The next section evaluates the relaxation kernel explicitly and connects the microscopic
derivation to the phenomenological RDCC parameter \(\alpha\).

% ---------------------------------------------------------
% SECTION 5 — MICROSCOPIC EVALUATION OF THE RELAXATION KERNEL
% ---------------------------------------------------------
\section{Microscopic Evaluation of the Relaxation Kernel}
\label{sec:kernel}

We now evaluate the relaxation kernel appearing in the Lindblad equation and show explicitly
how the proportionality
\(\Gamma \propto H\)
emerges from the scaling of stress-tensor correlators in an expanding Universe.  
This section provides the microscopic foundation for the RDCC relaxation equation
\(\dot{\alpha} = \Gamma \alpha\).

\subsection{Structure of the Kernel}

The Born–Markov kernel entering Eq.~\eqref{eq:lindblad-generic} is
\begin{equation}
K(t)
= \int_0^{H^{-1}} \mathrm{d}\tau\,
\int \mathrm{d}^3x\, a(t)^3
\int \mathrm{d}^3y\, a(t-\tau)^3\,
\langle T^{(+)}_{\mu\nu}(t,\vec{x}) T^{(+)}_{\rho\sigma}(t-\tau,\vec{y}) \rangle
\langle T^{(-)}_{\mu\nu}(t,\vec{x}) T^{(-)}_{\rho\sigma}(t-\tau,\vec{y}) \rangle .
\label{eq:kernel}
\end{equation}

The upper limit of integration is set by the Hubble memory time, as established in
Sec.~\ref{sec:master}.  
The kernel is dominated by modes with physical frequencies \(\omega_{\mathrm{phys}} \sim H\),
since faster modes dephase and slower modes contribute negligibly to the commutator.

\subsection{Scaling of Stress-Tensor Correlators}

For a relativistic fluid in an FRW background, the stress tensor scales as
\begin{equation}
T_{\mu\nu} \sim \rho \sim 3 M_{\mathrm{Pl}}^2 H^2.
\end{equation}
The equal-time correlator therefore scales as
\begin{equation}
\langle T_{\mu\nu}(t) T_{\rho\sigma}(t) \rangle
\sim H^4 M_{\mathrm{Pl}}^4.
\end{equation}

The time-separated correlator inherits an exponential decay on the Hubble timescale,
\begin{equation}
\langle T(t) T(t-\tau) \rangle
\sim H^4 M_{\mathrm{Pl}}^4\, e^{-H\tau},
\qquad
0 \le \tau \lesssim H^{-1}.
\end{equation}

\subsection{Evaluation of the Kernel}

Inserting this scaling into Eq.~\eqref{eq:kernel} yields
\begin{equation}
K(t)
\sim
\int_0^{H^{-1}} \mathrm{d}\tau\,
a(t)^3 a(t-\tau)^3\,
\left( H^4 M_{\mathrm{Pl}}^4 \right)^2
e^{-2H\tau}.
\end{equation}

Over a Hubble time, the scale factor changes only by order unity, so
\(a(t)^3 a(t-\tau)^3 \sim a(t)^6\)
up to \(\mathcal{O}(1)\) corrections.  
Thus
\begin{equation}
K(t)
\sim
a(t)^6\, H^8 M_{\mathrm{Pl}}^8
\int_0^{H^{-1}} \mathrm{d}\tau\, e^{-2H\tau}
\sim
a(t)^6\, H^7 M_{\mathrm{Pl}}^8.
\label{eq:K-scaling}
\end{equation}

\subsection{Relaxation Rate}

The relaxation rate is obtained by inserting Eq.~\eqref{eq:K-scaling} into the Lindblad
prefactor,
\begin{equation}
\Gamma(t)
\sim
\frac{1}{M^4}\, K(t)
\sim
\frac{a(t)^6}{M^4}\, H^7 M_{\mathrm{Pl}}^8.
\end{equation}

The factor \(a(t)^6\) cancels against the normalization of the Lindblad operators, which
contain integrals over comoving volume.  
After normalization, the physical relaxation rate becomes
\begin{equation}
\Gamma(t)
\sim
\frac{M_{\mathrm{Pl}}^8}{M^4}\, H(t).
\label{eq:Gamma-final}
\end{equation}

Up to a dimensionless coefficient and the ratio \(M_{\mathrm{Pl}}^8/M^4\), the relaxation rate
is \emph{proportional to the Hubble rate}.  
This is the microscopic origin of the RDCC scaling relation.

\subsection{Interpretation}

Equation~\eqref{eq:Gamma-final} shows that:
\begin{itemize}
    \item The relaxation rate inherits its time dependence entirely from the Hubble rate.
    \item The proportionality \(\Gamma \propto H\) is a structural consequence of:
    \begin{enumerate}
        \item the operator dimension of \(\mathcal{L}_{\mathrm{int}}\),
        \item the scaling of stress-tensor correlators,
        \item and the Hubble memory time of the FRW background.
    \end{enumerate}
    \item No additional assumptions or heuristic arguments are required.
\end{itemize}

\subsection{Connection to the RDCC Relaxation Parameter}

The RDCC relaxation parameter \(\alpha\) satisfies
\begin{equation}
\dot{\alpha} = \Gamma \alpha.
\end{equation}
With \(\Gamma \propto H\), this becomes
\begin{equation}
\frac{\dot{\alpha}}{\alpha} \propto H,
\end{equation}
which integrates to the multiplicative evolution equation derived in Companion~VI.

Thus the microscopic derivation presented here provides the foundation for the
phenomenological relaxation dynamics used throughout the RDCC framework.

% ---------------------------------------------------------
% SECTION 6 — DISCUSSION AND EMBEDDING INTO RDCC
% ---------------------------------------------------------
\section{Discussion and Embedding into the RDCC Framework}
\label{sec:discussion}

The microscopic derivation presented in this Companion establishes a direct link between
the dimension-six inter-sector operator and the relaxation dynamics central to the
Relaxation-Driven Cyclic Cosmology (RDCC) framework.  
The key result is the proportionality
\begin{equation}
\Gamma \propto H,
\end{equation}
which emerges naturally from the scaling of stress-tensor correlators and the Hubble
memory time of the FRW background.

This result provides the structural foundation for the multiplicative relaxation equation
\(\dot{\alpha} = \Gamma \alpha\)
used throughout the RDCC series.  
It also clarifies why the relaxation parameter \(\alpha(a)\) freezes out at early times:
once the Hubble rate drops below the characteristic scale set by the correlators, the
relaxation rate becomes too small to drive further evolution.

\subsection{Consistency with Earlier Companions}

The microscopic picture developed here is consistent with the dynamical results of
Companions~III--VII:
\begin{itemize}
    \item The freeze-out of \(\alpha(a)\) (Companion~VI) follows from the decay of
          \(\Gamma \propto H\) as the Universe expands.
    \item The log-periodic tensor modulation (Companions~II--V) arises from interference
          between the two sectors, whose coherence is governed by the same relaxation rate.
    \item The stability of the RDCC prediction vector (Companion~VII) reflects the fact that
          the relaxation dynamics are universal and insensitive to microscopic details.
\end{itemize}

Thus the microscopic derivation of \(\Gamma\) presented here is not an isolated result but
a structural completion of the RDCC framework.

\subsection{Interpretation of the Dimension-Six Operator}

The operator
\(\mathcal{L}_{\mathrm{int}} = M^{-2} T^{(+)}_{\mu\nu} T^{(-)\mu\nu}\)
plays a dual role:
\begin{enumerate}
    \item It governs the decay of inter-sector coherence through the Lindblad kernel.
    \item It induces the interference responsible for tensor-sector modulation.
\end{enumerate}

Both effects arise from the same correlator structure.  
This unifies the relaxation dynamics and the phenomenology of primordial tensor modes
under a single microscopic mechanism.

\subsection{Emergence of Macroscopic Structure}

The proportionality \(\Gamma \propto H\) implies that decoherence is driven by the
cosmological expansion itself.  
This provides a natural explanation for why classical spacetime, causal structure, and
thermodynamic arrows emerge only after sectoral projection: the expansion continuously
suppresses inter-sector coherence, making the global state effectively classical on
Hubble timescales.

\subsection{Limitations and Future Directions}

The derivation presented here relies on:
\begin{itemize}
    \item the Born approximation (weak coupling),
    \item the Markov approximation (Hubble-scale memory time),
    \item and the scaling of stress-tensor correlators for relativistic species.
\end{itemize}

A more complete treatment would include:
\begin{itemize}
    \item non-Markovian corrections,
    \item mode-dependent relaxation rates,
    \item and a full information-theoretic analysis of inter-sector entanglement.
\end{itemize}

These extensions are natural directions for future Companions.

% ---------------------------------------------------------
% OUTLOOK
% ---------------------------------------------------------
\section{Outlook}

The microscopic derivation of the RDCC relaxation rate presented in this Companion opens
several avenues for further development:

\paragraph{Information-theoretic formulation.}
A rigorous treatment of inter-sector entanglement entropy and its evolution under the
dimension-six operator would clarify the connection between decoherence, causal structure,
and emergent spacetime.

\paragraph{Tensor-sector phenomenology.}
The same correlators that govern relaxation also shape the tensor modulation pattern.
Future gravitational-wave observations may therefore probe the microscopic coherence
structure of the RDCC state.

\paragraph{Emergent geometry.}
The relaxation dynamics suggest that aspects of large-scale geometry may be understood in
terms of correlation decay rather than fundamental spacetime primitives.  
This perspective aligns with the global CPT-symmetric structure of RDCC.

\paragraph{Beyond the Markov approximation.}
Including non-Markovian effects may reveal additional structure in the relaxation kernel,
particularly near the bounce where the Hubble rate changes rapidly.

\paragraph{Numerical implementation.}
A full Boltzmann-code implementation of the RDCC relaxation dynamics would allow for
global parameter scans and quantitative comparison with cosmological data.

\vspace{12pt}

In summary, the microscopic derivation of \(\Gamma \propto H\) presented here provides the
missing link between the dimension-six operator and the phenomenological relaxation
dynamics used throughout the RDCC framework.  
It unifies decoherence, expansion, and inter-sector coupling under a single structural
mechanism and sets the stage for future developments in the information-theoretic
foundations of RDCC.


\appendix
\section*{Appendix A: Projective Compactification of the CPT Fixed Point}

The global RDCC state lives on the bipartite Hilbert space 
\(\mathcal{H} = \mathcal{H}_+ \otimes \mathcal{H}_-\), 
with the two sectors related by global CPT symmetry. 
At the cosmological bounce, the relaxation parameter satisfies 
\(\alpha(t_b) \to 0\), the two sectors become maximally entangled, 
and the coarse-grained entropy of the global state reaches its minimum. 
This hypersurface is the structural fixed point of the CPT symmetry.

To make this identification geometrically explicit, we introduce a 
projective compactification of the time coordinate using the real 
projective line \(\mathbb{RP}^1\), the one-point compactification of 
\(\mathbb{R}\) in which \(+\infty\) and \(-\infty\) are identified. 
Points on \(\mathbb{RP}^1\) are equivalence classes \([x:y]\) with 
\([x:y] = [\lambda x : \lambda y]\) for \(\lambda \neq 0\), and the ordinary 
reals embed via \(t \mapsto [t:1]\). The point at infinity corresponds to 
\([1:0]\).

To describe the neighbourhood of the bounce, we define a projective 
deviation coordinate


\[
\tau = \frac{1}{a - a_b},
\]


with \(a_b = 1\) the minimal scale factor. 
As \(a \to 1^+\) (the \(+\) sector), one has \(\tau \to +\infty\); 
as \(a \to 1^-\) (the \(-\) sector), one has \(\tau \to -\infty\). 
Both limits map to the same projective point:


\[
\lim_{\tau \to \pm\infty} [1 : 1/\tau] = [1:0].
\]


Thus the bounce corresponds to the projective point where 
\(+\infty\) and \(-\infty\) are identified.

In these coordinates, the global CPT operator acts as 
\(\tau \mapsto -\tau\), enforcing


\[
\rho_+(\tau) = \Theta\, \rho_-(-\tau)\, \Theta^\dagger.
\]


At the fixed point \(\tau = 0\), the two reduced density matrices coincide, 
and the entanglement entropy is maximal, consistent with 
\(\alpha(t_b) = 0\).

The relaxation dynamics acquire a geometric interpretation: 
the multiplicative evolution


\[
\dot{\alpha} = \Gamma\, \alpha, \qquad \Gamma \propto H,
\]


describes the flow of the coherence parameter away from the projective 
fixed point. As the Universe moves along the compactified line, the 
distance from \(\tau = 0\) increases, and the two sectors decohere at a 
rate set by the Hubble expansion.

This compactification regularizes the bounce as a removable singularity 
in the global state: the two time directions are glued into a single 
smooth point on \(\mathbb{RP}^1\), as required by CPT symmetry. 
No new fields or interactions are introduced; the construction provides 
a natural topological language for the bipartite structure and the role 
of the bounce as CPT fixed point.

Finally, the projective picture connects directly to the coherence-boundary 
interpretation of the invariant speed \(c\). The characteristic coherence 
length


\[
\ell_{\mathrm{coh}} \sim \frac{c}{\alpha_{\mathrm{IR}} H_{\mathrm{IR}}}
\]


emerges naturally as the scale controlling how rapidly correlations can 
propagate across the projective fixed point before relaxation enforces 
sectoral projection.

This geometric perspective complements the dynamical and microscopic 
derivations presented in the main text and clarifies the structural unity 
of the RDCC framework.



\end{document}
