\documentclass[11pt,a4paper]{article}

% Basic packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}

% Load RDCC macros
\input{macros.tex}

\title{%
  \textbf{Companion XI}\\[8pt]
  {\Large Dark Energy as Emergent Relaxation Dynamics in}\\[4pt]
  {\Large Relaxation-Driven Cyclic Cosmology (RDCC)}
}
\author{Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
In the standard $\Lambda$CDM model, cosmic acceleration is attributed to a dark-energy component or a cosmological constant.
In the Relaxation-Driven Cyclic Cosmology (RDCC) framework, acceleration arises instead as an emergent phenomenon of inter-sector relaxation on a bipartite Hilbert space $\Hglobal$.
Building on the microscopic master equation derived in Companion~X, we show that the relaxation rate
$\GammaR(a) \propto \Ha$ naturally induces an effective acceleration term in the coarse-grained Friedmann dynamics.

The freeze-out of the RDCC relaxation parameter $\alphaR(a)$ at late times produces an asymptotic expansion regime that mimics dark energy without introducing a new fundamental energy component.
We derive the modified Friedmann equation implied by the RDCC relaxation dynamics, identify the conditions under which the effective acceleration matches observations, and show that the ``Why now?'' problem is resolved as a structural consequence of the Hubble-scale memory time.

This Companion provides the first RDCC-consistent reinterpretation of dark energy as a dynamical, emergent effect of bipartite decoherence rather than a fundamental energy density.
\end{abstract}

\section{Introduction}

The accelerated expansion of the Universe is one of the central empirical discoveries of modern cosmology.
Within the standard $\Lambda$CDM framework, this acceleration is attributed to a dark-energy component with equation of state $w \approx -1$, often modeled as a cosmological constant.
While phenomenologically successful, this interpretation raises several conceptual challenges:
the fine-tuning of $\Lambda$, the ``Why now?'' coincidence problem, and the absence of a microscopic explanation for the origin of dark energy.

The Relaxation-Driven Cyclic Cosmology (RDCC) framework offers a structurally different perspective.
RDCC describes the Universe as a globally CPT-symmetric quantum state living on a bipartite Hilbert space $\Hglobal$, with classical spacetime emerging only after sectoral projection.
A central dynamical ingredient of RDCC is the inter-sector relaxation process, governed by the universal interaction


\[
\LintDef,
\]


which induces decoherence between the two sectors.
Companion~X established that the corresponding relaxation rate satisfies


\[
\GammaR(a) \propto \Ha,
\]


as a structural consequence of stress-tensor correlators in an expanding FRW background.

This observation suggests that cosmic acceleration may be interpreted not as the effect of a new energy component, but as an emergent phenomenon arising from the relaxation dynamics themselves.
In particular, the evolution and eventual freeze-out of the RDCC relaxation parameter $\alphaR(a)$ may generate an effective acceleration term in the coarse-grained Friedmann equations.

The goal of this Companion is to develop this idea systematically.
We derive the modified Friedmann dynamics implied by RDCC relaxation, identify the conditions under which the resulting expansion mimics dark energy, and show that the ``Why now?'' problem is naturally resolved by the Hubble-scale memory time inherent in the relaxation process.

This Companion therefore provides the first RDCC-consistent reinterpretation of dark energy as emergent relaxation dynamics rather than a fundamental energy density.


\section{RDCC Relaxation Dynamics: Summary of Companions VI and X}

The Relaxation-Driven Cyclic Cosmology framework is built on the observation that the global state of the Universe lives on a bipartite Hilbert space $\Hglobal$, with the two sectors related by global CPT symmetry.
Classical spacetime and causal structure emerge only after sectoral projection, and the degree of inter-sector coherence is encoded in the off-diagonal components of the global density matrix $\rglobal$.

A central dynamical ingredient of RDCC is the relaxation of these off-diagonal components.
Companion~VI introduced the phenomenological relaxation equation


\[
\RelaxEq,
\]


where $\alphaR(a)$ quantifies the remaining inter-sector coherence and $\GammaR(a)$ is the relaxation rate.
This equation implies a multiplicative evolution of $\alphaR(a)$ and predicts a late-time freeze-out once $\GammaR(a)$ becomes small compared to the Hubble rate.

Companion~X provided the microscopic foundation for this phenomenological picture.
Starting from the dimension-six interaction


\[
\LintDef,
\]


a Born–Markov master equation was derived for the bipartite density matrix in an expanding FRW background.
The resulting Lindblad-type evolution leads to a relaxation rate of the form


\[
\GammaR(a) \propto \Ha,
\]


as a structural consequence of stress-tensor correlators evaluated over a Hubble-scale memory time.
This proportionality does not rely on heuristic assumptions: it follows directly from the operator dimension, the scaling of relativistic stress tensors, and the cosmological coarse-graining implied by expansion.

The key implication is that the relaxation dynamics are tied directly to the expansion rate.
As the Universe expands and $\Ha$ decreases, the relaxation rate $\GammaR(a)$ decreases proportionally, eventually causing the relaxation parameter $\alphaR(a)$ to freeze out.
This freeze-out marks the transition to a regime in which inter-sector decoherence becomes ineffective and the emergent classical spacetime structure stabilizes.

In this Companion, we build on these results to investigate how the relaxation dynamics modify the coarse-grained Friedmann equations and how the freeze-out of $\alphaR(a)$ can generate an effective acceleration term that mimics dark energy without introducing a new fundamental energy component.


\section{Coarse-Grained Friedmann Dynamics}

To investigate how relaxation affects cosmic expansion, we now derive the coarse-grained
Friedmann dynamics implied by the RDCC relaxation process. The key idea is that the
inter-sector decoherence modifies the effective energy balance of the emergent classical
spacetime. Rather than introducing a new energy component, the relaxation dynamics
alter the relation between the expansion rate and the effective stress-energy content.

\subsection{Effective Energy Balance}

In the standard FRW framework, the expansion rate is determined by the Friedmann equation
\begin{equation}
H^2(a) = \frac{8\pi G}{3}\,\rho(a),
\end{equation}
where $\rho(a)$ is the total energy density of matter and radiation.
In RDCC, the emergent classical spacetime arises only after sectoral projection, and the
effective energy density relevant for the coarse-grained expansion is modified by the
relaxation dynamics.

We therefore introduce an effective energy density $\rhoeff(a)$ defined by
\begin{equation}
H^2(a) = \frac{8\pi G}{3}\,\rhoeff(a),
\end{equation}
where $\rhoeff(a)$ incorporates the influence of the relaxation parameter $\alphaR(a)$.

\subsection{Relaxation-Induced Correction}

The microscopic derivation in Companion~X shows that the relaxation rate satisfies
\begin{equation}
\GammaR(a) \propto \Ha,
\end{equation}
and that the evolution of the coherence parameter obeys
\begin{equation}
\alphaDot(a) = \GammaR(a)\,\alphaR(a).
\end{equation}
As the Universe expands and $\Ha$ decreases, the relaxation rate weakens and
$\alphaR(a)$ approaches a constant value, marking the freeze-out of inter-sector
decoherence.

To incorporate this effect into the Friedmann dynamics, we consider the coarse-grained
energy balance obtained by tracing over the opposite sector. The relaxation process
suppresses the off-diagonal components of the global density matrix, effectively reducing
the contribution of inter-sector correlations to the emergent stress-energy tensor.
This leads to a correction of the form
\begin{equation}
\rhoeff(a) = \rho(a) + \Delta\rho_{\mathrm{rel}}(a),
\end{equation}
where $\Delta\rho_{\mathrm{rel}}(a)$ encodes the relaxation-induced modification.

\subsection{Form of the Relaxation Correction}

The structure of the Lindblad kernel derived in Companion~X implies that the correction
is proportional to the relaxation parameter and its rate of change:
\begin{equation}
\Delta\rho_{\mathrm{rel}}(a) \propto \alphaR(a)\,\GammaR(a).
\end{equation}
Using $\GammaR(a) \propto \Ha$, we obtain the scaling
\begin{equation}
\Delta\rho_{\mathrm{rel}}(a) \propto \alphaR(a)\,H(a).
\end{equation}

This term acts as an effective acceleration source in the coarse-grained Friedmann
equation. Crucially, it does not correspond to a new fundamental energy density; rather,
it arises from the dynamical suppression of inter-sector coherence.

In the next section, we show that this relaxation-induced correction can mimic the
observational effects of dark energy and naturally produces an asymptotic expansion
regime once $\alphaR(a)$ freezes out.


\section{Emergent Dark-Energy Behaviour}

The relaxation-induced correction $\Delta\rho_{\mathrm{rel}}(a)$ modifies the coarse-grained
Friedmann dynamics in a way that can mimic the observational effects of dark energy.
In this section we analyse the conditions under which the RDCC relaxation term behaves
like an effective dark-energy component and drives accelerated expansion.

\subsection{Effective Dark-Energy Component}

We rewrite the coarse-grained Friedmann equation in the form
\begin{equation}
H^2(a) = \frac{8\pi G}{3}\,\big[\rho(a) + \rho_{\mathrm{DE}}^{\mathrm{eff}}(a)\big],
\end{equation}
where we identify
\begin{equation}
\rho_{\mathrm{DE}}^{\mathrm{eff}}(a) \equiv \Delta\rho_{\mathrm{rel}}(a).
\end{equation}
From the scaling relation
\begin{equation}
\Delta\rho_{\mathrm{rel}}(a) \propto \alphaR(a)\,H(a),
\end{equation}
we see that the effective dark-energy density is not a fundamental fluid, but a
dynamical quantity tied to the relaxation of inter-sector coherence.

To characterise its behaviour, we introduce an effective equation-of-state parameter
$w_{\mathrm{DE}}^{\mathrm{eff}}(a)$ via the continuity equation
\begin{equation}
\frac{\dd \rho_{\mathrm{DE}}^{\mathrm{eff}}}{\dd a}
= -\frac{3}{a}\big[1 + w_{\mathrm{DE}}^{\mathrm{eff}}(a)\big]\rho_{\mathrm{DE}}^{\mathrm{eff}}(a).
\end{equation}

\subsection{Acceleration Condition}

Cosmic acceleration occurs when the scale factor satisfies
\begin{equation}
\frac{\ddot{a}}{a} = -\frac{4\pi G}{3}\,\big[\rho_{\mathrm{tot}}(a) + 3p_{\mathrm{tot}}(a)\big] > 0,
\end{equation}
which, in the presence of an effective dark-energy component, requires
\begin{equation}
w_{\mathrm{DE}}^{\mathrm{eff}}(a) < -\frac{1}{3}
\end{equation}
once $\rho_{\mathrm{DE}}^{\mathrm{eff}}(a)$ becomes dynamically relevant.

The relaxation-induced term $\rho_{\mathrm{DE}}^{\mathrm{eff}}(a) \propto \alphaR(a)\,H(a)$
inherits its time dependence from both the expansion rate and the evolution of
$\alphaR(a)$.
During epochs in which $\GammaR(a) \propto \Ha$ is large, $\alphaR(a)$ evolves
significantly and the effective dark-energy density tracks the expansion.
As $\Ha$ decreases and the relaxation rate weakens, $\alphaR(a)$ approaches a constant
value and $\rho_{\mathrm{DE}}^{\mathrm{eff}}(a)$ tends toward an asymptotic regime.

\subsection{Asymptotic Regime and Dark-Energy Mimicry}

In the late-time regime where $\alphaR(a)$ has effectively frozen out, we can approximate
\begin{equation}
\alphaR(a) \simeq \alpha_{\infty} = \mathrm{const.}
\end{equation}
The effective dark-energy density then scales as
\begin{equation}
\rho_{\mathrm{DE}}^{\mathrm{eff}}(a) \propto H(a).
\end{equation}
In a regime where the expansion rate varies slowly, this behaviour can mimic an
approximately constant dark-energy density with $w_{\mathrm{DE}}^{\mathrm{eff}} \approx -1$,
reproducing the key phenomenological feature of $\Lambda$CDM without introducing a
fundamental cosmological constant.

The emergent dark-energy behaviour in RDCC is therefore a consequence of the interplay
between inter-sector relaxation, the Hubble-scale memory time, and the freeze-out of
$\alphaR(a)$.
In the next section, we show how this picture naturally addresses the ``Why now?''
problem by linking the onset of acceleration to the dynamical crossing of the relaxation
and expansion timescales.


\section{The ``Why now?'' Problem in RDCC}

One of the longstanding conceptual challenges of dark-energy cosmology is the
``Why now?'' problem: why does the onset of cosmic acceleration occur precisely
at the epoch in which matter and dark-energy densities are of comparable
magnitude? In $\Lambda$CDM this coincidence appears accidental, as the
cosmological constant is time-independent while the matter density redshifts as
$a^{-3}$.

In the RDCC framework, the timing of acceleration is not a coincidence but a
structural consequence of the relaxation dynamics. The key observation is that
the relaxation rate satisfies
\begin{equation}
\GammaR(a) \propto \Ha,
\end{equation}
so that the ratio of the relaxation and expansion timescales is
\begin{equation}
\frac{\GammaR(a)}{H(a)} = \mathrm{const.}
\end{equation}
up to dimensionless coefficients determined by the stress-tensor correlators.

\subsection{Crossing of Relaxation and Expansion Timescales}

The evolution of the coherence parameter $\alphaR(a)$ is governed by
\begin{equation}
\alphaDot(a) = \GammaR(a)\,\alphaR(a).
\end{equation}
As long as $\GammaR(a)$ is large compared to the rate of change of $H(a)$,
the relaxation process is efficient and $\alphaR(a)$ evolves significantly.
However, as the Universe expands and $H(a)$ decreases, the relaxation rate
weakens proportionally, eventually leading to a regime in which
\begin{equation}
\GammaR(a) \ll H(a).
\end{equation}
In this regime, the relaxation becomes ineffective and $\alphaR(a)$ freezes out.

The onset of freeze-out therefore occurs when the relaxation and expansion
timescales cross:
\begin{equation}
\GammaR(a_{\mathrm{fr}}) \sim H(a_{\mathrm{fr}}).
\end{equation}
This crossing is not fine-tuned; it is a direct consequence of the proportionality
$\GammaR(a) \propto H(a)$.

\subsection{Natural Timing of Acceleration}

The effective dark-energy density scales as
\begin{equation}
\rho_{\mathrm{DE}}^{\mathrm{eff}}(a) \propto \alphaR(a)\,H(a).
\end{equation}
Before freeze-out, $\alphaR(a)$ evolves rapidly and the effective dark-energy
density tracks the expansion rate. After freeze-out, $\alphaR(a)$ becomes
approximately constant, and the effective dark-energy density enters an
asymptotic regime that can mimic a cosmological constant.

The onset of acceleration therefore coincides with the freeze-out of
$\alphaR(a)$:
\begin{equation}
a_{\mathrm{acc}} \simeq a_{\mathrm{fr}}.
\end{equation}
This provides a natural explanation for the timing of cosmic acceleration:
it occurs precisely when the relaxation dynamics become inefficient, a
structural consequence of the Hubble-scale memory time.

\subsection{Resolution of the Coincidence Problem}

In RDCC, the ``Why now?'' problem is resolved without introducing a new
energy component or fine-tuning parameters. The coincidence between the
onset of acceleration and the present epoch arises because both are governed
by the same dynamical mechanism: the proportionality between the relaxation
rate and the Hubble expansion rate.

The RDCC framework therefore replaces the coincidence with a structural
relation:
\begin{equation}
\text{Acceleration begins when relaxation freezes out.}
\end{equation}
This provides a unified and dynamical explanation for the timing of cosmic
acceleration within the RDCC paradigm.


\section{Predictions and Observational Signatures}

The reinterpretation of dark energy as emergent relaxation dynamics leads to several
distinctive observational signatures. While the effective late-time behaviour of the
RDCC relaxation term can mimic a cosmological constant, the underlying dynamics differ
from $\Lambda$CDM in ways that may be testable with current or future observations.

\subsection{Deviation from a Constant Equation of State}

In $\Lambda$CDM, the dark-energy equation of state is strictly constant,
\begin{equation}
w_{\Lambda} = -1.
\end{equation}
In RDCC, the effective dark-energy density is given by
\begin{equation}
\rho_{\mathrm{DE}}^{\mathrm{eff}}(a) \propto \alphaR(a)\,H(a),
\end{equation}
which implies a dynamical equation-of-state parameter
\begin{equation}
w_{\mathrm{DE}}^{\mathrm{eff}}(a) = -1 + \mathcal{O}\!\left(\frac{\dd \ln H}{\dd \ln a}\right).
\end{equation}
During the freeze-out of $\alphaR(a)$, small deviations from $w = -1$ may occur.
These deviations are expected to be mild but could be detectable with high-precision
late-time cosmological probes.

\subsection{Late-Time Transition in the Expansion Rate}

The freeze-out of $\alphaR(a)$ induces a transition in the effective expansion dynamics.
Before freeze-out, the relaxation term tracks the Hubble rate; after freeze-out, it
approaches an asymptotic regime. This leads to a characteristic feature:
\begin{equation}
\frac{\dd H}{\dd a} \quad \text{changes slope near} \quad a \simeq a_{\mathrm{fr}}.
\end{equation}
Such a transition could manifest as a mild redshift-dependent deviation from the
$\Lambda$CDM expansion history, potentially observable in:
\begin{itemize}
    \item baryon acoustic oscillations (BAO),
    \item Type Ia supernovae,
    \item redshift drift measurements,
    \item and future standard-siren observations.
\end{itemize}

\subsection{Tensor-Sector Implications}

The relaxation dynamics are tied to the same inter-sector structure that generates
log-periodic modulation in the primordial tensor spectrum, as discussed in
Companions~II--V. If the effective dark-energy behaviour arises from the freeze-out
of $\alphaR(a)$, then the timing of this freeze-out may correlate with the amplitude
or phase of tensor-sector modulation.

This leads to a potential cross-correlation signature:
\begin{equation}
\text{Late-time acceleration} \quad \Longleftrightarrow \quad
\text{specific tensor-modulation phase}.
\end{equation}
Such a correlation would be a distinctive prediction of RDCC and would not arise in
$\Lambda$CDM or in models with fundamental dark-energy fields.

\subsection{Absence of a True Cosmological Constant}

A key observational consequence of RDCC is the absence of a fundamental cosmological
constant. Instead, the effective dark-energy density arises dynamically from the
relaxation process. This implies that:
\begin{equation}
\lim_{a \to \infty} \rho_{\mathrm{DE}}^{\mathrm{eff}}(a)
\quad \text{need not be strictly constant.}
\end{equation}
Future measurements of the late-time expansion rate, especially those sensitive to
the asymptotic behaviour of $H(a)$, may therefore distinguish RDCC from models with
a true cosmological constant.

\subsection{Summary of Observational Signatures}

The RDCC interpretation of dark energy predicts:
\begin{itemize}
    \item mild deviations from $w = -1$ during the freeze-out of $\alphaR(a)$,
    \item a characteristic transition in the slope of $H(a)$ near $a_{\mathrm{fr}}$,
    \item possible correlations between late-time acceleration and tensor-sector modulation,
    \item and the absence of a strictly constant asymptotic dark-energy density.
\end{itemize}

These signatures provide potential avenues for testing the RDCC framework and
distinguishing emergent relaxation dynamics from a fundamental cosmological constant.


\section{Discussion and Outlook}

In this Companion we have developed a reinterpretation of dark energy within the
Relaxation-Driven Cyclic Cosmology framework. Building on the microscopic
relaxation dynamics derived in Companion~X, we have shown that the suppression of
inter-sector coherence induces an effective modification of the coarse-grained
Friedmann equations. This modification behaves like a dynamical dark-energy
component, even though no new fundamental energy density is introduced.

The key result is that the relaxation rate satisfies
\begin{equation}
\GammaR(a) \propto \Ha,
\end{equation}
which ties the relaxation dynamics directly to the expansion rate. As the Universe
expands and $H(a)$ decreases, the relaxation becomes inefficient and the coherence
parameter $\alphaR(a)$ freezes out. This freeze-out naturally produces an
asymptotic expansion regime that can mimic a cosmological constant, providing a
structural explanation for late-time acceleration.

A particularly noteworthy consequence is the resolution of the ``Why now?''
problem. In RDCC, the onset of acceleration coincides with the freeze-out of
$\alphaR(a)$, which occurs when the relaxation and expansion timescales cross.
This timing is not fine-tuned but follows directly from the proportionality
$\GammaR(a) \propto H(a)$ and the Hubble-scale memory time of the relaxation
kernel.

The emergent dark-energy behaviour predicted by RDCC leads to several potential
observational signatures, including mild deviations from $w = -1$, a characteristic
transition in the slope of $H(a)$ near the freeze-out epoch, and correlations with
tensor-sector modulation. These signatures provide avenues for distinguishing RDCC
from $\Lambda$CDM and other dynamical dark-energy models.

\subsection*{Future Directions}

The results presented here open several directions for further development:

\begin{itemize}
    \item \textbf{Modified perturbation theory.}
    A full treatment of scalar, vector, and tensor perturbations in the presence of
    relaxation-induced corrections may reveal additional observational signatures.

    \item \textbf{Non-Markovian effects.}
    The derivation in Companion~X relies on the Born--Markov approximation.
    Including non-Markovian corrections may refine the effective dark-energy
    behaviour, especially near the freeze-out epoch.

    \item \textbf{Numerical implementation.}
    Incorporating the relaxation correction $\Delta\rho_{\mathrm{rel}}(a)$ into a
    Boltzmann code would allow for global parameter scans and quantitative
    comparison with cosmological data.

    \item \textbf{Information-theoretic interpretation.}
    The connection between decoherence, emergent spacetime, and effective
    acceleration suggests a deeper information-theoretic structure underlying
    cosmic expansion.
\end{itemize}

In summary, RDCC provides a unified and structurally motivated reinterpretation of
dark energy as emergent relaxation dynamics. This perspective integrates cosmic
acceleration into the same framework that governs inter-sector decoherence,
tensor-sector modulation, and the emergence of classical spacetime, offering a
coherent and testable alternative to models that introduce dark energy as a
fundamental component.


\begin{thebibliography}{99}

% - Standard Cosmology and Dark Energy -
\bibitem{Planck2018}
Planck Collaboration,
\textit{Planck 2018 results. VI. Cosmological parameters} (2018).

\bibitem{Riess1998}
A.~G.~Riess et al.,
\textit{Observational evidence from supernovae for an accelerating universe} (1998).

\bibitem{Perlmutter1999}
S.~Perlmutter et al.,
\textit{Measurements of $\Omega$ and $\Lambda$ from 42 high-redshift supernovae} (1999).

\bibitem{WeinbergLambda}
S.~Weinberg,
\textit{The cosmological constant problem}.

% - Dark Energy Reviews -
\bibitem{DEreview}
E.~J.~Copeland, M.~Sami, S.~Tsujikawa,
\textit{Dynamics of dark energy}.

% - Decoherence, Lindblad, and Open Quantum Systems -
\bibitem{BreuerPetruccione}
H.-P.~Breuer, F.~Petruccione,
\textit{The Theory of Open Quantum Systems}.

\bibitem{Lindblad1976}
G.~Lindblad,
\textit{On the generators of quantum dynamical semigroups}.

\bibitem{CalzettaHu}
E.~Calzetta, B.~L.~Hu,
\textit{Nonequilibrium Quantum Field Theory}.

% - Stress-Tensor Correlators and Cosmological Coarse-Graining -
\bibitem{Ford1997}
L.~H.~Ford,
\textit{Stress tensor fluctuations and passive quantum gravity}.

\bibitem{HuVerdaguer}
B.~L.~Hu, E.~Verdaguer,
\textit{Stochastic gravity: Theory and applications}.

% - RDCC Internal References -
\bibitem{RDCCFlagship}
M.~Lehmann,
\textit{RDCC Flagship Paper}.

\bibitem{RDCCCompVI}
M.~Lehmann,
\textit{Companion VI: Relaxation Dynamics in RDCC}.

\bibitem{RDCCCompX}
M.~Lehmann,
\textit{Companion X: Microscopic Derivation of the RDCC Relaxation Rate}.

\bibitem{RDCCCompIItoV}
M.~Lehmann,
\textit{Companions II–V: Tensor-Sector Modulation in RDCC}.

\end{thebibliography}


\section*{Acknowledgments}

The author thanks the scientific community for ongoing discussions on the nature of
cosmic acceleration and the conceptual foundations of cosmology. The development of
the RDCC framework has benefited from insights across quantum information theory,
open quantum systems, and modern cosmology. The author also acknowledges the value
of open-science platforms that support transparent dissemination and iterative
refinement of theoretical work.

\end{document}