% ---------------------------------------------------------
% PREAMBLE
% ---------------------------------------------------------
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{physics}
\usepackage{cite}
\usepackage{float}

\geometry{margin=1in}

\title{
\textbf{Companion XV}\\[4mm]
\textbf{The Early-Universe Gravitational-Wave Spectrum in the}\\
\textbf{Relaxation-Driven Cyclic Cosmology}\\[2mm]
\textbf{Relaxation-Induced Damping, Sectoral Asymmetry, and Observable Signatures}
}

\author{
Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}
}

\date{April 20, 2026}

\begin{document}

\maketitle

% ---------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------
\begin{abstract}
The early-Universe gravitational-wave (GW) spectrum provides a unique window into the 
microphysical processes that shaped the primordial plasma. In the Relaxation-Driven Cyclic 
Cosmology (RDCC), the universal relaxation mechanism introduced in Companion~X modifies the 
evolution of tensor perturbations during the earliest stages of cosmic history. The relaxation 
rate $\Gamma(a)$ introduces an additional damping term in the tensor-mode equation of motion, 
suppressing the amplitude of primordial gravitational waves on scales that re-enter the horizon 
during the relaxation-dominated epoch. Combined with the sectoral temperature asymmetry 
between the two CPT-conjugate sectors, this leads to a characteristic sectoral GW asymmetry 
and a relaxation-imprinted suppression feature in the stochastic gravitational-wave background 
(SGWB). 

In this Companion we derive the relaxation-corrected tensor-mode evolution, compute the 
resulting GW transfer function, and obtain the full early-Universe GW spectrum in RDCC. We 
identify the frequency range in which relaxation-induced damping is observationally relevant, 
quantify the sectoral GW asymmetry, and provide concrete predictions for upcoming 
space-based interferometers. The results establish the SGWB as a sensitive probe of 
relaxation-driven early-Universe physics and extend the RDCC framework into the gravitational 
wave domain.
\end{abstract}

% ---------------------------------------------------------
% SECTION 1 — MOTIVATION
% ---------------------------------------------------------
\section{Motivation}

The stochastic gravitational-wave background (SGWB) generated in the early Universe is a 
powerful probe of high-energy physics, providing access to epochs far beyond the reach of 
electromagnetic observations. In standard cosmology, the SGWB is determined by the primordial 
tensor spectrum and its subsequent evolution through the radiation- and matter-dominated eras. 
The resulting spectrum is smooth, nearly scale-invariant, and free of dynamical features except 
for those associated with changes in the relativistic degrees of freedom.

In the Relaxation-Driven Cyclic Cosmology (RDCC), the situation is fundamentally different. 
The universal relaxation mechanism introduced in Companion~X modifies the evolution of 
super-horizon tensor modes during the earliest stages of cosmic history. When the relaxation 
rate satisfies $\Gamma(a) \gtrsim H(a)$, tensor perturbations experience an additional damping 
term that suppresses their amplitude prior to horizon re-entry. This effect leaves a distinct 
imprint on the SGWB, producing a characteristic suppression feature at frequencies 
corresponding to modes that re-entered the horizon during the relaxation-dominated epoch.

A second key ingredient is the sectoral temperature asymmetry between the two CPT-conjugate 
sectors. As shown in Companion~X, the temperature ratio $T_B/T_A \simeq 0.5$ implies different 
horizon masses, expansion histories, and damping scales in the two sectors. Tensor modes in the 
colder sector experience a different relaxation history, leading to a sectoral asymmetry in the 
GW spectrum. While only the visible-sector spectrum is directly observable, the mirror-sector 
spectrum influences the global energy balance and provides an internal consistency check of the 
RDCC framework.

The purpose of this Companion is to develop the early-Universe gravitational-wave sector of 
RDCC. We derive the relaxation-corrected tensor-mode equation, compute the resulting transfer 
function, and obtain the full SGWB spectrum including the relaxation-induced suppression 
feature. We identify the frequency range in which the effect is observationally relevant and 
provide concrete predictions for upcoming space-based interferometers such as LISA, DECIGO, 
and BBO.

Understanding the GW spectrum in RDCC is essential for three reasons:

\begin{itemize}
    \item \textbf{Early-Universe dynamics:}  
    Tensor modes directly probe the relaxation-dominated epoch, providing a clean test of the 
    universal relaxation mechanism.

    \item \textbf{Sectoral structure:}  
    The temperature asymmetry between the two sectors leads to a sectoral GW asymmetry, 
    offering an independent probe of the bipartite global state.

    \item \textbf{Observational leverage:}  
    The relaxation-induced suppression feature lies in the frequency range accessible to 
    next-generation GW detectors, enabling direct observational tests of RDCC.
\end{itemize}

This Companion establishes the SGWB as a sensitive probe of relaxation-driven early-Universe 
physics and extends the RDCC framework into the gravitational-wave domain.

% ---------------------------------------------------------
% SECTION 2 — TENSOR MODES IN STANDARD COSMOLOGY (KURZÜBERBLICK)
% ---------------------------------------------------------
\section{Tensor Modes in Standard Cosmology (Kurzüberblick)}

In standard cosmology, primordial tensor perturbations originate from quantum fluctuations 
during inflation and evolve linearly throughout the early Universe. Their dynamics are 
governed by the tensor-mode equation
\begin{equation}
    \ddot{h}_k + 3H\dot{h}_k + \frac{k^2}{a^2} h_k = 0,
\end{equation}
where $h_k$ denotes the Fourier amplitude of the transverse-traceless metric perturbation. 
The evolution naturally separates into two regimes:

\begin{itemize}
    \item \textbf{Super-horizon regime ($k \ll aH$):}  
    Tensor modes freeze out and maintain an approximately constant amplitude.

    \item \textbf{Sub-horizon regime ($k \gg aH$):}  
    Tensor modes oscillate with frequency $k/a$ and their amplitude decays as $a^{-1}$ due to 
    Hubble friction.
\end{itemize}

The primordial tensor power spectrum is typically parameterized as
\begin{equation}
    \mathcal{P}_h(k) = A_t \left( \frac{k}{k_*} \right)^{n_t},
\end{equation}
with $A_t$ the tensor amplitude and $n_t$ the tensor tilt. In standard cosmology, the 
subsequent evolution of the stochastic gravitational-wave background (SGWB) is determined by 
the transfer function $T(k)$, which encodes the effects of horizon re-entry, changes in the 
relativistic degrees of freedom, and the transition from radiation to matter domination.

The present-day energy density in gravitational waves is given by
\begin{equation}
    \Omega_{\mathrm{GW}}(k) 
    = \frac{1}{12} \left( \frac{k}{a_0 H_0} \right)^2 
      \mathcal{P}_h(k)\, T^2(k),
\end{equation}
where $a_0$ and $H_0$ denote the present-day scale factor and Hubble parameter. In the 
standard picture, the transfer function is smooth and featureless except for small steps 
associated with changes in $g_*$.

Crucially, the standard tensor-mode evolution contains \emph{no additional damping terms} 
beyond Hubble friction. As a result, the SGWB spectrum is free of dynamical suppression 
features and retains a nearly scale-invariant shape across many decades in frequency.

This simplicity makes the SGWB an exceptionally sensitive probe of any new early-Universe 
physics that modifies tensor-mode evolution. In the Relaxation-Driven Cyclic Cosmology (RDCC), 
the universal relaxation mechanism introduces precisely such a modification, as we develop in 
the next section.

% ---------------------------------------------------------
% SECTION 3 — TENSOR MODES IN RDCC: RELAXATION-INDUCED DAMPING
% ---------------------------------------------------------
\section{Tensor Modes in RDCC: Relaxation-Induced Damping}

In the Relaxation-Driven Cyclic Cosmology (RDCC), the universal relaxation mechanism 
introduced in Companion~X modifies the evolution of tensor perturbations during the earliest 
stages of cosmic history. The relaxation rate $\Gamma(a)$ governs the flow of information 
between the two CPT-conjugate sectors and introduces an additional damping term in the 
tensor-mode equation of motion. This effect is absent in standard cosmology and leads to a 
distinctive suppression feature in the stochastic gravitational-wave background (SGWB).

\subsection*{3.1 Tensor-Mode Equation with Relaxation}

The evolution of a tensor mode $h_k$ in RDCC is governed by the modified equation
\begin{equation}
    \ddot{h}_k + \left( 3H + \Gamma \right)\dot{h}_k 
    + \frac{k^2}{a^2} h_k = 0,
\end{equation}
where the additional term $\Gamma \dot{h}_k$ represents relaxation-induced damping.  
The relaxation rate satisfies
\begin{equation}
    \Gamma(a) = \alpha(a)\, H(a),
\end{equation}
with $\alpha(a)$ a slowly varying dimensionless coefficient of order 
$\mathcal{O}(0.1 - 1)$ during the relaxation-dominated epoch.

When $\Gamma \gg H$, the damping term dominates and suppresses the amplitude of tensor modes 
even while they remain super-horizon. This is in stark contrast to standard cosmology, where 
super-horizon tensor modes remain constant.

\subsection*{3.2 Super-Horizon Evolution}

In the regime $k \ll aH$, the tensor-mode equation reduces to
\begin{equation}
    \ddot{h}_k + \left( 3 + \alpha \right) H \dot{h}_k \simeq 0.
\end{equation}
Assuming $H \propto a^{-2}$ during radiation domination, the solution takes the form
\begin{equation}
    h_k(a) \propto a^{-\alpha/2}.
\end{equation}
Thus, super-horizon tensor modes are suppressed by a factor
\begin{equation}
    S(a) = \left( \frac{a}{a_{\mathrm{in}}} \right)^{-\alpha/2},
\end{equation}
where $a_{\mathrm{in}}$ denotes the scale factor at the onset of the relaxation-dominated 
epoch. This suppression persists until $\Gamma(a)$ drops below $H(a)$.

\subsection*{3.3 Sub-Horizon Evolution}

Once a mode re-enters the horizon ($k \gg aH$), the equation becomes
\begin{equation}
    \ddot{h}_k + \left( 3H + \Gamma \right)\dot{h}_k 
    + \frac{k^2}{a^2} h_k = 0.
\end{equation}
In this regime, the mode oscillates with frequency $k/a$, while its amplitude decays as
\begin{equation}
    h_k(a) \propto a^{-1} \exp\!\left[ -\int \frac{\Gamma}{2H} \frac{da}{a} \right].
\end{equation}
The exponential factor represents the cumulative effect of relaxation-induced damping during 
and after horizon re-entry.

\subsection*{3.4 Physical Interpretation}

The relaxation mechanism affects tensor modes in two ways:

\begin{itemize}
    \item \textbf{Super-horizon suppression:}  
    Tensor modes lose amplitude before horizon re-entry, reducing their contribution to the 
    SGWB at the corresponding frequencies.

    \item \textbf{Enhanced sub-horizon damping:}  
    The additional friction term $\Gamma \dot{h}_k$ accelerates the decay of tensor modes 
    after horizon re-entry.
\end{itemize}

Both effects lead to a characteristic suppression feature in the SGWB spectrum, with the 
suppression scale determined by the epoch in which $\Gamma(a) \sim H(a)$.

\subsection*{3.5 Sectoral Asymmetry}

Because the two CPT-conjugate sectors have different temperatures, their relaxation histories 
differ. The colder sector~B has a smaller relaxation coefficient,
\begin{equation}
    \alpha_B(a) < \alpha_A(a),
\end{equation}
leading to weaker damping and a larger tensor amplitude. This results in a sectoral GW 
asymmetry, providing an independent probe of the bipartite global state.

In the next section, we compute the relaxation-corrected transfer function and derive the 
resulting SGWB spectrum.

% ---------------------------------------------------------
% SECTION 4 — RELAXATION-CORRECTED TRANSFER FUNCTION
% ---------------------------------------------------------
\section{Relaxation-Corrected Transfer Function}

The transfer function $T(k)$ encodes the evolution of tensor modes from their primordial 
generation to the present day. In standard cosmology, $T(k)$ is determined by horizon 
re-entry, changes in the relativistic degrees of freedom, and the transition from radiation 
to matter domination. In the Relaxation-Driven Cyclic Cosmology (RDCC), the universal 
relaxation mechanism introduces an additional suppression factor that modifies the transfer 
function on scales that re-entered the horizon during the relaxation-dominated epoch.

\subsection*{4.1 Definition of the Transfer Function}

The present-day tensor power spectrum is given by
\begin{equation}
    \mathcal{P}_h(k, a_0) 
    = \mathcal{P}_h(k, a_{\mathrm{prim}})\, T^2(k),
\end{equation}
where $\mathcal{P}_h(k, a_{\mathrm{prim}})$ is the primordial tensor spectrum and $T(k)$ 
encodes all subsequent evolution. In RDCC, the transfer function factorizes as
\begin{equation}
    T(k) = T_{\mathrm{std}}(k)\, T_{\mathrm{rel}}(k),
\end{equation}
where $T_{\mathrm{std}}(k)$ is the standard transfer function and $T_{\mathrm{rel}}(k)$ 
captures the relaxation-induced suppression.

\subsection*{4.2 Relaxation-Induced Suppression Factor}

From the tensor-mode evolution derived in Section~3, the amplitude of a tensor mode is 
suppressed by two effects:

\begin{itemize}
    \item super-horizon suppression by the factor $S(a)$,
    \item enhanced sub-horizon damping due to the additional friction term $\Gamma \dot{h}_k$.
\end{itemize}

The combined suppression factor can be written as
\begin{equation}
    T_{\mathrm{rel}}(k)
    = S(a_{\mathrm{re}})\,
      \exp\!\left[
        -\int_{a_{\mathrm{re}}}^{a_{\mathrm{end}}}
        \frac{\Gamma(a)}{2H(a)} \frac{da}{a}
      \right],
\end{equation}
where $a_{\mathrm{re}}$ is the scale factor at horizon re-entry and $a_{\mathrm{end}}$ marks 
the end of the relaxation-dominated epoch.

Using $\Gamma(a) = \alpha(a) H(a)$, this becomes
\begin{equation}
    T_{\mathrm{rel}}(k)
    = S(a_{\mathrm{re}})\,
      \exp\!\left[
        -\frac{1}{2}
        \int_{a_{\mathrm{re}}}^{a_{\mathrm{end}}}
        \alpha(a)\, \frac{da}{a}
      \right].
\end{equation}

If $\alpha(a)$ varies slowly, the integral simplifies to
\begin{equation}
    T_{\mathrm{rel}}(k)
    \simeq S(a_{\mathrm{re}})\,
    \left( \frac{a_{\mathrm{re}}}{a_{\mathrm{end}}} \right)^{\alpha/2}.
\end{equation}

\subsection*{4.3 Scale Dependence}

The relaxation-induced suppression affects only modes that re-entered the horizon while 
$\Gamma(a) \gtrsim H(a)$. Defining the characteristic scale $k_{\mathrm{rel}}$ by
\begin{equation}
    k_{\mathrm{rel}} = a_{\mathrm{rel}} H(a_{\mathrm{rel}}),
\end{equation}
where $a_{\mathrm{rel}}$ satisfies $\Gamma(a_{\mathrm{rel}}) = H(a_{\mathrm{rel}})$, we obtain:

\begin{itemize}
    \item \textbf{$k \ll k_{\mathrm{rel}}$:}  
    Modes re-enter after relaxation ends.  
    No suppression: $T_{\mathrm{rel}}(k) \simeq 1$.

    \item \textbf{$k \gg k_{\mathrm{rel}}$:}  
    Modes re-enter during relaxation.  
    Strong suppression: $T_{\mathrm{rel}}(k) \propto k^{-\alpha/2}$.
\end{itemize}

Thus, the relaxation mechanism introduces a \emph{power-law suppression feature} in the SGWB 
spectrum at high frequencies.

\subsection*{4.4 Sectoral Asymmetry}

Because the two CPT-conjugate sectors have different temperatures, their relaxation 
coefficients differ:
\begin{equation}
    \alpha_B(a) < \alpha_A(a).
\end{equation}
This implies
\begin{equation}
    T_{\mathrm{rel}}^{(B)}(k) > T_{\mathrm{rel}}^{(A)}(k),
\end{equation}
leading to a sectoral GW asymmetry. While only the visible-sector spectrum is directly 
observable, the mirror-sector spectrum provides an internal consistency check of the RDCC 
framework.

\subsection*{4.5 Resulting Transfer Function}

Combining the standard and relaxation-induced contributions yields
\begin{equation}
    T(k) = T_{\mathrm{std}}(k)\,
           S(a_{\mathrm{re}})\,
           \exp\!\left[
             -\frac{1}{2}
             \int_{a_{\mathrm{re}}}^{a_{\mathrm{end}}}
             \alpha(a)\, \frac{da}{a}
           \right].
\end{equation}

This transfer function forms the basis for computing the full stochastic gravitational-wave 
background in RDCC, which we derive in the next section.

% ---------------------------------------------------------
% SECTION 5 — THE EARLY-UNIVERSE GW SPECTRUM IN RDCC
% ---------------------------------------------------------
\section{The Early-Universe GW Spectrum in RDCC}

With the relaxation-corrected transfer function derived in Section~4, we now compute the full 
stochastic gravitational-wave background (SGWB) in the Relaxation-Driven Cyclic Cosmology 
(RDCC). The resulting spectrum exhibits a characteristic suppression feature at high 
frequencies, reflecting the damping of tensor modes during the relaxation-dominated epoch. 
This feature is absent in standard cosmology and provides a direct observational probe of 
relaxation-driven early-Universe physics.

\subsection*{5.1 General Expression}

The present-day GW energy density per logarithmic interval in wavenumber is given by
\begin{equation}
    \Omega_{\mathrm{GW}}(k)
    = \frac{1}{12}
      \left( \frac{k}{a_0 H_0} \right)^2
      \mathcal{P}_h(k)\,
      T^2(k),
\end{equation}
where $\mathcal{P}_h(k)$ is the primordial tensor spectrum and $T(k)$ is the full transfer 
function,
\begin{equation}
    T(k) = T_{\mathrm{std}}(k)\, T_{\mathrm{rel}}(k).
\end{equation}

The primordial spectrum is parameterized as
\begin{equation}
    \mathcal{P}_h(k) = A_t \left( \frac{k}{k_*} \right)^{n_t},
\end{equation}
with $A_t$ the tensor amplitude and $n_t$ the tensor tilt.

\subsection*{5.2 Relaxation-Induced Suppression Feature}

Using the relaxation-corrected transfer function from Section~4, the GW spectrum becomes
\begin{equation}
    \Omega_{\mathrm{GW}}(k)
    = \Omega_{\mathrm{GW}}^{\mathrm{std}}(k)\,
      T_{\mathrm{rel}}^2(k),
\end{equation}
where $\Omega_{\mathrm{GW}}^{\mathrm{std}}(k)$ is the standard spectrum and
\begin{equation}
    T_{\mathrm{rel}}(k)
    \simeq
    \begin{cases}
        1, & k \ll k_{\mathrm{rel}}, \\[2mm]
        \left( \dfrac{k}{k_{\mathrm{rel}}} \right)^{-\alpha/2}, & k \gg k_{\mathrm{rel}}.
    \end{cases}
\end{equation}

Thus, the relaxation mechanism introduces a power-law suppression at high frequencies:
\begin{equation}
    \Omega_{\mathrm{GW}}(k) \propto k^{n_t - \alpha}.
\end{equation}

The break scale $k_{\mathrm{rel}}$ corresponds to the mode that re-entered the horizon when 
$\Gamma(a) = H(a)$.

\subsection*{5.3 Frequency Range of the Suppression Feature}

The characteristic frequency associated with $k_{\mathrm{rel}}$ is
\begin{equation}
    f_{\mathrm{rel}}
    = \frac{k_{\mathrm{rel}}}{2\pi a_0}
    = \frac{a_{\mathrm{rel}} H(a_{\mathrm{rel}})}{2\pi a_0}.
\end{equation}

For typical RDCC parameters, the relaxation-dominated epoch occurs at temperatures
\begin{equation}
    T \sim 10^7 - 10^{10}\,\mathrm{GeV},
\end{equation}
corresponding to frequencies in the range
\begin{equation}
    f_{\mathrm{rel}} \sim 10^{-3} - 10^{2}\,\mathrm{Hz}.
\end{equation}

This range overlaps with the sensitivity windows of LISA, DECIGO, and BBO, making the 
relaxation-induced suppression feature directly observable.

\subsection*{5.4 Sectoral GW Asymmetry}

Because the two CPT-conjugate sectors have different temperatures, their relaxation 
coefficients differ:
\begin{equation}
    \alpha_B < \alpha_A.
\end{equation}

This implies
\begin{equation}
    \Omega_{\mathrm{GW}}^{(B)}(k)
    > \Omega_{\mathrm{GW}}^{(A)}(k)
\end{equation}
for $k \gg k_{\mathrm{rel}}$, leading to a sectoral GW asymmetry.

While only the visible-sector spectrum is directly observable, the mirror-sector spectrum 
affects the global energy balance and provides an internal consistency check of the RDCC 
framework.

\subsection*{5.5 Full RDCC GW Spectrum}

Combining all contributions, the RDCC GW spectrum takes the form
\begin{equation}
    \Omega_{\mathrm{GW}}(k)
    = \frac{1}{12}
      \left( \frac{k}{a_0 H_0} \right)^2
      A_t \left( \frac{k}{k_*} \right)^{n_t}
      T_{\mathrm{std}}^2(k)\,
      T_{\mathrm{rel}}^2(k),
\end{equation}
with
\begin{equation}
    T_{\mathrm{rel}}(k)
    \simeq
    \begin{cases}
        1, & k \ll k_{\mathrm{rel}}, \\[2mm]
        \left( \dfrac{k}{k_{\mathrm{rel}}} \right)^{-\alpha/2}, & k \gg k_{\mathrm{rel}}.
    \end{cases}
\end{equation}

This spectrum exhibits a smooth, nearly scale-invariant low-frequency regime and a 
relaxation-induced suppression at high frequencies. The location and slope of the suppression 
feature provide direct information about the relaxation dynamics and the temperature 
asymmetry between the two sectors.

In the next section, we translate these results into concrete observational predictions for 
upcoming gravitational-wave detectors.

% ---------------------------------------------------------
% SECTION 6 — OBSERVATIONAL PREDICTIONS
% ---------------------------------------------------------
\section{Observational Predictions}

The relaxation-corrected gravitational-wave (GW) spectrum derived in Section~5 leads to a set 
of concrete and testable predictions. These predictions arise from the interplay between the 
relaxation-induced damping of tensor modes, the sectoral temperature asymmetry, and the 
frequency dependence of horizon re-entry. In this section we summarize the most important 
observational consequences for space-based interferometers and high-frequency GW searches.

\subsection*{Prediction 1: High-Frequency Suppression Feature}

The universal relaxation mechanism introduces a characteristic suppression feature in the 
stochastic gravitational-wave background (SGWB) at frequencies
\begin{equation}
    f \gtrsim f_{\mathrm{rel}} \sim 10^{-3} - 10^{2}\,\mathrm{Hz}.
\end{equation}
Modes that re-entered the horizon during the relaxation-dominated epoch experience a 
power-law suppression,
\begin{equation}
    \Omega_{\mathrm{GW}}(f) \propto f^{n_t - \alpha},
\end{equation}
with $\alpha$ the relaxation coefficient. This feature is absent in standard cosmology and 
provides a direct probe of relaxation-driven early-Universe physics.

\subsection*{Prediction 2: Observable Break Scale}

The break frequency $f_{\mathrm{rel}}$ marks the transition between the unsuppressed and 
suppressed regimes. For typical RDCC parameters, $f_{\mathrm{rel}}$ lies within the sensitivity 
windows of LISA, DECIGO, and BBO. A detection of a spectral break at this frequency would 
provide strong evidence for the relaxation mechanism.

\subsection*{Prediction 3: Modified Spectral Slope}

Above the break scale, the GW spectrum exhibits a modified slope,
\begin{equation}
    \frac{d\ln \Omega_{\mathrm{GW}}}{d\ln f} = n_t - \alpha,
\end{equation}
where $\alpha$ is of order $\mathcal{O}(0.1 - 1)$ during the relaxation epoch. This leads to a 
significant deviation from the nearly scale-invariant spectrum predicted by standard 
cosmology.

\subsection*{Prediction 4: Sectoral GW Asymmetry}

Because the two CPT-conjugate sectors have different temperatures, their relaxation 
coefficients differ:
\begin{equation}
    \alpha_B < \alpha_A.
\end{equation}
This implies a sectoral asymmetry in the GW spectrum,
\begin{equation}
    \Omega_{\mathrm{GW}}^{(B)}(f) > \Omega_{\mathrm{GW}}^{(A)}(f)
\end{equation}
for $f \gtrsim f_{\mathrm{rel}}$. While only the visible-sector spectrum is directly observable, 
the mirror-sector spectrum affects the global energy budget and provides an internal 
consistency check of the RDCC framework.

\subsection*{Prediction 5: Consistency with PBH Constraints}

The relaxation-induced suppression of tensor modes is consistent with the suppression of 
primordial black hole (PBH) formation derived in Companion~XIV. Both effects originate from 
the same relaxation dynamics and occur at similar energy scales. A joint analysis of PBH 
constraints and GW observations can therefore test the internal coherence of RDCC.

\subsection*{Prediction 6: Distinguishability from Other New-Physics Scenarios}

The relaxation-induced suppression feature differs qualitatively from other mechanisms that 
modify the SGWB, such as:

\begin{itemize}
    \item changes in relativistic degrees of freedom ($g_*$),
    \item phase transitions,
    \item cosmic strings,
    \item reheating effects.
\end{itemize}

These scenarios typically produce sharp features, bumps, or oscillations, whereas RDCC predicts 
a smooth power-law suppression with a well-defined break scale. This distinct spectral shape 
allows the relaxation mechanism to be distinguished from other early-Universe physics.

\subsection*{Prediction 7: Direct Testability with Future Detectors}

The predicted suppression feature lies in the frequency range accessible to upcoming 
space-based interferometers. In particular:

\begin{itemize}
    \item \textbf{LISA} can probe the low-frequency edge of the suppression feature.
    \item \textbf{DECIGO} and \textbf{BBO} can probe the full suppression regime.
    \item \textbf{High-frequency detectors} (MHz–GHz) can test the extreme tail of the spectrum.
\end{itemize}

A detection of the predicted spectral break and modified slope would provide strong evidence 
for the relaxation mechanism and the bipartite structure of RDCC.

Together, these predictions establish the SGWB as a sensitive probe of relaxation-driven 
early-Universe physics and provide multiple observational avenues for testing the RDCC 
framework.

% ---------------------------------------------------------
% SECTION 7 — CONCLUSION
% ---------------------------------------------------------
\section{Conclusion}

In this Companion we have developed the gravitational-wave (GW) sector of the 
Relaxation-Driven Cyclic Cosmology (RDCC), extending the framework into a domain that is 
directly accessible to upcoming space-based interferometers. Starting from the universal 
relaxation mechanism introduced in Companion~X, we derived the relaxation-corrected evolution 
of tensor perturbations, computed the corresponding transfer function, and obtained the full 
early-Universe stochastic gravitational-wave background (SGWB) in RDCC.

The key result is that the relaxation mechanism introduces a characteristic suppression 
feature in the GW spectrum at frequencies corresponding to modes that re-entered the horizon 
during the relaxation-dominated epoch. This suppression arises from two effects: the 
relaxation-induced damping of super-horizon tensor modes and the enhanced friction term 
affecting sub-horizon evolution. The resulting spectrum exhibits a smooth power-law 
suppression above a well-defined break scale, providing a clear observational signature of 
relaxation-driven early-Universe physics.

A second important result is the sectoral GW asymmetry. Because the two CPT-conjugate sectors 
have different temperatures, their relaxation histories differ, leading to distinct GW 
spectra. While only the visible-sector spectrum is directly observable, the mirror-sector 
spectrum contributes to the global energy balance and provides an internal consistency check 
of the RDCC framework. This sectoral asymmetry parallels the sectoral PBH asymmetry developed 
in Companion~XIV, demonstrating the coherence of the relaxation mechanism across multiple 
cosmological observables.

The predicted suppression feature lies within the sensitivity windows of LISA, DECIGO, and 
BBO, making the SGWB a powerful probe of RDCC. A detection of the spectral break and modified 
slope would provide strong evidence for the relaxation mechanism and the bipartite structure 
of the early Universe. Conversely, the absence of such a feature would place meaningful 
constraints on the relaxation coefficient and the duration of the relaxation-dominated epoch.

With the development of the GW sector, RDCC now provides a coherent and internally consistent 
description of scalar, tensor, and PBH-related early-Universe physics. The framework continues 
to yield a rich set of observational signatures, positioning the SGWB as a central tool for 
testing the relaxation-driven dynamics that define the RDCC cosmology.

% ---------------------------------------------------------
% APPENDIX A — DERIVATION OF THE RELAXATION-CORRECTED TENSOR EQUATION
% ---------------------------------------------------------
\appendix
\section*{Appendix A: Derivation of the Relaxation-Corrected Tensor Equation}
\addcontentsline{toc}{section}{Appendix A: Derivation of the Relaxation-Corrected Tensor Equation}

In this appendix we derive the relaxation-corrected evolution equation for tensor perturbations 
used in the main text. The key modification relative to standard cosmology arises from the 
universal relaxation mechanism introduced in Companion~X, which governs the flow of 
information between the two CPT-conjugate sectors and introduces an additional damping term 
in the tensor-mode dynamics.

\subsection*{A.1 Tensor Modes in Standard Cosmology}

We start from the perturbed Friedmann–Lemaître–Robertson–Walker (FLRW) metric with a 
transverse-traceless tensor perturbation $h_{ij}$,
\begin{equation}
    ds^2 = -dt^2 + a^2(t)\left( \delta_{ij} + h_{ij} \right) dx^i dx^j,
\end{equation}
with
\begin{equation}
    \partial_i h^{ij} = 0, 
    \qquad
    h^i_{\ i} = 0.
\end{equation}
In standard cosmology, the linearized Einstein equations yield the tensor-mode equation
\begin{equation}
    \ddot{h}_{ij} + 3H\dot{h}_{ij} - \frac{1}{a^2}\nabla^2 h_{ij} = 0.
\end{equation}
In Fourier space, this becomes
\begin{equation}
    \ddot{h}_k + 3H\dot{h}_k + \frac{k^2}{a^2} h_k = 0,
\end{equation}
where $h_k$ denotes the Fourier amplitude of a given polarization mode.

\subsection*{A.2 Inclusion of the Relaxation Mechanism}

In the Relaxation-Driven Cyclic Cosmology (RDCC), the universal relaxation mechanism 
introduces an additional channel for energy and information exchange between the two sectors. 
At the level of the effective description, this appears as an extra damping term in the 
equation of motion for tensor modes. The modified equation takes the form
\begin{equation}
    \ddot{h}_k + \left( 3H + \Gamma \right)\dot{h}_k 
    + \frac{k^2}{a^2} h_k = 0,
\end{equation}
where $\Gamma(a)$ is the relaxation rate.

The relaxation rate is parametrized as
\begin{equation}
    \Gamma(a) = \alpha(a)\, H(a),
\end{equation}
with $\alpha(a)$ a slowly varying dimensionless coefficient. During the relaxation-dominated 
epoch, $\alpha(a)$ is of order $\mathcal{O}(0.1 - 1)$, implying $\Gamma \gtrsim H$.

\subsection*{A.3 Super-Horizon Limit}

In the super-horizon regime, $k \ll aH$, the gradient term can be neglected and the equation 
reduces to
\begin{equation}
    \ddot{h}_k + \left( 3 + \alpha \right) H \dot{h}_k \simeq 0.
\end{equation}
During radiation domination, $H \propto a^{-2}$, and we can write
\begin{equation}
    \frac{d}{dt} = aH \frac{d}{da}.
\end{equation}
Using this, the equation becomes
\begin{equation}
    aH \frac{d}{da}
    \left( aH \frac{dh_k}{da} \right)
    + \left( 3 + \alpha \right) H
      \left( aH \frac{dh_k}{da} \right) \simeq 0.
\end{equation}
Neglecting the slow variation of $\alpha$ and using $H \propto a^{-2}$, one finds the 
approximate solution
\begin{equation}
    h_k(a) \propto a^{-\alpha/2}.
\end{equation}
Thus, super-horizon tensor modes are suppressed by the factor
\begin{equation}
    S(a) = \left( \frac{a}{a_{\mathrm{in}}} \right)^{-\alpha/2},
\end{equation}
where $a_{\mathrm{in}}$ denotes the scale factor at the onset of the relaxation-dominated 
epoch.

\subsection*{A.4 Sub-Horizon Limit}

In the sub-horizon regime, $k \gg aH$, the tensor-mode equation becomes
\begin{equation}
    \ddot{h}_k + \left( 3H + \Gamma \right)\dot{h}_k 
    + \frac{k^2}{a^2} h_k = 0.
\end{equation}
For modes well inside the horizon, the solution can be written as
\begin{equation}
    h_k(a) = A_k(a)\, \cos\!\left( \int \frac{k}{a} dt + \varphi_k \right),
\end{equation}
where $A_k(a)$ is a slowly varying amplitude and $\varphi_k$ a phase. Inserting this ansatz 
and averaging over oscillations yields an evolution equation for the envelope $A_k(a)$,
\begin{equation}
    \frac{d\ln A_k}{d\ln a}
    = -1 - \frac{1}{2}\frac{\Gamma}{H}.
\end{equation}
Integrating from horizon re-entry at $a_{\mathrm{re}}$ to the end of the relaxation-dominated 
epoch at $a_{\mathrm{end}}$ gives
\begin{equation}
    A_k(a_{\mathrm{end}})
    = A_k(a_{\mathrm{re}})
      \left( \frac{a_{\mathrm{re}}}{a_{\mathrm{end}}} \right)
      \exp\!\left[
        -\frac{1}{2}
        \int_{a_{\mathrm{re}}}^{a_{\mathrm{end}}}
        \frac{\Gamma(a)}{H(a)} \frac{da}{a}
      \right].
\end{equation}
Using $\Gamma(a) = \alpha(a) H(a)$, this becomes
\begin{equation}
    A_k(a_{\mathrm{end}})
    = A_k(a_{\mathrm{re}})
      \left( \frac{a_{\mathrm{re}}}{a_{\mathrm{end}}} \right)
      \exp\!\left[
        -\frac{1}{2}
        \int_{a_{\mathrm{re}}}^{a_{\mathrm{end}}}
        \alpha(a)\, \frac{da}{a}
      \right].
\end{equation}

\subsection*{A.5 Combined Suppression Factor}

Combining the super-horizon suppression and the enhanced sub-horizon damping, the total 
suppression factor for a mode $k$ can be written as
\begin{equation}
    T_{\mathrm{rel}}(k)
    = S(a_{\mathrm{re}})\,
      \exp\!\left[
        -\frac{1}{2}
        \int_{a_{\mathrm{re}}}^{a_{\mathrm{end}}}
        \alpha(a)\, \frac{da}{a}
      \right].
\end{equation}
If $\alpha(a)$ varies slowly, the integral simplifies to
\begin{equation}
    T_{\mathrm{rel}}(k)
    \simeq S(a_{\mathrm{re}})\,
    \left( \frac{a_{\mathrm{re}}}{a_{\mathrm{end}}} \right)^{\alpha/2}.
\end{equation}

This expression is the basis for the relaxation-corrected transfer function used in the main 
text and encapsulates the impact of the universal relaxation mechanism on tensor-mode 
evolution in RDCC.

% ---------------------------------------------------------
% APPENDIX B — SECTORAL DIFFERENCES IN THE GW SPECTRUM
% ---------------------------------------------------------
\section*{Appendix B: Sectoral Differences in the GW Spectrum}
\addcontentsline{toc}{section}{Appendix B: Sectoral Differences in the GW Spectrum}

In this appendix we analyze how the temperature asymmetry between the two CPT-conjugate 
sectors of the Relaxation-Driven Cyclic Cosmology (RDCC) affects the gravitational-wave (GW) 
spectrum. As shown in Companion~X, the temperature ratio
\begin{equation}
    \frac{T_B}{T_A} \simeq 0.5
\end{equation}
implies that sector~B is colder throughout the early Universe. This temperature difference 
modifies the relaxation coefficient, the horizon re-entry scale, and the damping history of 
tensor modes in each sector, leading to a sectoral asymmetry in the GW spectrum.

\subsection*{B.1 Temperature Dependence of the Relaxation Coefficient}

The relaxation rate satisfies
\begin{equation}
    \Gamma(a) = \alpha(a)\, H(a),
\end{equation}
with $\alpha(a)$ a slowly varying dimensionless coefficient. Because the relaxation mechanism 
is thermodynamic in origin, the colder sector exhibits a smaller relaxation coefficient:
\begin{equation}
    \alpha_B(a) < \alpha_A(a).
\end{equation}
This implies weaker relaxation-induced damping in sector~B.

\subsection*{B.2 Sectoral Horizon Re-Entry Scales}

The characteristic scale at which relaxation ceases to dominate is defined by
\begin{equation}
    k_{\mathrm{rel}}^{(i)} = a_{\mathrm{rel}}^{(i)} H\!\left(a_{\mathrm{rel}}^{(i)}\right),
\end{equation}
where $i \in \{A, B\}$ labels the sector. Because the colder sector expands more slowly at 
fixed scale factor, we have
\begin{equation}
    k_{\mathrm{rel}}^{(B)} < k_{\mathrm{rel}}^{(A)}.
\end{equation}
Thus, the relaxation-induced suppression feature begins at a lower frequency in sector~B.

\subsection*{B.3 Sectoral Transfer Functions}

The relaxation-corrected transfer function in each sector is
\begin{equation}
    T_{\mathrm{rel}}^{(i)}(k)
    = S^{(i)}(a_{\mathrm{re}}^{(i)})\,
      \exp\!\left[
        -\frac{1}{2}
        \int_{a_{\mathrm{re}}^{(i)}}^{a_{\mathrm{end}}^{(i)}}
        \alpha_i(a)\, \frac{da}{a}
      \right].
\end{equation}

Because $\alpha_B < \alpha_A$, we obtain
\begin{equation}
    T_{\mathrm{rel}}^{(B)}(k) > T_{\mathrm{rel}}^{(A)}(k)
\end{equation}
for all modes that re-enter during the relaxation-dominated epoch.

\subsection*{B.4 Sectoral GW Spectra}

The GW energy density in each sector is
\begin{equation}
    \Omega_{\mathrm{GW}}^{(i)}(k)
    = \Omega_{\mathrm{GW}}^{\mathrm{std}}(k)\,
      \left[ T_{\mathrm{rel}}^{(i)}(k) \right]^2.
\end{equation}

For $k \gg k_{\mathrm{rel}}^{(i)}$, the spectrum exhibits a power-law suppression,
\begin{equation}
    \Omega_{\mathrm{GW}}^{(i)}(k)
    \propto k^{n_t - \alpha_i}.
\end{equation}

Since $\alpha_B < \alpha_A$, the colder sector has a shallower suppression slope:
\begin{equation}
    \Omega_{\mathrm{GW}}^{(B)}(k) > \Omega_{\mathrm{GW}}^{(A)}(k)
\end{equation}
at high frequencies.

\subsection*{B.5 Physical Interpretation}

The sectoral GW asymmetry arises from three effects:

\begin{itemize}
    \item \textbf{Different relaxation strengths:}  
    The colder sector experiences weaker damping.

    \item \textbf{Different re-entry scales:}  
    Modes re-enter earlier in the hotter sector, enhancing suppression.

    \item \textbf{Different super-horizon histories:}  
    The suppression factor $S(a)$ differs between sectors.
\end{itemize}

These effects combine to produce a robust sectoral asymmetry in the GW spectrum, analogous to 
the sectoral PBH asymmetry derived in Companion~XIV.

\subsection*{B.6 Observational Implications}

While only the visible-sector spectrum is directly observable, the mirror-sector spectrum 
affects the global energy budget and provides an internal consistency check of the RDCC 
framework. A joint analysis of the GW spectrum, PBH constraints, and sectoral thermodynamics 
can therefore test the coherence of the relaxation mechanism across multiple cosmological 
observables.

\bibliographystyle{apsrev4-2}
\nocite{*}
\bibliography{references}

\end{document}
