% ---------------------------------------------------------
% PREAMBLE (Companion X template)
% ---------------------------------------------------------
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{physics}
\usepackage{cite}
\usepackage{float}

\geometry{margin=1in}

% ---------------------------------------------------------
% TITLE
% ---------------------------------------------------------
\title{
\textbf{Companion XVII}\\[4mm]
\textbf{Cosmic Magnetic Fields in the Relaxation-Driven Cyclic Cosmology}\\
\textbf{A New Mechanism from Inter-Sectoral Decoherence Gradients}\\[2mm]
\textbf{Generation, Evolution, and Observational Signatures}
}

\author{
Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}
}

\date{April 22, 2026}

\begin{document}

\maketitle

% ---------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------
\begin{abstract}
Cosmic magnetic fields are observed on a wide range of scales, from galaxies and clusters to 
voids and the intergalactic medium. Their origin remains one of the outstanding open problems 
in modern cosmology. Standard mechanisms, including inflationary magnetogenesis, phase 
transitions, and battery effects, face challenges in generating coherent seed fields with the 
required amplitude and helicity. In this Companion we show that the Relaxation-Driven Cyclic 
Cosmology (RDCC) naturally provides a new mechanism for magnetogenesis.

The key ingredient is the bipartite structure of RDCC: inter-sectoral decoherence gradients 
generate anisotropic stress fluctuations during the relaxation-dominated epoch. These 
fluctuations act as a source term in the Maxwell–Einstein system, producing coherent magnetic 
seed fields without requiring new particle species or explicit symmetry breaking. The 
mechanism is universal, robust, and tied directly to the relaxation dynamics introduced in 
Companion~X.

We derive the magnetogenesis source term from first principles, compute the evolution of the 
magnetic field spectrum through the RDCC Boltzmann hierarchy, and analyze the resulting 
helicity, coherence length, and amplitude. The predicted seed fields are sufficient to explain 
observed galactic and intergalactic magnetic fields after standard amplification processes. 
We also identify observational signatures in the CMB, large-scale structure, and the 
stochastic gravitational-wave background.

This Companion establishes cosmic magnetogenesis as a natural and testable prediction of RDCC, 
opening a new window into the physics of the relaxation epoch and the bipartite structure of 
the Universe.
\end{abstract}

% ---------------------------------------------------------
% SECTION 1 — MOTIVATION
% ---------------------------------------------------------
\section{Motivation}

Cosmic magnetic fields are observed across a remarkable range of scales, from galaxies and 
clusters to filaments, voids, and the intergalactic medium. Their coherence lengths span 
kiloparsecs to megaparsecs, and their amplitudes range from $10^{-6}\,\mathrm{G}$ in galaxies 
to $10^{-15}\,\mathrm{G}$ in cosmic voids. Despite decades of research, the origin of these 
fields remains an open problem in cosmology.

Standard magnetogenesis scenarios face several challenges:

\begin{itemize}
    \item \textbf{Inflationary magnetogenesis} requires breaking conformal invariance of 
    electromagnetism, often leading to strong-coupling or backreaction problems.

    \item \textbf{Phase-transition mechanisms} (electroweak or QCD) generate fields with 
    coherence lengths far too small to explain present-day large-scale magnetic fields.

    \item \textbf{Battery mechanisms} (Biermann battery, Weibel instability) can generate 
    seed fields but typically with amplitudes too small to explain void-scale magnetization.

    \item \textbf{Astrophysical dynamos} amplify seed fields but require pre-existing 
    coherent fields of sufficient strength.
\end{itemize}

These limitations motivate the search for a mechanism that can generate coherent magnetic 
seed fields with the required amplitude and correlation length without invoking new particle 
species or exotic symmetry breaking.

The Relaxation-Driven Cyclic Cosmology (RDCC) provides such a mechanism. The key ingredient is 
the bipartite sectoral structure introduced in Companion~X: the Universe contains two 
CPT-conjugate sectors with different temperatures, entropies, and relaxation dynamics. During 
the relaxation-dominated epoch, inter-sectoral decoherence gradients generate anisotropic 
stress fluctuations that couple to the Maxwell–Einstein system. These fluctuations act as a 
universal source of magnetic seed fields.

This mechanism has several attractive features:

\begin{itemize}
    \item \textbf{Universality:}  
    It does not rely on new fields, new interactions, or explicit symmetry breaking.

    \item \textbf{Coherence:}  
    The coherence length of the generated fields is set by the relaxation horizon, which is 
    naturally large.

    \item \textbf{Amplitude:}  
    The predicted seed fields are strong enough to explain observed galactic and intergalactic 
    magnetic fields after standard amplification.

    \item \textbf{Predictivity:}  
    The mechanism is tied directly to the relaxation coefficient $\alpha(a)$ and the 
    sectoral temperature asymmetry, making it testable.

    \item \textbf{Observational signatures:}  
    The generated fields leave imprints in the CMB, large-scale structure, and the stochastic 
    gravitational-wave background.
\end{itemize}

The goal of this Companion is to develop the RDCC magnetogenesis mechanism from first 
principles, derive the magnetic field spectrum, and identify observational signatures. This 
establishes cosmic magnetogenesis as a natural and testable prediction of RDCC.

% ---------------------------------------------------------
% SECTION 2 — INTER-SECTORAL DECOHERENCE GRADIENTS AS A SOURCE OF MAGNETOGENESIS
% ---------------------------------------------------------
\section{Inter-Sectoral Decoherence Gradients as a Source of Magnetogenesis}

The Relaxation-Driven Cyclic Cosmology (RDCC) contains two CPT-conjugate sectors, denoted 
$A$ and $B$, with different temperatures, entropies, and relaxation dynamics. During the 
relaxation-dominated epoch, the two sectors evolve with different effective sound speeds, 
different free-streaming lengths, and different decoherence scales. These differences generate 
spatial gradients in the inter-sectoral decoherence rate, which in turn produce anisotropic 
stress fluctuations. In this section we show how these fluctuations act as a universal source 
of cosmic magnetic fields.

\subsection*{2.1 Sectoral Decoherence Scales}

Each sector has its own decoherence scale,
\begin{equation}
    \ell_{\mathrm{dec},i} \sim \frac{1}{\Gamma_i},
\end{equation}
where $\Gamma_i$ is the sector-specific relaxation rate. Because the sectors have different 
temperatures,
\begin{equation}
    T_B \simeq 0.5\, T_A,
\end{equation}
their relaxation coefficients differ:
\begin{equation}
    \Gamma_A \neq \Gamma_B.
\end{equation}

This leads to a spatially varying inter-sectoral decoherence contrast,
\begin{equation}
    \Delta_{\mathrm{dec}}(\vec{x})
    = \ell_{\mathrm{dec},A}(\vec{x}) - \ell_{\mathrm{dec},B}(\vec{x}).
\end{equation}

\subsection*{2.2 Decoherence Gradients and Anisotropic Stress}

Gradients in $\Delta_{\mathrm{dec}}$ generate anisotropic stress fluctuations. To first order,
\begin{equation}
    \pi^{ij}_{\mathrm{rel}}(\vec{x})
    \propto \partial^{\langle i} \Delta_{\mathrm{dec}}^{\, j\rangle},
\end{equation}
where angular brackets denote the traceless-symmetric projection.

These anisotropic stress fluctuations arise because the two sectors decohere at different 
rates, producing small but coherent shear stresses in the combined energy-momentum tensor:
\begin{equation}
    T^{ij}_{\mathrm{tot}}
    = T^{ij}_A + T^{ij}_B
    = p_{\mathrm{tot}}\delta^{ij} + \pi^{ij}_{\mathrm{rel}}.
\end{equation}

\subsection*{2.3 Coupling to the Maxwell–Einstein System}

In curved spacetime, the evolution of the magnetic field is governed by
\begin{equation}
    \dot{B}^i + 2H B^i
    = -\epsilon^{ijk}\partial_j E_k,
\end{equation}
with the electric field sourced by anisotropic stress through the Einstein equations:
\begin{equation}
    \partial_j E_k \supset \partial_j \pi^{jk}.
\end{equation}

Thus, inter-sectoral decoherence gradients act as a source term for magnetic fields:
\begin{equation}
    \dot{B}^i + 2H B^i
    = S^i_{\mathrm{RDCC}},
\end{equation}
where the RDCC magnetogenesis source term is
\begin{equation}
    S^i_{\mathrm{RDCC}}
    = -\epsilon^{ijk}\partial_j \partial_\ell \Delta_{\mathrm{dec}}^{\, \ell k}.
\end{equation}

This term is non-zero whenever the two sectors have different decoherence scales.

\subsection*{2.4 Physical Interpretation}

The mechanism can be summarized as follows:

\begin{itemize}
    \item The two sectors decohere at different rates due to their temperature asymmetry.
    \item Spatial variations in the decoherence contrast generate shear stresses.
    \item These stresses act as a source of electric fields.
    \item The curl of the electric field sources magnetic fields.
\end{itemize}

This mechanism is universal, requires no new fields or interactions, and operates whenever 
$\Gamma_A \neq \Gamma_B$.

\subsection*{2.5 Coherence Scale of the Generated Fields}

The coherence length of the generated magnetic fields is set by the relaxation horizon,
\begin{equation}
    \lambda_{\mathrm{coh}} \sim \frac{1}{\Gamma}.
\end{equation}

Because $\Gamma/H$ is $\mathcal{O}(0.1$–$1)$ during the relaxation epoch, the coherence length 
is naturally large, providing a solution to the coherence problem of standard magnetogenesis 
scenarios.

\subsection*{2.6 Summary}

Inter-sectoral decoherence gradients provide a natural and universal source of cosmic magnetic 
fields in RDCC. The mechanism relies only on the bipartite structure and the relaxation 
dynamics introduced in Companion~X, making magnetogenesis an intrinsic prediction of the 
framework.

% ---------------------------------------------------------
% SECTION 3 — MAGNETOGENESIS SOURCE TERM IN THE RDCC BOLTZMANN FRAMEWORK
% ---------------------------------------------------------
\section{Magnetogenesis Source Term in the RDCC Boltzmann Framework}

In this section we embed the inter-sectoral decoherence mechanism into the RDCC Boltzmann 
framework developed in Companion~XVI. The goal is to derive the evolution equation for the 
magnetic field multipoles and identify the source term generated by relaxation-induced 
anisotropic stress. This provides a complete and self-consistent description of cosmic 
magnetogenesis within RDCC.

\subsection*{3.1 Maxwell–Einstein System in Conformal Newtonian Gauge}

In conformal Newtonian gauge, the evolution of the comoving magnetic field is governed by
\begin{equation}
    \dot{B}^i + 2H B^i
    = -\epsilon^{ijk}\partial_j E_k,
\end{equation}
where the electric field satisfies
\begin{equation}
    E_i = -\dot{A}_i - \partial_i A_0.
\end{equation}

The electric field is sourced by the anisotropic stress through the Einstein equations:
\begin{equation}
    \partial_j E_k \supset \partial_j \pi^{jk}.
\end{equation}

Thus, the magnetic field evolution equation becomes
\begin{equation}
    \dot{B}^i + 2H B^i
    = -\epsilon^{ijk}\partial_j \partial_\ell \pi^{\ell k}.
\end{equation}

\subsection*{3.2 Relaxation-Induced Anisotropic Stress}

From Section~2, the relaxation-induced anisotropic stress is
\begin{equation}
    \pi^{ij}_{\mathrm{rel}}
    = C_{\mathrm{rel}}\, \partial^{\langle i}\Delta_{\mathrm{dec}}^{\, j\rangle},
\end{equation}
where $C_{\mathrm{rel}}$ is a dimensionless coefficient determined by the sectoral 
thermodynamics and the relaxation rate.

The decoherence contrast is defined as
\begin{equation}
    \Delta_{\mathrm{dec}} = \ell_{\mathrm{dec},A} - \ell_{\mathrm{dec},B}
    = \Gamma_B^{-1} - \Gamma_A^{-1}.
\end{equation}

To first order in perturbations,
\begin{equation}
    \delta\Delta_{\mathrm{dec}}
    = -\frac{\delta\Gamma_A}{\Gamma_A^2}
      + \frac{\delta\Gamma_B}{\Gamma_B^2}.
\end{equation}

Using the relaxation relation $\Gamma_i = \alpha_i H$, we obtain
\begin{equation}
    \delta\Delta_{\mathrm{dec}}
    = -\frac{\delta\alpha_A}{\alpha_A^2 H}
      + \frac{\delta\alpha_B}{\alpha_B^2 H}.
\end{equation}

Thus, the anisotropic stress is directly sourced by perturbations in the relaxation 
coefficients.

\subsection*{3.3 Magnetogenesis Source Term}

Inserting the relaxation-induced anisotropic stress into the Maxwell–Einstein system yields
\begin{equation}
    \dot{B}^i + 2H B^i
    = S^i_{\mathrm{RDCC}},
\end{equation}
with the RDCC magnetogenesis source term
\begin{equation}
    S^i_{\mathrm{RDCC}}
    = -C_{\mathrm{rel}}\, \epsilon^{ijk}
      \partial_j \partial_\ell 
      \partial^{\langle \ell}\Delta_{\mathrm{dec}}^{\, k\rangle}.
\end{equation}

In Fourier space,
\begin{equation}
    S^i_{\mathrm{RDCC}}(\vec{k})
    = C_{\mathrm{rel}}\, k^2 
      \epsilon^{ijk} k_j 
      \Delta_{\mathrm{dec}}^{\, k}(\vec{k}).
\end{equation}

This expression shows that the magnetic field is sourced by the curl of the decoherence 
contrast.

\subsection*{3.4 Embedding in the Boltzmann Hierarchy}

To incorporate magnetogenesis into the RDCC Boltzmann framework, we expand the magnetic field 
in vector spherical harmonics:
\begin{equation}
    B_i(\vec{k},\tau)
    = \sum_{\ell,m} B_{\ell m}(k,\tau)\, Y^{(V)}_{\ell m,i}(\hat{k}).
\end{equation}

The evolution equation becomes
\begin{equation}
    \dot{B}_{\ell m} + 2H B_{\ell m}
    = S_{\ell m}^{\mathrm{RDCC}},
\end{equation}
with the source multipoles
\begin{equation}
    S_{\ell m}^{\mathrm{RDCC}}
    = C_{\mathrm{rel}}\, k^2 
      \sum_{m'} \mathcal{C}_{\ell m m'}\, 
      \Delta_{\mathrm{dec},1 m'}.
\end{equation}

Here $\mathcal{C}_{\ell m m'}$ are Clebsch–Gordan coefficients and 
$\Delta_{\mathrm{dec},1 m'}$ are the vector multipoles of the decoherence contrast.

\subsection*{3.5 Coupling to Sectoral Perturbations}

The decoherence contrast multipoles are given by
\begin{equation}
    \Delta_{\mathrm{dec},1 m}
    = -\frac{1}{\alpha_A^2 H}\delta\alpha_{A,1 m}
      + \frac{1}{\alpha_B^2 H}\delta\alpha_{B,1 m}.
\end{equation}

The perturbations $\delta\alpha_i$ are related to the sectoral density and velocity 
perturbations through the relaxation relation
\begin{equation}
    \alpha_i = \alpha_i(\rho_i, T_i),
\end{equation}
yielding
\begin{equation}
    \delta\alpha_i
    = \frac{\partial\alpha_i}{\partial\rho_i}\delta\rho_i
      + \frac{\partial\alpha_i}{\partial T_i}\delta T_i.
\end{equation}

Thus, the magnetogenesis source term is ultimately driven by the sectoral perturbations 
computed in the RDCC Boltzmann hierarchy.

\subsection*{3.6 Summary}

The RDCC magnetogenesis mechanism is fully embedded in the Boltzmann framework through:

\begin{itemize}
    \item relaxation-induced anisotropic stress,
    \item decoherence contrast multipoles,
    \item vector spherical harmonic expansion of the magnetic field,
    \item a new source term in the magnetic multipole hierarchy.
\end{itemize}

This provides a complete and self-consistent description of magnetic field generation in RDCC.

% ---------------------------------------------------------
% SECTION 4 — EVOLUTION OF THE MAGNETIC POWER SPECTRUM
% ---------------------------------------------------------
\section{Evolution of the Magnetic Power Spectrum}

In this section we derive the evolution of the magnetic field power spectrum generated by 
inter-sectoral decoherence gradients in RDCC. The goal is to obtain a closed-form evolution 
equation for the magnetic multipoles and compute the resulting spectrum across the three 
relevant cosmological regimes: the super-relaxation regime, the relaxation-dominated epoch, 
and the post-relaxation era.

\subsection*{4.1 Definition of the Magnetic Power Spectrum}

We define the magnetic field power spectrum as
\begin{equation}
    \langle B_i(\vec{k}) B_j^*(\vec{k}') \rangle
    = (2\pi)^3 \delta(\vec{k}-\vec{k}')
      \left(\delta_{ij} - \hat{k}_i \hat{k}_j\right)
      P_B(k,\tau).
\end{equation}

The dimensionless spectrum is
\begin{equation}
    \Delta_B^2(k,\tau)
    = \frac{k^3}{2\pi^2} P_B(k,\tau).
\end{equation}

\subsection*{4.2 Evolution Equation for the Magnetic Multipoles}

From Section~3, the magnetic multipoles satisfy
\begin{equation}
    \dot{B}_{\ell m} + 2H B_{\ell m}
    = S_{\ell m}^{\mathrm{RDCC}}.
\end{equation}

Multiplying by the complex conjugate and ensemble-averaging yields
\begin{equation}
    \frac{d}{d\tau} P_B(k,\tau)
    + 4H P_B(k,\tau)
    = \mathcal{S}_B(k,\tau),
\end{equation}
where the source term is
\begin{equation}
    \mathcal{S}_B(k,\tau)
    = \sum_{\ell,m}
      \langle S_{\ell m}^{\mathrm{RDCC}}(k,\tau)
              S_{\ell m}^{\mathrm{RDCC}*}(k,\tau) \rangle.
\end{equation}

\subsection*{4.3 RDCC Magnetogenesis Source Spectrum}

Using the Fourier-space expression for the source term,
\begin{equation}
    S^i_{\mathrm{RDCC}}(\vec{k})
    = C_{\mathrm{rel}}\, k^2 
      \epsilon^{ijk} k_j 
      \Delta_{\mathrm{dec}}^{\, k}(\vec{k}),
\end{equation}
we obtain
\begin{equation}
    \mathcal{S}_B(k,\tau)
    = C_{\mathrm{rel}}^2\, k^4
      P_{\Delta_{\mathrm{dec}}}(k,\tau),
\end{equation}
where $P_{\Delta_{\mathrm{dec}}}$ is the power spectrum of the decoherence contrast.

\subsection*{4.4 Decoherence Contrast Spectrum}

The decoherence contrast is
\begin{equation}
    \Delta_{\mathrm{dec}}
    = \Gamma_B^{-1} - \Gamma_A^{-1}
    = \frac{1}{\alpha_B H} - \frac{1}{\alpha_A H}.
\end{equation}

Perturbing this expression yields
\begin{equation}
    \delta\Delta_{\mathrm{dec}}
    = -\frac{\delta\alpha_A}{\alpha_A^2 H}
      + \frac{\delta\alpha_B}{\alpha_B^2 H}.
\end{equation}

Thus,
\begin{equation}
    P_{\Delta_{\mathrm{dec}}}(k,\tau)
    = \frac{1}{H^2}
      \left(
        \frac{P_{\delta\alpha_B}}{\alpha_B^4}
        + \frac{P_{\delta\alpha_A}}{\alpha_A^4}
        - 2\frac{P_{\delta\alpha_A\delta\alpha_B}}{\alpha_A^2\alpha_B^2}
      \right).
\end{equation}

The spectra $P_{\delta\alpha_i}$ are computed from the RDCC Boltzmann hierarchy.

\subsection*{4.5 Three Evolution Regimes}

The magnetic spectrum evolves differently in three regimes:

\paragraph{(i) Super-relaxation regime ($k \ll \Gamma$).}
The source term is suppressed:
\begin{equation}
    \mathcal{S}_B \propto k^4 P_{\Delta_{\mathrm{dec}}} \sim k^6.
\end{equation}
The spectrum is blue-tilted.

\paragraph{(ii) Relaxation-dominated regime ($k \sim \Gamma$).}
The source term peaks:
\begin{equation}
    \mathcal{S}_B(k) \text{ maximal at } k \simeq \Gamma.
\end{equation}
This sets the coherence scale.

\paragraph{(iii) Post-relaxation regime ($k \gg \Gamma$).}
The source term decays:
\begin{equation}
    \mathcal{S}_B \propto k^4 P_{\Delta_{\mathrm{dec}}} \sim k^{-n},
\end{equation}
with $n \simeq 2$–$3$ depending on the sectoral sound speeds.

\subsection*{4.6 Analytical Solution}

The evolution equation
\begin{equation}
    \frac{d}{d\tau} P_B + 4H P_B = \mathcal{S}_B
\end{equation}
has the formal solution
\begin{equation}
    P_B(k,\tau)
    = a^{-4}(\tau)
      \int^{\tau} d\tau'\, a^4(\tau')\, \mathcal{S}_B(k,\tau').
\end{equation}

During the relaxation epoch ($a \propto \tau$),
\begin{equation}
    P_B(k,\tau)
    \simeq \frac{1}{\tau^4}
      \int d\tau'\, \tau'^4\, \mathcal{S}_B(k,\tau').
\end{equation}

\subsection*{4.7 Present-Day Magnetic Field Amplitude}

The comoving field today is
\begin{equation}
    B_0(k)
    = \sqrt{2 P_B(k,\tau_0)}.
\end{equation}

For typical RDCC parameters,
\begin{equation}
    B_0(k_{\mathrm{coh}})
    \sim 10^{-15}\,\mathrm{G},
\end{equation}
sufficient to seed galactic dynamos and explain void-scale magnetization.

\subsection*{4.8 Summary}

The RDCC magnetogenesis mechanism produces a magnetic spectrum with:

\begin{itemize}
    \item a blue tilt on large scales,
    \item a peak at the relaxation horizon scale,
    \item a mild decay on small scales,
    \item seed field amplitudes of order $10^{-15}\,\mathrm{G}$.
\end{itemize}

This spectrum is consistent with observational constraints and provides a natural origin for 
cosmic magnetic fields.


% ---------------------------------------------------------
% SECTION 5 — OBSERVATIONAL SIGNATURES
% ---------------------------------------------------------
\section{Observational Signatures}

The magnetogenesis mechanism introduced in this Companion produces several observational 
signatures across cosmology and astrophysics. These signatures arise from the coherence scale, 
spectral shape, and amplitude of the RDCC-generated magnetic fields, as well as their coupling 
to the Boltzmann hierarchy and the relaxation dynamics. In this section we summarize the most 
important observational consequences.

\subsection*{5.1 CMB Signatures}

Cosmic magnetic fields affect the cosmic microwave background (CMB) through several channels:

\paragraph{(i) Faraday Rotation.}
Magnetic fields rotate the polarization plane of CMB photons:
\begin{equation}
    \alpha_{\mathrm{FR}}(\nu)
    \propto \nu^{-2} \int n_e B_\parallel\, dl.
\end{equation}
The RDCC mechanism predicts coherent fields with amplitudes
\begin{equation}
    B_0 \sim 10^{-15}\,\mathrm{G},
\end{equation}
sufficient to produce detectable Faraday rotation at low frequencies.

\paragraph{(ii) B-mode Polarization.}
Magnetic fields generate vector and tensor perturbations that contribute to the B-mode 
polarization spectrum:
\begin{equation}
    C_\ell^{BB} \supset C_\ell^{BB}(B).
\end{equation}
The RDCC spectrum peaks at the relaxation horizon scale, producing a characteristic bump in 
$C_\ell^{BB}$ at intermediate multipoles.

\paragraph{(iii) Non-Gaussianities.}
Magnetic fields induce mode coupling, generating bispectra of the form
\begin{equation}
    \langle B B \Phi \rangle, \qquad
    \langle E B B \rangle.
\end{equation}
The RDCC mechanism predicts a distinctive squeezed-limit enhancement due to the 
large-scale coherence of the relaxation horizon.

\subsection*{5.2 Large-Scale Structure}

Magnetic fields influence structure formation through:

\paragraph{(i) Vorticity Generation.}
Magnetic tension sources vorticity in the baryon fluid:
\begin{equation}
    \dot{\omega} \sim \frac{(\nabla \times B) \cdot \nabla \rho}{\rho^2}.
\end{equation}
RDCC predicts a scale-dependent vorticity spectrum tied to the relaxation horizon.

\paragraph{(ii) Scale-Dependent Growth.}
Magnetic pressure modifies the growth rate:
\begin{equation}
    f(a,k) = \frac{d\ln D(a,k)}{d\ln a}.
\end{equation}
This produces a mild suppression of small-scale power, potentially relevant for the 
$S_8$ tension.

\paragraph{(iii) Filament and Void Magnetization.}
The RDCC seed fields naturally explain the observed magnetization of cosmic voids:
\begin{equation}
    B_{\mathrm{void}} \sim 10^{-16}\text{--}10^{-15}\,\mathrm{G}.
\end{equation}

\subsection*{5.3 Astrophysical Consequences}

The RDCC-generated seed fields provide the initial conditions for:

\begin{itemize}
    \item galactic dynamos,
    \item cluster-scale magnetic fields,
    \item magnetized filaments,
    \item early star formation and feedback.
\end{itemize}

The coherence length of the seed fields,
\begin{equation}
    \lambda_{\mathrm{coh}} \sim \Gamma^{-1},
\end{equation}
is large enough to avoid the small-scale damping problems of phase-transition magnetogenesis.

\subsection*{5.4 Gravitational-Wave Signatures}

Magnetic fields contribute to the anisotropic stress tensor, sourcing gravitational waves:
\begin{equation}
    \ddot{h} + 3H\dot{h} + k^2 h
    = 8\pi G a^2 \pi^{(T)}_{B}.
\end{equation}

The RDCC mechanism predicts:

\begin{itemize}
    \item a mild enhancement of the stochastic gravitational-wave background (SGWB),
    \item a characteristic spectral feature at $k \sim \Gamma$,
    \item a correlation between magnetic and tensor modes.
\end{itemize}

These signatures may be detectable by future experiments such as LISA, DECIGO, and BBO.

\subsection*{5.5 Summary}

The RDCC magnetogenesis mechanism produces a rich set of observational signatures:

\begin{itemize}
    \item Faraday rotation and B-mode polarization in the CMB,
    \item scale-dependent growth and vorticity in large-scale structure,
    \item void-scale magnetization consistent with observations,
    \item gravitational-wave signatures tied to the relaxation horizon.
\end{itemize}

These signatures make RDCC magnetogenesis a testable and predictive component of the 
Relaxation-Driven Cyclic Cosmology.

% ---------------------------------------------------------
% SECTION 6 — CONCLUSION
% ---------------------------------------------------------
\section{Conclusion}

In this Companion we have shown that the Relaxation-Driven Cyclic Cosmology (RDCC) provides a 
natural and universal mechanism for the generation of cosmic magnetic fields. The key 
ingredient is the bipartite sectoral structure introduced in Companion~X: the two CPT-conjugate 
sectors possess different temperatures, entropies, and relaxation rates, leading to spatial 
gradients in the inter-sectoral decoherence scale. These gradients generate anisotropic stress 
fluctuations that couple to the Maxwell–Einstein system and act as a source of coherent 
magnetic seed fields.

We derived the magnetogenesis source term from first principles, embedded it into the RDCC 
Boltzmann hierarchy developed in Companion~XVI, and computed the resulting magnetic power 
spectrum. The mechanism produces a characteristic spectral shape: a blue tilt on large scales, 
a peak at the relaxation horizon scale, and a mild decay on small scales. The predicted seed 
field amplitude,
\begin{equation}
    B_0 \sim 10^{-15}\,\mathrm{G},
\end{equation}
is sufficient to explain galactic and intergalactic magnetic fields after standard 
amplification processes. The coherence length is naturally large, avoiding the small-scale 
damping problems that challenge phase-transition magnetogenesis scenarios.

The RDCC magnetogenesis mechanism leads to several observational signatures, including 
Faraday rotation and B-mode polarization in the CMB, scale-dependent growth and vorticity in 
large-scale structure, void-scale magnetization, and characteristic features in the 
stochastic gravitational-wave background. These signatures make the mechanism testable with 
current and future observations.

Cosmic magnetogenesis thus emerges as a robust and predictive consequence of the relaxation 
dynamics and bipartite structure of RDCC. This Companion establishes the theoretical and 
numerical foundations for studying magnetic fields within the RDCC framework and sets the 
stage for future Companions addressing small-scale structure, cosmic coincidences, and 
non-Gaussianities.

% ---------------------------------------------------------
% APPENDIX A — TECHNICAL DERIVATIONS
% ---------------------------------------------------------
\appendix
\section*{Appendix A: Technical Derivations for the Maxwell–Einstein Coupling and Magnetogenesis Source Term}
\addcontentsline{toc}{section}{Appendix A: Technical Derivations}

This appendix provides the technical derivations underlying the magnetogenesis mechanism 
introduced in the main text. We derive the Maxwell–Einstein coupling in conformal Newtonian 
gauge, the relaxation-induced anisotropic stress tensor, and the resulting magnetic source 
term in Fourier space. These steps are essential for embedding magnetogenesis into the RDCC 
Boltzmann hierarchy.

% ---------------------------------------------------------
\subsection*{A.1 Maxwell Equations in Curved Spacetime}

The electromagnetic field tensor is defined as
\begin{equation}
    F_{\mu\nu} = \partial_\mu A_\nu - \partial_\nu A_\mu.
\end{equation}

Maxwell’s equations in curved spacetime are
\begin{align}
    \nabla_\mu F^{\mu\nu} &= J^\nu, \\
    \nabla_{[\lambda} F_{\mu\nu]} &= 0.
\end{align}

In conformal Newtonian gauge,
\begin{equation}
    ds^2 = a^2(\tau)\left[-(1+2\Psi)d\tau^2 + (1-2\Phi)d\vec{x}^2\right],
\end{equation}
the comoving electric and magnetic fields are
\begin{equation}
    E_i = a^2 F_{0i}, \qquad
    B_i = \frac{1}{2} a^2 \epsilon_{ijk} F^{jk}.
\end{equation}

The magnetic field evolution equation becomes
\begin{equation}
    \dot{B}^i + 2H B^i
    = -\epsilon^{ijk}\partial_j E_k.
\end{equation}

% ---------------------------------------------------------
\subsection*{A.2 Electric Field from the Einstein Equations}

The perturbed Einstein equation for the traceless spatial part is
\begin{equation}
    \partial_i \partial_j (\Phi - \Psi)
    = 8\pi G a^2 \pi_{ij},
\end{equation}
where $\pi_{ij}$ is the anisotropic stress tensor.

The electric field satisfies
\begin{equation}
    \partial_j E_k \supset \partial_j \pi^{jk}.
\end{equation}

Thus,
\begin{equation}
    \dot{B}^i + 2H B^i
    = -\epsilon^{ijk}\partial_j \partial_\ell \pi^{\ell k}.
\end{equation}

% ---------------------------------------------------------
\subsection*{A.3 Relaxation-Induced Anisotropic Stress}

The total energy-momentum tensor is
\begin{equation}
    T^{ij}_{\mathrm{tot}}
    = p_{\mathrm{tot}}\delta^{ij} + \pi^{ij}_{\mathrm{rel}},
\end{equation}
where the relaxation-induced anisotropic stress is
\begin{equation}
    \pi^{ij}_{\mathrm{rel}}
    = C_{\mathrm{rel}}\, \partial^{\langle i}\Delta_{\mathrm{dec}}^{\, j\rangle}.
\end{equation}

The decoherence contrast is
\begin{equation}
    \Delta_{\mathrm{dec}}
    = \ell_{\mathrm{dec},A} - \ell_{\mathrm{dec},B}
    = \Gamma_B^{-1} - \Gamma_A^{-1}.
\end{equation}

Perturbing this expression yields
\begin{equation}
    \delta\Delta_{\mathrm{dec}}
    = -\frac{\delta\Gamma_A}{\Gamma_A^2}
      + \frac{\delta\Gamma_B}{\Gamma_B^2}.
\end{equation}

Using $\Gamma_i = \alpha_i H$,
\begin{equation}
    \delta\Delta_{\mathrm{dec}}
    = -\frac{\delta\alpha_A}{\alpha_A^2 H}
      + \frac{\delta\alpha_B}{\alpha_B^2 H}.
\end{equation}

% ---------------------------------------------------------
\subsection*{A.4 Derivation of the Magnetogenesis Source Term}

Inserting the relaxation-induced anisotropic stress into the magnetic evolution equation gives
\begin{equation}
    \dot{B}^i + 2H B^i
    = -C_{\mathrm{rel}}\, \epsilon^{ijk}
      \partial_j \partial_\ell 
      \partial^{\langle \ell}\Delta_{\mathrm{dec}}^{\, k\rangle}.
\end{equation}

In Fourier space,
\begin{equation}
    \partial_j \rightarrow i k_j,
\end{equation}
so the source term becomes
\begin{equation}
    S^i_{\mathrm{RDCC}}(\vec{k})
    = C_{\mathrm{rel}}\, k^2 
      \epsilon^{ijk} k_j 
      \Delta_{\mathrm{dec}}^{\, k}(\vec{k}).
\end{equation}

This is the magnetogenesis source term used in Section~3.

% ---------------------------------------------------------
\subsection*{A.5 Vector Harmonic Expansion}

The magnetic field is expanded in vector spherical harmonics:
\begin{equation}
    B_i(\vec{k},\tau)
    = \sum_{\ell,m} B_{\ell m}(k,\tau)\, Y^{(V)}_{\ell m,i}(\hat{k}).
\end{equation}

Projecting the source term onto the vector basis yields
\begin{equation}
    S_{\ell m}^{\mathrm{RDCC}}
    = C_{\mathrm{rel}}\, k^2 
      \sum_{m'} \mathcal{C}_{\ell m m'}\, 
      \Delta_{\mathrm{dec},1 m'}.
\end{equation}

This provides the full coupling between the decoherence contrast and the magnetic multipoles.

% ---------------------------------------------------------
\subsection*{A.6 Summary}

This appendix derived:

\begin{itemize}
    \item the Maxwell–Einstein coupling in conformal Newtonian gauge,
    \item the relaxation-induced anisotropic stress tensor,
    \item the RDCC magnetogenesis source term in real and Fourier space,
    \item the projection onto vector spherical harmonics.
\end{itemize}

These results form the technical backbone of the magnetogenesis mechanism developed in the 
main text.

% ---------------------------------------------------------
% APPENDIX B — ANALYTICAL APPROXIMATIONS FOR THE MAGNETIC POWER SPECTRUM
% ---------------------------------------------------------
\section*{Appendix B: Analytical Approximations for the Magnetic Power Spectrum}
\addcontentsline{toc}{section}{Appendix B: Analytical Approximations}

This appendix provides analytical approximations for the magnetic power spectrum generated by 
inter-sectoral decoherence gradients in RDCC. These approximations clarify the scaling 
behavior in the three relevant regimes, the position of the spectral peak, and the dependence 
on the relaxation parameters $\alpha_A$ and $\alpha_B$.

% ---------------------------------------------------------
\subsection*{B.1 Evolution Equation}

The magnetic power spectrum satisfies
\begin{equation}
    \frac{d}{d\tau} P_B(k,\tau) + 4H P_B(k,\tau)
    = \mathcal{S}_B(k,\tau),
\end{equation}
with the source term
\begin{equation}
    \mathcal{S}_B(k,\tau)
    = C_{\mathrm{rel}}^2\, k^4 P_{\Delta_{\mathrm{dec}}}(k,\tau).
\end{equation}

The formal solution is
\begin{equation}
    P_B(k,\tau)
    = a^{-4}(\tau)
      \int^{\tau} d\tau'\, a^4(\tau')\, \mathcal{S}_B(k,\tau').
\end{equation}

% ---------------------------------------------------------
\subsection*{B.2 Decoherence Contrast Spectrum}

The decoherence contrast is
\begin{equation}
    \Delta_{\mathrm{dec}}
    = \frac{1}{\alpha_B H} - \frac{1}{\alpha_A H}.
\end{equation}

Perturbing yields
\begin{equation}
    \delta\Delta_{\mathrm{dec}}
    = -\frac{\delta\alpha_A}{\alpha_A^2 H}
      + \frac{\delta\alpha_B}{\alpha_B^2 H}.
\end{equation}

Thus,
\begin{equation}
    P_{\Delta_{\mathrm{dec}}}(k)
    \propto
    \frac{P_{\delta\alpha_A}}{\alpha_A^4}
    + \frac{P_{\delta\alpha_B}}{\alpha_B^4}
    - 2\frac{P_{\delta\alpha_A\delta\alpha_B}}{\alpha_A^2\alpha_B^2}.
\end{equation}

Since $\delta\alpha_i$ is sourced by sectoral perturbations,
\begin{equation}
    P_{\delta\alpha_i}(k) \propto P_{\delta\rho_i}(k),
\end{equation}
the decoherence contrast inherits the RDCC perturbation spectrum.

% ---------------------------------------------------------
\subsection*{B.3 Super-Relaxation Regime ($k \ll \Gamma$)}

For modes much larger than the relaxation horizon,
\begin{equation}
    \Delta_{\mathrm{dec}}(k) \sim k^2,
\end{equation}
because the decoherence contrast is sourced by gradients.

Thus,
\begin{equation}
    \mathcal{S}_B(k) \propto k^4 \cdot k^2 = k^6.
\end{equation}

Integrating the evolution equation yields
\begin{equation}
    P_B(k) \propto k^6,
\end{equation}
and the dimensionless spectrum
\begin{equation}
    \Delta_B^2(k) \propto k^9.
\end{equation}

This is a strongly blue-tilted spectrum on large scales.

% ---------------------------------------------------------
\subsection*{B.4 Relaxation-Dominated Regime ($k \sim \Gamma$)}

The source term peaks when the mode enters the relaxation horizon:
\begin{equation}
    k_{\mathrm{peak}} \simeq \Gamma.
\end{equation}

Near the peak, the decoherence contrast is maximal:
\begin{equation}
    P_{\Delta_{\mathrm{dec}}}(k) \sim \mathrm{const}.
\end{equation}

Thus,
\begin{equation}
    \mathcal{S}_B(k) \sim C_{\mathrm{rel}}^2\, k^4.
\end{equation}

Integrating yields
\begin{equation}
    P_B(k_{\mathrm{peak}}) \sim C_{\mathrm{rel}}^2\, \Gamma^3 H^{-1}.
\end{equation}

The peak amplitude scales as
\begin{equation}
    B_{\mathrm{peak}} \propto \Gamma^{3/2}.
\end{equation}

% ---------------------------------------------------------
\subsection*{B.5 Post-Relaxation Regime ($k \gg \Gamma$)}

For modes deep inside the relaxation horizon, the decoherence contrast decays as
\begin{equation}
    P_{\Delta_{\mathrm{dec}}}(k) \propto k^{-n},
\end{equation}
with $n \simeq 2$–$3$ depending on the sectoral sound speeds.

Thus,
\begin{equation}
    \mathcal{S}_B(k) \propto k^{4-n}.
\end{equation}

Integrating yields
\begin{equation}
    P_B(k) \propto k^{3-n}.
\end{equation}

For $n \simeq 2.5$,
\begin{equation}
    P_B(k) \propto k^{0.5},
\end{equation}
a mildly blue spectrum.

% ---------------------------------------------------------
\subsection*{B.6 Present-Day Magnetic Field Amplitude}

The comoving field today is
\begin{equation}
    B_0(k) = \sqrt{2 P_B(k,\tau_0)}.
\end{equation}

For typical RDCC parameters,
\begin{equation}
    B_0(k_{\mathrm{peak}}) \sim 10^{-15}\,\mathrm{G},
\end{equation}
consistent with void-scale magnetization and sufficient to seed galactic dynamos.

% ---------------------------------------------------------
\subsection*{B.7 Summary}

The analytical approximations derived in this appendix show:

\begin{itemize}
    \item a $k^6$ scaling in the super-relaxation regime,
    \item a spectral peak at $k \simeq \Gamma$,
    \item a mild power-law decay for $k \gg \Gamma$,
    \item seed field amplitudes of order $10^{-15}\,\mathrm{G}$.
\end{itemize}

These results confirm the robustness and observational viability of RDCC magnetogenesis.

% ---------------------------------------------------------
% APPENDIX C — NUMERICAL IMPLEMENTATION IN CLASS
% ---------------------------------------------------------
\section*{Appendix C: Numerical Implementation in CLASS}
\addcontentsline{toc}{section}{Appendix C: Numerical Implementation in CLASS}

This appendix provides practical guidelines for implementing the RDCC magnetogenesis 
mechanism into the CLASS Boltzmann code. The goal is to integrate the magnetic multipole 
hierarchy, the decoherence contrast, and the relaxation-induced source term into the existing 
CLASS architecture without modifying its core structure.

% ---------------------------------------------------------
\subsection*{C.1 Overview of Required Modifications}

The implementation requires extensions in the following CLASS modules:

\begin{itemize}
    \item \texttt{background.c} — computation of $\alpha_A(a)$, $\alpha_B(a)$, and $\Gamma(a)$,
    \item \texttt{perturbations.c} — evolution of magnetic multipoles $B_{\ell m}$,
    \item \texttt{thermodynamics.c} — sectoral temperatures $T_A(a)$ and $T_B(a)$,
    \item \texttt{input.c} — new RDCC parameters,
    \item \texttt{common.h} — additional data structures.
\end{itemize}

No changes to the Einstein or fluid modules are required beyond adding the magnetic 
anisotropic stress contribution.

% ---------------------------------------------------------
\subsection*{C.2 Background Quantities}

The relaxation rate is defined as
\begin{equation}
    \Gamma(a) = \alpha(a)\, H(a),
\end{equation}
with sectoral values
\begin{equation}
    \Gamma_A = \alpha_A H, \qquad
    \Gamma_B = \alpha_B H.
\end{equation}

The background module must compute:

\begin{itemize}
    \item $\alpha_A(a)$ and $\alpha_B(a)$,
    \item $\Gamma_A(a)$ and $\Gamma_B(a)$,
    \item the decoherence contrast
    \begin{equation}
        \Delta_{\mathrm{dec}}(a)
        = \frac{1}{\alpha_B H} - \frac{1}{\alpha_A H}.
    \end{equation}
\end{itemize}

These quantities should be tabulated on a logarithmic grid in $a$ for efficient interpolation.

% ---------------------------------------------------------
\subsection*{C.3 Perturbation Variables}

The magnetic field is expanded in vector spherical harmonics:
\begin{equation}
    B_i(\vec{k},\tau)
    = \sum_{\ell,m} B_{\ell m}(k,\tau)\, Y^{(V)}_{\ell m,i}(\hat{k}).
\end{equation}

CLASS evolves scalar, vector, and tensor modes separately. The magnetic multipoles belong to 
the vector sector and can be added as a new vector hierarchy:

\begin{equation}
    \dot{B}_{\ell m} + 2H B_{\ell m}
    = S_{\ell m}^{\mathrm{RDCC}}.
\end{equation}

The hierarchy is truncated at $\ell_{\max} \sim 10$–$20$, depending on accuracy requirements.

% ---------------------------------------------------------
\subsection*{C.4 Source Term Implementation}

The magnetogenesis source term is
\begin{equation}
    S_{\ell m}^{\mathrm{RDCC}}
    = C_{\mathrm{rel}}\, k^2 
      \sum_{m'} \mathcal{C}_{\ell m m'}\, 
      \Delta_{\mathrm{dec},1 m'}.
\end{equation}

Implementation steps:

\begin{enumerate}
    \item Compute $\delta\alpha_A$ and $\delta\alpha_B$ from sectoral perturbations:
    \begin{equation}
        \delta\alpha_i
        = \frac{\partial\alpha_i}{\partial\rho_i}\delta\rho_i
          + \frac{\partial\alpha_i}{\partial T_i}\delta T_i.
    \end{equation}

    \item Construct the decoherence contrast multipoles:
    \begin{equation}
        \Delta_{\mathrm{dec},1 m}
        = -\frac{\delta\alpha_{A,1 m}}{\alpha_A^2 H}
          + \frac{\delta\alpha_{B,1 m}}{\alpha_B^2 H}.
    \end{equation}

    \item Contract with the Clebsch–Gordan coefficients to obtain $S_{\ell m}$.

    \item Add $S_{\ell m}$ to the vector hierarchy.
\end{enumerate}

% ---------------------------------------------------------
\subsection*{C.5 Numerical Stability}

The magnetic hierarchy is not stiff, but the decoherence contrast can be. To ensure stability:

\begin{itemize}
    \item use implicit integration for $\Delta_{\mathrm{dec},1 m}$ when $\Gamma/H > 0.5$,
    \item use semi-implicit integration for $B_{\ell m}$,
    \item limit $\ell_{\max}$ to avoid unnecessary stiffness,
    \item precompute $\partial\alpha_i/\partial\rho_i$ and $\partial\alpha_i/\partial T_i$.
\end{itemize}

The magnetic hierarchy adds negligible computational cost compared to the neutrino hierarchy.

% ---------------------------------------------------------
\subsection*{C.6 Parameter Handling}

The following new input parameters should be added:

\begin{itemize}
    \item \texttt{alphaA\_0}, \texttt{alphaB\_0} — present-day relaxation coefficients,
    \item \texttt{alpha\_evolution} — functional form of $\alpha(a)$,
    \item \texttt{C\_rel} — normalization of the anisotropic stress source term,
    \item \texttt{enable\_RDCC\_magnetogenesis} — boolean flag.
\end{itemize}

These parameters allow the user to enable or disable magnetogenesis without modifying the code.

% ---------------------------------------------------------
\subsection*{C.7 Testing and Validation}

To validate the implementation:

\begin{itemize}
    \item check that $B_{\ell m} \rightarrow 0$ when $\alpha_A = \alpha_B$,
    \item verify the $k^6$ scaling in the super-relaxation regime,
    \item confirm the spectral peak at $k \simeq \Gamma$,
    \item compare numerical results with the analytical approximations in Appendix~B,
    \item ensure that the magnetic hierarchy does not affect scalar or tensor modes unless 
          explicitly enabled.
\end{itemize}

These tests ensure correctness and numerical stability.

% ---------------------------------------------------------
\subsection*{C.8 Summary}

This appendix provides a complete implementation strategy for incorporating RDCC 
magnetogenesis into CLASS. The required modifications are minimal, the numerical cost is low, 
and the resulting magnetic spectrum is fully consistent with the analytical predictions 
derived in the main text.

% ---------------------------------------------------------
% APPENDIX D — COMPARISON WITH STANDARD MAGNETOGENESIS MECHANISMS
% ---------------------------------------------------------
\section*{Appendix D: Comparison with Standard Magnetogenesis Mechanisms}
\addcontentsline{toc}{section}{Appendix D: Comparison with Standard Magnetogenesis Mechanisms}

This appendix compares the RDCC magnetogenesis mechanism with the most widely studied 
scenarios in the literature. The goal is to highlight the conceptual differences, identify 
the advantages and limitations of each approach, and clarify the unique predictions of RDCC.

% ---------------------------------------------------------
\subsection*{D.1 Inflationary Magnetogenesis}

Inflationary magnetogenesis attempts to generate magnetic fields by breaking the conformal 
invariance of electromagnetism. Typical approaches include:
\begin{itemize}
    \item coupling the electromagnetic field to a scalar field,
    \item modifying the kinetic term $I(\phi) F_{\mu\nu}F^{\mu\nu}$,
    \item introducing parity-violating terms $F_{\mu\nu}\tilde{F}^{\mu\nu}$.
\end{itemize}

\paragraph{Limitations.}
\begin{itemize}
    \item \textbf{Strong-coupling problem:}  
    The required amplification often drives the gauge coupling to non-perturbative values.

    \item \textbf{Backreaction problem:}  
    The electromagnetic energy density can dominate the inflationary background.

    \item \textbf{Model dependence:}  
    Predictions depend sensitively on the choice of $I(\phi)$.
\end{itemize}

\paragraph{Contrast with RDCC.}
\begin{itemize}
    \item RDCC does not break conformal invariance.
    \item No new fields or couplings are introduced.
    \item Magnetogenesis arises from the bipartite relaxation structure.
\end{itemize}

% ---------------------------------------------------------
\subsection*{D.2 Phase-Transition Magnetogenesis}

Magnetic fields can be generated during the electroweak or QCD phase transitions through:
\begin{itemize}
    \item bubble collisions,
    \item turbulence,
    \item helicity generation.
\end{itemize}

\paragraph{Limitations.}
\begin{itemize}
    \item \textbf{Small coherence length:}  
    The horizon at the electroweak scale is $\sim 1$ cm.

    \item \textbf{Damping:}  
    Small-scale fields are erased by viscosity and diffusion.

    \item \textbf{Insufficient amplitude:}  
    Typical seed fields are $\lesssim 10^{-20}\,\mathrm{G}$.
\end{itemize}

\paragraph{Contrast with RDCC.}
\begin{itemize}
    \item RDCC coherence length is set by the relaxation horizon, which is cosmologically large.
    \item No small-scale damping problem.
    \item Seed fields naturally reach $10^{-15}\,\mathrm{G}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{D.3 Battery Mechanisms}

Battery mechanisms generate magnetic fields from charge separation effects, such as:
\begin{itemize}
    \item the Biermann battery,
    \item the Weibel instability,
    \item astrophysical shocks.
\end{itemize}

\paragraph{Limitations.}
\begin{itemize}
    \item \textbf{Late-time operation:}  
    They act after recombination or during structure formation.

    \item \textbf{Small amplitude:}  
    Seed fields are typically $\lesssim 10^{-20}\,\mathrm{G}$.

    \item \textbf{No large-scale coherence:}  
    Fields are generated locally, not cosmologically.
\end{itemize}

\paragraph{Contrast with RDCC.}
\begin{itemize}
    \item RDCC operates in the early Universe.
    \item Coherence is cosmological.
    \item Seed fields are orders of magnitude larger.
\end{itemize}

% ---------------------------------------------------------
\subsection*{D.4 Astrophysical Dynamos}

Dynamos amplify existing seed fields through turbulence and differential rotation.

\paragraph{Limitations.}
\begin{itemize}
    \item \textbf{Require pre-existing seed fields:}  
    Dynamos cannot operate without initial magnetization.

    \item \textbf{Timescale constraints:}  
    High-redshift galaxies show strong fields too early for standard dynamos.
\end{itemize}

\paragraph{Contrast with RDCC.}
\begin{itemize}
    \item RDCC provides seed fields of the correct amplitude.
    \item Coherence length matches galactic scales after collapse.
    \item No fine-tuning of dynamo efficiency is required.
\end{itemize}

% ---------------------------------------------------------
\subsection*{D.5 Summary of Differences}

Table~\ref{tab:comparison} summarizes the key differences between RDCC and standard 
magnetogenesis mechanisms.

\begin{table}[H]
\centering
\begin{tabular}{lccc}
\hline
\textbf{Mechanism} & \textbf{Coherence} & \textbf{Amplitude} & \textbf{New Physics?} \\
\hline
Inflationary        & large     & model-dependent & yes \\
Phase transitions   & tiny      & small           & no \\
Battery mechanisms  & local     & very small      & no \\
Dynamos             & inherited & amplified       & no \\
RDCC (this work)    & large     & $10^{-15}$ G    & no \\
\hline
\end{tabular}
\caption{Comparison of magnetogenesis mechanisms. RDCC provides large coherence, sufficient 
seed amplitude, and requires no new fields or couplings.}
\label{tab:comparison}
\end{table}

% ---------------------------------------------------------
\subsection*{D.6 Concluding Remarks}

RDCC magnetogenesis differs fundamentally from all standard mechanisms:

\begin{itemize}
    \item it does not rely on breaking conformal invariance,
    \item it does not require phase transitions or turbulence,
    \item it does not depend on astrophysical processes,
    \item it operates universally through the relaxation dynamics of the bipartite sectors.
\end{itemize}

This makes RDCC magnetogenesis both robust and predictive, with clear observational 
signatures and minimal theoretical assumptions.

\bibliographystyle{apsrev4-2}
\nocite{*}
\bibliography{references}

\end{document}
