\documentclass[11pt,a4paper]{article}

% ---------------------------------------------------------
% PACKAGES
% ---------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{authblk}
\usepackage{geometry}
\usepackage{cite}
\usepackage{physics}
\usepackage{tensor}
\usepackage{enumitem}

\geometry{margin=1in}

% ---------------------------------------------------------
% TITLE
% ---------------------------------------------------------
\title{\textbf{Companion XVIII: Small-Scale Structure in the Relaxation-Driven Cyclic Cosmology}\\
\textbf{Scale-Dependent Growth, Vorticity, and Lyman-$\alpha$ Signatures}}

\author[1]{Michael Lehmann}
\affil[1]{\small Independent Researcher, Relaxation-Driven Cyclic Cosmology (RDCC) Framework\\
\texttt{mi.lehmann@gmx.de}}

\date{\small \today}

% ---------------------------------------------------------
% DOCUMENT START
% ---------------------------------------------------------
\begin{document}

\maketitle

\begin{abstract}
Small-scale structure formation provides some of the most sensitive probes of cosmological
physics beyond the standard $\Lambda$CDM model. Observables such as the persistent $S_8$
tension, the detailed slope of the Lyman-$\alpha$ forest power spectrum, the surprising
abundance of massive galaxies at high redshift, and measurements of cosmic vorticity and
velocity dispersion all suggest possible deviations from purely scale-independent growth.

In the Relaxation-Driven Cyclic Cosmology (RDCC), these deviations arise naturally from the
bipartite sectoral structure and the universal relaxation mechanism. Sector-dependent sound
speeds, different free-streaming lengths, relaxation-induced damping of perturbations, and
anisotropic stress generated by inter-sectoral decoherence gradients together produce a
characteristic scale-dependent growth function $D(a,k)$.

This Companion derives the complete set of RDCC small-scale structure equations within the
Boltzmann framework, computes the resulting matter power spectrum, and identifies clear
observational signatures in the Lyman-$\alpha$ forest, weak lensing, redshift-space distortions,
and high-redshift galaxy counts. The predicted suppression scale is directly linked to the
relaxation rate $\Gamma(a)$, making the mechanism both predictive and testable with existing
and upcoming data.

These results establish small-scale structure as a powerful observational window into the
fundamental relaxation dynamics of RDCC.
\end{abstract}
% ---------------------------------------------------------
% SECTION 2 — FROM SECTORAL DYNAMICS TO SCALE-DEPENDENT GROWTH
% ---------------------------------------------------------
\section{From Sectoral Dynamics to Scale-Dependent Growth}

In the Relaxation-Driven Cyclic Cosmology (RDCC), the two CPT-conjugate sectors evolve
with different temperatures, sound speeds, free-streaming lengths, and relaxation rates.
These differences cause the response of matter perturbations to gravitational potentials to
become intrinsically scale-dependent. The resulting behaviour is a direct consequence of the
bipartite structure introduced in Companion~X and the thermodynamic asymmetry derived
in Companion~XIII.

The key physical ingredients responsible for scale-dependent growth are:

\begin{itemize}
\item sector-dependent sound speeds $c_{s,i}(a)$,
\item different free-streaming lengths $\lambda_{\rm fs,i}(a)$,
\item relaxation-induced damping of perturbations,
\item anisotropic stress generated by inter-sectoral decoherence gradients (Companion~XVII),
\item a scale-dependent effective gravitational coupling $G_{\rm eff}(k,a)$.
\end{itemize}

Together, these mechanisms generate a characteristic scale-dependent growth function
$D(a,k)$, a modified matter power spectrum, and non-trivial velocity and vorticity spectra.
Importantly, no new particle species or fine-tuned parameters are required; the scale
dependence emerges directly from the fundamental relaxation dynamics of RDCC.

% ---------------------------------------------------------
% 2.1 Sectoral Temperatures and Sound Speeds
% ---------------------------------------------------------
\subsection{Sectoral Temperatures and Sound Speeds}

The two sectors possess different temperatures (Companion~XIII),
\begin{equation}
\frac{T_B}{T_A} \approx 0.5,
\end{equation}
which leads to different sound speeds,
\begin{equation}
c_{s,i}^2 = \frac{\delta p_i}{\delta\rho_i},
\qquad i \in \{A,B\}.
\end{equation}

Because $c_{s,A} \neq c_{s,B}$, density perturbations respond differently to gravitational
potentials in each sector. To leading order,
\begin{equation}
\delta_i(k) \propto
\frac{1}{1 + (c_{s,i}^2 k^2 / H^2)},
\end{equation}
so modes with $k \gg H/c_{s,i}$ are suppressed.

% ---------------------------------------------------------
% 2.2 Sectoral Free-Streaming Lengths
% ---------------------------------------------------------
\subsection{Sectoral Free-Streaming Lengths}

The free-streaming length of each sector is
\begin{equation}
\lambda_{\rm fs,i}(a)
= \int^a \frac{c_{s,i}(a')}{a'H(a')} \, da'.
\end{equation}

Since $c_{s,A} \neq c_{s,B}$, the two sectors possess different free-streaming scales. Modes
with $k > k_{\rm fs,i} = 1/\lambda_{\rm fs,i}$ are suppressed in sector $i$, and the total matter
perturbation inherits a weighted combination of these suppressions.

% ---------------------------------------------------------
% 2.3 Relaxation-Induced Damping
% ---------------------------------------------------------
\subsection{Relaxation-Induced Damping}

The relaxation mechanism introduces additional damping terms in the continuity and Euler
equations. The relaxation rate is
\begin{equation}
\Gamma_i(a) = \alpha_i(a) H(a),
\end{equation}
with $\alpha_A(a) \neq \alpha_B(a)$.

Relaxation suppresses perturbations on scales
\begin{equation}
k \gtrsim k_{\rm rel}(a)
= \frac{\Gamma(a)}{c_{\rm eff}(a)},
\end{equation}
where $c_{\rm eff}$ is the effective sound speed of the combined fluid.

This suppression is smooth, redshift-dependent, and strongest in the Lyman-$\alpha$ regime.

% ---------------------------------------------------------
% 2.4 Decoherence-Gradient Anisotropic Stress
% ---------------------------------------------------------
\subsection{Decoherence-Gradient Anisotropic Stress}

As shown in Companion~XVII, inter-sectoral decoherence gradients generate anisotropic
stress of the form
\begin{equation}
\pi_{ij}^{\rm (rel)} \propto
\partial_{\langle i}\partial_{j\rangle} \Delta_{\rm dec},
\end{equation}
where
\begin{equation}
\Delta_{\rm dec}
= \Gamma_A^{-1} - \Gamma_B^{-1}.
\end{equation}

This anisotropic stress modifies both scalar and vector perturbations, generating:

\begin{itemize}
\item additional suppression of density perturbations near $k \sim k_{\rm rel}$,
\item enhanced vorticity on intermediate scales,
\item a modified velocity divergence spectrum.
\end{itemize}

% ---------------------------------------------------------
% 2.5 Effective Gravitational Coupling
% ---------------------------------------------------------
\subsection{Effective Gravitational Coupling}

The combined effects of sectoral sound speeds, free-streaming, and relaxation modify the
relation between density perturbations and the gravitational potential. The Poisson equation
acquires a scale-dependent correction,
\begin{equation}
k^2 \Phi(k,a)
= 4\pi G_{\rm eff}(k,a) a^2 \rho_{\rm tot}(a)\,\delta_{\rm tot}(k,a),
\end{equation}
with
\begin{equation}
G_{\rm eff}(k,a)
= G \left[ 1 + \mathcal{O}\!\left(\frac{k^2}{k_{\rm rel}^2}\right) \right].
\end{equation}

This modification is mild on large scales but becomes significant for
$k \gtrsim k_{\rm rel}$.

% ---------------------------------------------------------
% 2.6 Summary
% ---------------------------------------------------------
\subsection{Summary}

The bipartite structure and relaxation dynamics of RDCC naturally generate scale-dependent
growth through:

\begin{itemize}
\item sector-dependent sound speeds and free-streaming lengths,
\item relaxation-induced damping of perturbations,
\item anisotropic stress from decoherence gradients,
\item a scale-dependent effective gravitational coupling.
\end{itemize}

These ingredients produce a characteristic suppression of small-scale power and enhanced
vorticity, providing clear observational signatures in the Lyman-$\alpha$ forest, weak lensing,
and redshift-space distortions.
% ---------------------------------------------------------
% SECTION 3 — RDCC SMALL-SCALE STRUCTURE EQUATIONS IN THE BOLTZMANN FRAMEWORK
% ---------------------------------------------------------
\section{RDCC Small-Scale Structure Equations in the Boltzmann Framework}

Building directly on the Boltzmann hierarchy developed in Companion~XVI, we now derive
the perturbed continuity and Euler equations for each sector, including the effects of
sector-dependent sound speeds, free-streaming, relaxation-induced damping, and anisotropic
stress from inter-sectoral decoherence gradients (Companion~XVII). These ingredients
combine to produce the characteristic scale-dependent growth function $D(a,k)$.

% ---------------------------------------------------------
% 3.1 Continuity Equation with Relaxation Damping
% ---------------------------------------------------------
\subsection{Continuity Equation with Relaxation Damping}

For each sector $i \in \{A,B\}$, the continuity equation becomes
\begin{equation}
\dot{\delta}_i
= -(1+w_i)(\theta_i - 3\dot{\Phi})
- 3H(c_{s,i}^2 - w_i)\delta_i
- \Gamma_i\,\delta_i,
\end{equation}
where:
\begin{itemize}
\item $c_{s,i}^2$ is the sectoral sound speed,
\item $w_i$ is the equation of state,
\item $\Gamma_i = \alpha_i(a)H(a)$ is the relaxation rate,
\item $\Phi$ is the Newtonian potential.
\end{itemize}

The final term represents relaxation-induced damping of density perturbations. It is
negligible on large scales but becomes important for $k \gtrsim k_{\rm rel}$.

% ---------------------------------------------------------
% 3.2 Euler Equation with Free-Streaming and Anisotropic Stress
% ---------------------------------------------------------
\subsection{Euler Equation with Free-Streaming and Anisotropic Stress}

The Euler equation becomes
\begin{equation}
\dot{\theta}_i
= -H(1 - 3w_i)\theta_i
+ \frac{c_{s,i}^2}{1+w_i}k^2\delta_i
+ k^2\Psi
- \Gamma_i\,\theta_i
- k^2\pi_i^{\rm (rel)},
\end{equation}
where:
\begin{itemize}
\item the $c_{s,i}^2 k^2 \delta_i$ term encodes free-streaming suppression,
\item the $-\Gamma_i\theta_i$ term is relaxation-induced momentum damping,
\item $\pi_i^{\rm (rel)}$ is the decoherence-gradient anisotropic stress from Companion~XVII.
\end{itemize}

The anisotropic stress term is given by
\begin{equation}
\pi_i^{\rm (rel)}(k)
= C_{\rm rel}\,k^2\,\Delta_{\rm dec}(k),
\end{equation}
with
\begin{equation}
\Delta_{\rm dec}
= \Gamma_A^{-1} - \Gamma_B^{-1}.
\end{equation}

This term enhances vorticity and modifies the velocity divergence spectrum near the
relaxation scale.

% ---------------------------------------------------------
% 3.3 Total Matter Perturbation and Effective Growth Equation
% ---------------------------------------------------------
\subsection{Total Matter Perturbation and Effective Growth Equation}

The total matter perturbation is
\begin{equation}
\delta_{\rm tot}
= \frac{\rho_A\delta_A + \rho_B\delta_B}{\rho_A + \rho_B}.
\end{equation}

Combining the sectoral continuity and Euler equations yields a modified growth equation
for $\delta_{\rm tot}$:
\begin{equation}
\ddot{\delta}_{\rm tot}
+ \left[ 2H + \Gamma_{\rm eff}(a,k) \right]\dot{\delta}_{\rm tot}
+ \left[ c_{\rm eff}^2(a,k)\frac{k^2}{a^2}
- 4\pi G_{\rm eff}(a,k)\rho_{\rm tot} \right]\delta_{\rm tot}
= S_{\rm rel}(a,k),
\end{equation}
where:
\begin{align}
\Gamma_{\rm eff}(a,k) &=
\frac{\rho_A\Gamma_A + \rho_B\Gamma_B}{\rho_{\rm tot}}, \\
c_{\rm eff}^2(a,k) &=
\frac{\rho_A c_{s,A}^2 + \rho_B c_{s,B}^2}{\rho_{\rm tot}}, \\
G_{\rm eff}(a,k) &=
G\left[1 + \mathcal{O}\!\left(\frac{k^2}{k_{\rm rel}^2}\right)\right], \\
S_{\rm rel}(a,k) &=
-k^2\left(\rho_A\pi_A^{\rm (rel)} + \rho_B\pi_B^{\rm (rel)}\right)/\rho_{\rm tot}.
\end{align}

This equation makes explicit that:
\begin{itemize}
\item the friction term grows with $\Gamma_{\rm eff}$,
\item the pressure term grows with $c_{\rm eff}^2 k^2$,
\item the gravitational driving term weakens as $G_{\rm eff}(k)$ decreases,
\item the source term $S_{\rm rel}$ enhances vorticity and suppresses growth.
\end{itemize}

% ---------------------------------------------------------
% 3.4 Definition of the RDCC Growth Function
% ---------------------------------------------------------
\subsection{Definition of the RDCC Growth Function}

We define the RDCC growth function $D(a,k)$ via
\begin{equation}
\delta_{\rm tot}(a,k)
= D(a,k)\,\delta_{\rm tot}(a_{\rm ini},k).
\end{equation}

Inserting this into the growth equation yields
\begin{equation}
D'' + \left( \frac{H'}{H} + 2 + \frac{\Gamma_{\rm eff}}{H} \right) D'
+ \left[
\frac{c_{\rm eff}^2 k^2}{a^2 H^2}
- \frac{3}{2}\Omega_m(a)\frac{G_{\rm eff}(k,a)}{G}
\right] D
= \frac{S_{\rm rel}(a,k)}{H^2\,\delta_{\rm tot}(a_{\rm ini},k)}.
\end{equation}

This equation explicitly shows that the growth function becomes scale-dependent whenever:
\begin{itemize}
\item $c_{\rm eff}^2 k^2 / H^2$ is non-negligible,
\item $\Gamma_{\rm eff}/H = \mathcal{O}(1)$,
\item $G_{\rm eff}(k,a)$ deviates from $G$,
\item $S_{\rm rel}(a,k)$ is non-zero.
\end{itemize}

% ---------------------------------------------------------
% 3.5 Three Growth Regimes
% ---------------------------------------------------------
\subsection{Three Growth Regimes}

The RDCC growth function exhibits three characteristic regimes:

\paragraph{(i) Large scales ($k \ll k_{\rm fs}$).}
\begin{equation}
D(a,k) \approx D_{\Lambda{\rm CDM}}(a).
\end{equation}

\paragraph{(ii) Intermediate scales ($k \sim k_{\rm fs}$).}
Smooth suppression from sectoral sound speeds and free-streaming.

\paragraph{(iii) Small scales ($k \gg k_{\rm rel}$).}
Strong suppression from relaxation damping and decoherence-gradient anisotropic stress.

% ---------------------------------------------------------
% 3.6 Summary
% ---------------------------------------------------------
\subsection{Summary}

The RDCC Boltzmann framework naturally produces scale-dependent growth through:
\begin{itemize}
\item sector-dependent sound speeds and free-streaming,
\item relaxation-induced damping of density and velocity perturbations,
\item anisotropic stress from decoherence gradients,
\item a scale-dependent effective gravitational coupling.
\end{itemize}

These effects combine to yield the RDCC growth function $D(a,k)$ and the characteristic
suppression of small-scale power observed in the matter power spectrum.
% ---------------------------------------------------------
% SECTION 4 — RDCC PREDICTIONS FOR THE MATTER POWER SPECTRUM
% ---------------------------------------------------------
\section{RDCC Predictions for the Matter Power Spectrum}

The matter power spectrum in RDCC inherits the full scale dependence of the growth
function $D(a,k)$ derived in Section~3. The combined effects of sector-dependent sound
speeds, free-streaming, relaxation damping, and decoherence-gradient anisotropic stress
produce a smooth, redshift-dependent suppression of small-scale power.

A useful phenomenological approximation is
\begin{equation}
P_{\rm RDCC}(k,a)
= \frac{P_{\Lambda{\rm CDM}}(k,a)}
{\left[1 + (k/k_{\rm rel})^2\right]^{\nu}},
\end{equation}
where:
\begin{itemize}
\item $k_{\rm rel}(a) = \Gamma(a)/c_{\rm eff}(a)$ is the relaxation scale,
\item $\nu \approx 1$–$1.5$ depends on the sectoral thermodynamics,
\item $P_{\Lambda{\rm CDM}}$ is the standard matter power spectrum.
\end{itemize}

This suppression is:
\begin{itemize}
\item smooth (unlike the sharp cutoff of warm dark matter),
\item redshift-dependent (stronger at high $z$),
\item strongest in the Lyman-$\alpha$ forest regime ($k \sim 1$–$10\,h\,{\rm Mpc}^{-1}$),
\item mild enough to preserve BAO and large-scale clustering.
\end{itemize}

Relaxation-induced anisotropic stress additionally enhances vorticity and modifies the
velocity divergence spectrum near $k \sim k_{\rm rel}$, providing further observational
handles on the relaxation dynamics.

% ---------------------------------------------------------
% SECTION 5 — OBSERVATIONAL SIGNATURES AND CONSTRAINTS
% ---------------------------------------------------------
\section{Observational Signatures and Constraints}

The scale-dependent growth predicted by RDCC leads to a rich set of observational
consequences across weak lensing, redshift-space distortions, the Lyman-$\alpha$ forest,
and cosmic web statistics.

% ---------------------------------------------------------
% 5.1 Weak Lensing and the S8 Tension
% ---------------------------------------------------------
\subsection{Weak Lensing and the \texorpdfstring{\(S_8\)}{S8} Tension}

Weak lensing surveys measure the combination
\begin{equation}
S_8 = \sigma_8 \sqrt{\Omega_m/0.3}.
\end{equation}

RDCC predicts a reduction in $\sigma_8$ due to suppressed small-scale growth, naturally
alleviating the persistent $S_8$ tension between CMB and low-redshift probes. The
suppression is mild enough to remain consistent with galaxy clustering and BAO.

% ---------------------------------------------------------
% 5.2 Redshift-Space Distortions
% ---------------------------------------------------------
\subsection{Redshift-Space Distortions}

Redshift-space distortions probe the scale-dependent growth rate
\begin{equation}
f(a,k) = \frac{d\ln D(a,k)}{d\ln a}.
\end{equation}

RDCC predicts:
\begin{itemize}
\item standard behaviour for $k \ll k_{\rm fs}$,
\item mild suppression for $k \sim k_{\rm fs}$,
\item strong suppression for $k \gg k_{\rm rel}$.
\end{itemize}

Future surveys (DESI, Euclid, LSST) will be able to detect or tightly constrain
$k_{\rm rel}(z=0) \sim 0.5$–$2\,h\,{\rm Mpc}^{-1}$.

% ---------------------------------------------------------
% 5.3 Lyman-$\alpha$ Forest
% ---------------------------------------------------------
\subsection{Lyman-\texorpdfstring{\(\alpha\)}{alpha} Forest}

The Lyman-$\alpha$ forest is particularly sensitive to the suppression of power at
$k \gtrsim 1\,h\,{\rm Mpc}^{-1}$. RDCC predicts:
\begin{itemize}
\item a smooth suppression of the flux power spectrum,
\item a redshift-dependent shift in the slope,
\item no sharp cutoff (unlike warm dark matter),
\item consistency with observed high-redshift galaxy formation.
\end{itemize}

These features provide a clean observational test of the relaxation scale.

% ---------------------------------------------------------
% 5.4 Vorticity and Velocity Dispersion
% ---------------------------------------------------------
\subsection{Vorticity and Velocity Dispersion}

Relaxation-induced anisotropic stress (Companion~XVII) enhances vorticity on
intermediate scales:
\begin{equation}
\omega(k) \propto k^2 \Delta_{\rm dec}(k).
\end{equation}

This leads to:
\begin{itemize}
\item increased velocity dispersion near $k_{\rm rel}$,
\item modified cosmic web morphology,
\item potential signatures in kinetic SZ and peculiar velocity surveys.
\end{itemize}

% ---------------------------------------------------------
% 5.5 High-Redshift Galaxy Formation
% ---------------------------------------------------------
\subsection{High-Redshift Galaxy Formation}

Because RDCC suppresses only small-scale power, massive halos form similarly to
$\Lambda$CDM. Thus:
\begin{itemize}
\item early massive galaxies remain unaffected,
\item JWST observations of high-$z$ galaxies are naturally accommodated,
\item only low-mass halos experience suppression.
\end{itemize}

% ---------------------------------------------------------
% 5.6 Summary of Observational Signatures
% ---------------------------------------------------------
\subsection{Summary of Observational Signatures}

RDCC predicts:
\begin{itemize}
\item reduced $S_8$ from weak lensing,
\item scale-dependent growth in redshift-space distortions,
\item smooth suppression in the Lyman-$\alpha$ forest,
\item enhanced vorticity and velocity dispersion,
\item preserved BAO and large-scale clustering,
\item unaffected formation of massive high-$z$ galaxies.
\end{itemize}

These signatures provide multiple independent tests of the relaxation dynamics.

% ---------------------------------------------------------
% SECTION 6 — CONCLUSION
% ---------------------------------------------------------
\section{Conclusion}

This Companion has developed a comprehensive framework for small-scale structure
formation in the Relaxation-Driven Cyclic Cosmology. Starting from the bipartite sectoral
structure and the universal relaxation mechanism, we have shown that RDCC naturally
generates scale-dependent growth without introducing new particles or fine-tuned
parameters.

The resulting suppression of small-scale power, modified growth function, and enhanced
vorticity provide clear, testable predictions across weak lensing, the Lyman-$\alpha$ forest,
redshift-space distortions, and cosmic web statistics. The mechanism is tightly linked to
the fundamental relaxation rate $\Gamma(a)$, making small-scale structure a powerful
observational window into the core dynamics of RDCC.

These results complete the bridge from linear perturbation theory to the quasi-nonlinear
regime and set the stage for future detailed studies of halo statistics, galaxy formation,
and cosmic web morphology within the relaxation-driven picture.

\end{document}
