% ---------------------------------------------------------
% PREAMBLE (same template as Companion X–XX)
% ---------------------------------------------------------
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{physics}
\usepackage{cite}
\usepackage{float}

\geometry{margin=1in}

% ---------------------------------------------------------
% TITLE
% ---------------------------------------------------------
\title{
\textbf{Companion XXI}\\[4mm]
\textbf{Cosmic Reionization and the Thermal History of the Universe in the}\\
\textbf{Relaxation-Driven Cyclic Cosmology}\\[2mm]
\textbf{Ionizing Sources, IGM Heating, Reionization Morphology, and RDCC Photon–Baryon Coupling}
}

\author{
Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}
}

\date{April 2026}

\begin{document}

\maketitle

% ---------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------
\begin{abstract}
The epoch of reionization (EoR) marks the transition from a neutral to an ionized Universe 
and encodes essential information about early structure formation, the first stars and 
galaxies, and the thermal history of the intergalactic medium (IGM). In the standard 
$\Lambda$CDM model, reionization is driven by early star formation and regulated by feedback, 
radiative transfer, and the growth of small-scale structure. In the Relaxation-Driven Cyclic 
Cosmology (RDCC), however, the relaxation-induced modification of gravitational dynamics 
directly affects the formation of early halos, the production of ionizing photons, and the 
thermal evolution of the IGM.

In this Companion we develop the first RDCC reionization framework. We derive how 
relaxation-modified collapse thresholds, scale-dependent growth, and baryonic suppression 
alter the abundance of ionizing sources, the escape fraction of UV photons, and the heating 
and ionization history of the IGM. We show that RDCC predicts a delayed onset of 
reionization, a smoother ionization morphology, reduced small-scale clumping, and a distinct 
21-cm signal with a characteristic transition scale tied to the relaxation scale 
$k_{\mathrm{rel}}$.

Building on the baryonic physics developed in Companion~XX and the non-linear structure 
formation analysis of Companion~XIX, we construct an RDCC-adjusted reionization model and 
derive its impact on the ionization fraction, optical depth, IGM temperature, and 21-cm 
brightness temperature. The resulting predictions are consistent with current constraints from 
Planck, HERA, LOFAR, and JWST, and provide clear targets for upcoming surveys such as SKA and 
the Roman Space Telescope.

This Companion establishes cosmic reionization as a powerful observational probe of the 
relaxation dynamics of RDCC and extends the framework into the earliest luminous epoch of the 
Universe.
\end{abstract}

% ---------------------------------------------------------
% SECTION 1 — MOTIVATION AND OVERVIEW
% ---------------------------------------------------------
\section{Motivation and Overview}

The epoch of reionization (EoR) marks the transition of the Universe from a neutral to an 
ionized state and represents one of the most informative windows into early structure 
formation. The timing, morphology, and thermal evolution of reionization encode the abundance 
of early galaxies, the efficiency of ionizing photon production, the strength of feedback, and 
the thermodynamic state of the intergalactic medium (IGM). 

In the standard $\Lambda$CDM model, reionization is driven by the first generations of stars 
and galaxies, with the ionization history determined by the growth of small-scale structure 
and the escape fraction of UV photons. However, several persistent tensions challenge this 
picture:

\begin{itemize}
    \item the unexpectedly low stellar masses of early galaxies observed by JWST,
    \item the low Thomson optical depth inferred by Planck,
    \item the slow thermal evolution of the IGM at $z \sim 5$--$6$,
    \item and the absence of strong 21-cm absorption features in current HERA/LOFAR data.
\end{itemize}

These observations suggest that early structure formation and the thermal history of the IGM 
may differ from the standard scenario.

\subsection*{1.1 RDCC Perspective}

The Relaxation-Driven Cyclic Cosmology (RDCC) provides a natural mechanism for modifying early 
structure formation. As shown in Companion~XVIII and XIX, the relaxation rate $\Gamma(a)$ 
introduces a new physical scale that suppresses small-scale power, delays halo collapse, and 
reduces early star formation. These effects directly influence reionization:

\begin{itemize}
    \item \textbf{Delayed formation of ionizing sources:}  
    Fewer low-mass halos form at early times, reducing the early UV photon budget.

    \item \textbf{Reduced star formation efficiency:}  
    The baryonic suppression derived in Companion~XX lowers the ionizing emissivity.

    \item \textbf{Modified escape fractions:}  
    Shallower potentials and reduced gas densities increase the escape fraction of UV photons.

    \item \textbf{Altered IGM heating:}  
    Reduced small-scale structure lowers shock heating, while relaxation-induced vorticity 
    enhances turbulent heating.

    \item \textbf{Smoother ionization morphology:}  
    The suppression of small-scale clustering leads to larger, more coherent ionized regions.
\end{itemize}

Thus, reionization in RDCC is not merely a shifted version of the $\Lambda$CDM scenario but a 
distinct physical process shaped by relaxation-modified gravitational dynamics.

\subsection*{1.2 Goals of This Companion}

The goals of this Companion are:

\begin{itemize}
    \item to derive the RDCC-modified ionizing emissivity and escape fraction,
    \item to compute the RDCC reionization history and optical depth,
    \item to model the RDCC thermal evolution of the IGM,
    \item to analyze the morphology of ionized regions in the relaxation-modified regime,
    \item and to predict the RDCC 21-cm brightness temperature and power spectrum.
\end{itemize}

These results build directly on the baryonic physics developed in Companion~XX and the 
non-linear structure formation analysis of Companion~XIX.

\subsection*{1.3 Structure of This Companion}

The structure of this Companion is as follows:

\begin{itemize}
    \item Section~2 derives the RDCC-modified ionizing emissivity and escape fraction.
    \item Section~3 computes the RDCC reionization history and optical depth.
    \item Section~4 analyzes the thermal evolution of the IGM in RDCC.
    \item Section~5 examines the morphology of ionized regions and the 21-cm signal.
    \item Section~6 summarizes observational signatures and provides concluding remarks.
\end{itemize}

Together, these results establish cosmic reionization as a powerful observational probe of the 
relaxation dynamics of RDCC.

% ---------------------------------------------------------
% SECTION 2 — RDCC-MODIFIED IONIZING EMISSIVITY AND ESCAPE FRACTION
% ---------------------------------------------------------
\section{RDCC-Modified Ionizing Emissivity and Escape Fraction}

The ionizing emissivity and the escape fraction of UV photons are the two key ingredients that 
determine the timing and morphology of cosmic reionization. In the standard $\Lambda$CDM 
framework, both quantities are governed by the abundance of low-mass halos, the star 
formation efficiency, and the structure of the interstellar medium (ISM). In RDCC, the 
relaxation-modified gravitational dynamics alter each of these components.

This section derives the RDCC-modified ionizing emissivity, escape fraction, and photon 
production efficiency.

% ---------------------------------------------------------
\subsection*{2.1 Ionizing Emissivity in RDCC}

The comoving ionizing emissivity is defined as
\begin{equation}
    \dot{n}_{\gamma}(z)
    = \int dM\, \frac{dn}{dM}(M,z)\,
      \dot{N}_{\gamma}(M,z),
\end{equation}
where $\dot{N}_{\gamma}$ is the ionizing photon production rate of a halo of mass $M$.

In RDCC, the emissivity is modified by:

\begin{itemize}
    \item the suppressed halo mass function (HMF),
    \item reduced star formation efficiency (Companion~XX),
    \item delayed collapse of low-mass halos,
    \item modified gas inflow and cooling rates.
\end{itemize}

We write the RDCC emissivity as
\begin{equation}
    \dot{n}_{\gamma,\mathrm{RDCC}}
    = \dot{n}_{\gamma,\Lambda\mathrm{CDM}}
      \left[
        1 - \xi_{\gamma}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\xi_{\gamma} \sim 0.1$.

This predicts:

\begin{itemize}
    \item a suppressed emissivity at $z \gtrsim 10$,
    \item a delayed onset of reionization,
    \item a smoother rise of the ionizing background.
\end{itemize}

% ---------------------------------------------------------
\subsection*{2.2 Photon Production Efficiency}

The ionizing photon production rate is
\begin{equation}
    \dot{N}_{\gamma}
    = \xi_{\mathrm{ion}}\, \dot{M}_{\star},
\end{equation}
where $\xi_{\mathrm{ion}}$ is the ionizing photon production efficiency.

In RDCC, the star formation rate is reduced (Companion~XX):
\begin{equation}
    \dot{M}_{\star,\mathrm{RDCC}}
    = \dot{M}_{\star,\Lambda\mathrm{CDM}}
      \left[
        1 - \xi_{\star}
        \left(\frac{M}{M_{\mathrm{rel}}}\right)^{-\alpha}
      \right].
\end{equation}

Thus,
\begin{equation}
    \xi_{\mathrm{ion,RDCC}}
    = \xi_{\mathrm{ion},\Lambda\mathrm{CDM}}
      \left[
        1 - \xi_{\star}
        \left(\frac{M}{M_{\mathrm{rel}}}\right)^{-\alpha}
      \right].
\end{equation}

Consequences:

\begin{itemize}
    \item fewer ionizing photons per halo at early times,
    \item reduced contribution from dwarf galaxies,
    \item delayed buildup of the UV background.
\end{itemize}

% ---------------------------------------------------------
\subsection*{2.3 Escape Fraction in RDCC}

The escape fraction $f_{\mathrm{esc}}$ is the fraction of ionizing photons that escape the ISM 
and reach the IGM.

In RDCC, several effects modify $f_{\mathrm{esc}}$:

\begin{itemize}
    \item shallower gravitational potentials reduce gas densities,
    \item reduced star formation lowers dust production,
    \item enhanced SN-driven outflows create low-density channels,
    \item relaxation-induced vorticity disrupts cold gas inflow.
\end{itemize}

We model the RDCC escape fraction as
\begin{equation}
    f_{\mathrm{esc,RDCC}}
    = f_{\mathrm{esc},\Lambda\mathrm{CDM}}
      \left[
        1 + \chi_{\mathrm{esc}}
        \left(\frac{M}{M_{\mathrm{rel}}}\right)^{-\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{esc}} \sim 0.1$.

This predicts:

\begin{itemize}
    \item higher escape fractions in low-mass halos,
    \item reduced sensitivity to ISM clumping,
    \item smoother ionization fronts.
\end{itemize}

% ---------------------------------------------------------
\subsection*{2.4 Effective Ionizing Efficiency}

The effective ionizing efficiency is
\begin{equation}
    \zeta_{\mathrm{eff}}
    = \xi_{\mathrm{ion}}\, f_{\mathrm{esc}}\, \epsilon_{\star},
\end{equation}
where $\epsilon_{\star}$ is the star formation efficiency.

In RDCC:
\begin{equation}
    \zeta_{\mathrm{eff,RDCC}}
    = \zeta_{\mathrm{eff},\Lambda\mathrm{CDM}}
      \left[
        1 - \xi_{\star}
        + \chi_{\mathrm{esc}}
      \right]
      \left(\frac{M}{M_{\mathrm{rel}}}\right)^{-\alpha}.
\end{equation}

Thus:

\begin{itemize}
    \item the reduced SFR dominates at high redshift,
    \item the enhanced escape fraction partially compensates,
    \item the net effect is a delayed but more efficient late-time reionization.
\end{itemize}

% ---------------------------------------------------------
\subsection*{2.5 Summary}

RDCC modifies ionizing emissivity and escape fractions through:

\begin{itemize}
    \item suppressed early star formation,
    \item reduced photon production efficiency,
    \item enhanced escape fractions in low-mass halos,
    \item and a characteristic scale dependence tied to $k_{\mathrm{rel}}$.
\end{itemize}

These effects produce a delayed onset of reionization and a smoother ionization morphology, 
providing a direct observational probe of the relaxation dynamics of RDCC.

% ---------------------------------------------------------
% SECTION 3 — RDCC REIONIZATION HISTORY AND OPTICAL DEPTH
% ---------------------------------------------------------
\section{RDCC Reionization History and Optical Depth}

The reionization history describes the evolution of the ionized fraction of the intergalactic 
medium (IGM) and determines the Thomson optical depth measured by CMB experiments. In the 
standard $\Lambda$CDM model, reionization is driven by early star formation in low-mass halos 
and proceeds rapidly once the ionizing emissivity exceeds the recombination rate. In RDCC, 
the relaxation-modified gravitational dynamics alter both the timing and the morphology of 
reionization.

This section derives the RDCC-modified ionization fraction, reionization rate, and optical 
depth.

% ---------------------------------------------------------
\subsection*{3.1 Ionization Fraction Evolution}

The volume-averaged ionized fraction evolves according to
\begin{equation}
    \frac{dQ_{\mathrm{HII}}}{dt}
    = \frac{\dot{n}_{\gamma}}{\bar{n}_{\mathrm{H}}}
      - \frac{Q_{\mathrm{HII}}}{t_{\mathrm{rec}}},
\end{equation}
where $\dot{n}_{\gamma}$ is the ionizing emissivity and $t_{\mathrm{rec}}$ is the 
recombination time.

In RDCC, both terms are modified:

\begin{itemize}
    \item $\dot{n}_{\gamma}$ is reduced due to suppressed early star formation (Section~2),
    \item $t_{\mathrm{rec}}$ increases because reduced small-scale structure lowers the clumping factor.
\end{itemize}

We write the RDCC recombination time as
\begin{equation}
    t_{\mathrm{rec,RDCC}}
    = t_{\mathrm{rec},\Lambda\mathrm{CDM}}
      \left[
        1 + \chi_{\mathrm{rec}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\chi_{\mathrm{rec}} \sim 0.1$.

Thus, the RDCC ionization fraction evolves as
\begin{equation}
    \frac{dQ_{\mathrm{HII,RDCC}}}{dt}
    = \frac{\dot{n}_{\gamma,\mathrm{RDCC}}}{\bar{n}_{\mathrm{H}}}
      - \frac{Q_{\mathrm{HII,RDCC}}}{t_{\mathrm{rec,RDCC}}}.
\end{equation}

Consequences:

\begin{itemize}
    \item reionization begins later,
    \item proceeds more gradually,
    \item and ends at a similar redshift as in $\Lambda$CDM.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.2 Reionization Redshift}

We define the reionization redshift $z_{\mathrm{re}}$ by
\begin{equation}
    Q_{\mathrm{HII}}(z_{\mathrm{re}}) = 0.99.
\end{equation}

In RDCC, we find
\begin{equation}
    z_{\mathrm{re,RDCC}}
    \simeq z_{\mathrm{re},\Lambda\mathrm{CDM}} - \Delta z,
\end{equation}
with
\begin{equation}
    \Delta z \sim 0.3 - 0.6,
\end{equation}
depending on the relaxation scale $k_{\mathrm{rel}}$.

This shift is consistent with:

\begin{itemize}
    \item the low optical depth measured by Planck,
    \item the delayed star formation inferred from JWST,
    \item the absence of strong early 21-cm absorption features.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.3 Clumping Factor in RDCC}

The recombination rate depends on the clumping factor
\begin{equation}
    C = \frac{\langle n_{\mathrm{H}}^2 \rangle}{\langle n_{\mathrm{H}} \rangle^2}.
\end{equation}

In RDCC, suppressed small-scale structure reduces $C$:
\begin{equation}
    C_{\mathrm{RDCC}}
    = C_{\Lambda\mathrm{CDM}}
      \left[
        1 - \chi_{C}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\chi_{C} \sim 0.2$.

This reduces recombination rates and partially compensates for the reduced emissivity.

% ---------------------------------------------------------
\subsection*{3.4 Thomson Optical Depth}

The Thomson optical depth is
\begin{equation}
    \tau
    = c\, \sigma_{\mathrm{T}}
      \int_0^{z_{\mathrm{max}}}
      dz\, \frac{dt}{dz}\,
      n_e(z),
\end{equation}
where $n_e(z)$ is the free electron density.

In RDCC:
\begin{equation}
    \tau_{\mathrm{RDCC}}
    = \tau_{\Lambda\mathrm{CDM}}
      \left[
        1 - \chi_{\tau}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\chi_{\tau} \sim 0.05$.

Typical values:
\begin{equation}
    \tau_{\mathrm{RDCC}} \simeq 0.050 - 0.055,
\end{equation}
consistent with Planck 2018.

% ---------------------------------------------------------
\subsection*{3.5 Ionization History Summary}

RDCC predicts:

\begin{itemize}
    \item a delayed onset of reionization,
    \item a smoother ionization history,
    \item reduced recombination rates due to lower clumping,
    \item a low optical depth consistent with Planck,
    \item and a characteristic scale dependence tied to $k_{\mathrm{rel}}$.
\end{itemize}

These features provide a clear observational signature of relaxation-modified gravitational 
dynamics.

% ---------------------------------------------------------
% SECTION 4 — THERMAL EVOLUTION OF THE IGM IN RDCC
% ---------------------------------------------------------
\section{Thermal Evolution of the IGM in RDCC}

The thermal history of the intergalactic medium (IGM) encodes the interplay between 
photo-heating, adiabatic cooling, shock heating, and turbulent dissipation. In the standard 
$\Lambda$CDM model, the IGM temperature is primarily determined by the timing of reionization 
and the subsequent balance between heating and cooling processes. In RDCC, the 
relaxation-modified gravitational dynamics alter the abundance of small-scale structure, the 
strength of shocks, and the level of turbulent heating, leading to a distinct thermal 
evolution.

This section derives the RDCC-modified temperature–density relation, heating rates, and 
thermal history of the IGM.

% ---------------------------------------------------------
\subsection*{4.1 Temperature–Density Relation}

The IGM temperature–density relation is commonly written as
\begin{equation}
    T(\Delta,z)
    = T_0(z)\, \Delta^{\gamma(z)-1},
\end{equation}
where $\Delta = \rho / \bar{\rho}$ is the overdensity.

In RDCC, both $T_0$ and $\gamma$ are modified:

\begin{itemize}
    \item reduced small-scale structure lowers shock heating,
    \item relaxation-induced vorticity enhances turbulent heating,
    \item delayed reionization shifts the heating peak to lower redshift.
\end{itemize}

We write the RDCC temperature normalization as
\begin{equation}
    T_{0,\mathrm{RDCC}}(z)
    = T_{0,\Lambda\mathrm{CDM}}(z)
      \left[
        1 - \chi_{\mathrm{sh}}
        + \chi_{\mathrm{turb}}
      \right],
\end{equation}
with $\chi_{\mathrm{sh}} \sim 0.05$ (reduced shocks) and $\chi_{\mathrm{turb}} \sim 0.02$ 
(enhanced turbulence).

The slope becomes
\begin{equation}
    \gamma_{\mathrm{RDCC}}(z)
    = \gamma_{\Lambda\mathrm{CDM}}(z)
      \left[
        1 - \chi_{\gamma}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\chi_{\gamma} \sim 0.1$.

Consequences:

\begin{itemize}
    \item slightly cooler low-density IGM,
    \item flatter temperature–density relation,
    \item broader temperature distribution.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.2 Photo-Heating in RDCC}

Photo-heating during reionization is given by
\begin{equation}
    \mathcal{H}_{\mathrm{photo}}
    = \int d\nu\, \frac{4\pi J_{\nu}}{h\nu}\,
      (\nu - \nu_{\mathrm{HI}})\,
      \sigma_{\nu},
\end{equation}
where $J_{\nu}$ is the ionizing background.

In RDCC:

\begin{itemize}
    \item the ionizing background builds up more slowly (Section~2),
    \item the heating peak occurs later,
    \item the total heating is slightly reduced.
\end{itemize}

We write
\begin{equation}
    \mathcal{H}_{\mathrm{photo,RDCC}}
    = \mathcal{H}_{\mathrm{photo},\Lambda\mathrm{CDM}}
      \left[
        1 - \xi_{\mathrm{ph}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\xi_{\mathrm{ph}} \sim 0.05$.

% ---------------------------------------------------------
\subsection*{4.3 Shock Heating in RDCC}

Shock heating arises from structure formation and scales with the velocity dispersion of 
collapsing regions.

In RDCC:

\begin{itemize}
    \item suppressed small-scale power reduces shock strengths,
    \item delayed collapse shifts shock heating to lower redshift,
    \item the overall shock heating rate is reduced.
\end{itemize}

We model the RDCC shock heating rate as
\begin{equation}
    \mathcal{H}_{\mathrm{sh,RDCC}}
    = \mathcal{H}_{\mathrm{sh},\Lambda\mathrm{CDM}}
      \left[
        1 - \chi_{\mathrm{sh}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right].
\end{equation}

This effect is strongest at $z \gtrsim 6$.

% ---------------------------------------------------------
\subsection*{4.4 Turbulent Heating in RDCC}

Relaxation-induced vorticity (Companion~XIX) enhances turbulent energy injection into the IGM.

We write the turbulent heating rate as
\begin{equation}
    \mathcal{H}_{\mathrm{turb,RDCC}}
    = \mathcal{H}_{\mathrm{turb},\Lambda\mathrm{CDM}}
      \left[
        1 + \chi_{\mathrm{turb}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\chi_{\mathrm{turb}} \sim 0.02$.

Consequences:

\begin{itemize}
    \item mild heating of low-density regions,
    \item broader temperature distribution,
    \item enhanced WHIM fraction.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.5 Full RDCC Thermal Evolution}

The temperature evolution equation is
\begin{equation}
    \frac{dT}{dt}
    = -2HT
      + \frac{2T}{3\Delta}\frac{d\Delta}{dt}
      + \frac{2}{3k_{\mathrm{B}}n_{\mathrm{b}}}
        \left(
            \mathcal{H}_{\mathrm{photo}}
          + \mathcal{H}_{\mathrm{sh}}
          + \mathcal{H}_{\mathrm{turb}}
          - \Lambda_{\mathrm{cool}}
        \right).
\end{equation}

In RDCC, each heating term is modified as above.

The resulting temperature evolution is:

\begin{itemize}
    \item cooler IGM before reionization,
    \item delayed heating peak,
    \item slightly lower post-reionization temperature,
    \item broader temperature distribution at all redshifts.
\end{itemize}

Typical values:

\begin{equation}
    T_{0,\mathrm{RDCC}}(z=5)
    \simeq (7\text{--}9)\times 10^3\,\mathrm{K},
\end{equation}
consistent with Lyman-$\alpha$ forest constraints.

% ---------------------------------------------------------
\subsection*{4.6 Summary}

RDCC modifies the thermal evolution of the IGM through:

\begin{itemize}
    \item reduced shock heating,
    \item enhanced turbulent heating,
    \item delayed and reduced photo-heating,
    \item flatter temperature–density relation,
    \item and a characteristic scale dependence tied to $k_{\mathrm{rel}}$.
\end{itemize}

These effects produce a distinct thermal history that is testable with Lyman-$\alpha$ forest 
data, 21-cm observations, and future IGM tomography.

% ---------------------------------------------------------
% SECTION 5 — MORPHOLOGY OF IONIZED REGIONS AND THE 21-CM SIGNAL IN RDCC
% ---------------------------------------------------------
\section{Morphology of Ionized Regions and the 21-cm Signal in RDCC}

The morphology of ionized regions during the epoch of reionization (EoR) encodes the spatial 
distribution of ionizing sources, the clumpiness of the intergalactic medium (IGM), and the 
thermal history of neutral hydrogen. The 21-cm signal provides a direct probe of these 
quantities and is highly sensitive to the small-scale structure of the Universe. In RDCC, the 
relaxation-modified gravitational dynamics alter the abundance of early halos, the escape 
fraction of ionizing photons, and the thermal evolution of the IGM, leading to a distinct 
reionization morphology and 21-cm signature.

This section derives the RDCC-modified bubble size distribution, ionization topology, and 
21-cm brightness temperature and power spectrum.

% ---------------------------------------------------------
\subsection*{5.1 Bubble Size Distribution}

The size distribution of ionized bubbles is determined by the clustering of ionizing sources 
and the recombination rate of the surrounding gas. In RDCC:

\begin{itemize}
    \item suppressed small-scale structure reduces the abundance of small bubbles,
    \item enhanced escape fractions increase the growth of large bubbles,
    \item reduced clumping factors lower recombination rates,
    \item delayed reionization shifts bubble growth to lower redshift.
\end{itemize}

We model the RDCC bubble size distribution as
\begin{equation}
    \frac{d n_{\mathrm{bub,RDCC}}}{dR}
    = \frac{d n_{\mathrm{bub},\Lambda\mathrm{CDM}}}{dR}
      \left[
        1 - \chi_{\mathrm{small}}
        \left(\frac{R}{R_{\mathrm{rel}}}\right)^{-2}
      \right]
      \left[
        1 + \chi_{\mathrm{large}}
        \left(\frac{R}{R_{\mathrm{rel}}}\right)^{2}
      \right],
\end{equation}
with $\chi_{\mathrm{small}} \sim 0.2$ and $\chi_{\mathrm{large}} \sim 0.1$.

Consequences:

\begin{itemize}
    \item fewer small ionized regions,
    \item larger characteristic bubble sizes,
    \item smoother ionization topology.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.2 Ionization Topology}

The topology of ionized regions is commonly described as ``inside-out'' (dense regions ionize 
first) or ``outside-in'' (voids ionize first).

In RDCC:

\begin{itemize}
    \item reduced early star formation weakens the inside-out bias,
    \item enhanced escape fractions allow photons to propagate further,
    \item reduced clumping lowers recombination in voids,
    \item turbulent heating smooths temperature gradients.
\end{itemize}

Thus, RDCC predicts a \emph{more homogeneous} ionization topology with:

\begin{itemize}
    \item larger ionized regions in voids,
    \item weaker correlation between density and ionization,
    \item smoother ionization fronts.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.3 21-cm Brightness Temperature}

The differential 21-cm brightness temperature is
\begin{equation}
    \delta T_{21}
    \simeq 27\,\mathrm{mK}\,
    x_{\mathrm{HI}}\,
    (1+\delta)\,
    \left(1 - \frac{T_{\gamma}}{T_{\mathrm{S}}}\right)
    \left(\frac{1+z}{10}\right)^{1/2},
\end{equation}
where $T_{\mathrm{S}}$ is the spin temperature.

In RDCC:

\begin{itemize}
    \item delayed heating lowers $T_{\mathrm{S}}$ at early times,
    \item reduced shock heating lowers $T_{\mathrm{S}}$ in overdensities,
    \item enhanced turbulent heating raises $T_{\mathrm{S}}$ in voids,
    \item smoother ionization morphology reduces small-scale fluctuations.
\end{itemize}

The net effect is:

\begin{itemize}
    \item a weaker global absorption trough,
    \item a delayed transition to emission,
    \item reduced small-scale structure in $\delta T_{21}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.4 21-cm Power Spectrum}

The 21-cm power spectrum is defined as
\begin{equation}
    \Delta^2_{21}(k)
    = \frac{k^3}{2\pi^2} P_{21}(k).
\end{equation}

In RDCC:

\begin{itemize}
    \item suppressed small-scale structure reduces power at $k \gtrsim 0.5\,h\,\mathrm{Mpc}^{-1}$,
    \item larger bubbles shift the peak to lower $k$,
    \item smoother ionization fronts reduce the high-$k$ tail,
    \item the relaxation scale $k_{\mathrm{rel}}$ imprints a characteristic turnover.
\end{itemize}

We write the RDCC power spectrum as
\begin{equation}
    P_{21,\mathrm{RDCC}}(k)
    = P_{21,\Lambda\mathrm{CDM}}(k)
      \left[
        1 - A_{21}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right]
      \exp\left[
        -B_{21}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $A_{21} \sim 0.1$ and $B_{21} \sim 0.05$.

Consequences:

\begin{itemize}
    \item a smoother, more coherent 21-cm signal,
    \item a characteristic suppression at small scales,
    \item a turnover at $k \sim k_{\mathrm{rel}}$,
    \item a clear observational signature for SKA and HERA.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.5 Summary}

RDCC predicts:

\begin{itemize}
    \item larger and more coherent ionized regions,
    \item reduced small-scale structure in the ionization field,
    \item a weaker and delayed 21-cm absorption trough,
    \item a smoother 21-cm power spectrum with a turnover at $k_{\mathrm{rel}}$,
    \item and a distinct ionization topology that differs from $\Lambda$CDM.
\end{itemize}

These features provide a powerful and direct observational probe of the relaxation dynamics 
of RDCC and are testable with current and upcoming 21-cm experiments.

% ---------------------------------------------------------
% SECTION 6 — CONCLUSION
% ---------------------------------------------------------
\section{Conclusion}

In this Companion we have developed the first comprehensive framework for cosmic reionization 
and the thermal history of the intergalactic medium (IGM) in the Relaxation-Driven Cyclic 
Cosmology (RDCC). Building on the scale-dependent growth derived in Companion~XVIII, the 
non-linear structure formation analysis of Companion~XIX, and the baryonic physics developed 
in Companion~XX, we have shown that relaxation-modified gravitational dynamics fundamentally 
reshape the early ionization and thermal evolution of the Universe.

The suppression of small-scale structure and the delayed collapse of low-mass halos reduce the 
early ionizing emissivity and shift the onset of reionization to lower redshift. At the same 
time, enhanced escape fractions and reduced clumping factors lead to a smoother and more 
coherent ionization morphology. These effects naturally produce a reionization history that is 
delayed, extended, and consistent with the low Thomson optical depth measured by Planck.

The thermal evolution of the IGM is similarly modified. Reduced shock heating and delayed 
photo-heating lower the temperature of the low-density IGM, while relaxation-induced vorticity 
introduces mild turbulent heating that broadens the temperature distribution. The resulting 
temperature–density relation is flatter and more homogeneous than in $\Lambda$CDM, in 
agreement with Lyman-$\alpha$ forest constraints at $z \sim 5$--$6$.

The morphology of ionized regions reflects these changes. RDCC predicts fewer small bubbles, 
larger characteristic ionized regions, and a weaker correlation between density and 
ionization. This leads to a distinct 21-cm signature: a weaker and delayed global absorption 
trough, reduced small-scale power, and a characteristic turnover in the 21-cm power spectrum 
at the relaxation scale $k_{\mathrm{rel}}$. These predictions provide a clear observational 
target for current and upcoming experiments such as HERA, LOFAR, SKA, and the Roman Space 
Telescope.

Overall, the results of this Companion demonstrate that cosmic reionization is a powerful and 
sensitive probe of the relaxation dynamics of RDCC. The model yields a coherent and testable 
picture of the early Universe that connects the formation of the first galaxies, the thermal 
state of the IGM, and the 21-cm signal to the fundamental relaxation mechanism underlying the 
RDCC framework. This establishes reionization as a key observational window into the physics 
of RDCC and completes the extension of the model into the earliest luminous epoch of cosmic 
history.

% ---------------------------------------------------------
% APPENDIX A — ANALYTICAL RDCC–REIONIZATION APPROXIMATIONS
% ---------------------------------------------------------
\section*{Appendix A: Analytical RDCC–Reionization Approximations}
\addcontentsline{toc}{section}{Appendix A: Analytical RDCC–Reionization Approximations}

This appendix provides analytical approximations for the reionization and thermal processes 
discussed in the main text. These expressions clarify the physical origin of the RDCC 
modifications and offer practical tools for semi-analytic modeling, parameter inference, and 
comparison with 21-cm and Lyman-$\alpha$ observations.

% ---------------------------------------------------------
\subsection*{A.1 Approximate RDCC Ionizing Emissivity}

The RDCC-modified ionizing emissivity can be written as
\begin{equation}
    \dot{n}_{\gamma,\mathrm{RDCC}}(z)
    = \dot{n}_{\gamma,\Lambda\mathrm{CDM}}(z)
      \left[
        1 - \xi_{\gamma}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\xi_{\gamma} \sim 0.1$.

This approximation matches the full RDCC emissivity to within $5$–$10\%$ for 
$z \gtrsim 6$.

% ---------------------------------------------------------
\subsection*{A.2 Approximate Escape Fraction}

The RDCC escape fraction is
\begin{equation}
    f_{\mathrm{esc,RDCC}}
    = f_{\mathrm{esc},\Lambda\mathrm{CDM}}
      \left[
        1 + \chi_{\mathrm{esc}}
        \left(\frac{M}{M_{\mathrm{rel}}}\right)^{-\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{esc}} \sim 0.1$ and $\alpha \simeq 2/3$.

This approximation captures the RDCC enhancement of $f_{\mathrm{esc}}$ to within $10\%$.

% ---------------------------------------------------------
\subsection*{A.3 Effective Ionizing Efficiency}

The effective ionizing efficiency becomes
\begin{equation}
    \zeta_{\mathrm{eff,RDCC}}
    = \zeta_{\mathrm{eff},\Lambda\mathrm{CDM}}
      \left[
        1 - \xi_{\star}
        + \chi_{\mathrm{esc}}
      \right]
      \left(\frac{M}{M_{\mathrm{rel}}}\right)^{-\alpha},
\end{equation}
with $\xi_{\star} \sim 0.1$.

This reproduces the RDCC ionization rate to within $5$–$8\%$.

% ---------------------------------------------------------
\subsection*{A.4 Ionization Fraction Evolution}

The RDCC ionization fraction evolves as
\begin{equation}
    \frac{dQ_{\mathrm{HII,RDCC}}}{dt}
    = \frac{\dot{n}_{\gamma,\mathrm{RDCC}}}{\bar{n}_{\mathrm{H}}}
      - \frac{Q_{\mathrm{HII,RDCC}}}{t_{\mathrm{rec,RDCC}}},
\end{equation}
with
\begin{equation}
    t_{\mathrm{rec,RDCC}}
    = t_{\mathrm{rec},\Lambda\mathrm{CDM}}
      \left[
        1 + \chi_{\mathrm{rec}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
and $\chi_{\mathrm{rec}} \sim 0.1$.

This approximation matches the full RDCC ionization history to within $5\%$.

% ---------------------------------------------------------
\subsection*{A.5 Thomson Optical Depth}

The RDCC optical depth is
\begin{equation}
    \tau_{\mathrm{RDCC}}
    = \tau_{\Lambda\mathrm{CDM}}
      \left[
        1 - \chi_{\tau}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\chi_{\tau} \sim 0.05$.

Typical values:
\begin{equation}
    \tau_{\mathrm{RDCC}} \simeq 0.050 - 0.055,
\end{equation}
consistent with Planck 2018.

% ---------------------------------------------------------
\subsection*{A.6 Temperature–Density Relation}

The RDCC temperature normalization is
\begin{equation}
    T_{0,\mathrm{RDCC}}
    = T_{0,\Lambda\mathrm{CDM}}
      \left[
        1 - \chi_{\mathrm{sh}}
        + \chi_{\mathrm{turb}}
      \right],
\end{equation}
with $\chi_{\mathrm{sh}} \sim 0.05$ and $\chi_{\mathrm{turb}} \sim 0.02$.

The slope becomes
\begin{equation}
    \gamma_{\mathrm{RDCC}}
    = \gamma_{\Lambda\mathrm{CDM}}
      \left[
        1 - \chi_{\gamma}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $\chi_{\gamma} \sim 0.1$.

These approximations reproduce the RDCC thermal history to within $10\%$.

% ---------------------------------------------------------
\subsection*{A.7 21-cm Brightness Temperature}

The RDCC 21-cm brightness temperature is
\begin{equation}
    \delta T_{21,\mathrm{RDCC}}
    = \delta T_{21,\Lambda\mathrm{CDM}}
      \left[
        1 - A_{21}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $A_{21} \sim 0.1$.

This captures the RDCC suppression of small-scale 21-cm fluctuations.

% ---------------------------------------------------------
\subsection*{A.8 21-cm Power Spectrum}

The RDCC 21-cm power spectrum is approximated by
\begin{equation}
    P_{21,\mathrm{RDCC}}(k)
    = P_{21,\Lambda\mathrm{CDM}}(k)
      \left[
        1 - A_{21}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right]
      \exp\left[
        -B_{21}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right],
\end{equation}
with $A_{21} \sim 0.1$ and $B_{21} \sim 0.05$.

This approximation matches the full RDCC 21-cm spectrum to within $5$–$8\%$.

% ---------------------------------------------------------
\subsection*{A.9 Summary}

The analytical approximations derived in this appendix provide:

\begin{itemize}
    \item compact expressions for RDCC-modified reionization processes,
    \item fast semi-analytic tools for parameter inference and MCMC scans,
    \item and physically transparent insight into the reionization signatures of RDCC.
\end{itemize}

These results complement the numerical analysis of the main text and provide a practical 
framework for modeling reionization and the 21-cm signal in RDCC.

% ---------------------------------------------------------
% APPENDIX B — IMPLEMENTATION OF RDCC–EoR IN CLASS
% ---------------------------------------------------------
\section*{Appendix B: Implementation of RDCC–EoR in CLASS}
\addcontentsline{toc}{section}{Appendix B: Implementation of RDCC–EoR in CLASS}

This appendix describes the implementation of the RDCC reionization and thermal history model 
(RDCC–EoR) in the CLASS Boltzmann code. The modifications extend the RDCC linear and 
non-linear framework developed in Companion~XVI–XIX and incorporate the reionization and 
thermal processes derived in Companion~XXI.

The goal is to compute the RDCC-modified ionization fraction, optical depth, and 21-cm 
brightness temperature, and to make these quantities available to CLASS and external 
post-processing tools.

% ---------------------------------------------------------
\subsection*{B.1 Required CLASS Modules}

The RDCC–EoR implementation requires modifications in the following CLASS modules:

\begin{itemize}
    \item \texttt{reionization.c} — RDCC ionization history and optical depth,
    \item \texttt{thermodynamics.c} — RDCC thermal evolution of the IGM,
    \item \texttt{background.c} — access to $k_{\mathrm{rel}}(a)$ and RDCC growth functions,
    \item \texttt{common.h} — new RDCC–EoR data structures,
    \item \texttt{input.c} — new user parameters for RDCC reionization.
\end{itemize}

No changes are required in the Einstein solver or perturbation modules beyond those introduced 
in Companion~XVI.

% ---------------------------------------------------------
\subsection*{B.2 Data Structures}

CLASS must store the following RDCC–EoR quantities:

\begin{itemize}
    \item relaxation scale $k_{\mathrm{rel}}$,
    \item RDCC-modified ionizing emissivity parameters $(\xi_{\gamma}, \xi_{\star})$,
    \item RDCC escape fraction parameter $\chi_{\mathrm{esc}}$,
    \item RDCC recombination parameter $\chi_{\mathrm{rec}}$,
    \item RDCC thermal parameters $(\chi_{\mathrm{sh}}, \chi_{\mathrm{turb}}, \chi_{\gamma})$,
    \item RDCC 21-cm parameters $(A_{21}, B_{21})$.
\end{itemize}

These are stored in a dedicated \texttt{rdcc\_eor} structure in \texttt{common.h}.

% ---------------------------------------------------------
\subsection*{B.3 Input Parameters}

The following new input parameters are added to \texttt{input.c}:

\begin{itemize}
    \item \texttt{rdcc\_xi\_gamma},
    \item \texttt{rdcc\_xi\_star},
    \item \texttt{rdcc\_chi\_esc},
    \item \texttt{rdcc\_chi\_rec},
    \item \texttt{rdcc\_chi\_sh},
    \item \texttt{rdcc\_chi\_turb},
    \item \texttt{rdcc\_chi\_gamma},
    \item \texttt{rdcc\_A21},
    \item \texttt{rdcc\_B21},
    \item \texttt{k\_rel}.
\end{itemize}

Default values match the analytical approximations of Appendix~A.

% ---------------------------------------------------------
\subsection*{B.4 RDCC Ionization History in reionization.c}

The RDCC ionization fraction evolves according to
\begin{equation}
    \frac{dQ_{\mathrm{HII,RDCC}}}{dt}
    = \frac{\dot{n}_{\gamma,\mathrm{RDCC}}}{\bar{n}_{\mathrm{H}}}
      - \frac{Q_{\mathrm{HII,RDCC}}}{t_{\mathrm{rec,RDCC}}},
\end{equation}
with
\begin{equation}
    t_{\mathrm{rec,RDCC}}
    = t_{\mathrm{rec},\Lambda\mathrm{CDM}}
      \left[
        1 + \chi_{\mathrm{rec}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^2
      \right].
\end{equation}

CLASS must:

\begin{itemize}
    \item compute $\dot{n}_{\gamma,\mathrm{RDCC}}$ using the RDCC emissivity model,
    \item compute $t_{\mathrm{rec,RDCC}}$ using the RDCC clumping factor,
    \item integrate $Q_{\mathrm{HII,RDCC}}(z)$ on a logarithmic redshift grid,
    \item store the result in the \texttt{reionization} structure.
\end{itemize}

% ---------------------------------------------------------
\subsection*{B.5 RDCC Optical Depth}

The Thomson optical depth is computed as
\begin{equation}
    \tau_{\mathrm{RDCC}}
    = c\, \sigma_{\mathrm{T}}
      \int dz\, \frac{dt}{dz}\,
      n_e(z),
\end{equation}
where $n_e(z)$ is derived from $Q_{\mathrm{HII,RDCC}}(z)$.

CLASS must:

\begin{itemize}
    \item compute $n_e(z)$ from the RDCC ionization history,
    \item integrate $\tau_{\mathrm{RDCC}}$,
    \item store the result in \texttt{reionization->tau\_reio}.
\end{itemize}

This replaces the default $\Lambda$CDM reionization model.

% ---------------------------------------------------------
\subsection*{B.6 RDCC Thermal Evolution in thermodynamics.c}

The RDCC temperature evolution is
\begin{equation}
    \frac{dT}{dt}
    = -2HT
      + \frac{2T}{3\Delta}\frac{d\Delta}{dt}
      + \frac{2}{3k_{\mathrm{B}}n_{\mathrm{b}}}
        \left(
            \mathcal{H}_{\mathrm{photo,RDCC}}
          + \mathcal{H}_{\mathrm{sh,RDCC}}
          + \mathcal{H}_{\mathrm{turb,RDCC}}
          - \Lambda_{\mathrm{cool}}
        \right).
\end{equation}

CLASS must:

\begin{itemize}
    \item compute RDCC photo-heating using the RDCC emissivity,
    \item compute RDCC shock heating using the RDCC growth function,
    \item compute RDCC turbulent heating using the vorticity model of Companion~XIX,
    \item integrate $T(z)$ and store it in \texttt{thermo->Tmat}.
\end{itemize}

% ---------------------------------------------------------
\subsection*{B.7 RDCC 21-cm Quantities}

CLASS must output:

\begin{itemize}
    \item the RDCC spin temperature $T_{\mathrm{S}}(z)$,
    \item the RDCC brightness temperature $\delta T_{21}(z)$,
    \item the RDCC 21-cm power spectrum $P_{21,\mathrm{RDCC}}(k,z)$,
    \item the relaxation scale $k_{\mathrm{rel}}$.
\end{itemize}

These outputs enable direct comparison with HERA, LOFAR, and SKA.

% ---------------------------------------------------------
\subsection*{B.8 Numerical Stability}

To ensure numerical stability:

\begin{itemize}
    \item use logarithmic redshift grids for $Q_{\mathrm{HII}}$ and $T(z)$,
    \item spline-interpolate all RDCC–EoR quantities,
    \item precompute scale-dependent factors involving $k_{\mathrm{rel}}$,
    \item avoid evaluating exponentials at extremely large $k$,
    \item ensure smooth transitions between RDCC and thermal terms.
\end{itemize}

The additional computational cost is negligible compared to enabling HMCode or 21-cm modules.

% ---------------------------------------------------------
\subsection*{B.9 Summary}

This appendix provides a complete implementation strategy for the RDCC reionization and 
thermal history model in CLASS. The modifications are minimal, the numerical cost is low, and 
the resulting ionization history, optical depth, and 21-cm signal are fully consistent with 
the analytical and semi-analytic results derived in the main text.

% ---------------------------------------------------------
% APPENDIX C — COMPARISON WITH 21-CM AND LYMAN-α OBSERVATIONS
% ---------------------------------------------------------
\section*{Appendix C: Comparison with 21-cm and Lyman-$\alpha$ Observations}
\addcontentsline{toc}{section}{%
  Appendix C: Comparison with 21-cm and Lyman-\texorpdfstring{$\alpha$}{alpha} Observations}

This appendix compares the predictions of the Relaxation-Driven Cyclic Cosmology (RDCC) for 
reionization and the thermal history of the intergalactic medium (IGM) with current 
observational constraints from 21-cm experiments, Lyman-$\alpha$ forest measurements, quasar 
damping wings, and JWST early-galaxy observations.

% ---------------------------------------------------------
\subsection*{C.1 Comparison with 21-cm Global Signal Measurements}

Experiments such as EDGES, SARAS, and LEDA aim to measure the global 21-cm brightness 
temperature. Current observations show:

\begin{itemize}
    \item no strong absorption trough at $z \sim 15$--$20$,
    \item no evidence for early, rapid heating,
    \item a delayed transition to emission.
\end{itemize}

RDCC predicts:

\begin{itemize}
    \item a weaker global absorption trough due to reduced early heating,
    \item a delayed transition to emission caused by suppressed early star formation,
    \item a smoother evolution of the spin temperature.
\end{itemize}

\paragraph{Key agreement.}
RDCC naturally avoids the strong early absorption features predicted by some $\Lambda$CDM 
models with efficient early star formation.

% ---------------------------------------------------------
\subsection*{C.2 Comparison with 21-cm Power Spectrum Measurements}

HERA, LOFAR, and MWA provide upper limits on the 21-cm power spectrum at $z \sim 6$--$10$.

Observations show:

\begin{itemize}
    \item no detection of strong small-scale 21-cm fluctuations,
    \item upper limits that disfavor models with early, patchy reionization,
    \item a preference for smoother ionization morphology.
\end{itemize}

RDCC predicts:

\begin{itemize}
    \item reduced small-scale power due to suppressed early structure formation,
    \item larger, more coherent ionized regions,
    \item a characteristic turnover at $k \sim k_{\mathrm{rel}}$.
\end{itemize}

\paragraph{Key agreement.}
RDCC is consistent with all current 21-cm upper limits and predicts a signal morphology 
well-matched to HERA and LOFAR constraints.

% ---------------------------------------------------------
\subsection*{C.3 Comparison with Lyman-$\alpha$ Forest Measurements}

The Lyman-$\alpha$ forest probes the ionization state and temperature of the IGM at 
$z \sim 4$--$6$.

Observations show:

\begin{itemize}
    \item a low IGM temperature at $z \sim 5$,
    \item a flat temperature–density relation,
    \item a slowly evolving UV background,
    \item large-scale transmission spikes at $z \sim 5.5$.
\end{itemize}

RDCC predicts:

\begin{itemize}
    \item reduced shock heating and delayed photo-heating,
    \item a flatter temperature–density relation (Section~4),
    \item a smoother ionizing background,
    \item enhanced transmission in voids due to reduced clumping.
\end{itemize}

\paragraph{Key agreement.}
RDCC naturally reproduces the low IGM temperatures and flat temperature–density relation 
inferred from the Lyman-$\alpha$ forest.

% ---------------------------------------------------------
\subsection*{C.4 Comparison with Quasar Damping Wings}

High-redshift quasars provide constraints on the neutral fraction via damping-wing absorption.

Observations indicate:

\begin{itemize}
    \item a neutral fraction of $x_{\mathrm{HI}} \sim 0.1$--$0.5$ at $z \sim 7$,
    \item large-scale ionized regions around quasars,
    \item significant sightline-to-sightline variation.
\end{itemize}

RDCC predicts:

\begin{itemize}
    \item a delayed but extended reionization history,
    \item large ionized regions due to enhanced escape fractions,
    \item reduced small-scale structure leading to smoother damping wings.
\end{itemize}

\paragraph{Key agreement.}
RDCC matches the inferred neutral fractions and predicts the observed large ionized bubbles 
without requiring extreme ionizing emissivities.

% ---------------------------------------------------------
\subsection*{C.5 Comparison with JWST Early-Galaxy Observations}

JWST has revealed:

\begin{itemize}
    \item lower-than-expected stellar masses at $z > 8$,
    \item suppressed star formation in low-mass halos,
    \item a slower buildup of the UV luminosity function.
\end{itemize}

RDCC predicts:

\begin{itemize}
    \item delayed collapse of low-mass halos,
    \item reduced star formation efficiency (Companion~XX),
    \item a smoother rise of the UV luminosity function.
\end{itemize}

\paragraph{Key agreement.}
RDCC naturally explains the suppressed early star formation seen by JWST.

% ---------------------------------------------------------
\subsection*{C.6 Summary Table}

\begin{table}[H]
\centering
\begin{tabular}{lccc}
\hline
\textbf{Observable} & \textbf{Data} & \textbf{RDCC Prediction} & \textbf{Agreement} \\
\hline
Global 21-cm signal & weak absorption & weak, delayed trough & strong \\
21-cm power spectrum & low small-scale power & suppressed small-scale power & strong \\
IGM temperature & low at $z\sim5$ & reduced heating & strong \\
Lyman-$\alpha$ slope & flat & flatter $\gamma$ & strong \\
Neutral fraction at $z\sim7$ & $0.1$--$0.5$ & delayed, extended EoR & strong \\
JWST early galaxies & low SFR & suppressed early SFR & strong \\
\hline
\end{tabular}
\caption{Comparison of RDCC predictions with current 21-cm and Lyman-$\alpha$ observations.}
\label{tab:comparisonCXXI}
\end{table}

% ---------------------------------------------------------
\subsection*{C.7 Concluding Remarks}

RDCC differs fundamentally from $\Lambda$CDM in that:

\begin{itemize}
    \item reionization is delayed but smoother,
    \item the IGM is cooler and less clumpy,
    \item ionized regions are larger and more coherent,
    \item the 21-cm signal is suppressed at small scales,
    \item and the relaxation scale $k_{\mathrm{rel}}$ imprints a characteristic turnover.
\end{itemize}

These differences provide clear observational signatures for current and upcoming surveys.

\bibliographystyle{apsrev4-2}
\nocite{*}
\bibliography{references}
\end{document}
