% ---------------------------------------------------------
% PREAMBLE (same template as Companion X–XXIV)
% ---------------------------------------------------------
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{physics}
\usepackage{cite}
\usepackage{float}

\geometry{margin=1in}

% ---------------------------------------------------------
% TITLE
% ---------------------------------------------------------
\title{
\textbf{Companion XXV}\\[4mm]
\textbf{The 21-cm Signature of the Relaxation-Driven Cyclic Cosmology}\\
\textbf{Reionization Topology, X-ray Heating, and Large-Scale Structure}
}

\author{
Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}
}

\date{April 2026}

\begin{document}

\maketitle

% ---------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------
\begin{abstract}
The 21-cm signal from neutral hydrogen provides a uniquely sensitive probe of the thermal, 
ionization, and structural evolution of the early Universe. Upcoming observations from SKA, 
HERA, and LOFAR will map the morphology of reionization, the timing of X-ray heating, and the 
large-scale distribution of ionized and neutral regions. These measurements offer a powerful 
test of the Relaxation-Driven Cyclic Cosmology (RDCC), whose modified small-scale structure, 
delayed fragmentation, and enhanced early black hole and quasar activity fundamentally alter 
the topology of reionization.

Building on the early-galaxy physics of Companion~XXII, the black hole formation framework of 
Companion~XXIII, and the quasar radiative feedback model of Companion~XXIV, we derive the 
RDCC predictions for the global 21-cm signal, the 21-cm power spectrum, and the morphology of 
ionization bubbles. We show that RDCC produces smoother, more coherent ionization structures, 
delayed but stronger X-ray heating, and a characteristic suppression of small-scale 21-cm 
power due to the relaxation-modified halo mass function.

This Companion establishes the 21-cm signal as a direct observational probe of the relaxation 
scale $k_{\mathrm{rel}}$ and provides clear predictions for SKA, HERA, LOFAR, JWST, and 
cross-correlation studies with early galaxies and quasars.
\end{abstract}

% ---------------------------------------------------------
% SECTION 1 — MOTIVATION AND OVERVIEW
% ---------------------------------------------------------
\section{Motivation and Overview}

The 21-cm signal from neutral hydrogen provides a uniquely sensitive probe of the thermal, 
ionization, and structural evolution of the early Universe. Unlike galaxy surveys or quasar 
observations, which trace only the brightest objects, the 21-cm line maps the entire 
intergalactic medium (IGM), including faint galaxies, diffuse gas, and the topology of 
reionization. Upcoming observations from SKA, HERA, and LOFAR will measure the global 21-cm 
signal, the 21-cm power spectrum, and the morphology of ionized and neutral regions across 
cosmic time.

These measurements offer a powerful test of the Relaxation-Driven Cyclic Cosmology (RDCC). 
The suppression of small-scale power, delayed fragmentation, and enhanced early black hole and 
quasar activity predicted by RDCC fundamentally alter the timing, morphology, and coherence of 
reionization. As shown in Companion~XXII, early star formation is delayed and smoother; in 
Companion~XXIII, black hole seeds are more massive and grow more efficiently; and in 
Companion~XXIV, early quasars produce larger and more coherent ionization bubbles.

The 21-cm signal is therefore a direct tracer of the relaxation scale $k_{\mathrm{rel}}$ and 
provides a unique opportunity to test the RDCC framework.

% ---------------------------------------------------------
\subsection*{1.1 Observational Context}

Current and upcoming experiments probe different aspects of the 21-cm signal:

\begin{itemize}
    \item \textbf{HERA} measures the 21-cm power spectrum at $z \sim 6$–$20$.
    \item \textbf{LOFAR} constrains large-scale ionization structure at $z \sim 8$–$10$.
    \item \textbf{SKA} will map the full 3D morphology of reionization.
    \item \textbf{JWST} provides complementary constraints on early galaxies and quasars.
    \item \textbf{Cross-correlations} (21-cm × galaxies × quasars) will probe bubble topology.
\end{itemize}

These observations are sensitive to:

\begin{itemize}
    \item the timing of X-ray heating,
    \item the growth of ionized bubbles,
    \item the clustering of early sources,
    \item the small-scale structure of the IGM,
    \item the global reionization history.
\end{itemize}

Each of these quantities is modified in RDCC.

% ---------------------------------------------------------
\subsection*{1.2 RDCC Perspective}

The Relaxation-Driven Cyclic Cosmology modifies early structure formation through the 
scale-dependent relaxation rate $\Gamma(a)$, which suppresses small-scale power and delays the 
collapse of low-mass halos. As shown in Companion~XVIII and Companion~XIX, this leads to:

\begin{itemize}
    \item reduced small-scale clumping,
    \item smoother density fields,
    \item delayed but more coherent star formation,
    \item enhanced early black hole and quasar activity,
    \item larger and smoother ionization bubbles.
\end{itemize}

These effects directly influence the 21-cm signal:

\begin{itemize}
    \item \textbf{Global signal:} delayed absorption trough, smoother heating transition.
    \item \textbf{Power spectrum:} suppressed small-scale power, enhanced large-scale modes.
    \item \textbf{Bubble topology:} larger, more coherent ionized regions.
    \item \textbf{Cross-correlations:} stronger 21-cm × quasar anticorrelation.
\end{itemize}

Thus, RDCC predicts a distinctive 21-cm signature that differs qualitatively from $\Lambda$CDM.

% ---------------------------------------------------------
\subsection*{1.3 Goals of This Companion}

The goals of this Companion are:

\begin{itemize}
    \item to derive the RDCC predictions for the global 21-cm signal,
    \item to compute the RDCC-modified 21-cm power spectrum,
    \item to model the morphology of ionization bubbles in RDCC,
    \item to analyze the impact of early quasars on X-ray heating and reionization,
    \item and to compare RDCC predictions with SKA, HERA, LOFAR, and JWST.
\end{itemize}

These results build directly on the early-galaxy, black hole, and quasar physics developed in 
Companion~XXII–XXIV.

% ---------------------------------------------------------
\subsection*{1.4 Structure of This Companion}

The structure of this Companion is as follows:

\begin{itemize}
    \item Section~2 derives the RDCC global 21-cm signal.
    \item Section~3 computes the RDCC 21-cm power spectrum.
    \item Section~4 analyzes the morphology of ionization bubbles.
    \item Section~5 develops the RDCC X-ray heating model.
    \item Section~6 compares RDCC predictions with SKA, HERA, LOFAR, and JWST.
\end{itemize}

Together, these results establish the 21-cm signal as a direct observational probe of the 
relaxation dynamics of RDCC.

% ---------------------------------------------------------
% SECTION 2 — THE RDCC GLOBAL 21-cm SIGNAL
% ---------------------------------------------------------
\section{The RDCC Global 21-cm Signal}

The global 21-cm signal traces the contrast between the spin temperature $T_{\mathrm{s}}$ of 
neutral hydrogen and the CMB temperature $T_{\gamma}$. It is sensitive to the thermal and 
ionization history of the intergalactic medium (IGM), the timing of X-ray heating, and the 
formation of the first stars, black holes, and quasars. In the Relaxation-Driven Cyclic 
Cosmology (RDCC), the suppression of small-scale structure and the delayed collapse of 
low-mass halos modify each of these processes, leading to a distinctive global 21-cm signature.

The differential brightness temperature is
\begin{equation}
    \delta T_{21}(z)
    \simeq 27\,x_{\mathrm{HI}}(z)
    \left(1 - \frac{T_{\gamma}(z)}{T_{\mathrm{s}}(z)}\right)
    \left(\frac{1+z}{10}\right)^{1/2}
    \mathrm{mK},
\end{equation}
where $x_{\mathrm{HI}}$ is the neutral fraction.

RDCC modifies $\delta T_{21}(z)$ through:

\begin{itemize}
    \item delayed star formation (Companion~XXII),
    \item enhanced early black hole and quasar activity (Companion~XXIII–XXIV),
    \item smoother density fields and reduced clumping (Companion~XVIII–XIX),
    \item modified X-ray heating and ionization rates.
\end{itemize}

These effects reshape the three canonical phases of the global 21-cm signal.

% ---------------------------------------------------------
\subsection*{2.1 Phase I: The Absorption Trough}

The absorption trough occurs when $T_{\mathrm{s}} < T_{\gamma}$ and the IGM is cold. In 
$\Lambda$CDM, this phase is driven by Lyman-$\alpha$ coupling from the first stars.

In RDCC:

\begin{itemize}
    \item delayed fragmentation postpones the onset of Lyman-$\alpha$ coupling,
    \item smoother halo assembly reduces early star formation rates,
    \item the IGM remains colder for longer due to reduced X-ray heating.
\end{itemize}

Thus, the absorption trough is:

\begin{itemize}
    \item \textbf{delayed} relative to $\Lambda$CDM,
    \item \textbf{shallower} due to reduced early coupling,
    \item \textbf{broader} due to smoother early structure formation.
\end{itemize}

A useful RDCC approximation is
\begin{equation}
    z_{\mathrm{abs,RDCC}}
    \simeq z_{\mathrm{abs}}
    \left[
        1 - \chi_{\mathrm{abs}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
    \right],
\end{equation}
with $\chi_{\mathrm{abs}} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{2.2 Phase II: X-ray Heating}

The heating transition occurs when $T_{\mathrm{s}}$ rises above $T_{\gamma}$ due to X-ray 
heating from early black holes, X-ray binaries, and quasars.

In RDCC:

\begin{itemize}
    \item early black hole seeds are more massive (Companion~XXIII),
    \item quasars ignite earlier and are more luminous (Companion~XXIV),
    \item reduced clumping increases the mean free path of X-rays,
    \item smoother density fields reduce shadowing.
\end{itemize}

Thus, X-ray heating is:

\begin{itemize}
    \item \textbf{delayed} relative to $\Lambda$CDM due to delayed star formation,
    \item \textbf{stronger} once it begins due to enhanced quasar activity,
    \item \textbf{smoother} due to reduced small-scale structure.
\end{itemize}

The heating rate becomes
\begin{equation}
    \Gamma_{\mathrm{X,RDCC}}
    = \Gamma_{\mathrm{X}}
      \left[
        1 + \chi_{\mathrm{X}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{X}} \sim 0.2$.

% ---------------------------------------------------------
\subsection*{2.3 Phase III: The Emission Era}

Once $T_{\mathrm{s}} \gg T_{\gamma}$, the 21-cm signal enters emission. The amplitude is 
determined by the neutral fraction $x_{\mathrm{HI}}$ and the thermal state of the IGM.

In RDCC:

\begin{itemize}
    \item larger ionization bubbles reduce $x_{\mathrm{HI}}$ earlier,
    \item smoother ionization fronts produce more coherent emission,
    \item enhanced quasar activity accelerates late-time heating.
\end{itemize}

The emission amplitude becomes
\begin{equation}
    \delta T_{21,\mathrm{RDCC}}
    \simeq \delta T_{21}
    \left[
        1 - \chi_{\mathrm{ion}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
    \right],
\end{equation}
with $\chi_{\mathrm{ion}} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{2.4 RDCC Global Signal Summary}

RDCC predicts:

\begin{itemize}
    \item a \textbf{delayed and broadened absorption trough},
    \item a \textbf{delayed but stronger heating transition},
    \item a \textbf{smoother and more coherent emission era},
    \item a \textbf{characteristic suppression of small-scale features},
    \item a \textbf{direct imprint of the relaxation scale $k_{\mathrm{rel}}$}.
\end{itemize}

These features provide clear observational signatures for SKA, HERA, and LOFAR and set the 
stage for the RDCC 21-cm power spectrum developed in Section~3.

% ---------------------------------------------------------
% SECTION 3 — THE RDCC 21-cm POWER SPECTRUM
% ---------------------------------------------------------
\section{The RDCC 21-cm Power Spectrum}

The 21-cm power spectrum $P_{21}(k,z)$ encodes the spatial fluctuations of the neutral 
hydrogen distribution during cosmic dawn and reionization. It is sensitive to the underlying 
matter power spectrum, the timing of X-ray heating, the growth of ionized bubbles, and the 
clustering of early galaxies and quasars. In the Relaxation-Driven Cyclic Cosmology (RDCC), 
the suppression of small-scale power and the smoother assembly of early structures modify each 
of these components, leading to a distinctive 21-cm power spectrum.

The brightness temperature fluctuation is
\begin{equation}
    \delta T_{21}(\mathbf{x},z)
    = \bar{T}_{21}(z)\,
      \left[
        \delta_{\mathrm{b}}
        - \delta_{x_{\mathrm{HI}}}
        + \delta_{\mathrm{vel}}
        + \delta_{T_{\mathrm{s}}}
      \right],
\end{equation}
where $\delta_{\mathrm{b}}$ is the baryon overdensity, $\delta_{x_{\mathrm{HI}}}$ the neutral 
fraction fluctuation, $\delta_{\mathrm{vel}}$ the velocity gradient term, and 
$\delta_{T_{\mathrm{s}}}$ the spin-temperature fluctuation.

RDCC modifies each of these contributions.

% ---------------------------------------------------------
\subsection*{3.1 RDCC Matter Power Spectrum Contribution}

The baryonic term traces the matter power spectrum:
\begin{equation}
    P_{21}^{\mathrm{b}}(k,z)
    = \bar{T}_{21}^{2}(z)\, P_{\mathrm{m}}(k,z).
\end{equation}

In RDCC, the matter power spectrum is suppressed at small scales:
\begin{equation}
    P_{\mathrm{m,RDCC}}(k,z)
    = P_{\mathrm{m}}(k,z)
      \left[
        1 - \chi_{\mathrm{rel}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{rel}} \sim 0.1$–$0.2$.

Thus:

\begin{itemize}
    \item small-scale 21-cm power is suppressed,
    \item large-scale modes are enhanced due to increased bias,
    \item the turnover scale traces $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.2 Ionization Field Contribution}

Ionization fluctuations dominate the 21-cm signal during reionization:
\begin{equation}
    P_{21}^{x}(k,z)
    = \bar{T}_{21}^{2}(z)\, P_{x_{\mathrm{HI}}}(k,z).
\end{equation}

In RDCC:

\begin{itemize}
    \item ionization bubbles are larger (Companion~XXIV),
    \item bubble walls are smoother due to reduced clumping,
    \item bubble growth is more coherent due to smoother source clustering.
\end{itemize}

A useful RDCC approximation is
\begin{equation}
    P_{x_{\mathrm{HI}},\mathrm{RDCC}}(k,z)
    = P_{x_{\mathrm{HI}}}(k,z)
      \exp\!\left[
        -\left(\frac{k}{k_{\mathrm{bub}}}\right)^{2}
      \right],
\end{equation}
with
\begin{equation}
    k_{\mathrm{bub}}^{-1}
    \sim R_{\mathrm{ion,RDCC}}
    \sim 1.5\text{--}2 \times R_{\mathrm{ion}}.
\end{equation}

Thus, RDCC predicts:

\begin{itemize}
    \item suppressed small-scale ionization power,
    \item enhanced large-scale bubble correlations,
    \item a characteristic shift of the bubble peak to lower $k$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.3 X-ray Heating Contribution}

During cosmic dawn, X-ray heating fluctuations dominate:
\begin{equation}
    P_{21}^{T}(k,z)
    = \bar{T}_{21}^{2}(z)\, P_{T_{\mathrm{s}}}(k,z).
\end{equation}

In RDCC:

\begin{itemize}
    \item X-ray heating is delayed (Section~2),
    \item heating is stronger once it begins (Companion~XXIV),
    \item heating is smoother due to reduced small-scale structure.
\end{itemize}

A simple RDCC model is
\begin{equation}
    P_{T_{\mathrm{s}},\mathrm{RDCC}}(k,z)
    = P_{T_{\mathrm{s}}}(k,z)
      \left[
        1 - \chi_{\mathrm{heat}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{heat}} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{3.4 Velocity Gradient Contribution}

The velocity gradient term is
\begin{equation}
    P_{21}^{\mathrm{vel}}(k,z)
    = \mu^{4}\, \bar{T}_{21}^{2}(z)\, P_{\mathrm{m}}(k,z),
\end{equation}
with $\mu = \hat{k}\cdot\hat{n}$.

In RDCC:

\begin{itemize}
    \item the velocity field is unchanged at large scales,
    \item small-scale velocity gradients are suppressed,
    \item redshift-space distortions are reduced at $k > k_{\mathrm{rel}}$.
\end{itemize}

Thus,
\begin{equation}
    P_{21,\mathrm{RDCC}}^{\mathrm{vel}}
    = P_{21}^{\mathrm{vel}}
      \left[
        1 - \chi_{\mathrm{vel}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{3.5 Full RDCC 21-cm Power Spectrum}

Combining all contributions:
\begin{equation}
    P_{21,\mathrm{RDCC}}(k,z)
    = P_{21}^{\mathrm{b}}
    + P_{21}^{x}
    + P_{21}^{T}
    + P_{21}^{\mathrm{vel}},
\end{equation}
with each term replaced by its RDCC-modified form.

The resulting spectrum exhibits:

\begin{itemize}
    \item suppressed small-scale power ($k > k_{\mathrm{rel}}$),
    \item enhanced large-scale modes due to increased bias,
    \item a shift of the bubble peak to lower $k$,
    \item smoother heating and ionization fluctuations,
    \item a characteristic turnover tracing $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.6 Summary}

RDCC modifies the 21-cm power spectrum through:

\begin{itemize}
    \item suppression of small-scale matter fluctuations,
    \item larger and smoother ionization bubbles,
    \item delayed but stronger X-ray heating,
    \item reduced redshift-space distortions,
    \item enhanced large-scale clustering of early sources.
\end{itemize}

These features provide clear observational signatures for SKA, HERA, and LOFAR and motivate 
the detailed bubble morphology analysis of Section~4.

% ---------------------------------------------------------
% SECTION 4 — IONIZATION BUBBLE MORPHOLOGY IN RDCC
% ---------------------------------------------------------
\section{Ionization Bubble Morphology in RDCC}

The morphology of ionized regions during reionization encodes the spatial distribution of 
ionizing sources, the clustering of early galaxies and quasars, and the small-scale structure 
of the intergalactic medium (IGM). In the Relaxation-Driven Cyclic Cosmology (RDCC), the 
suppression of small-scale power, delayed fragmentation, and enhanced early quasar activity 
fundamentally alter the topology of ionization bubbles. This section derives the RDCC 
predictions for bubble sizes, shapes, clustering, and percolation.

% ---------------------------------------------------------
\subsection*{4.1 Bubble Growth in RDCC}

Ionization bubbles grow according to
\begin{equation}
    R_{\mathrm{ion}}
    = \left(
        \frac{3 \dot{N}_{\gamma}}
             {4\pi \alpha_{B} n_{\mathrm{H}}^{2}}
      \right)^{1/3},
\end{equation}
where $\dot{N}_{\gamma}$ is the ionizing photon rate.

In RDCC:

\begin{itemize}
    \item quasars ignite earlier and are more luminous (Companion~XXIV),
    \item early galaxies form later but more coherently (Companion~XXII),
    \item reduced clumping increases the mean free path of ionizing photons,
    \item smoother density fields reduce shadowing and bubble fragmentation.
\end{itemize}

Thus, the RDCC bubble radius becomes
\begin{equation}
    R_{\mathrm{ion,RDCC}}
    = R_{\mathrm{ion}}
      \left[
        1 + \chi_{R}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{R} \sim 0.2$.

Typical RDCC values:
\begin{equation}
    R_{\mathrm{ion,RDCC}} \sim 1.5\text{--}2 \times R_{\mathrm{ion}}.
\end{equation}

% ---------------------------------------------------------
\subsection*{4.2 Bubble Size Distribution}

The bubble size distribution (BSD) is defined as
\begin{equation}
    \frac{dn}{dR}(R,z).
\end{equation}

In $\Lambda$CDM, the BSD is broad and dominated by small bubbles at early times.

In RDCC:

\begin{itemize}
    \item suppressed small-scale structure reduces the number of small bubbles,
    \item enhanced quasar activity increases the number of large bubbles,
    \item smoother source clustering narrows the BSD.
\end{itemize}

A useful RDCC approximation is
\begin{equation}
    \frac{dn}{dR}_{\mathrm{RDCC}}
    = \frac{dn}{dR}
      \exp\!\left[
        -\left(\frac{R_{\mathrm{min}}}{R}\right)^{\beta}
      \right],
\end{equation}
with $R_{\mathrm{min}} \sim 0.5\,R_{\mathrm{ion,RDCC}}$.

% ---------------------------------------------------------
\subsection*{4.3 Bubble Clustering and Bias}

The bubble bias is defined as
\begin{equation}
    b_{\mathrm{bub}}
    = \frac{\delta_{\mathrm{bub}}}{\delta_{\mathrm{m}}}.
\end{equation}

In RDCC:

\begin{itemize}
    \item galaxy bias is enhanced (Companion~XXII),
    \item quasar bias is strongly enhanced (Companion~XXIV),
    \item bubble clustering inherits both contributions.
\end{itemize}

Thus,
\begin{equation}
    b_{\mathrm{bub,RDCC}}
    = b_{\mathrm{bub}}
      \left[
        1 + \chi_{\mathrm{bias}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{bias}} \sim 0.1$–$0.2$.

This leads to:

\begin{itemize}
    \item stronger large-scale bubble correlations,
    \item enhanced 21-cm power at low $k$,
    \item earlier percolation of ionized regions.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.4 Bubble Shape and Topology}

Bubble shapes are characterized by Minkowski functionals:

\begin{itemize}
    \item $V_{0}$ — volume,
    \item $V_{1}$ — surface area,
    \item $V_{2}$ — mean curvature,
    \item $V_{3}$ — Euler characteristic.
\end{itemize}

In RDCC:

\begin{itemize}
    \item reduced small-scale structure smooths bubble surfaces,
    \item larger bubbles reduce curvature variations,
    \item smoother density fields reduce topological complexity.
\end{itemize}

Thus:

\begin{align}
    V_{1,\mathrm{RDCC}} &< V_{1}, \\
    V_{2,\mathrm{RDCC}} &< V_{2}, \\
    V_{3,\mathrm{RDCC}} &\rightarrow 0 \quad \text{earlier}.
\end{align}

This implies:

\begin{itemize}
    \item fewer small-scale features,
    \item smoother bubble boundaries,
    \item earlier transition to a percolated ionized network.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.5 Percolation in RDCC}

Percolation occurs when ionized regions connect to form a spanning cluster.

In RDCC:

\begin{itemize}
    \item larger bubbles accelerate percolation,
    \item enhanced bias increases bubble overlap,
    \item smoother density fields reduce fragmentation.
\end{itemize}

Thus, the percolation redshift becomes
\begin{equation}
    z_{\mathrm{perc,RDCC}}
    \simeq z_{\mathrm{perc}}
    \left[
        1 + \chi_{\mathrm{perc}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
    \right],
\end{equation}
with $\chi_{\mathrm{perc}} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{4.6 Summary}

RDCC modifies ionization bubble morphology through:

\begin{itemize}
    \item larger and smoother ionization bubbles,
    \item suppressed small-scale bubble population,
    \item enhanced bubble clustering and bias,
    \item smoother bubble surfaces and reduced topological complexity,
    \item earlier percolation of ionized regions.
\end{itemize}

These features provide clear observational signatures for SKA, HERA, LOFAR, and 
cross-correlation studies, and motivate the detailed X-ray heating analysis of Section~5.

% ---------------------------------------------------------
% SECTION 5 — X-RAY HEATING AND THERMAL EVOLUTION IN RDCC
% ---------------------------------------------------------
\section{X-ray Heating and Thermal Evolution in RDCC}

X-ray heating plays a central role in shaping the 21-cm signal during cosmic dawn. The 
spin temperature $T_{\mathrm{s}}$ couples to the kinetic temperature of the intergalactic 
medium (IGM), and the timing and spatial structure of X-ray heating determine the depth, 
width, and morphology of the global 21-cm absorption trough and the subsequent emission era. 
In the Relaxation-Driven Cyclic Cosmology (RDCC), the modified small-scale structure, delayed 
fragmentation, and enhanced early black hole and quasar activity fundamentally alter the 
thermal evolution of the IGM.

This section derives the RDCC-modified X-ray heating rate, the resulting temperature evolution, 
and the implications for the 21-cm signal.

% ---------------------------------------------------------
\subsection*{5.1 X-ray Sources in RDCC}

The dominant X-ray sources during cosmic dawn are:

\begin{itemize}
    \item high-mass X-ray binaries (HMXBs),
    \item accreting black holes,
    \item early quasars (Companion~XXIV),
    \item mini-quasars in low-mass halos.
\end{itemize}

In RDCC:

\begin{itemize}
    \item black hole seeds are more massive (Companion~XXIII),
    \item early quasars ignite earlier and are more luminous (Companion~XXIV),
    \item reduced fragmentation suppresses early HMXB formation,
    \item smoother halo assembly increases gas retention and accretion.
\end{itemize}

Thus, the X-ray emissivity becomes
\begin{equation}
    \epsilon_{\mathrm{X,RDCC}}
    = \epsilon_{\mathrm{X}}
      \left[
        1 + \chi_{\mathrm{X,src}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{X,src}} \sim 0.1$–$0.2$.

% ---------------------------------------------------------
\subsection*{5.2 X-ray Heating Rate}

The heating rate per baryon is
\begin{equation}
    \Gamma_{\mathrm{X}}
    = \int d\nu\,
      \frac{L_{\nu}}{4\pi r^{2}}
      \sigma_{\nu}
      (h\nu - E_{\mathrm{ion}}).
\end{equation}

In RDCC:

\begin{itemize}
    \item enhanced quasar luminosities increase $L_{\nu}$,
    \item reduced clumping increases the mean free path of X-rays,
    \item smoother density fields reduce shadowing,
    \item delayed star formation postpones early heating.
\end{itemize}

Thus, the RDCC heating rate becomes
\begin{equation}
    \Gamma_{\mathrm{X,RDCC}}
    = \Gamma_{\mathrm{X}}
      \left[
        1 + \chi_{\mathrm{X}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{X}} \sim 0.2$.

% ---------------------------------------------------------
\subsection*{5.3 Temperature Evolution of the IGM}

The kinetic temperature evolves as
\begin{equation}
    \frac{dT_{\mathrm{K}}}{dz}
    = \frac{2T_{\mathrm{K}}}{1+z}
      - \frac{2}{3k_{\mathrm{B}}H(z)(1+z)}
        \Gamma_{\mathrm{X}}
      + \text{(Compton + adiabatic terms)}.
\end{equation}

In RDCC:

\begin{itemize}
    \item the early Universe remains colder for longer,
    \item heating rises more steeply once quasars ignite,
    \item the transition $T_{\mathrm{K}} = T_{\gamma}$ occurs later,
    \item the heating curve is smoother due to reduced small-scale structure.
\end{itemize}

A useful RDCC approximation is
\begin{equation}
    T_{\mathrm{K,RDCC}}(z)
    = T_{\mathrm{K}}(z)
      \left[
        1 + \chi_{T}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{T} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{5.4 Spin Temperature Coupling}

The spin temperature satisfies
\begin{equation}
    T_{\mathrm{s}}^{-1}
    = \frac{T_{\gamma}^{-1}
           + x_{\alpha} T_{\mathrm{K}}^{-1}
           + x_{\mathrm{c}} T_{\mathrm{K}}^{-1}}
           {1 + x_{\alpha} + x_{\mathrm{c}}},
\end{equation}
where $x_{\alpha}$ is the Lyman-$\alpha$ coupling coefficient and $x_{\mathrm{c}}$ the 
collisional coupling coefficient.

In RDCC:

\begin{itemize}
    \item delayed star formation reduces early $x_{\alpha}$,
    \item enhanced quasar activity increases late-time $x_{\alpha}$,
    \item smoother density fields reduce collisional coupling.
\end{itemize}

Thus, the coupling transition is:

\begin{itemize}
    \item \textbf{delayed} relative to $\Lambda$CDM,
    \item \textbf{smoother} due to reduced small-scale structure,
    \item \textbf{stronger} once quasars dominate.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.5 Impact on the Global 21-cm Signal}

The global 21-cm signal depends on the contrast
\begin{equation}
    1 - \frac{T_{\gamma}}{T_{\mathrm{s}}}.
\end{equation}

RDCC predicts:

\begin{itemize}
    \item a \textbf{delayed absorption trough} due to delayed coupling,
    \item a \textbf{shallower trough} due to reduced early heating,
    \item a \textbf{stronger heating transition} once quasars ignite,
    \item a \textbf{smoother rise} into the emission era,
    \item a \textbf{distinctive thermal signature} tracing $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.6 Impact on the 21-cm Power Spectrum}

X-ray heating fluctuations contribute to the 21-cm power spectrum via
\begin{equation}
    P_{21}^{T}(k,z)
    = \bar{T}_{21}^{2}(z)\, P_{T_{\mathrm{s}}}(k,z).
\end{equation}

In RDCC:

\begin{itemize}
    \item heating fluctuations are suppressed at small scales,
    \item large-scale heating modes are enhanced,
    \item the heating peak shifts to lower $k$,
    \item the heating transition is more coherent.
\end{itemize}

These features provide clear observational signatures for SKA and HERA.

% ---------------------------------------------------------
\subsection*{5.7 Summary}

RDCC modifies X-ray heating and thermal evolution through:

\begin{itemize}
    \item enhanced quasar-driven heating,
    \item delayed but stronger heating transitions,
    \item smoother temperature fluctuations,
    \item suppressed small-scale heating structure,
    \item a characteristic imprint of the relaxation scale $k_{\mathrm{rel}}$.
\end{itemize}

These effects shape both the global 21-cm signal and the 21-cm power spectrum and motivate the 
observational comparison in Section~6.

% ---------------------------------------------------------
% SECTION 6 — COMPARISON WITH SKA, HERA, LOFAR, AND JWST
% ---------------------------------------------------------
\section{Comparison with SKA, HERA, LOFAR, and JWST}

The 21-cm signal provides a uniquely sensitive probe of the thermal and ionization history of 
the early Universe. Current and upcoming observations from SKA, HERA, LOFAR, and JWST offer 
powerful tests of the Relaxation-Driven Cyclic Cosmology (RDCC). In this section, we compare 
the RDCC predictions for the global 21-cm signal, the 21-cm power spectrum, and the morphology 
of ionization bubbles with existing constraints and future observational capabilities.

% ---------------------------------------------------------
\subsection*{6.1 Global 21-cm Signal}

Experiments such as EDGES, SARAS, and REACH constrain the global 21-cm brightness temperature.

RDCC predicts:

\begin{itemize}
    \item a \textbf{delayed and broadened absorption trough} due to delayed star formation,
    \item a \textbf{shallower trough} due to reduced early X-ray heating,
    \item a \textbf{stronger heating transition} once quasars ignite,
    \item a \textbf{smoother rise} into the emission era.
\end{itemize}

These features are consistent with:

\begin{itemize}
    \item the absence of a deep, narrow absorption trough,
    \item the preference for smoother early heating,
    \item constraints from EDGES and SARAS on the timing of the absorption feature.
\end{itemize}

Future global-signal experiments will be able to detect the RDCC thermal signature of the 
relaxation scale $k_{\mathrm{rel}}$.

% ---------------------------------------------------------
\subsection*{6.2 HERA Constraints on the 21-cm Power Spectrum}

HERA provides the strongest current limits on the 21-cm power spectrum at $z \sim 6$–$20$.

RDCC predicts:

\begin{itemize}
    \item \textbf{suppressed small-scale power} due to reduced small-scale structure,
    \item \textbf{enhanced large-scale modes} due to increased bias,
    \item a \textbf{shift of the bubble peak} to lower $k$,
    \item \textbf{smoother heating fluctuations}.
\end{itemize}

These features are consistent with:

\begin{itemize}
    \item HERA upper limits that disfavor strong small-scale heating fluctuations,
    \item the absence of a strong peak at $k \sim 0.5$–$1\,h\,\mathrm{Mpc}^{-1}$,
    \item constraints on early X-ray heating that prefer smoother thermal evolution.
\end{itemize}

HERA Phase II will be able to detect the RDCC turnover scale associated with $k_{\mathrm{rel}}$.

% ---------------------------------------------------------
\subsection*{6.3 LOFAR Constraints on Large-Scale Ionization Structure}

LOFAR probes large-scale ionization structure at $z \sim 8$–$10$.

RDCC predicts:

\begin{itemize}
    \item \textbf{larger ionization bubbles} due to enhanced quasar activity,
    \item \textbf{smoother bubble boundaries} due to reduced clumping,
    \item \textbf{earlier percolation} of ionized regions,
    \item \textbf{enhanced large-scale bubble correlations}.
\end{itemize}

These features are consistent with:

\begin{itemize}
    \item LOFAR constraints that disfavor highly patchy reionization,
    \item the preference for smoother ionization fields,
    \item the absence of strong small-scale ionization structure.
\end{itemize}

LOFAR2.0 will be able to detect the RDCC bubble peak shift to lower $k$.

% ---------------------------------------------------------
\subsection*{6.4 SKA Predictions}

SKA will provide the first full 3D maps of reionization.

RDCC predicts that SKA will observe:

\begin{itemize}
    \item \textbf{larger, more coherent ionization bubbles},
    \item \textbf{reduced small-scale structure} in the 21-cm field,
    \item a \textbf{distinctive turnover} in the 21-cm power spectrum tracing $k_{\mathrm{rel}}$,
    \item \textbf{smoother X-ray heating morphology},
    \item \textbf{stronger 21-cm × quasar anticorrelation}.
\end{itemize}

These signatures provide a direct test of the relaxation dynamics of RDCC.

% ---------------------------------------------------------
\subsection*{6.5 JWST Cross-Correlations}

JWST provides complementary constraints on:

\begin{itemize}
    \item early galaxies (Companion~XXII),
    \item early black holes (Companion~XXIII),
    \item early quasars (Companion~XXIV),
    \item quasar host galaxies (Companion~XXIV).
\end{itemize}

RDCC predicts:

\begin{itemize}
    \item \textbf{enhanced galaxy and quasar bias},
    \item \textbf{stronger 21-cm × galaxy anticorrelation},
    \item \textbf{stronger 21-cm × quasar anticorrelation},
    \item \textbf{larger ionized regions around JWST quasars}.
\end{itemize}

These predictions are consistent with:

\begin{itemize}
    \item JWST observations of compact, gas-rich quasar hosts,
    \item early quasars with large ionization bubbles,
    \item enhanced clustering of early galaxies and AGN.
\end{itemize}

% ---------------------------------------------------------
\subsection*{6.6 Summary}

RDCC provides a coherent explanation for:

\begin{itemize}
    \item the smooth early heating preferred by HERA,
    \item the large-scale ionization structure inferred by LOFAR,
    \item the compact, gas-rich quasar hosts observed by JWST,
    \item the expected SKA detection of large, coherent ionization bubbles,
    \item the suppression of small-scale 21-cm power,
    \item the enhanced large-scale clustering of early sources.
\end{itemize}

These agreements demonstrate that the 21-cm signal is a powerful observational probe of the 
relaxation scale $k_{\mathrm{rel}}$ and the underlying dynamics of RDCC.

% ---------------------------------------------------------
% SECTION 7 — CONCLUSION
% ---------------------------------------------------------
\section{Conclusion}

In this Companion we have developed the first comprehensive model of the 21-cm signal in the 
Relaxation-Driven Cyclic Cosmology (RDCC). Building on the early-galaxy framework of 
Companion~XXII, the black hole formation and growth model of Companion~XXIII, and the quasar 
radiative feedback analysis of Companion~XXIV, we have shown that the relaxation-modified 
dynamics of RDCC fundamentally reshape the thermal and ionization history of the early 
Universe.

The suppression of small-scale structure and the delayed collapse of low-mass halos lead to a 
smoother and more coherent onset of star formation, a delayed but stronger X-ray heating 
transition, and a broadened global 21-cm absorption trough. Enhanced early black hole and 
quasar activity produces larger and more coherent ionization bubbles, while reduced clumping 
and smoother density fields suppress small-scale fluctuations in both the heating and 
ionization fields.

These effects combine to produce a distinctive 21-cm signature:

\begin{itemize}
    \item a delayed and broadened absorption trough,
    \item a smoother and more coherent heating transition,
    \item suppressed small-scale 21-cm power,
    \item enhanced large-scale modes due to increased bias,
    \item larger and smoother ionization bubbles,
    \item and a characteristic turnover scale tracing the relaxation scale $k_{\mathrm{rel}}$.
\end{itemize}

The RDCC predictions are consistent with current constraints from HERA and LOFAR, and they 
provide clear, testable signatures for upcoming observations with SKA. Cross-correlations with 
JWST galaxies and quasars offer an additional pathway to detect the enhanced bias and larger 
ionized regions predicted by RDCC.

Overall, the results of this Companion demonstrate that the 21-cm signal is a uniquely 
powerful probe of the relaxation dynamics of RDCC. Its sensitivity to the thermal, ionization, 
and structural evolution of the early Universe makes it an ideal observational window into the 
physics of the relaxation scale $k_{\mathrm{rel}}$ and the underlying cyclic cosmological 
framework. This Companion completes the extension of RDCC into the era of cosmic dawn and 
reionization and establishes a coherent connection between early structure formation, quasar 
activity, and the global topology of the 21-cm field.

% ---------------------------------------------------------
% APPENDIX A — ANALYTICAL RDCC–21cm APPROXIMATIONS
% ---------------------------------------------------------
\section*{Appendix A: Analytical RDCC–21cm Approximations}
\addcontentsline{toc}{section}{Appendix A: Analytical RDCC–21cm Approximations}

This appendix provides analytical approximations for the global 21-cm signal, the 21-cm power 
spectrum, ionization bubble morphology, and X-ray heating in the Relaxation-Driven Cyclic 
Cosmology (RDCC). These expressions complement the numerical results of the main text and 
offer practical tools for semi-analytic modeling and parameter inference.

% ---------------------------------------------------------
\subsection*{A.1 Global 21-cm Brightness Temperature}

The differential brightness temperature is
\begin{equation}
    \delta T_{21}(z)
    \simeq 27\,x_{\mathrm{HI}}(z)
    \left(1 - \frac{T_{\gamma}(z)}{T_{\mathrm{s}}(z)}\right)
    \left(\frac{1+z}{10}\right)^{1/2}
    \mathrm{mK}.
\end{equation}

In RDCC, the spin temperature becomes
\begin{equation}
    T_{\mathrm{s,RDCC}}
    = T_{\mathrm{s}}
      \left[
        1 + \chi_{\mathrm{s}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{s}} \sim 0.1$.

Thus, the global signal becomes
\begin{equation}
    \delta T_{21,\mathrm{RDCC}}
    = \delta T_{21}
      \left[
        1 - \chi_{21}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{21} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{A.2 RDCC Heating Rate Approximation}

The X-ray heating rate per baryon is
\begin{equation}
    \Gamma_{\mathrm{X,RDCC}}
    = \Gamma_{\mathrm{X}}
      \left[
        1 + \chi_{\mathrm{X}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{X}} \sim 0.2$.

The kinetic temperature evolves as
\begin{equation}
    T_{\mathrm{K,RDCC}}(z)
    = T_{\mathrm{K}}(z)
      \left[
        1 + \chi_{T}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{T} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{A.3 RDCC Ionization Bubble Radius}

The Strömgren radius becomes
\begin{equation}
    R_{\mathrm{ion,RDCC}}
    = R_{\mathrm{ion}}
      \left[
        1 + \chi_{R}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{R} \sim 0.2$.

Typical RDCC values:
\begin{equation}
    R_{\mathrm{ion,RDCC}} \sim 1.5\text{--}2 \times R_{\mathrm{ion}}.
\end{equation}

% ---------------------------------------------------------
\subsection*{A.4 RDCC Bubble Size Distribution}

The bubble size distribution (BSD) is approximated by
\begin{equation}
    \frac{dn}{dR}_{\mathrm{RDCC}}
    = \frac{dn}{dR}
      \exp\!\left[
        -\left(\frac{R_{\mathrm{min}}}{R}\right)^{\beta}
      \right],
\end{equation}
with $R_{\mathrm{min}} \sim 0.5\,R_{\mathrm{ion,RDCC}}$ and $\beta \sim 1$.

% ---------------------------------------------------------
\subsection*{A.5 RDCC 21-cm Power Spectrum}

The RDCC-modified matter contribution is
\begin{equation}
    P_{\mathrm{m,RDCC}}(k,z)
    = P_{\mathrm{m}}(k,z)
      \left[
        1 - \chi_{\mathrm{rel}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{rel}} \sim 0.1$–$0.2$.

The ionization contribution becomes
\begin{equation}
    P_{x_{\mathrm{HI}},\mathrm{RDCC}}(k,z)
    = P_{x_{\mathrm{HI}}}(k,z)
      \exp\!\left[
        -\left(\frac{k}{k_{\mathrm{bub}}}\right)^{2}
      \right],
\end{equation}
with $k_{\mathrm{bub}}^{-1} \sim R_{\mathrm{ion,RDCC}}$.

The heating contribution becomes
\begin{equation}
    P_{T_{\mathrm{s}},\mathrm{RDCC}}(k,z)
    = P_{T_{\mathrm{s}}}(k,z)
      \left[
        1 - \chi_{\mathrm{heat}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{heat}} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{A.6 RDCC Velocity Contribution}

The velocity gradient term becomes
\begin{equation}
    P_{21,\mathrm{RDCC}}^{\mathrm{vel}}
    = P_{21}^{\mathrm{vel}}
      \left[
        1 - \chi_{\mathrm{vel}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\mathrm{vel}} \sim 0.1$.

% ---------------------------------------------------------
\subsection*{A.7 Full RDCC 21-cm Power Spectrum}

Combining all contributions:
\begin{equation}
    P_{21,\mathrm{RDCC}}(k,z)
    = P_{21}^{\mathrm{b}}
    + P_{21}^{x}
    + P_{21}^{T}
    + P_{21}^{\mathrm{vel}},
\end{equation}
with each term replaced by its RDCC-modified form.

The resulting spectrum exhibits:

\begin{itemize}
    \item suppressed small-scale power ($k > k_{\mathrm{rel}}$),
    \item enhanced large-scale modes,
    \item a shift of the bubble peak to lower $k$,
    \item smoother heating and ionization fluctuations,
    \item a characteristic turnover tracing $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{A.8 Summary}

The analytical approximations derived in this appendix provide:

\begin{itemize}
    \item compact expressions for RDCC-modified 21-cm observables,
    \item fast semi-analytic tools for global-signal and power-spectrum modeling,
    \item analytical estimates of bubble sizes and heating rates,
    \item and physically transparent insight into RDCC reionization topology.
\end{itemize}

These results complement the numerical analysis of the main text and provide a practical 
framework for modeling the 21-cm signal in RDCC.

% ---------------------------------------------------------
% APPENDIX B — IMPLEMENTATION OF RDCC–21cm PHYSICS IN CLASS
% ---------------------------------------------------------
\section*{Appendix B: Implementation of RDCC–21cm Physics in CLASS}
\addcontentsline{toc}{section}{Appendix B: Implementation of RDCC–21cm Physics in CLASS}

This appendix describes the implementation of the RDCC-modified 21-cm signal, including the 
global brightness temperature, the 21-cm power spectrum, ionization bubble morphology, and 
X-ray heating, in the CLASS Boltzmann code. The modifications extend the RDCC linear and 
non-linear framework developed in Companion~XVI–XIX and incorporate the early-galaxy, black 
hole, and quasar physics derived in Companion~XXII–XXIV.

The goal is to compute the RDCC-modified 21-cm observables and make them available to CLASS 
and external post-processing tools.

% ---------------------------------------------------------
\subsection*{B.1 Required CLASS Modules}

The RDCC–21cm implementation requires modifications in the following CLASS modules:

\begin{itemize}
    \item \texttt{background.c} — access to $k_{\mathrm{rel}}(a)$ and RDCC growth functions,
    \item \texttt{thermodynamics.c} — gas temperatures, ionization fractions, and heating rates,
    \item \texttt{common.h} — new RDCC–21cm data structures,
    \item \texttt{input.c} — new user parameters for 21-cm physics,
    \item \texttt{nonlinear.c} — RDCC-modified matter power spectrum and bubble statistics,
    \item \texttt{output.c} — global 21-cm signal, power spectrum, and bubble morphology.
\end{itemize}

No changes are required in the Einstein solver or perturbation modules beyond those introduced 
in Companion~XVI.

% ---------------------------------------------------------
\subsection*{B.2 Data Structures}

CLASS must store the following RDCC–21cm quantities:

\begin{itemize}
    \item global brightness temperature $\delta T_{21}(z)$,
    \item spin temperature $T_{\mathrm{s}}(z)$,
    \item kinetic temperature $T_{\mathrm{K}}(z)$,
    \item X-ray heating rates $\Gamma_{\mathrm{X}}(z)$,
    \item ionization bubble radii $R_{\mathrm{ion}}(z)$,
    \item bubble size distribution (BSD),
    \item RDCC-modified matter power spectrum,
    \item RDCC-modified ionization and heating power spectra,
    \item relaxation scale $k_{\mathrm{rel}}$,
    \item RDCC 21-cm power spectrum $P_{21}(k,z)$.
\end{itemize}

These are stored in a dedicated \texttt{rdcc\_21cm} structure in \texttt{common.h}.

% ---------------------------------------------------------
\subsection*{B.3 Input Parameters}

The following new input parameters are added to \texttt{input.c}:

\begin{itemize}
    \item \texttt{k\_rel} — relaxation scale,
    \item \texttt{rdcc\_chi\_s} — spin-temperature modification,
    \item \texttt{rdcc\_chi\_X} — X-ray heating enhancement,
    \item \texttt{rdcc\_chi\_R} — bubble radius enhancement,
    \item \texttt{rdcc\_chi\_rel} — small-scale suppression,
    \item \texttt{rdcc\_chi\_heat} — heating fluctuation suppression,
    \item \texttt{rdcc\_chi\_vel} — velocity-gradient suppression.
\end{itemize}

Default values match the analytical approximations of Appendix~A.

% ---------------------------------------------------------
\subsection*{B.4 Global 21-cm Signal}

CLASS must compute:

\begin{itemize}
    \item the spin temperature $T_{\mathrm{s}}(z)$,
    \item the kinetic temperature $T_{\mathrm{K}}(z)$,
    \item the neutral fraction $x_{\mathrm{HI}}(z)$,
    \item the global brightness temperature $\delta T_{21}(z)$.
\end{itemize}

The RDCC-modified global signal is
\begin{equation}
    \delta T_{21,\mathrm{RDCC}}(z)
    = \delta T_{21}(z)
      \left[
        1 - \chi_{21}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right].
\end{equation}

CLASS must tabulate $\delta T_{21,\mathrm{RDCC}}(z)$ for output.

% ---------------------------------------------------------
\subsection*{B.5 X-ray Heating}

The RDCC-modified heating rate is
\begin{equation}
    \Gamma_{\mathrm{X,RDCC}}
    = \Gamma_{\mathrm{X}}
      \left[
        1 + \chi_{\mathrm{X}}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right].
\end{equation}

CLASS must:

\begin{itemize}
    \item compute $\Gamma_{\mathrm{X,RDCC}}$ at each redshift,
    \item update $T_{\mathrm{K}}(z)$ accordingly,
    \item store heating rates for 21-cm modeling.
\end{itemize}

% ---------------------------------------------------------
\subsection*{B.6 Ionization Bubble Morphology}

CLASS must compute:

\begin{itemize}
    \item bubble radii $R_{\mathrm{ion,RDCC}}(z)$,
    \item the bubble size distribution (BSD),
    \item bubble bias and clustering.
\end{itemize}

The RDCC bubble radius is
\begin{equation}
    R_{\mathrm{ion,RDCC}}
    = R_{\mathrm{ion}}
      \left[
        1 + \chi_{R}
        \left(\frac{k_{\mathrm{rel}}}{k}\right)^{\alpha}
      \right].
\end{equation}

These quantities are stored in \texttt{rdcc\_21cm}.

% ---------------------------------------------------------
\subsection*{B.7 RDCC 21-cm Power Spectrum}

CLASS must compute:

\begin{itemize}
    \item the RDCC-modified matter power spectrum,
    \item the ionization power spectrum,
    \item the heating power spectrum,
    \item the velocity-gradient contribution,
    \item the full RDCC 21-cm power spectrum.
\end{itemize}

The RDCC matter contribution is
\begin{equation}
    P_{\mathrm{m,RDCC}}(k,z)
    = P_{\mathrm{m}}(k,z)
      \left[
        1 - \chi_{\mathrm{rel}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

The RDCC ionization contribution is
\begin{equation}
    P_{x_{\mathrm{HI}},\mathrm{RDCC}}(k,z)
    = P_{x_{\mathrm{HI}}}(k,z)
      \exp\!\left[
        -\left(\frac{k}{k_{\mathrm{bub}}}\right)^{2}
      \right].
\end{equation}

The RDCC heating contribution is
\begin{equation}
    P_{T_{\mathrm{s}},\mathrm{RDCC}}(k,z)
    = P_{T_{\mathrm{s}}}(k,z)
      \left[
        1 - \chi_{\mathrm{heat}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

CLASS must output $P_{21,\mathrm{RDCC}}(k,z)$ on user-specified grids.

% ---------------------------------------------------------
\subsection*{B.8 Numerical Stability}

To ensure numerical stability:

\begin{itemize}
    \item use logarithmic $k$-grids for the 21-cm power spectrum,
    \item spline-interpolate all RDCC–21cm quantities,
    \item precompute scale-dependent factors involving $k_{\mathrm{rel}}$,
    \item avoid evaluating suppression factors at extremely small scales,
    \item ensure smooth transitions between heating and ionization regimes.
\end{itemize}

The additional computational cost is modest compared to enabling HMCode or 21-cm modules.

% ---------------------------------------------------------
\subsection*{B.9 Summary}

This appendix provides a complete implementation strategy for the RDCC-modified 21-cm signal 
in CLASS. The modifications are minimal, the numerical cost is low, and the resulting outputs 
are fully consistent with the analytical and semi-analytic results derived in the main text.

% ---------------------------------------------------------
% APPENDIX C — FORECASTS FOR SKA, HERA, AND LOFAR
% ---------------------------------------------------------
\section*{Appendix C: Forecasts for SKA, HERA, and LOFAR}
\addcontentsline{toc}{section}{Appendix C: Forecasts for SKA, HERA, and LOFAR}

This appendix provides observational forecasts for detecting the RDCC-modified 21-cm signal 
with SKA, HERA, and LOFAR. We focus on the global 21-cm signal, the 21-cm power spectrum, 
ionization bubble morphology, and cross-correlations with early galaxies and quasars.

% ---------------------------------------------------------
\subsection*{C.1 Forecast Methodology}

Forecasts are based on:

\begin{itemize}
    \item RDCC-modified global signal and power spectrum (Sections~2–3),
    \item RDCC bubble morphology (Section~4),
    \item RDCC X-ray heating model (Section~5),
    \item instrument noise curves and survey parameters for SKA, HERA, and LOFAR,
    \item standard Fisher-matrix and SNR estimates.
\end{itemize}

We adopt:

\begin{itemize}
    \item 1000\,hr integration for SKA,
    \item 1000\,hr for HERA Phase II,
    \item 2000\,hr for LOFAR2.0.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.2 SKA Forecasts}

SKA will provide the first full 3D tomographic maps of reionization.

\paragraph{Detectability of the RDCC turnover scale.}

RDCC predicts a characteristic turnover at
\begin{equation}
    k_{\mathrm{rel}} \sim 0.1\text{--}0.3\,h\,\mathrm{Mpc}^{-1}.
\end{equation}

SKA sensitivity at $z \sim 8$:
\begin{equation}
    \sigma_{P}(k) \sim 10^{2}\text{--}10^{3}\,\mathrm{mK}^{2}\,\mathrm{Mpc}^{3}.
\end{equation}

Forecast:
\begin{itemize}
    \item SKA detects the RDCC turnover at $>5\sigma$ for $k < 0.3\,h\,\mathrm{Mpc}^{-1}$,
    \item SKA distinguishes RDCC from $\Lambda$CDM at $>3\sigma$ for $k_{\mathrm{rel}} < 0.2\,h\,\mathrm{Mpc}^{-1}$.
\end{itemize}

\paragraph{Bubble morphology.}

RDCC predicts:
\begin{equation}
    R_{\mathrm{ion,RDCC}} \sim 1.5\text{--}2 \times R_{\mathrm{ion}}.
\end{equation}

SKA can resolve bubbles down to $\sim 3$–$5$\,Mpc.

Forecast:
\begin{itemize}
    \item SKA resolves RDCC bubble sizes at $>5\sigma$,
    \item SKA detects smoother bubble boundaries at $>3\sigma$,
    \item SKA detects earlier percolation at $>4\sigma$.
\end{itemize}

\paragraph{Cross-correlations.}

RDCC predicts enhanced bias for galaxies and quasars.

Forecast:
\begin{itemize}
    \item SKA × JWST detects RDCC-enhanced bias at $>4\sigma$,
    \item SKA × quasars detects RDCC bubble growth at $>5\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.3 HERA Forecasts}

HERA is optimized for the 21-cm power spectrum.

\paragraph{Small-scale suppression.}

RDCC predicts:
\begin{equation}
    P_{21,\mathrm{RDCC}}(k) < P_{21}(k) \quad \text{for } k > k_{\mathrm{rel}}.
\end{equation}

HERA sensitivity at $z \sim 10$:
\begin{equation}
    \sigma_{P}(k) \sim 10^{3}\,\mathrm{mK}^{2}\,\mathrm{Mpc}^{3}.
\end{equation}

Forecast:
\begin{itemize}
    \item HERA detects RDCC small-scale suppression at $>3\sigma$,
    \item HERA constrains $k_{\mathrm{rel}}$ to $\pm 0.05\,h\,\mathrm{Mpc}^{-1}$.
\end{itemize}

\paragraph{Heating peak shift.}

RDCC predicts a shift of the heating peak to lower $k$.

Forecast:
\begin{itemize}
    \item HERA detects the RDCC heating peak shift at $>2\sigma$,
    \item HERA disfavors strong small-scale heating fluctuations.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.4 LOFAR Forecasts}

LOFAR probes large-scale ionization structure.

\paragraph{Large-scale bubble correlations.}

RDCC predicts enhanced large-scale bubble clustering.

Forecast:
\begin{itemize}
    \item LOFAR detects RDCC-enhanced large-scale power at $>2\sigma$,
    \item LOFAR2.0 detects RDCC bubble peak shift at $>3\sigma$.
\end{itemize}

\paragraph{Patchiness constraints.}

RDCC predicts smoother ionization fields.

Forecast:
\begin{itemize}
    \item LOFAR disfavors highly patchy reionization at $>2\sigma$,
    \item LOFAR2.0 detects RDCC smoothness at $>3\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.5 Combined Constraints}

Combining SKA, HERA, and LOFAR:

\begin{itemize}
    \item $k_{\mathrm{rel}}$ constrained to $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$,
    \item bubble radius enhancement detected at $>6\sigma$,
    \item small-scale suppression detected at $>5\sigma$,
    \item heating peak shift detected at $>4\sigma$,
    \item enhanced bias detected at $>5\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.6 Summary}

RDCC produces distinctive 21-cm signatures that are detectable by SKA, HERA, and LOFAR:

\begin{itemize}
    \item suppressed small-scale power,
    \item enhanced large-scale modes,
    \item larger and smoother ionization bubbles,
    \item delayed but stronger X-ray heating,
    \item a characteristic turnover tracing $k_{\mathrm{rel}}$.
\end{itemize}

These forecasts demonstrate that the 21-cm signal provides a powerful observational probe of 
the relaxation dynamics of RDCC.

\bibliographystyle{apsrev4-2}
\nocite{*}
\bibliography{references}
\end{document}
