% ---------------------------------------------------------
% PREAMBLE (same template as Companion X–XXV)
% ---------------------------------------------------------
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{physics}
\usepackage{cite}
\usepackage{float}

\geometry{margin=1in}

% ---------------------------------------------------------
% TITLE
% ---------------------------------------------------------
\title{
\textbf{Companion XXVI}\\[4mm]
\textbf{Gravitational Lensing in the Relaxation-Driven Cyclic Cosmology}\\
\textbf{CMB Lensing, Galaxy Lensing, and Large-Scale Potential Smoothing}
}

\author{
Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}
}

\date{April 2026}

\begin{document}

\maketitle

% ---------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------
\begin{abstract}
Gravitational lensing provides one of the most powerful probes of the large-scale matter 
distribution in the Universe. Measurements of CMB lensing, galaxy–galaxy lensing, cosmic 
shear, and lensing convergence maps from CMB-S4, Euclid, Roman, LSST, and SKA offer 
unprecedented sensitivity to the growth of structure, the clustering of matter, and the 
small-scale potential fluctuations that shape the cosmic web.

In the Relaxation-Driven Cyclic Cosmology (RDCC), the suppression of small-scale power, the 
smoother assembly of halos, and the enhanced large-scale coherence of gravitational 
potentials fundamentally alter the lensing signal. Building on the linear and non-linear 
structure formation framework of Companion~XVI–XIX, we derive the RDCC predictions for CMB 
lensing, galaxy lensing, and cosmic shear. We show that RDCC produces smoother lensing 
potentials, reduced small-scale convergence power, enhanced large-scale shear correlations, 
and a characteristic turnover scale tracing the relaxation scale $k_{\mathrm{rel}}$.

This Companion establishes gravitational lensing as a precision probe of the relaxation 
dynamics of RDCC and provides clear predictions for upcoming surveys including CMB-S4, 
Euclid, Roman, LSST, and SKA.
\end{abstract}

% ---------------------------------------------------------
% SECTION 1 — MOTIVATION AND OVERVIEW
% ---------------------------------------------------------
\section{Motivation and Overview}

Gravitational lensing provides one of the most powerful and model-independent probes of the 
large-scale matter distribution in the Universe. Measurements of CMB lensing, galaxy–galaxy 
lensing, cosmic shear, and lensing convergence maps from CMB-S4, Euclid, Roman, LSST, and SKA 
offer unprecedented sensitivity to the growth of structure, the clustering of matter, and the 
small-scale gravitational potentials that shape the cosmic web.

In the Relaxation-Driven Cyclic Cosmology (RDCC), the suppression of small-scale power, the 
smoother assembly of halos, and the enhanced large-scale coherence of gravitational potentials 
fundamentally alter the lensing signal. These effects arise from the scale-dependent 
relaxation rate $\Gamma(a)$, which modifies the growth of structure and the non-linear 
evolution of density fields. As shown in Companion~XVIII and Companion~XIX, RDCC predicts 
smoother small-scale structure, reduced halo concentrations, and enhanced large-scale bias. 
These modifications directly impact lensing observables.

This Companion develops the RDCC predictions for CMB lensing, galaxy lensing, and cosmic 
shear, and compares them with current and upcoming observational constraints.

% ---------------------------------------------------------
\subsection*{1.1 Observational Context}

Modern lensing surveys probe the matter distribution across a wide range of scales:

\begin{itemize}
    \item \textbf{CMB lensing} (Planck, ACT, SPT, CMB-S4) measures the integrated potential 
    along the line of sight to $z \sim 1100$.
    \item \textbf{Cosmic shear} (Euclid, Roman, LSST) maps the projected matter distribution 
    at $z \sim 0.5$–$2$.
    \item \textbf{Galaxy–galaxy lensing} probes halo profiles and bias.
    \item \textbf{Strong lensing} constrains small-scale substructure.
    \item \textbf{SKA lensing} will provide radio-based shear maps with exquisite precision.
\end{itemize}

These observations are sensitive to:

\begin{itemize}
    \item the matter power spectrum $P(k,z)$,
    \item the growth rate of structure $f(a)$,
    \item halo concentrations and profiles,
    \item the abundance of small-scale subhalos,
    \item the coherence of large-scale gravitational potentials.
\end{itemize}

Each of these quantities is modified in RDCC.

% ---------------------------------------------------------
\subsection*{1.2 RDCC Perspective}

The Relaxation-Driven Cyclic Cosmology modifies structure formation through the 
scale-dependent relaxation rate $\Gamma(a)$, which suppresses small-scale power and enhances 
large-scale coherence. As shown in Companion~XVI–XIX, this leads to:

\begin{itemize}
    \item reduced small-scale clustering,
    \item smoother gravitational potentials,
    \item lower halo concentrations,
    \item suppressed subhalo abundance,
    \item enhanced large-scale bias.
\end{itemize}

These effects directly influence lensing observables:

\begin{itemize}
    \item \textbf{CMB lensing:} reduced small-scale convergence power, enhanced large-scale 
    modes, smoother lensing potential.
    \item \textbf{Cosmic shear:} suppressed small-scale shear, enhanced large-scale 
    correlations, reduced shape noise from small-scale structure.
    \item \textbf{Galaxy–galaxy lensing:} lower halo concentrations, smoother mass profiles, 
    enhanced large-scale bias.
    \item \textbf{Strong lensing:} reduced subhalo lensing anomalies.
\end{itemize}

Thus, gravitational lensing provides a direct observational probe of the relaxation scale 
$k_{\mathrm{rel}}$.

% ---------------------------------------------------------
\subsection*{1.3 Goals of This Companion}

The goals of this Companion are:

\begin{itemize}
    \item to derive the RDCC predictions for CMB lensing,
    \item to compute the RDCC cosmic shear and galaxy–galaxy lensing signals,
    \item to analyze RDCC modifications to halo profiles and substructure,
    \item to model the RDCC lensing potential and convergence field,
    \item and to compare RDCC predictions with CMB-S4, Euclid, Roman, LSST, and SKA.
\end{itemize}

These results build directly on the RDCC structure formation framework developed in 
Companion~XVI–XIX.

% ---------------------------------------------------------
\subsection*{1.4 Structure of This Companion}

The structure of this Companion is as follows:

\begin{itemize}
    \item Section~2 derives the RDCC lensing potential and convergence field.
    \item Section~3 computes the RDCC CMB lensing power spectrum.
    \item Section~4 analyzes cosmic shear and galaxy–galaxy lensing in RDCC.
    \item Section~5 develops the RDCC halo profile and substructure model.
    \item Section~6 compares RDCC predictions with CMB-S4, Euclid, Roman, LSST, and SKA.
\end{itemize}

Together, these results establish gravitational lensing as a precision probe of the relaxation 
dynamics of RDCC.

% ---------------------------------------------------------
% SECTION 2 — THE RDCC LENSING POTENTIAL AND CONVERGENCE FIELD
% ---------------------------------------------------------
\section{The RDCC Lensing Potential and Convergence Field}

Gravitational lensing is determined by the projected gravitational potential along the line of 
sight. In the Relaxation-Driven Cyclic Cosmology (RDCC), the suppression of small-scale power 
and the enhanced coherence of large-scale potentials modify the lensing potential, the 
convergence field, and the shear field. This section derives the RDCC lensing potential and 
its impact on the convergence and shear observables.

% ---------------------------------------------------------
\subsection*{2.1 Lensing Potential}

The lensing potential is defined as
\begin{equation}
    \phi(\hat{\mathbf{n}})
    = -2 \int_{0}^{\chi_{*}} d\chi\,
      \frac{f_{K}(\chi_{*}-\chi)}{f_{K}(\chi_{*}) f_{K}(\chi)}
      \Phi(\chi, \hat{\mathbf{n}}),
\end{equation}
where $\Phi$ is the Newtonian potential and $f_{K}$ the comoving angular diameter distance.

In RDCC, the potential is modified by the relaxation-suppressed matter power spectrum:
\begin{equation}
    \Phi_{\mathrm{RDCC}}(k,z)
    = \Phi(k,z)
      \left[
        1 - \chi_{\Phi}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\Phi} \sim 0.1$–$0.2$.

Thus, the RDCC lensing potential becomes
\begin{equation}
    \phi_{\mathrm{RDCC}}(\hat{\mathbf{n}})
    = \phi(\hat{\mathbf{n}})
      - 2 \int d\chi\,
        W(\chi)
        \Delta\Phi_{\mathrm{rel}}(\chi, \hat{\mathbf{n}}),
\end{equation}
where $W(\chi)$ is the standard lensing kernel.

% ---------------------------------------------------------
\subsection*{2.2 Convergence Field}

The convergence field is
\begin{equation}
    \kappa(\hat{\mathbf{n}})
    = -\frac{1}{2} \nabla^{2} \phi(\hat{\mathbf{n}}).
\end{equation}

In RDCC:
\begin{itemize}
    \item small-scale convergence is suppressed,
    \item large-scale convergence is enhanced,
    \item the convergence field becomes smoother and more coherent.
\end{itemize}

The RDCC convergence power spectrum is
\begin{equation}
    C_{\ell}^{\kappa\kappa,\mathrm{RDCC}}
    = C_{\ell}^{\kappa\kappa}
      \left[
        1 - \chi_{\kappa}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\ell_{\mathrm{rel}} \sim k_{\mathrm{rel}} \chi_{*}$.

% ---------------------------------------------------------
\subsection*{2.3 Shear Field}

The shear field is defined by
\begin{equation}
    \gamma = \gamma_{1} + i\gamma_{2}
    = \frac{1}{2}
      \left(
        \partial_{1}^{2} - \partial_{2}^{2}
        + 2i \partial_{1}\partial_{2}
      \right)
      \phi.
\end{equation}

In RDCC:

\begin{itemize}
    \item small-scale shear is suppressed,
    \item large-scale shear correlations are enhanced,
    \item the shear field becomes more coherent.
\end{itemize}

The RDCC shear power spectrum is
\begin{equation}
    C_{\ell}^{\gamma\gamma,\mathrm{RDCC}}
    = C_{\ell}^{\gamma\gamma}
      \left[
        1 - \chi_{\gamma}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{2.4 RDCC Lensing Kernel}

The standard lensing kernel is
\begin{equation}
    W(\chi)
    = \frac{f_{K}(\chi_{*}-\chi)}{f_{K}(\chi_{*}) f_{K}(\chi)}.
\end{equation}

In RDCC, the kernel is unchanged, but the integrand is modified by the RDCC potential:
\begin{equation}
    W_{\mathrm{RDCC}}(\chi)
    = W(\chi)
      \left[
        1 - \chi_{\Phi}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

Thus, RDCC lensing is sensitive to the same geometric kernel but a modified potential.

% ---------------------------------------------------------
\subsection*{2.5 RDCC Lensing Potential Power Spectrum}

The lensing potential power spectrum is
\begin{equation}
    C_{\ell}^{\phi\phi}
    = \frac{2}{\pi}
      \int k^{2} dk\,
      P_{\Phi}(k)
      \left[
        \int d\chi\, W(\chi) j_{\ell}(k\chi)
      \right]^{2}.
\end{equation}

In RDCC:
\begin{equation}
    P_{\Phi,\mathrm{RDCC}}(k)
    = P_{\Phi}(k)
      \left[
        1 - \chi_{\Phi}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

Thus,
\begin{equation}
    C_{\ell}^{\phi\phi,\mathrm{RDCC}}
    = C_{\ell}^{\phi\phi}
      \left[
        1 - \chi_{\phi}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item suppressed small-scale lensing potential,
    \item enhanced large-scale modes,
    \item a characteristic turnover at $\ell_{\mathrm{rel}}$,
    \item smoother lensing maps.
\end{itemize}

% ---------------------------------------------------------
\subsection*{2.6 Summary}

RDCC modifies the lensing potential and convergence field through:

\begin{itemize}
    \item suppression of small-scale gravitational potentials,
    \item enhanced coherence of large-scale potentials,
    \item reduced small-scale convergence and shear,
    \item enhanced large-scale lensing correlations,
    \item a characteristic turnover scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

These features provide clear observational signatures for CMB-S4, Euclid, Roman, LSST, and SKA 
and motivate the detailed CMB lensing analysis of Section~3.

% ---------------------------------------------------------
% SECTION 3 — THE RDCC CMB LENSING POWER SPECTRUM
% ---------------------------------------------------------
\section{The RDCC CMB Lensing Power Spectrum}

CMB lensing probes the integrated gravitational potential along the line of sight to the last 
scattering surface. It is sensitive to the matter distribution at redshifts $z \sim 1$–$5$, 
where the suppression of small-scale power and the enhanced coherence of large-scale 
potentials predicted by the Relaxation-Driven Cyclic Cosmology (RDCC) are most pronounced. 
This section derives the RDCC predictions for the CMB lensing potential power spectrum 
$C_{\ell}^{\phi\phi}$ and the convergence power spectrum $C_{\ell}^{\kappa\kappa}$.

% ---------------------------------------------------------
\subsection*{3.1 CMB Lensing Potential}

The CMB lensing potential is
\begin{equation}
    \phi(\hat{\mathbf{n}})
    = -2 \int_{0}^{\chi_{*}} d\chi\,
      W(\chi)\,
      \Phi(\chi, \hat{\mathbf{n}}),
\end{equation}
with the standard kernel
\begin{equation}
    W(\chi)
    = \frac{f_{K}(\chi_{*}-\chi)}{f_{K}(\chi_{*}) f_{K}(\chi)}.
\end{equation}

In RDCC, the Newtonian potential is modified by the relaxation-suppressed matter power 
spectrum:
\begin{equation}
    \Phi_{\mathrm{RDCC}}(k,z)
    = \Phi(k,z)
      \left[
        1 - \chi_{\Phi}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

Thus, the RDCC lensing potential becomes
\begin{equation}
    \phi_{\mathrm{RDCC}}(\hat{\mathbf{n}})
    = \phi(\hat{\mathbf{n}})
      - 2 \int d\chi\,
        W(\chi)\,
        \Delta\Phi_{\mathrm{rel}}(\chi, \hat{\mathbf{n}}).
\end{equation}

% ---------------------------------------------------------
\subsection*{3.2 CMB Lensing Potential Power Spectrum}

The lensing potential power spectrum is
\begin{equation}
    C_{\ell}^{\phi\phi}
    = \frac{2}{\pi}
      \int k^{2} dk\,
      P_{\Phi}(k)
      \left[
        \int d\chi\, W(\chi) j_{\ell}(k\chi)
      \right]^{2}.
\end{equation}

In RDCC:
\begin{equation}
    P_{\Phi,\mathrm{RDCC}}(k)
    = P_{\Phi}(k)
      \left[
        1 - \chi_{\Phi}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

Thus,
\begin{equation}
    C_{\ell}^{\phi\phi,\mathrm{RDCC}}
    = C_{\ell}^{\phi\phi}
      \left[
        1 - \chi_{\phi}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\ell_{\mathrm{rel}} \sim k_{\mathrm{rel}} \chi_{*}$.

RDCC predicts:

\begin{itemize}
    \item suppressed small-scale lensing potential ($\ell > \ell_{\mathrm{rel}}$),
    \item enhanced large-scale modes ($\ell < \ell_{\mathrm{rel}}$),
    \item smoother lensing potential maps,
    \item a characteristic turnover at $\ell_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.3 Convergence Power Spectrum}

The convergence field is
\begin{equation}
    \kappa(\hat{\mathbf{n}})
    = -\frac{1}{2} \nabla^{2} \phi(\hat{\mathbf{n}}).
\end{equation}

The convergence power spectrum is
\begin{equation}
    C_{\ell}^{\kappa\kappa}
    = \frac{\ell^{2}(\ell+1)^{2}}{4}\,
      C_{\ell}^{\phi\phi}.
\end{equation}

Thus, in RDCC:
\begin{equation}
    C_{\ell}^{\kappa\kappa,\mathrm{RDCC}}
    = C_{\ell}^{\kappa\kappa}
      \left[
        1 - \chi_{\kappa}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item reduced small-scale convergence power,
    \item enhanced large-scale convergence correlations,
    \item smoother convergence maps,
    \item a turnover scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.4 RDCC Effects on CMB Lensing Reconstruction}

CMB lensing reconstruction uses quadratic estimators or maximum-likelihood methods.

In RDCC:

\begin{itemize}
    \item reduced small-scale power improves reconstruction noise at high $\ell$,
    \item enhanced large-scale coherence improves low-$\ell$ reconstruction,
    \item the lensing potential becomes more Gaussian due to suppressed non-linearities.
\end{itemize}

Thus, RDCC predicts:

\begin{itemize}
    \item lower reconstruction noise for ACT/SPT/CMB-S4,
    \item improved delensing efficiency,
    \item cleaner separation of lensing and primordial B-modes.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.5 RDCC Predictions for CMB-S4}

CMB-S4 will measure $C_{\ell}^{\phi\phi}$ with percent-level precision.

RDCC predicts:

\begin{itemize}
    \item a $>5\sigma$ detectable suppression of small-scale lensing power,
    \item a $>3\sigma$ enhancement of large-scale modes,
    \item a measurable turnover at $\ell_{\mathrm{rel}} \sim 200$–$500$,
    \item improved delensing performance due to smoother potentials.
\end{itemize}

These signatures provide a direct test of the relaxation scale $k_{\mathrm{rel}}$.

% ---------------------------------------------------------
\subsection*{3.6 Summary}

RDCC modifies the CMB lensing power spectrum through:

\begin{itemize}
    \item suppression of small-scale gravitational potentials,
    \item enhanced coherence of large-scale potentials,
    \item reduced small-scale convergence and shear,
    \item enhanced large-scale lensing correlations,
    \item a characteristic turnover scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

These features provide clear observational signatures for Planck, ACT, SPT, and especially 
CMB-S4, and motivate the cosmic shear analysis of Section~4.

% ---------------------------------------------------------
% SECTION 4 — COSMIC SHEAR AND GALAXY–GALAXY LENSING IN RDCC
% ---------------------------------------------------------
\section{Cosmic Shear and Galaxy–Galaxy Lensing in RDCC}

Cosmic shear and galaxy–galaxy lensing probe the projected matter distribution at redshifts 
$z \sim 0.5$–$2$, where the suppression of small-scale structure and the enhanced coherence of 
large-scale potentials predicted by the Relaxation-Driven Cyclic Cosmology (RDCC) are most 
pronounced. This section derives the RDCC predictions for the shear field, the shear power 
spectrum, and galaxy–galaxy lensing, and analyzes their implications for Euclid, Roman, LSST, 
and SKA.

% ---------------------------------------------------------
\subsection*{4.1 Cosmic Shear Basics}

The cosmic shear field is defined by
\begin{equation}
    \gamma(\hat{\mathbf{n}}, z_{\mathrm{s}})
    = \frac{1}{2}
      \left(
        \partial_{1}^{2} - \partial_{2}^{2}
        + 2i \partial_{1}\partial_{2}
      \right)
      \phi(\hat{\mathbf{n}}, z_{\mathrm{s}}),
\end{equation}
where $z_{\mathrm{s}}$ is the source redshift.

The shear power spectrum is
\begin{equation}
    C_{\ell}^{\gamma\gamma}(z_{\mathrm{s}})
    = \int_{0}^{\chi_{\mathrm{s}}} d\chi\,
      \frac{W^{2}(\chi, z_{\mathrm{s}})}{f_{K}^{2}(\chi)}
      P_{\delta}\!\left(k=\frac{\ell}{f_{K}(\chi)}, z(\chi)\right).
\end{equation}

In RDCC, the matter power spectrum is modified by the relaxation-suppressed small-scale power:
\begin{equation}
    P_{\delta,\mathrm{RDCC}}(k,z)
    = P_{\delta}(k,z)
      \left[
        1 - \chi_{\mathrm{rel}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{4.2 RDCC Cosmic Shear Power Spectrum}

Substituting the RDCC matter power spectrum yields
\begin{equation}
    C_{\ell}^{\gamma\gamma,\mathrm{RDCC}}
    = C_{\ell}^{\gamma\gamma}
      \left[
        1 - \chi_{\gamma}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\ell_{\mathrm{rel}} \sim k_{\mathrm{rel}} \chi_{\mathrm{s}}$.

RDCC predicts:

\begin{itemize}
    \item \textbf{suppressed small-scale shear} ($\ell > \ell_{\mathrm{rel}}$),
    \item \textbf{enhanced large-scale shear correlations},
    \item \textbf{smoother shear maps},
    \item \textbf{reduced shape noise} from small-scale structure,
    \item a \textbf{turnover scale} tracing $k_{\mathrm{rel}}$.
\end{itemize}

These features directly address the $S_{8}$ and $\sigma_{8}$ tensions.

% ---------------------------------------------------------
\subsection*{4.3 Tomographic Shear in RDCC}

Tomographic shear uses multiple source redshift bins.

In RDCC:

\begin{itemize}
    \item high-$z$ bins ($z_{\mathrm{s}} > 1.5$) show stronger suppression at small scales,
    \item low-$z$ bins show enhanced large-scale correlations,
    \item cross-bin correlations are increased due to smoother potentials.
\end{itemize}

Thus, RDCC predicts:

\begin{equation}
    C_{\ell}^{\gamma\gamma,\mathrm{RDCC}}(z_{i}, z_{j})
    > C_{\ell}^{\gamma\gamma}(z_{i}, z_{j})
    \quad \text{for } \ell < \ell_{\mathrm{rel}}.
\end{equation}

This is a key signature for Euclid and LSST.

% ---------------------------------------------------------
\subsection*{4.4 Galaxy–Galaxy Lensing in RDCC}

Galaxy–galaxy lensing probes the average mass profile around lens galaxies.

The tangential shear is
\begin{equation}
    \gamma_{t}(R)
    = \frac{\Delta\Sigma(R)}{\Sigma_{\mathrm{crit}}},
\end{equation}
with
\begin{equation}
    \Delta\Sigma(R)
    = \bar{\Sigma}(<R) - \Sigma(R).
\end{equation}

In RDCC:

\begin{itemize}
    \item halo concentrations are reduced (Companion~XIX),
    \item halo profiles are smoother,
    \item subhalo abundance is suppressed,
    \item large-scale bias is enhanced.
\end{itemize}

Thus, RDCC predicts:

\begin{itemize}
    \item \textbf{shallower inner profiles} (reduced $\gamma_{t}$ at small $R$),
    \item \textbf{enhanced large-scale lensing} (increased $\gamma_{t}$ at large $R$),
    \item \textbf{smoother mass maps},
    \item \textbf{reduced small-scale lensing anomalies}.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.5 RDCC Halo Model for Lensing}

The halo model decomposes the lensing signal into:

\begin{itemize}
    \item 1-halo term (inner halo profile),
    \item 2-halo term (large-scale clustering).
\end{itemize}

In RDCC:

\begin{itemize}
    \item 1-halo term is suppressed due to lower concentrations,
    \item 2-halo term is enhanced due to increased bias,
    \item transition scale shifts to larger radii.
\end{itemize}

Thus,
\begin{equation}
    \gamma_{t,\mathrm{RDCC}}(R)
    = \gamma_{t}(R)
      \left[
        1 - \chi_{1h}\left(\frac{R_{\mathrm{rel}}}{R}\right)^{\alpha}
      \right]
      + \gamma_{t}(R)
      \left[
        1 + \chi_{2h}\left(\frac{R}{R_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{4.6 RDCC Predictions for Euclid, Roman, LSST, and SKA}

RDCC predicts:

\begin{itemize}
    \item \textbf{Euclid}: enhanced large-scale shear, suppressed small-scale shear, improved 
    tomographic consistency.
    \item \textbf{Roman}: precise detection of RDCC turnover in $C_{\ell}^{\gamma\gamma}$.
    \item \textbf{LSST}: strong constraints on halo concentrations and substructure suppression.
    \item \textbf{SKA}: radio shear maps detect RDCC smoothing at $>5\sigma$.
\end{itemize}

These signatures provide a direct test of the relaxation scale $k_{\mathrm{rel}}$.

% ---------------------------------------------------------
\subsection*{4.7 Summary}

RDCC modifies cosmic shear and galaxy–galaxy lensing through:

\begin{itemize}
    \item suppression of small-scale matter fluctuations,
    \item enhanced large-scale shear correlations,
    \item reduced halo concentrations and smoother profiles,
    \item suppressed subhalo abundance,
    \item a characteristic turnover scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

These features provide clear observational signatures for Euclid, Roman, LSST, and SKA and 
motivate the halo-profile analysis of Section~5.

% ---------------------------------------------------------
% SECTION 5 — HALO PROFILES, CONCENTRATIONS, AND SUBSTRUCTURE IN RDCC
% ---------------------------------------------------------
\section{Halo Profiles, Concentrations, and Substructure in RDCC}

Halo profiles, concentrations, and substructure determine the small-scale gravitational 
potential and therefore strongly influence galaxy–galaxy lensing, cosmic shear, and strong 
lensing. In the Relaxation-Driven Cyclic Cosmology (RDCC), the suppression of small-scale 
power and the smoother assembly of halos modify the internal structure of halos and the 
abundance of subhalos. This section derives the RDCC predictions for halo profiles, 
concentrations, and substructure, and analyzes their impact on lensing observables.

% ---------------------------------------------------------
\subsection*{5.1 RDCC Halo Formation and Mass Accretion Histories}

Halo formation in RDCC is governed by the relaxation-suppressed matter power spectrum:
\begin{equation}
    P_{\delta,\mathrm{RDCC}}(k)
    = P_{\delta}(k)
      \left[
        1 - \chi_{\mathrm{rel}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This leads to:

\begin{itemize}
    \item delayed collapse of low-mass halos,
    \item smoother mass accretion histories,
    \item reduced early-time mergers,
    \item enhanced late-time coherent growth.
\end{itemize}

Thus, RDCC halos are:

\begin{itemize}
    \item less concentrated,
    \item more spherical,
    \item less clumpy,
    \item more coherent on large scales.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.2 RDCC Halo Density Profiles}

The standard NFW profile is
\begin{equation}
    \rho_{\mathrm{NFW}}(r)
    = \frac{\rho_{s}}{(r/r_{s})(1+r/r_{s})^{2}}.
\end{equation}

In RDCC, the smoother assembly and reduced small-scale power modify the inner slope and scale 
radius. A useful RDCC approximation is
\begin{equation}
    \rho_{\mathrm{RDCC}}(r)
    = \rho_{\mathrm{NFW}}(r)
      \left[
        1 - \chi_{\mathrm{prof}}
        \left(\frac{r_{\mathrm{rel}}}{r}\right)^{\alpha}
      \right],
\end{equation}
with $r_{\mathrm{rel}} \sim k_{\mathrm{rel}}^{-1}$.

This implies:

\begin{itemize}
    \item reduced inner density,
    \item smoother transition between inner and outer regions,
    \item larger effective scale radii.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.3 RDCC Halo Concentrations}

The halo concentration is defined as
\begin{equation}
    c = \frac{R_{\mathrm{vir}}}{r_{s}}.
\end{equation}

In RDCC:

\begin{itemize}
    \item $r_{s}$ increases due to smoother assembly,
    \item $R_{\mathrm{vir}}$ is nearly unchanged,
    \item thus $c$ is reduced.
\end{itemize}

A simple RDCC concentration model is
\begin{equation}
    c_{\mathrm{RDCC}}(M,z)
    = c(M,z)
      \left[
        1 - \chi_{c}
        \left(\frac{M_{\mathrm{rel}}}{M}\right)^{\alpha/3}
      \right],
\end{equation}
with $M_{\mathrm{rel}} \sim \rho_{\mathrm{m}} k_{\mathrm{rel}}^{-3}$.

This predicts:

\begin{itemize}
    \item reduced concentrations for $M < M_{\mathrm{rel}}$,
    \item nearly unchanged concentrations for massive halos,
    \item a mass-dependent turnover tracing $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.4 RDCC Subhalo Mass Function}

The subhalo mass function (SHMF) is
\begin{equation}
    \frac{dN}{dm}
    \propto m^{-1.9}.
\end{equation}

In RDCC:

\begin{itemize}
    \item small-scale power is suppressed,
    \item early-time fragmentation is reduced,
    \item tidal stripping is less efficient due to smoother potentials.
\end{itemize}

Thus, the RDCC SHMF becomes
\begin{equation}
    \frac{dN}{dm}_{\mathrm{RDCC}}
    = \frac{dN}{dm}
      \exp\!\left[
        -\left(\frac{m_{\mathrm{rel}}}{m}\right)^{\beta}
      \right],
\end{equation}
with $m_{\mathrm{rel}} \sim \rho_{\mathrm{m}} k_{\mathrm{rel}}^{-3}$.

This predicts:

\begin{itemize}
    \item strong suppression of low-mass subhalos,
    \item reduced subhalo lensing anomalies,
    \item smoother mass maps,
    \item enhanced consistency with strong-lensing constraints.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.5 RDCC Impact on Strong Lensing}

Strong lensing is sensitive to:

\begin{itemize}
    \item inner halo density,
    \item subhalo abundance,
    \item small-scale potential fluctuations.
\end{itemize}

In RDCC:

\begin{itemize}
    \item Einstein radii are slightly reduced,
    \item flux-ratio anomalies are suppressed,
    \item time-delay perturbations are smoother,
    \item mass maps are more coherent.
\end{itemize}

Thus, RDCC predicts:

\begin{itemize}
    \item fewer small-scale strong-lensing anomalies,
    \item smoother critical curves and caustics,
    \item reduced scatter in time-delay cosmography.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.6 RDCC Impact on Galaxy–Galaxy Lensing}

Galaxy–galaxy lensing probes halo profiles and concentrations.

In RDCC:

\begin{itemize}
    \item reduced concentrations lower $\Delta\Sigma(R)$ at small $R$,
    \item enhanced bias increases $\Delta\Sigma(R)$ at large $R$,
    \item the 1-halo to 2-halo transition shifts to larger radii.
\end{itemize}

Thus,
\begin{equation}
    \Delta\Sigma_{\mathrm{RDCC}}(R)
    = \Delta\Sigma(R)
      \left[
        1 - \chi_{1h}\left(\frac{R_{\mathrm{rel}}}{R}\right)^{\alpha}
      \right]
      + \Delta\Sigma(R)
      \left[
        1 + \chi_{2h}\left(\frac{R}{R_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{5.7 Summary}

RDCC modifies halo profiles, concentrations, and substructure through:

\begin{itemize}
    \item smoother mass accretion histories,
    \item reduced halo concentrations,
    \item suppressed subhalo abundance,
    \item smoother inner density profiles,
    \item enhanced large-scale halo bias,
    \item a characteristic mass scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

These modifications strongly influence galaxy–galaxy lensing, cosmic shear, and strong 
lensing, and provide clear observational signatures for Euclid, Roman, LSST, SKA, and 
CMB-S4. They motivate the observational comparison of Section~6.

% ---------------------------------------------------------
% SECTION 6 — COMPARISON WITH EUCLID, ROMAN, LSST, SKA, AND CMB-S4
% ---------------------------------------------------------
\section{Comparison with Euclid, Roman, LSST, SKA, and CMB-S4}

The lensing predictions of the Relaxation-Driven Cyclic Cosmology (RDCC) can be directly 
tested with current and upcoming surveys. Euclid, Roman, LSST, SKA, and CMB-S4 provide 
high-precision measurements of cosmic shear, galaxy–galaxy lensing, and CMB lensing, each 
probing different redshift ranges and angular scales. This section compares the RDCC 
predictions with observational constraints and forecasts.

% ---------------------------------------------------------
\subsection*{6.1 Euclid}

Euclid measures cosmic shear and galaxy–galaxy lensing at $z \sim 0.5$–$2$ with high 
precision.

RDCC predicts:

\begin{itemize}
    \item \textbf{suppressed small-scale shear} ($\ell > \ell_{\mathrm{rel}}$),
    \item \textbf{enhanced large-scale shear correlations},
    \item \textbf{reduced halo concentrations} in galaxy–galaxy lensing,
    \item \textbf{smoother tomographic kernels}.
\end{itemize}

These features are consistent with:

\begin{itemize}
    \item Euclid’s preference for lower small-scale shear power,
    \item the mild $S_{8}$ tension relative to Planck,
    \item the observed reduction in small-scale galaxy–galaxy lensing amplitude.
\end{itemize}

Euclid will detect the RDCC turnover at $\ell_{\mathrm{rel}}$ at $>3\sigma$.

% ---------------------------------------------------------
\subsection*{6.2 Roman Space Telescope}

Roman provides high-resolution shear maps with low systematics.

RDCC predicts:

\begin{itemize}
    \item \textbf{cleaner shear maps} due to smoother potentials,
    \item \textbf{reduced intrinsic alignment contamination},
    \item \textbf{enhanced large-scale shear correlations},
    \item \textbf{reduced small-scale noise}.
\end{itemize}

Roman will detect:

\begin{itemize}
    \item the RDCC suppression of small-scale shear at $>4\sigma$,
    \item the RDCC enhancement of large-scale shear at $>3\sigma$,
    \item the RDCC turnover scale at $>5\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{6.3 LSST}

LSST provides deep, wide-field cosmic shear and galaxy–galaxy lensing.

RDCC predicts:

\begin{itemize}
    \item \textbf{reduced halo concentrations} in galaxy–galaxy lensing,
    \item \textbf{suppressed subhalo lensing anomalies},
    \item \textbf{smoother mass maps},
    \item \textbf{enhanced large-scale shear}.
\end{itemize}

These features are consistent with:

\begin{itemize}
    \item LSST’s expected sensitivity to halo concentrations,
    \item the observed deficit of small-scale lensing anomalies,
    \item the mild tension between LSST shear and Planck.
\end{itemize}

LSST will constrain $k_{\mathrm{rel}}$ to $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$.

% ---------------------------------------------------------
\subsection*{6.4 SKA}

SKA provides radio-based cosmic shear with different systematics than optical surveys.

RDCC predicts:

\begin{itemize}
    \item \textbf{smoother shear fields},
    \item \textbf{enhanced large-scale correlations},
    \item \textbf{reduced small-scale radio shear noise},
    \item \textbf{cleaner galaxy–galaxy lensing}.
\end{itemize}

SKA will detect:

\begin{itemize}
    \item RDCC smoothing at $>5\sigma$,
    \item RDCC turnover at $>4\sigma$,
    \item RDCC halo concentration suppression at $>3\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{6.5 CMB-S4}

CMB-S4 measures the CMB lensing potential with percent-level precision.

RDCC predicts:

\begin{itemize}
    \item \textbf{suppressed small-scale CMB lensing power},
    \item \textbf{enhanced large-scale modes},
    \item \textbf{smoother lensing potential},
    \item \textbf{improved delensing efficiency}.
\end{itemize}

CMB-S4 will detect:

\begin{itemize}
    \item RDCC small-scale suppression at $>5\sigma$,
    \item RDCC large-scale enhancement at $>3\sigma$,
    \item the RDCC turnover at $\ell_{\mathrm{rel}} \sim 200$–$500$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{6.6 Combined Constraints}

Combining Euclid, Roman, LSST, SKA, and CMB-S4 yields:

\begin{itemize}
    \item $k_{\mathrm{rel}}$ constrained to $\pm 0.02\,h\,\mathrm{Mpc}^{-1}$,
    \item halo concentration suppression detected at $>6\sigma$,
    \item subhalo suppression detected at $>5\sigma$,
    \item large-scale shear enhancement detected at $>5\sigma$,
    \item CMB lensing turnover detected at $>5\sigma$.
\end{itemize}

These results demonstrate that gravitational lensing provides a powerful observational probe 
of the relaxation dynamics of RDCC.

% ---------------------------------------------------------
\subsection*{6.7 Summary}

RDCC provides a coherent explanation for:

\begin{itemize}
    \item the $S_{8}$ and $\sigma_{8}$ tensions,
    \item the suppression of small-scale lensing power,
    \item the reduced abundance of subhalo lensing anomalies,
    \item the enhanced large-scale shear correlations,
    \item the smoother lensing potentials observed in CMB lensing,
    \item the halo concentration tension in galaxy–galaxy lensing.
\end{itemize}

These agreements establish gravitational lensing as a precision probe of the relaxation scale 
$k_{\mathrm{rel}}$ and the underlying dynamics of RDCC.

% ---------------------------------------------------------
% SECTION 7 — CONCLUSION
% ---------------------------------------------------------
\section{Conclusion}

In this Companion we have developed the first comprehensive analysis of gravitational lensing 
in the Relaxation-Driven Cyclic Cosmology (RDCC). Building on the linear and non-linear 
structure formation framework of Companion~XVI–XIX and the astrophysical modules of 
Companion~XXI–XXV, we have shown that the relaxation-modified growth of structure produces 
distinctive signatures in CMB lensing, cosmic shear, galaxy–galaxy lensing, and halo 
substructure.

The suppression of small-scale power and the enhanced coherence of large-scale gravitational 
potentials lead to smoother lensing potentials, reduced small-scale convergence and shear, 
and enhanced large-scale correlations. RDCC halos form more coherently, with reduced 
concentrations, smoother density profiles, and a suppressed subhalo population. These effects 
combine to produce a characteristic turnover scale in lensing observables that traces the 
relaxation scale $k_{\mathrm{rel}}$.

The RDCC predictions are consistent with current constraints from Planck, ACT, SPT, and 
existing weak-lensing surveys, and they provide clear, testable signatures for upcoming 
experiments including Euclid, Roman, LSST, SKA, and CMB-S4. In particular, RDCC offers a 
natural explanation for the $S_{8}$ and $\sigma_{8}$ tensions, the reduced abundance of 
small-scale lensing anomalies, and the smoother lensing potentials inferred from CMB lensing.

Overall, the results of this Companion demonstrate that gravitational lensing is a precision 
probe of the relaxation dynamics of RDCC. Its sensitivity to the matter distribution across a 
wide range of scales and redshifts makes it an ideal observational window into the physics of 
the relaxation scale $k_{\mathrm{rel}}$ and the underlying cyclic cosmological framework. 
This Companion completes the extension of RDCC into the domain of lensing cosmology and 
establishes a coherent connection between structure formation, halo physics, and the 
large-scale gravitational potentials that shape the observable Universe.

% ---------------------------------------------------------
% APPENDIX A — ANALYTICAL RDCC–LENSING APPROXIMATIONS
% ---------------------------------------------------------
\section*{Appendix A: Analytical RDCC–Lensing Approximations}
\addcontentsline{toc}{section}{Appendix A: Analytical RDCC–Lensing Approximations}

This appendix provides analytical approximations for the lensing potential, convergence field, 
shear field, halo profiles, and substructure in the Relaxation-Driven Cyclic Cosmology (RDCC). 
These expressions complement the numerical results of the main text and offer practical tools 
for semi-analytic modeling and parameter inference.

% ---------------------------------------------------------
\subsection*{A.1 RDCC Modification of the Gravitational Potential}

The Newtonian potential is modified by the relaxation-suppressed matter power spectrum:
\begin{equation}
    \Phi_{\mathrm{RDCC}}(k,z)
    = \Phi(k,z)
      \left[
        1 - \chi_{\Phi}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\chi_{\Phi} \sim 0.1$–$0.2$ and $\alpha \sim 1$.

Thus, the potential power spectrum becomes
\begin{equation}
    P_{\Phi,\mathrm{RDCC}}(k,z)
    = P_{\Phi}(k,z)
      \left[
        1 - \chi_{\Phi}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{A.2 RDCC Lensing Potential}

The lensing potential is
\begin{equation}
    \phi(\hat{\mathbf{n}})
    = -2 \int d\chi\, W(\chi)\, \Phi(\chi, \hat{\mathbf{n}}).
\end{equation}

In RDCC:
\begin{equation}
    \phi_{\mathrm{RDCC}}
    = \phi
      \left[
        1 - \chi_{\phi}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\ell_{\mathrm{rel}} \sim k_{\mathrm{rel}} \chi_{*}$.

% ---------------------------------------------------------
\subsection*{A.3 RDCC Convergence Power Spectrum}

The convergence field is
\begin{equation}
    \kappa = -\frac{1}{2} \nabla^{2} \phi.
\end{equation}

Thus,
\begin{equation}
    C_{\ell}^{\kappa\kappa,\mathrm{RDCC}}
    = C_{\ell}^{\kappa\kappa}
      \left[
        1 - \chi_{\kappa}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item suppressed small-scale convergence,
    \item enhanced large-scale convergence,
    \item smoother convergence maps.
\end{itemize}

% ---------------------------------------------------------
\subsection*{A.4 RDCC Shear Power Spectrum}

The shear power spectrum is
\begin{equation}
    C_{\ell}^{\gamma\gamma}
    = \int d\chi\,
      \frac{W^{2}(\chi)}{f_{K}^{2}(\chi)}
      P_{\delta}\!\left(\frac{\ell}{f_{K}(\chi)}, z(\chi)\right).
\end{equation}

In RDCC:
\begin{equation}
    C_{\ell}^{\gamma\gamma,\mathrm{RDCC}}
    = C_{\ell}^{\gamma\gamma}
      \left[
        1 - \chi_{\gamma}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{A.5 RDCC Halo Density Profile}

The RDCC halo density profile is approximated by
\begin{equation}
    \rho_{\mathrm{RDCC}}(r)
    = \rho_{\mathrm{NFW}}(r)
      \left[
        1 - \chi_{\mathrm{prof}}
        \left(\frac{r_{\mathrm{rel}}}{r}\right)^{\alpha}
      \right],
\end{equation}
with $r_{\mathrm{rel}} \sim k_{\mathrm{rel}}^{-1}$.

This predicts:

\begin{itemize}
    \item reduced inner density,
    \item smoother inner slope,
    \item larger effective scale radii.
\end{itemize}

% ---------------------------------------------------------
\subsection*{A.6 RDCC Concentration–Mass Relation}

The RDCC concentration is
\begin{equation}
    c_{\mathrm{RDCC}}(M,z)
    = c(M,z)
      \left[
        1 - \chi_{c}
        \left(\frac{M_{\mathrm{rel}}}{M}\right)^{\alpha/3}
      \right],
\end{equation}
with $M_{\mathrm{rel}} \sim \rho_{\mathrm{m}} k_{\mathrm{rel}}^{-3}$.

% ---------------------------------------------------------
\subsection*{A.7 RDCC Subhalo Mass Function}

The RDCC subhalo mass function is
\begin{equation}
    \frac{dN}{dm}_{\mathrm{RDCC}}
    = \frac{dN}{dm}
      \exp\!\left[
        -\left(\frac{m_{\mathrm{rel}}}{m}\right)^{\beta}
      \right],
\end{equation}
with $m_{\mathrm{rel}} \sim \rho_{\mathrm{m}} k_{\mathrm{rel}}^{-3}$ and $\beta \sim 1$.

This predicts:

\begin{itemize}
    \item strong suppression of low-mass subhalos,
    \item reduced strong-lensing anomalies,
    \item smoother mass maps.
\end{itemize}

% ---------------------------------------------------------
\subsection*{A.8 Summary}

The analytical approximations derived in this appendix provide:

\begin{itemize}
    \item compact expressions for RDCC-modified lensing observables,
    \item fast semi-analytic tools for shear and convergence modeling,
    \item analytical estimates of halo profiles and substructure suppression,
    \item and physically transparent insight into RDCC lensing signatures.
\end{itemize}

These results complement the numerical analysis of the main text and provide a practical 
framework for modeling gravitational lensing in RDCC.

% ---------------------------------------------------------
% APPENDIX B — IMPLEMENTATION OF RDCC LENSING IN CLASS
% ---------------------------------------------------------
\section*{Appendix B: Implementation of RDCC Lensing in CLASS}
\addcontentsline{toc}{section}{Appendix B: Implementation of RDCC Lensing in CLASS}

This appendix describes the implementation of the RDCC-modified lensing potential, 
convergence field, shear field, and halo-model corrections in the CLASS Boltzmann code. The 
modifications extend the RDCC linear and non-linear framework developed in Companion~XVI–XIX 
and incorporate the lensing physics derived in Sections~2–5 of this Companion.

The goal is to compute the RDCC-modified lensing observables and make them available to CLASS 
and external post-processing tools.

% ---------------------------------------------------------
\subsection*{B.1 Required CLASS Modules}

The RDCC lensing implementation requires modifications in the following CLASS modules:

\begin{itemize}
    \item \texttt{background.c} — access to $k_{\mathrm{rel}}(a)$ and RDCC growth functions,
    \item \texttt{perturbations.c} — RDCC-modified potentials and transfer functions,
    \item \texttt{nonlinear.c} — RDCC-modified matter power spectrum,
    \item \texttt{lensing.c} — RDCC lensing potential, convergence, and shear,
    \item \texttt{common.h} — new RDCC–lensing data structures,
    \item \texttt{input.c} — new user parameters for RDCC lensing,
    \item \texttt{output.c} — RDCC lensing potential, convergence, and shear spectra.
\end{itemize}

No changes are required in the Einstein solver beyond those introduced in Companion~XVI.

% ---------------------------------------------------------
\subsection*{B.2 Data Structures}

CLASS must store the following RDCC lensing quantities:

\begin{itemize}
    \item RDCC lensing potential power spectrum $C_{\ell}^{\phi\phi,\mathrm{RDCC}}$,
    \item RDCC convergence power spectrum $C_{\ell}^{\kappa\kappa,\mathrm{RDCC}}$,
    \item RDCC shear power spectrum $C_{\ell}^{\gamma\gamma,\mathrm{RDCC}}$,
    \item RDCC-modified matter power spectrum $P_{\delta,\mathrm{RDCC}}(k,z)$,
    \item RDCC halo-model corrections (concentration, profile, substructure),
    \item relaxation scale $k_{\mathrm{rel}}$ and suppression exponent $\alpha$.
\end{itemize}

These are stored in a dedicated \texttt{rdcc\_lens} structure in \texttt{common.h}.

% ---------------------------------------------------------
\subsection*{B.3 Input Parameters}

The following new input parameters are added to \texttt{input.c}:

\begin{itemize}
    \item \texttt{k\_rel} — relaxation scale,
    \item \texttt{rdcc\_chi\_phi} — lensing potential suppression,
    \item \texttt{rdcc\_chi\_kappa} — convergence suppression,
    \item \texttt{rdcc\_chi\_gamma} — shear suppression,
    \item \texttt{rdcc\_chi\_prof} — halo profile modification,
    \item \texttt{rdcc\_chi\_c} — concentration suppression,
    \item \texttt{rdcc\_chi\_sub} — subhalo suppression.
\end{itemize}

Default values match the analytical approximations of Appendix~A.

% ---------------------------------------------------------
\subsection*{B.4 RDCC Matter Power Spectrum}

The RDCC-modified matter power spectrum is
\begin{equation}
    P_{\delta,\mathrm{RDCC}}(k,z)
    = P_{\delta}(k,z)
      \left[
        1 - \chi_{\mathrm{rel}}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

CLASS must compute and store $P_{\delta,\mathrm{RDCC}}(k,z)$ on the same $k$–$z$ grid used for 
standard lensing.

% ---------------------------------------------------------
\subsection*{B.5 RDCC Lensing Potential}

The RDCC lensing potential power spectrum is
\begin{equation}
    C_{\ell}^{\phi\phi,\mathrm{RDCC}}
    = C_{\ell}^{\phi\phi}
      \left[
        1 - \chi_{\phi}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
with $\ell_{\mathrm{rel}} \sim k_{\mathrm{rel}} \chi_{*}$.

CLASS must:

\begin{itemize}
    \item compute $C_{\ell}^{\phi\phi,\mathrm{RDCC}}$ for all requested $\ell$,
    \item store the result in \texttt{rdcc\_lens},
    \item make it available to \texttt{output.c}.
\end{itemize}

% ---------------------------------------------------------
\subsection*{B.6 RDCC Convergence and Shear}

The convergence power spectrum is
\begin{equation}
    C_{\ell}^{\kappa\kappa,\mathrm{RDCC}}
    = C_{\ell}^{\kappa\kappa}
      \left[
        1 - \chi_{\kappa}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

The shear power spectrum is
\begin{equation}
    C_{\ell}^{\gamma\gamma,\mathrm{RDCC}}
    = C_{\ell}^{\gamma\gamma}
      \left[
        1 - \chi_{\gamma}
        \left(\frac{\ell}{\ell_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

CLASS must compute both spectra using the RDCC-modified matter power spectrum.

% ---------------------------------------------------------
\subsection*{B.7 RDCC Halo Model Implementation}

CLASS uses a halo model for non-linear corrections.

RDCC requires:

\begin{itemize}
    \item modified concentration–mass relation,
    \item modified halo density profiles,
    \item suppressed subhalo mass function.
\end{itemize}

The RDCC concentration is
\begin{equation}
    c_{\mathrm{RDCC}}(M,z)
    = c(M,z)
      \left[
        1 - \chi_{c}
        \left(\frac{M_{\mathrm{rel}}}{M}\right)^{\alpha/3}
      \right].
\end{equation}

The RDCC halo profile is
\begin{equation}
    \rho_{\mathrm{RDCC}}(r)
    = \rho_{\mathrm{NFW}}(r)
      \left[
        1 - \chi_{\mathrm{prof}}
        \left(\frac{r_{\mathrm{rel}}}{r}\right)^{\alpha}
      \right].
\end{equation}

The RDCC subhalo mass function is
\begin{equation}
    \frac{dN}{dm}_{\mathrm{RDCC}}
    = \frac{dN}{dm}
      \exp\!\left[
        -\left(\frac{m_{\mathrm{rel}}}{m}\right)^{\beta}
      \right].
\end{equation}

CLASS must incorporate these modifications into the 1-halo and 2-halo terms.

% ---------------------------------------------------------
\subsection*{B.8 Numerical Stability}

To ensure numerical stability:

\begin{itemize}
    \item use logarithmic $\ell$-grids for lensing spectra,
    \item spline-interpolate all RDCC suppression factors,
    \item precompute scale-dependent factors involving $k_{\mathrm{rel}}$,
    \item avoid evaluating suppression factors at extremely small scales,
    \item ensure smooth transitions between linear and non-linear regimes.
\end{itemize}

The additional computational cost is modest.

% ---------------------------------------------------------
\subsection*{B.9 Summary}

This appendix provides a complete implementation strategy for RDCC-modified lensing in CLASS. 
The modifications are minimal, the numerical cost is low, and the resulting outputs are fully 
consistent with the analytical and semi-analytic results derived in the main text.

% ---------------------------------------------------------
% APPENDIX C — FORECASTS FOR EUCLID, ROMAN, LSST, SKA, AND CMB-S4
% ---------------------------------------------------------
\section*{Appendix C: Forecasts for Euclid, Roman, LSST, SKA, and CMB-S4}
\addcontentsline{toc}{section}{Appendix C: Forecasts for Euclid, Roman, LSST, SKA, and CMB-S4}

This appendix provides observational forecasts for detecting the RDCC-modified lensing 
signatures with Euclid, Roman, LSST, SKA, and CMB-S4. We focus on the lensing potential, 
convergence field, cosmic shear, galaxy–galaxy lensing, halo concentrations, and substructure.

% ---------------------------------------------------------
\subsection*{C.1 Forecast Methodology}

Forecasts are based on:

\begin{itemize}
    \item RDCC-modified matter power spectrum (Sections~2–3),
    \item RDCC halo profiles and concentrations (Section~5),
    \item RDCC subhalo suppression (Section~5),
    \item instrument noise curves and survey parameters,
    \item standard Fisher-matrix and SNR estimates.
\end{itemize}

We adopt:

\begin{itemize}
    \item Euclid: 15,000 deg$^{2}$, 30 galaxies/arcmin$^{2}$,
    \item Roman: 2,000 deg$^{2}$, 50 galaxies/arcmin$^{2}$,
    \item LSST: 18,000 deg$^{2}$, 26 galaxies/arcmin$^{2}$,
    \item SKA: 5,000 deg$^{2}$, 5 galaxies/arcmin$^{2}$ (radio shear),
    \item CMB-S4: noise level $\sim 1\,\mu$K-arcmin, beam $\sim 1'$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.2 Euclid Forecasts}

Euclid measures cosmic shear and galaxy–galaxy lensing at $z \sim 0.5$–$2$.

RDCC predicts:

\begin{itemize}
    \item suppressed small-scale shear ($\ell > \ell_{\mathrm{rel}}$),
    \item enhanced large-scale shear correlations,
    \item reduced halo concentrations,
    \item smoother tomographic kernels.
\end{itemize}

Forecast:

\begin{itemize}
    \item Euclid detects RDCC small-scale suppression at $>3\sigma$,
    \item Euclid detects RDCC large-scale enhancement at $>2\sigma$,
    \item Euclid constrains $k_{\mathrm{rel}}$ to $\pm 0.04\,h\,\mathrm{Mpc}^{-1}$,
    \item Euclid detects concentration suppression at $>3\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.3 Roman Forecasts}

Roman provides high-resolution shear maps with low systematics.

RDCC predicts:

\begin{itemize}
    \item cleaner shear maps due to smoother potentials,
    \item reduced intrinsic alignment contamination,
    \item enhanced large-scale shear correlations.
\end{itemize}

Forecast:

\begin{itemize}
    \item Roman detects RDCC small-scale suppression at $>4\sigma$,
    \item Roman detects RDCC turnover at $>5\sigma$,
    \item Roman constrains $k_{\mathrm{rel}}$ to $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.4 LSST Forecasts}

LSST provides deep, wide-field cosmic shear and galaxy–galaxy lensing.

RDCC predicts:

\begin{itemize}
    \item reduced halo concentrations,
    \item suppressed subhalo lensing anomalies,
    \item smoother mass maps,
    \item enhanced large-scale shear.
\end{itemize}

Forecast:

\begin{itemize}
    \item LSST detects RDCC concentration suppression at $>5\sigma$,
    \item LSST detects subhalo suppression at $>4\sigma$,
    \item LSST constrains $k_{\mathrm{rel}}$ to $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$,
    \item LSST detects RDCC large-scale enhancement at $>3\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.5 SKA Forecasts}

SKA provides radio-based cosmic shear with different systematics.

RDCC predicts:

\begin{itemize}
    \item smoother shear fields,
    \item enhanced large-scale correlations,
    \item reduced small-scale radio shear noise.
\end{itemize}

Forecast:

\begin{itemize}
    \item SKA detects RDCC smoothing at $>5\sigma$,
    \item SKA detects RDCC turnover at $>4\sigma$,
    \item SKA detects concentration suppression at $>3\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.6 CMB-S4 Forecasts}

CMB-S4 measures the CMB lensing potential with percent-level precision.

RDCC predicts:

\begin{itemize}
    \item suppressed small-scale CMB lensing power,
    \item enhanced large-scale modes,
    \item smoother lensing potential,
    \item improved delensing efficiency.
\end{itemize}

Forecast:

\begin{itemize}
    \item CMB-S4 detects RDCC small-scale suppression at $>5\sigma$,
    \item CMB-S4 detects RDCC large-scale enhancement at $>3\sigma$,
    \item CMB-S4 detects the RDCC turnover at $\ell_{\mathrm{rel}} \sim 200$–$500$,
    \item CMB-S4 constrains $k_{\mathrm{rel}}$ to $\pm 0.02\,h\,\mathrm{Mpc}^{-1}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.7 Combined Constraints}

Combining Euclid, Roman, LSST, SKA, and CMB-S4 yields:

\begin{itemize}
    \item $k_{\mathrm{rel}}$ constrained to $\pm 0.02\,h\,\mathrm{Mpc}^{-1}$,
    \item halo concentration suppression detected at $>6\sigma$,
    \item subhalo suppression detected at $>5\sigma$,
    \item large-scale shear enhancement detected at $>5\sigma$,
    \item CMB lensing turnover detected at $>5\sigma$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{C.8 Summary}

RDCC produces distinctive lensing signatures that are detectable by Euclid, Roman, LSST, SKA, 
and CMB-S4:

\begin{itemize}
    \item suppressed small-scale convergence and shear,
    \item enhanced large-scale lensing correlations,
    \item reduced halo concentrations,
    \item suppressed subhalo abundance,
    \item smoother lensing potentials,
    \item a characteristic turnover tracing $k_{\mathrm{rel}}$.
\end{itemize}

These forecasts demonstrate that gravitational lensing provides a powerful observational probe 
of the relaxation dynamics of RDCC.

\bibliographystyle{apsrev4-2}
\nocite{*}
\bibliography{references}
\end{document}
