% ---------------------------------------------------------
% PREAMBLE (same template as Companion X–XXVI)
% ---------------------------------------------------------
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{physics}
\usepackage{cite}
\usepackage{float}

\geometry{margin=1in}

% ---------------------------------------------------------
% TITLE
% ---------------------------------------------------------
\title{
\textbf{Companion XXVII}\\[4mm]
\textbf{Redshift-Space Distortions and Peculiar Velocities}\\
\textbf{in the Relaxation-Driven Cyclic Cosmology}
}

\author{
Michael Lehmann\\
Independent Researcher\\
\texttt{mi.lehmann@gmx.de}
}

\date{April 2026}

\begin{document}

\maketitle

% ---------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------
\begin{abstract}
Redshift-space distortions (RSD) and peculiar velocities provide a powerful probe of the 
growth of structure, the coherence of large-scale flows, and the small-scale velocity 
dispersion of galaxies and halos. Modern spectroscopic surveys such as DESI, Euclid, Roman, 
4MOST, PFS, and SKA deliver high-precision measurements of the anisotropic galaxy power 
spectrum and correlation function, enabling percent-level constraints on the growth rate 
$f\sigma_{8}$.

In the Relaxation-Driven Cyclic Cosmology (RDCC), the suppression of small-scale power, the 
enhanced coherence of large-scale gravitational potentials, and the scale-dependent 
relaxation rate $\Gamma(a)$ modify the peculiar velocity field and the redshift-space galaxy 
distribution. RDCC predicts smoother velocity fields, reduced small-scale velocity 
dispersion, enhanced large-scale flows, and a characteristic turnover scale tracing the 
relaxation scale $k_{\mathrm{rel}}$.

This Companion develops the RDCC predictions for RSD, the velocity divergence power spectrum, 
the anisotropic galaxy power spectrum, and the multipole moments of the redshift-space 
correlation function. We show that RDCC produces distinctive signatures in the monopole, 
quadrupole, and hexadecapole moments, and we provide forecasts for DESI, Euclid, Roman, and 
SKA. These results establish peculiar velocities and RSD as precision probes of the 
relaxation dynamics of RDCC.
\end{abstract}

% ---------------------------------------------------------
% SECTION 1 — MOTIVATION AND OVERVIEW
% ---------------------------------------------------------
\section{Motivation and Overview}

Redshift-space distortions (RSD) and peculiar velocities provide a direct probe of the growth 
of structure in the Universe. Unlike gravitational lensing, which measures the projected 
matter distribution, RSD are sensitive to the velocity field and therefore to the dynamical 
evolution of density perturbations. Modern spectroscopic surveys such as DESI, Euclid, Roman, 
4MOST, PFS, and SKA deliver high-precision measurements of the anisotropic galaxy power 
spectrum and correlation function, enabling percent-level constraints on the growth rate 
$f\sigma_{8}$.

In the Relaxation-Driven Cyclic Cosmology (RDCC), the suppression of small-scale power, the 
enhanced coherence of large-scale gravitational potentials, and the scale-dependent 
relaxation rate $\Gamma(a)$ modify the peculiar velocity field and the redshift-space galaxy 
distribution. RDCC predicts smoother velocity fields, reduced small-scale velocity 
dispersion, enhanced large-scale flows, and a characteristic turnover scale tracing the 
relaxation scale $k_{\mathrm{rel}}$.

This Companion develops the RDCC predictions for RSD, the velocity divergence power spectrum, 
the anisotropic galaxy power spectrum, and the multipole moments of the redshift-space 
correlation function. We show that RDCC produces distinctive signatures in the monopole, 
quadrupole, and hexadecapole moments, and we provide forecasts for DESI, Euclid, Roman, and 
SKA.

% ---------------------------------------------------------
\subsection*{1.1 Observational Context}

Redshift-space distortions arise because galaxy redshifts contain contributions from both the 
Hubble flow and peculiar velocities. These velocities distort the observed clustering pattern 
in redshift space, producing:

\begin{itemize}
    \item \textbf{Kaiser effect} — large-scale coherent infall enhances clustering along the 
    line of sight.
    \item \textbf{Fingers of God (FoG)} — small-scale random motions elongate structures in 
    redshift space.
    \item \textbf{Anisotropic power spectrum} — the galaxy power spectrum becomes a function 
    of both $k$ and $\mu$.
\end{itemize}

Modern surveys measure:

\begin{itemize}
    \item the anisotropic galaxy power spectrum $P(k,\mu)$,
    \item the multipole moments $P_{\ell}(k)$,
    \item the redshift-space correlation function $\xi(s,\mu)$,
    \item the growth rate $f\sigma_{8}$,
    \item the velocity divergence power spectrum $P_{\theta\theta}(k)$.
\end{itemize}

These observables are highly sensitive to the growth of structure and the velocity field.

% ---------------------------------------------------------
\subsection*{1.2 RDCC Perspective}

The Relaxation-Driven Cyclic Cosmology modifies structure formation through the 
scale-dependent relaxation rate $\Gamma(a)$, which suppresses small-scale power and enhances 
large-scale coherence. As shown in Companion~XVI–XIX, this leads to:

\begin{itemize}
    \item reduced small-scale clustering,
    \item smoother gravitational potentials,
    \item enhanced large-scale bias,
    \item suppressed small-scale velocity dispersion,
    \item enhanced large-scale coherent flows.
\end{itemize}

These effects directly influence RSD observables:

\begin{itemize}
    \item \textbf{Kaiser effect:} enhanced large-scale infall due to smoother potentials.
    \item \textbf{FoG effect:} reduced small-scale velocity dispersion.
    \item \textbf{Multipoles:} suppressed small-scale quadrupole and hexadecapole.
    \item \textbf{Velocity divergence:} smoother $P_{\theta\theta}(k)$ with a turnover at 
    $k_{\mathrm{rel}}$.
\end{itemize}

Thus, RSD provide a direct observational probe of the relaxation scale $k_{\mathrm{rel}}$.

% ---------------------------------------------------------
\subsection*{1.3 Goals of This Companion}

The goals of this Companion are:

\begin{itemize}
    \item to derive the RDCC predictions for the peculiar velocity field,
    \item to compute the RDCC velocity divergence power spectrum,
    \item to analyze the RDCC anisotropic galaxy power spectrum,
    \item to derive the RDCC multipole moments of the correlation function,
    \item and to compare RDCC predictions with DESI, Euclid, Roman, and SKA.
\end{itemize}

These results build directly on the RDCC structure formation framework developed in 
Companion~XVI–XIX.

% ---------------------------------------------------------
\subsection*{1.4 Structure of This Companion}

The structure of this Companion is as follows:

\begin{itemize}
    \item Section~2 derives the RDCC peculiar velocity field and velocity divergence spectrum.
    \item Section~3 computes the RDCC anisotropic galaxy power spectrum.
    \item Section~4 analyzes the RDCC multipole moments of the correlation function.
    \item Section~5 develops the RDCC FoG model and small-scale velocity dispersion.
    \item Section~6 compares RDCC predictions with DESI, Euclid, Roman, and SKA.
\end{itemize}

Together, these results establish RSD as a precision probe of the relaxation dynamics of RDCC.

% ---------------------------------------------------------
% SECTION 2 — THE RDCC PECULIAR VELOCITY FIELD AND VELOCITY DIVERGENCE SPECTRUM
% ---------------------------------------------------------
\section{The RDCC Peculiar Velocity Field and Velocity Divergence Spectrum}

Redshift-space distortions arise from peculiar velocities, which trace the dynamical response 
of matter to gravitational potentials. In the Relaxation-Driven Cyclic Cosmology (RDCC), the 
suppression of small-scale power and the enhanced coherence of large-scale potentials modify 
the peculiar velocity field, the velocity divergence, and the redshift-space galaxy 
distribution. This section derives the RDCC peculiar velocity field and the velocity 
divergence power spectrum.

% ---------------------------------------------------------
\subsection*{2.1 Peculiar Velocities in RDCC}

The peculiar velocity field is defined by
\begin{equation}
    \mathbf{v}(\mathbf{x},a)
    = aHf(a)\,\nabla^{-1}\delta(\mathbf{x},a),
\end{equation}
where $f(a)$ is the growth rate.

In RDCC, the growth rate becomes scale-dependent:
\begin{equation}
    f_{\mathrm{RDCC}}(k,a)
    = f(a)
      \left[
        1 - \chi_{f}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

Thus, the RDCC peculiar velocity field is
\begin{equation}
    \mathbf{v}_{\mathrm{RDCC}}(\mathbf{k},a)
    = \mathbf{v}(\mathbf{k},a)
      \left[
        1 - \chi_{f}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item suppressed small-scale velocities,
    \item enhanced large-scale coherent flows,
    \item smoother velocity fields.
\end{itemize}

% ---------------------------------------------------------
\subsection*{2.2 Velocity Divergence Field}

The velocity divergence is defined as
\begin{equation}
    \theta(\mathbf{x},a)
    = \nabla \cdot \mathbf{v}(\mathbf{x},a).
\end{equation}

In Fourier space:
\begin{equation}
    \theta(\mathbf{k},a)
    = -aHf(a)\,\delta(\mathbf{k},a).
\end{equation}

In RDCC:
\begin{equation}
    \theta_{\mathrm{RDCC}}(\mathbf{k},a)
    = -aHf_{\mathrm{RDCC}}(k,a)\,\delta(\mathbf{k},a).
\end{equation}

Thus,
\begin{equation}
    \theta_{\mathrm{RDCC}}(\mathbf{k},a)
    = \theta(\mathbf{k},a)
      \left[
        1 - \chi_{f}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{2.3 Velocity Divergence Power Spectrum}

The velocity divergence power spectrum is
\begin{equation}
    P_{\theta\theta}(k,a)
    = (aH)^{2} f^{2}(a)\, P_{\delta\delta}(k,a).
\end{equation}

In RDCC:
\begin{equation}
    P_{\theta\theta,\mathrm{RDCC}}(k,a)
    = P_{\theta\theta}(k,a)
      \left[
        1 - \chi_{\theta}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item suppressed small-scale velocity divergence,
    \item enhanced large-scale divergence,
    \item a turnover at $k_{\mathrm{rel}}$,
    \item smoother velocity divergence maps.
\end{itemize}

% ---------------------------------------------------------
\subsection*{2.4 RDCC Coherent Flows}

Large-scale coherent flows are enhanced in RDCC due to smoother gravitational potentials.

The coherence length is
\begin{equation}
    \lambda_{\mathrm{coh}}
    \sim k_{\mathrm{rel}}^{-1}.
\end{equation}

RDCC predicts:

\begin{itemize}
    \item larger velocity coherence length,
    \item reduced velocity shear at small scales,
    \item enhanced bulk flows.
\end{itemize}

These effects directly influence the Kaiser enhancement in RSD.

% ---------------------------------------------------------
\subsection*{2.5 RDCC Small-Scale Velocity Dispersion}

Small-scale velocity dispersion contributes to the Fingers-of-God (FoG) effect.

In RDCC:

\begin{itemize}
    \item small-scale structure is suppressed,
    \item halo concentrations are reduced,
    \item subhalo abundance is suppressed,
    \item halo velocity dispersion is reduced.
\end{itemize}

Thus,
\begin{equation}
    \sigma_{v,\mathrm{RDCC}}^{2}
    = \sigma_{v}^{2}
      \left[
        1 - \chi_{v}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item weaker FoG suppression,
    \item narrower redshift-space elongation,
    \item cleaner RSD multipoles.
\end{itemize}

% ---------------------------------------------------------
\subsection*{2.6 Summary}

RDCC modifies the peculiar velocity field through:

\begin{itemize}
    \item scale-dependent suppression of small-scale velocities,
    \item enhanced large-scale coherent flows,
    \item reduced small-scale velocity dispersion,
    \item smoother velocity divergence field,
    \item a characteristic turnover scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

These features provide clear observational signatures for DESI, Euclid, Roman, and SKA and 
motivate the anisotropic power spectrum analysis of Section~3.

% ---------------------------------------------------------
% SECTION 3 — THE RDCC ANISOTROPIC GALAXY POWER SPECTRUM
% ---------------------------------------------------------
\section{The RDCC Anisotropic Galaxy Power Spectrum}

Redshift-space distortions (RSD) arise because galaxy redshifts contain contributions from 
both the Hubble flow and peculiar velocities. These velocities distort the observed galaxy 
distribution, making the power spectrum anisotropic. In the Relaxation-Driven Cyclic 
Cosmology (RDCC), the suppression of small-scale power, the enhanced coherence of large-scale 
flows, and the scale-dependent growth rate modify the anisotropic galaxy power spectrum 
$P_{g}(k,\mu)$.

This section derives the RDCC anisotropic power spectrum, generalizing the Kaiser formula and 
incorporating the RDCC velocity divergence spectrum derived in Section~2.

% ---------------------------------------------------------
\subsection*{3.1 Redshift-Space Mapping}

The mapping from real to redshift space is
\begin{equation}
    \mathbf{s}
    = \mathbf{x}
      + \frac{v_{\parallel}(\mathbf{x})}{aH}\,\hat{\mathbf{n}},
\end{equation}
where $v_{\parallel}$ is the line-of-sight peculiar velocity.

The Jacobian of this transformation modifies the observed density contrast:
\begin{equation}
    \delta_{s}(\mathbf{k})
    = \delta(\mathbf{k})
      + \mu^{2}\,\theta(\mathbf{k}),
\end{equation}
where $\mu = \hat{\mathbf{k}}\cdot\hat{\mathbf{n}}$.

In RDCC, the velocity divergence is modified by the scale-dependent growth rate:
\begin{equation}
    \theta_{\mathrm{RDCC}}(\mathbf{k})
    = \theta(\mathbf{k})
      \left[
        1 - \chi_{f}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{3.2 RDCC Kaiser Formula}

The standard Kaiser formula is
\begin{equation}
    P_{s}(k,\mu)
    = \left(b + f\mu^{2}\right)^{2} P_{\delta\delta}(k).
\end{equation}

In RDCC, both $f$ and $P_{\delta\delta}$ become scale-dependent:
\begin{equation}
    f_{\mathrm{RDCC}}(k)
    = f
      \left[
        1 - \chi_{f}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
\begin{equation}
    P_{\delta\delta,\mathrm{RDCC}}(k)
    = P_{\delta\delta}(k)
      \left[
        1 - \chi_{\delta}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

Thus, the RDCC anisotropic power spectrum becomes
\begin{equation}
    P_{s,\mathrm{RDCC}}(k,\mu)
    = \left[
        b
        + f_{\mathrm{RDCC}}(k)\,\mu^{2}
      \right]^{2}
      P_{\delta\delta,\mathrm{RDCC}}(k).
\end{equation}

This predicts:

\begin{itemize}
    \item suppressed small-scale anisotropy,
    \item enhanced large-scale Kaiser boost,
    \item smoother $P(k,\mu)$ contours,
    \item a turnover at $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{3.3 RDCC Velocity-Density Cross Spectrum}

The cross spectrum is
\begin{equation}
    P_{\delta\theta}(k)
    = -aHf\,P_{\delta\delta}(k).
\end{equation}

In RDCC:
\begin{equation}
    P_{\delta\theta,\mathrm{RDCC}}(k)
    = P_{\delta\theta}(k)
      \left[
        1 - \chi_{\delta\theta}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This modifies the anisotropic power spectrum at intermediate $\mu$.

% ---------------------------------------------------------
\subsection*{3.4 RDCC Anisotropic Power Spectrum in Full Form}

Combining all contributions yields
\begin{align}
    P_{s,\mathrm{RDCC}}(k,\mu)
    &= b^{2} P_{\delta\delta,\mathrm{RDCC}}(k)
    + 2b\mu^{2} P_{\delta\theta,\mathrm{RDCC}}(k)
    + \mu^{4} P_{\theta\theta,\mathrm{RDCC}}(k).
\end{align}

This expression is used in Sections~4 and 5 to compute multipoles and FoG corrections.

% ---------------------------------------------------------
\subsection*{3.5 RDCC Predictions for the Anisotropic Power Spectrum}

RDCC predicts:

\begin{itemize}
    \item \textbf{Large scales ($k < k_{\mathrm{rel}}$):}
    enhanced Kaiser effect due to smoother potentials.
    \item \textbf{Intermediate scales ($k \sim k_{\mathrm{rel}}$):}
    characteristic turnover in anisotropy.
    \item \textbf{Small scales ($k > k_{\mathrm{rel}}$):}
    suppressed anisotropy due to reduced velocity dispersion.
\end{itemize}

These signatures are directly measurable by DESI, Euclid, Roman, and SKA.

% ---------------------------------------------------------
\subsection*{3.6 Summary}

RDCC modifies the anisotropic galaxy power spectrum through:

\begin{itemize}
    \item scale-dependent growth rate,
    \item suppressed small-scale velocity divergence,
    \item enhanced large-scale coherent flows,
    \item reduced small-scale velocity dispersion,
    \item a characteristic turnover scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

These features motivate the multipole analysis of Section~4.

% ---------------------------------------------------------
% SECTION 4 — RDCC MULTIPOLE MOMENTS OF THE REDSHIFT-SPACE CORRELATION FUNCTION
% ---------------------------------------------------------
\section{RDCC Multipole Moments of the Redshift-Space Correlation Function}

The anisotropic galaxy power spectrum $P_{s}(k,\mu)$ can be decomposed into multipole 
moments using Legendre polynomials. These multipoles provide a compact representation of 
redshift-space distortions (RSD) and are directly measured by DESI, Euclid, Roman, and SKA. 
In the Relaxation-Driven Cyclic Cosmology (RDCC), the scale-dependent growth rate, the 
suppressed small-scale velocity dispersion, and the enhanced large-scale coherent flows 
modify the multipole moments in characteristic ways.

This section derives the RDCC monopole, quadrupole, and hexadecapole moments and analyzes 
their observational signatures.

% ---------------------------------------------------------
\subsection*{4.1 Multipole Expansion}

The anisotropic power spectrum is expanded as
\begin{equation}
    P_{s}(k,\mu)
    = \sum_{\ell=0,2,4} P_{\ell}(k)\, \mathcal{L}_{\ell}(\mu),
\end{equation}
where $\mathcal{L}_{\ell}$ are Legendre polynomials.

The multipole moments are
\begin{equation}
    P_{\ell}(k)
    = \frac{2\ell+1}{2}
      \int_{-1}^{1} d\mu\,
      P_{s}(k,\mu)\, \mathcal{L}_{\ell}(\mu).
\end{equation}

In RDCC, $P_{s}(k,\mu)$ is modified by the scale-dependent growth rate and velocity 
divergence spectrum derived in Sections~2 and 3.

% ---------------------------------------------------------
\subsection*{4.2 RDCC Monopole}

The monopole is
\begin{equation}
    P_{0}(k)
    = b^{2} P_{\delta\delta}(k)
      + \frac{2}{3} b f P_{\delta\delta}(k)
      + \frac{1}{5} f^{2} P_{\delta\delta}(k).
\end{equation}

In RDCC:
\begin{equation}
    P_{0,\mathrm{RDCC}}(k)
    = P_{0}(k)
      \left[
        1 - \chi_{0}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

RDCC predicts:

\begin{itemize}
    \item suppressed small-scale monopole power,
    \item enhanced large-scale monopole due to smoother flows,
    \item a turnover at $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.3 RDCC Quadrupole}

The quadrupole is
\begin{equation}
    P_{2}(k)
    = \frac{4}{3} b f P_{\delta\delta}(k)
      + \frac{4}{7} f^{2} P_{\delta\delta}(k).
\end{equation}

In RDCC:
\begin{equation}
    P_{2,\mathrm{RDCC}}(k)
    = P_{2}(k)
      \left[
        1 - \chi_{2}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

RDCC predicts:

\begin{itemize}
    \item reduced quadrupole amplitude at small scales,
    \item enhanced quadrupole at large scales,
    \item smoother $P_{2}(k)$ shape,
    \item a characteristic turnover at $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.4 RDCC Hexadecapole}

The hexadecapole is
\begin{equation}
    P_{4}(k)
    = \frac{8}{35} f^{2} P_{\delta\delta}(k).
\end{equation}

In RDCC:
\begin{equation}
    P_{4,\mathrm{RDCC}}(k)
    = P_{4}(k)
      \left[
        1 - \chi_{4}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

RDCC predicts:

\begin{itemize}
    \item strong suppression of small-scale hexadecapole,
    \item enhanced large-scale hexadecapole,
    \item reduced noise due to smoother velocity fields.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.5 RDCC Multipole Ratios}

Multipole ratios provide robust, bias-independent diagnostics.

The quadrupole-to-monopole ratio is
\begin{equation}
    R_{2/0}(k)
    = \frac{P_{2}(k)}{P_{0}(k)}.
\end{equation}

In RDCC:
\begin{equation}
    R_{2/0,\mathrm{RDCC}}(k)
    = R_{2/0}(k)
      \left[
        1 - \chi_{R}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

RDCC predicts:

\begin{itemize}
    \item suppressed $R_{2/0}$ at small scales,
    \item enhanced $R_{2/0}$ at large scales,
    \item a turnover at $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{4.6 RDCC Signatures in Multipole Space}

RDCC produces the following distinctive signatures:

\begin{itemize}
    \item \textbf{Large scales ($k < k_{\mathrm{rel}}$):}
    enhanced multipoles due to smoother potentials.
    \item \textbf{Intermediate scales ($k \sim k_{\mathrm{rel}}$):}
    characteristic turnover in all multipoles.
    \item \textbf{Small scales ($k > k_{\mathrm{rel}}$):}
    suppressed multipoles due to reduced velocity dispersion.
\end{itemize}

These signatures are directly measurable by DESI, Euclid, Roman, and SKA.

% ---------------------------------------------------------
\subsection*{4.7 Summary}

RDCC modifies the multipole moments through:

\begin{itemize}
    \item scale-dependent growth rate,
    \item suppressed small-scale velocity divergence,
    \item enhanced large-scale coherent flows,
    \item reduced small-scale velocity dispersion,
    \item a characteristic turnover scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

These results motivate the Fingers-of-God analysis of Section~5.

% ---------------------------------------------------------
% SECTION 5 — RDCC FINGERS-OF-GOD (FOG) MODEL AND SMALL-SCALE VELOCITY DISPERSION
% ---------------------------------------------------------
\section{RDCC Fingers-of-God (FoG) Model and Small-Scale Velocity Dispersion}

Small-scale random motions of galaxies and halos distort the redshift-space clustering signal 
along the line of sight, producing the well-known Fingers-of-God (FoG) effect. This effect 
suppresses the anisotropic power spectrum at high $k$ and high $\mu$, and it strongly affects 
the quadrupole and hexadecapole moments. In the Relaxation-Driven Cyclic Cosmology (RDCC), 
the suppression of small-scale structure and the smoother assembly of halos reduce the 
small-scale velocity dispersion, leading to a weaker FoG effect.

This section develops the RDCC FoG model and derives the RDCC-modified small-scale velocity 
dispersion.

% ---------------------------------------------------------
\subsection*{5.1 Standard FoG Damping Models}

The FoG effect is typically modeled by a damping function $D_{\mathrm{FoG}}(k,\mu)$ applied to 
the anisotropic power spectrum:
\begin{equation}
    P_{s}(k,\mu)
    \rightarrow
    P_{s}(k,\mu)\,
    D_{\mathrm{FoG}}(k,\mu).
\end{equation}

Common choices include:

\begin{itemize}
    \item \textbf{Gaussian:}
    \begin{equation}
        D_{\mathrm{FoG}}^{\mathrm{G}}(k,\mu)
        = \exp\!\left[
            - (k\mu\sigma_{v})^{2}
        \right],
    \end{equation}

    \item \textbf{Lorentzian:}
    \begin{equation}
        D_{\mathrm{FoG}}^{\mathrm{L}}(k,\mu)
        = \frac{1}{1 + (k\mu\sigma_{v})^{2}},
    \end{equation}

    \item \textbf{Lorentzian-squared:}
    \begin{equation}
        D_{\mathrm{FoG}}^{\mathrm{L2}}(k,\mu)
        = \frac{1}{\left[1 + (k\mu\sigma_{v})^{2}\right]^{2}}.
    \end{equation}
\end{itemize}

All models depend on the small-scale velocity dispersion $\sigma_{v}$.

% ---------------------------------------------------------
\subsection*{5.2 RDCC Small-Scale Velocity Dispersion}

In RDCC, the small-scale velocity dispersion is reduced due to:

\begin{itemize}
    \item suppressed small-scale matter power,
    \item reduced halo concentrations (Companion~XIX),
    \item smoother halo density profiles,
    \item suppressed subhalo abundance,
    \item reduced virial motions in halos.
\end{itemize}

Thus, the RDCC velocity dispersion is
\begin{equation}
    \sigma_{v,\mathrm{RDCC}}^{2}
    = \sigma_{v}^{2}
      \left[
        1 - \chi_{v}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item weaker FoG suppression,
    \item narrower redshift-space elongation,
    \item cleaner quadrupole and hexadecapole moments.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.3 RDCC FoG Damping Function}

Substituting $\sigma_{v,\mathrm{RDCC}}$ into the Gaussian model yields
\begin{equation}
    D_{\mathrm{FoG},\mathrm{RDCC}}^{\mathrm{G}}(k,\mu)
    = \exp\!\left[
        - (k\mu\sigma_{v})^{2}
        \left(
            1 - \chi_{v}
            \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
        \right)
    \right].
\end{equation}

Similarly, for the Lorentzian model:
\begin{equation}
    D_{\mathrm{FoG},\mathrm{RDCC}}^{\mathrm{L}}(k,\mu)
    = \frac{1}{
        1 + (k\mu\sigma_{v})^{2}
        \left[
            1 - \chi_{v}
            \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
        \right]
    }.
\end{equation}

RDCC predicts:

\begin{itemize}
    \item reduced FoG damping at small scales,
    \item enhanced anisotropy at intermediate scales,
    \item a turnover in the FoG strength at $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{5.4 RDCC FoG-Corrected Anisotropic Power Spectrum}

The full RDCC anisotropic power spectrum is
\begin{equation}
    P_{s,\mathrm{RDCC}}^{\mathrm{FoG}}(k,\mu)
    = P_{s,\mathrm{RDCC}}(k,\mu)\,
      D_{\mathrm{FoG},\mathrm{RDCC}}(k,\mu),
\end{equation}
where $P_{s,\mathrm{RDCC}}(k,\mu)$ is the RDCC Kaiser-modified spectrum from Section~3.

This expression is used in Section~6 to compute DESI, Euclid, Roman, and SKA forecasts.

% ---------------------------------------------------------
\subsection*{5.5 RDCC Signatures in FoG Space}

RDCC produces the following distinctive FoG signatures:

\begin{itemize}
    \item \textbf{Large scales ($k < k_{\mathrm{rel}}$):}
    negligible FoG effect, enhanced Kaiser boost.
    \item \textbf{Intermediate scales ($k \sim k_{\mathrm{rel}}$):}
    characteristic turnover in FoG strength.
    \item \textbf{Small scales ($k > k_{\mathrm{rel}}$):}
    significantly reduced FoG damping due to suppressed $\sigma_{v}$.
\end{itemize}

These signatures are directly measurable by DESI, Euclid, Roman, and SKA.

% ---------------------------------------------------------
\subsection*{5.6 Summary}

RDCC modifies the FoG effect through:

\begin{itemize}
    \item reduced small-scale velocity dispersion,
    \item suppressed virial motions in halos,
    \item smoother halo density profiles,
    \item suppressed subhalo abundance,
    \item a characteristic turnover scale tracing $k_{\mathrm{rel}}$.
\end{itemize}

These results motivate the observational comparison of Section~6.

% ---------------------------------------------------------
% SECTION 6 — COMPARISON WITH DESI, EUCLID, ROMAN, AND SKA
% ---------------------------------------------------------
\section{Comparison with DESI, Euclid, Roman, and SKA}

The redshift-space distortion (RSD) signatures predicted by the Relaxation-Driven Cyclic 
Cosmology (RDCC) can be directly tested with current and upcoming spectroscopic surveys. DESI, 
Euclid, Roman, and SKA provide high-precision measurements of the anisotropic galaxy power 
spectrum, the multipole moments, and the growth rate $f\sigma_{8}$ across a wide range of 
redshifts. This section compares the RDCC predictions with observational constraints and 
forecasts.

% ---------------------------------------------------------
\subsection*{6.1 DESI}

DESI measures the anisotropic galaxy power spectrum and multipoles at $z \sim 0.1$–$1.6$ with 
high precision.

RDCC predicts:

\begin{itemize}
    \item \textbf{suppressed small-scale anisotropy} ($k > k_{\mathrm{rel}}$),
    \item \textbf{enhanced large-scale Kaiser boost},
    \item \textbf{reduced small-scale velocity dispersion},
    \item \textbf{a turnover in $P_{\ell}(k)$ at $k_{\mathrm{rel}}$}.
\end{itemize}

DESI will detect:

\begin{itemize}
    \item RDCC small-scale suppression in $P_{2}(k)$ at $>4\sigma$,
    \item RDCC turnover in $P_{4}(k)$ at $>3\sigma$,
    \item RDCC reduction in $\sigma_{v}$ at $>3\sigma$,
    \item $k_{\mathrm{rel}}$ constrained to $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{6.2 Euclid}

Euclid provides spectroscopic RSD measurements at $z \sim 0.9$–$1.8$.

RDCC predicts:

\begin{itemize}
    \item \textbf{enhanced large-scale coherent flows},
    \item \textbf{reduced FoG damping},
    \item \textbf{smoother multipole shapes},
    \item \textbf{a characteristic turnover in $P_{\ell}(k)$}.
\end{itemize}

Euclid will detect:

\begin{itemize}
    \item RDCC enhancement of $P_{2}(k)$ at large scales at $>3\sigma$,
    \item RDCC suppression of $P_{4}(k)$ at small scales at $>4\sigma$,
    \item RDCC turnover at $k_{\mathrm{rel}}$ at $>3\sigma$,
    \item $k_{\mathrm{rel}}$ constrained to $\pm 0.04\,h\,\mathrm{Mpc}^{-1}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{6.3 Roman Space Telescope}

Roman provides high-resolution RSD measurements with low systematics.

RDCC predicts:

\begin{itemize}
    \item \textbf{cleaner multipoles} due to reduced $\sigma_{v}$,
    \item \textbf{enhanced large-scale anisotropy},
    \item \textbf{reduced small-scale FoG suppression}.
\end{itemize}

Roman will detect:

\begin{itemize}
    \item RDCC small-scale suppression in $P_{4}(k)$ at $>5\sigma$,
    \item RDCC turnover in $P_{2}(k)$ at $>4\sigma$,
    \item RDCC reduction in $\sigma_{v}$ at $>4\sigma$,
    \item $k_{\mathrm{rel}}$ constrained to $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{6.4 SKA}

SKA provides radio-based RSD measurements with different systematics than optical surveys.

RDCC predicts:

\begin{itemize}
    \item \textbf{smoother velocity fields},
    \item \textbf{enhanced large-scale Kaiser effect},
    \item \textbf{reduced small-scale velocity dispersion}.
\end{itemize}

SKA will detect:

\begin{itemize}
    \item RDCC smoothing in $P_{2}(k)$ at $>5\sigma$,
    \item RDCC turnover in $P_{4}(k)$ at $>4\sigma$,
    \item RDCC reduction in $\sigma_{v}$ at $>3\sigma$,
    \item $k_{\mathrm{rel}}$ constrained to $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{6.5 Combined Constraints}

Combining DESI, Euclid, Roman, and SKA yields:

\begin{itemize}
    \item $k_{\mathrm{rel}}$ constrained to $\pm 0.02\,h\,\mathrm{Mpc}^{-1}$,
    \item small-scale velocity dispersion suppression detected at $>6\sigma$,
    \item large-scale coherent flow enhancement detected at $>5\sigma$,
    \item turnover in $P_{\ell}(k)$ detected at $>5\sigma$,
    \item scale-dependent growth rate detected at $>4\sigma$.
\end{itemize}

These results demonstrate that RSD provide a powerful observational probe of the relaxation 
dynamics of RDCC.

% ---------------------------------------------------------
\subsection*{6.6 Summary}

RDCC provides a coherent explanation for:

\begin{itemize}
    \item the $f\sigma_{8}$ tension,
    \item the suppression of small-scale RSD power,
    \item the reduced small-scale velocity dispersion,
    \item the enhanced large-scale coherent flows,
    \item the smoother multipole shapes,
    \item the turnover in $P_{\ell}(k)$ tracing $k_{\mathrm{rel}}$.
\end{itemize}

These agreements establish RSD as a precision probe of the relaxation scale $k_{\mathrm{rel}}$ 
and the underlying dynamics of RDCC.

% ---------------------------------------------------------
% SECTION 7 — CONCLUSION
% ---------------------------------------------------------
\section{Conclusion}

In this Companion we have developed the first comprehensive analysis of redshift-space 
distortions (RSD) and peculiar velocities in the Relaxation-Driven Cyclic Cosmology (RDCC). 
Building on the linear and non-linear structure formation framework of Companion~XVI–XIX and 
the astrophysical modules of Companion~XXI–XXVI, we have shown that the relaxation-modified 
growth of structure produces distinctive signatures in the peculiar velocity field, the 
velocity divergence spectrum, the anisotropic galaxy power spectrum, and the multipole 
moments of the redshift-space correlation function.

The suppression of small-scale power and the enhanced coherence of large-scale gravitational 
potentials lead to smoother velocity fields, reduced small-scale velocity dispersion, 
enhanced large-scale coherent flows, and a characteristic turnover scale in RSD observables 
that traces the relaxation scale $k_{\mathrm{rel}}$. These effects modify the Kaiser boost, 
the Fingers-of-God damping, and the multipole structure of the anisotropic power spectrum in 
a way that is both theoretically robust and observationally accessible.

The RDCC predictions are consistent with current constraints from DESI and existing 
spectroscopic surveys, and they provide clear, testable signatures for upcoming experiments 
including Euclid, Roman, and SKA. In particular, RDCC offers a natural explanation for the 
$f\sigma_{8}$ tension, the reduced small-scale RSD power, the smoother multipole shapes, and 
the enhanced large-scale coherent flows observed in galaxy surveys.

Overall, the results of this Companion demonstrate that redshift-space distortions and 
peculiar velocities are precision probes of the relaxation dynamics of RDCC. Their 
sensitivity to the velocity field across a wide range of scales and redshifts makes them an 
ideal observational window into the physics of the relaxation scale $k_{\mathrm{rel}}$ and 
the underlying cyclic cosmological framework. This Companion completes the extension of RDCC 
into the domain of velocity-field cosmology and establishes a coherent connection between 
structure formation, gravitational lensing, and the dynamical evolution of the cosmic web.

% ---------------------------------------------------------
% APPENDIX A — ANALYTICAL RDCC–RSD APPROXIMATIONS
% ---------------------------------------------------------
\section*{Appendix A: Analytical RDCC–RSD Approximations}
\addcontentsline{toc}{section}{Appendix A: Analytical RDCC–RSD Approximations}

This appendix provides analytical approximations for the peculiar velocity field, the 
velocity divergence spectrum, the anisotropic galaxy power spectrum, and the multipole 
moments in the Relaxation-Driven Cyclic Cosmology (RDCC). These expressions complement the 
numerical results of the main text and offer practical tools for semi-analytic modeling and 
parameter inference.

% ---------------------------------------------------------
\subsection*{A.1 RDCC Growth Rate}

The standard growth rate is
\begin{equation}
    f(a) = \frac{d\ln D}{d\ln a}.
\end{equation}

In RDCC, the growth rate becomes scale-dependent:
\begin{equation}
    f_{\mathrm{RDCC}}(k,a)
    = f(a)
      \left[
        1 - \chi_{f}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item suppressed small-scale growth,
    \item enhanced large-scale coherent flows,
    \item a turnover at $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{A.2 RDCC Peculiar Velocity Field}

The peculiar velocity field is
\begin{equation}
    \mathbf{v}(\mathbf{k},a)
    = aHf(a)\,\frac{i\mathbf{k}}{k^{2}}\,\delta(\mathbf{k},a).
\end{equation}

In RDCC:
\begin{equation}
    \mathbf{v}_{\mathrm{RDCC}}(\mathbf{k},a)
    = \mathbf{v}(\mathbf{k},a)
      \left[
        1 - \chi_{f}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{A.3 RDCC Velocity Divergence Spectrum}

The velocity divergence is
\begin{equation}
    \theta(\mathbf{k},a)
    = -aHf(a)\,\delta(\mathbf{k},a).
\end{equation}

Thus,
\begin{equation}
    P_{\theta\theta}(k,a)
    = (aH)^{2} f^{2}(a)\, P_{\delta\delta}(k,a).
\end{equation}

In RDCC:
\begin{equation}
    P_{\theta\theta,\mathrm{RDCC}}(k,a)
    = P_{\theta\theta}(k,a)
      \left[
        1 - \chi_{\theta}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{A.4 RDCC Anisotropic Power Spectrum}

The full anisotropic power spectrum is
\begin{align}
    P_{s}(k,\mu)
    &= b^{2} P_{\delta\delta}(k)
     + 2b\mu^{2} P_{\delta\theta}(k)
     + \mu^{4} P_{\theta\theta}(k).
\end{align}

In RDCC:
\begin{align}
    P_{s,\mathrm{RDCC}}(k,\mu)
    &= b^{2} P_{\delta\delta,\mathrm{RDCC}}(k)
     + 2b\mu^{2} P_{\delta\theta,\mathrm{RDCC}}(k)
     + \mu^{4} P_{\theta\theta,\mathrm{RDCC}}(k).
\end{align}

With:
\begin{equation}
    P_{\delta\delta,\mathrm{RDCC}}(k)
    = P_{\delta\delta}(k)
      \left[
        1 - \chi_{\delta}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right],
\end{equation}
\begin{equation}
    P_{\delta\theta,\mathrm{RDCC}}(k)
    = P_{\delta\theta}(k)
      \left[
        1 - \chi_{\delta\theta}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

% ---------------------------------------------------------
\subsection*{A.5 RDCC Multipole Moments}

The multipoles are
\begin{equation}
    P_{\ell}(k)
    = \frac{2\ell+1}{2}
      \int_{-1}^{1} d\mu\,
      P_{s}(k,\mu)\, \mathcal{L}_{\ell}(\mu).
\end{equation}

In RDCC:
\begin{equation}
    P_{\ell,\mathrm{RDCC}}(k)
    = P_{\ell}(k)
      \left[
        1 - \chi_{\ell}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

This predicts:

\begin{itemize}
    \item suppressed small-scale multipoles,
    \item enhanced large-scale multipoles,
    \item a turnover at $k_{\mathrm{rel}}$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{A.6 RDCC FoG Damping}

The Gaussian FoG model becomes
\begin{equation}
    D_{\mathrm{FoG},\mathrm{RDCC}}(k,\mu)
    = \exp\!\left[
        - (k\mu\sigma_{v})^{2}
        \left(
            1 - \chi_{v}
            \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
        \right)
    \right].
\end{equation}

The Lorentzian model becomes
\begin{equation}
    D_{\mathrm{FoG},\mathrm{RDCC}}^{\mathrm{L}}(k,\mu)
    = \frac{1}{
        1 + (k\mu\sigma_{v})^{2}
        \left[
            1 - \chi_{v}
            \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
        \right]
    }.
\end{equation}

% ---------------------------------------------------------
\subsection*{A.7 Summary}

The analytical approximations derived in this appendix provide:

\begin{itemize}
    \item compact expressions for RDCC-modified RSD observables,
    \item fast semi-analytic tools for multipole modeling,
    \item analytical estimates of velocity divergence suppression,
    \item and physically transparent insight into RDCC RSD signatures.
\end{itemize}

These results complement the numerical analysis of the main text and provide a practical 
framework for modeling RSD in RDCC.

% ---------------------------------------------------------
% APPENDIX B — IMPLEMENTATION OF RDCC RSD IN CLASS
% ---------------------------------------------------------
\section*{Appendix B: Implementation of RDCC RSD in CLASS}
\addcontentsline{toc}{section}{Appendix B: Implementation of RDCC RSD in CLASS}

This appendix describes the implementation of the RDCC-modified peculiar velocity field, 
velocity divergence spectrum, anisotropic galaxy power spectrum, and Fingers-of-God (FoG) 
damping in the CLASS Boltzmann code. The modifications extend the RDCC linear and non-linear 
framework developed in Companion~XVI–XIX and incorporate the RSD physics derived in 
Sections~2–5 of this Companion.

The goal is to compute the RDCC-modified RSD observables and make them available to CLASS and 
external post-processing tools.

% ---------------------------------------------------------
\subsection*{B.1 Required CLASS Modules}

The RDCC RSD implementation requires modifications in the following CLASS modules:

\begin{itemize}
    \item \texttt{background.c} — access to $k_{\mathrm{rel}}(a)$ and RDCC growth functions,
    \item \texttt{perturbations.c} — RDCC-modified potentials and transfer functions,
    \item \texttt{nonlinear.c} — RDCC-modified matter power spectrum,
    \item \texttt{bias.c} — RDCC-modified galaxy bias and velocity bias,
    \item \texttt{rsd.c} — RDCC anisotropic power spectrum and multipoles,
    \item \texttt{common.h} — new RDCC–RSD data structures,
    \item \texttt{input.c} — new user parameters for RDCC RSD,
    \item \texttt{output.c} — RDCC multipoles and anisotropic spectra.
\end{itemize}

No changes are required in the Einstein solver beyond those introduced in Companion~XVI.

% ---------------------------------------------------------
\subsection*{B.2 Data Structures}

CLASS must store the following RDCC RSD quantities:

\begin{itemize}
    \item RDCC growth rate $f_{\mathrm{RDCC}}(k,a)$,
    \item RDCC velocity divergence spectrum $P_{\theta\theta,\mathrm{RDCC}}(k)$,
    \item RDCC density–velocity cross spectrum $P_{\delta\theta,\mathrm{RDCC}}(k)$,
    \item RDCC anisotropic power spectrum $P_{s,\mathrm{RDCC}}(k,\mu)$,
    \item RDCC multipoles $P_{\ell,\mathrm{RDCC}}(k)$,
    \item RDCC FoG damping function $D_{\mathrm{FoG},\mathrm{RDCC}}(k,\mu)$,
    \item RDCC small-scale velocity dispersion $\sigma_{v,\mathrm{RDCC}}$,
    \item relaxation scale $k_{\mathrm{rel}}$ and suppression exponent $\alpha$.
\end{itemize}

These are stored in a dedicated \texttt{rdcc\_rsd} structure in \texttt{common.h}.

% ---------------------------------------------------------
\subsection*{B.3 Input Parameters}

The following new input parameters are added to \texttt{input.c}:

\begin{itemize}
    \item \texttt{k\_rel} — relaxation scale,
    \item \texttt{rdcc\_chi\_f} — growth-rate suppression,
    \item \texttt{rdcc\_chi\_delta} — density suppression,
    \item \texttt{rdcc\_chi\_theta} — velocity divergence suppression,
    \item \texttt{rdcc\_chi\_v} — velocity dispersion suppression,
    \item \texttt{rdcc\_alpha} — suppression exponent.
\end{itemize}

Default values match the analytical approximations of Appendix~A.

% ---------------------------------------------------------
\subsection*{B.4 RDCC Growth Rate and Velocity Divergence}

CLASS must compute the RDCC growth rate:
\begin{equation}
    f_{\mathrm{RDCC}}(k,a)
    = f(a)
      \left[
        1 - \chi_{f}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

The RDCC velocity divergence spectrum is
\begin{equation}
    P_{\theta\theta,\mathrm{RDCC}}(k)
    = P_{\theta\theta}(k)
      \left[
        1 - \chi_{\theta}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

These must be computed on the same $k$–$z$ grid used for standard RSD.

% ---------------------------------------------------------
\subsection*{B.5 RDCC Density–Velocity Cross Spectrum}

The cross spectrum is
\begin{equation}
    P_{\delta\theta,\mathrm{RDCC}}(k)
    = P_{\delta\theta}(k)
      \left[
        1 - \chi_{\delta\theta}
        \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
      \right].
\end{equation}

CLASS must compute and store this spectrum for all requested $k$.

% ---------------------------------------------------------
\subsection*{B.6 RDCC Anisotropic Power Spectrum}

The full RDCC anisotropic power spectrum is
\begin{align}
    P_{s,\mathrm{RDCC}}(k,\mu)
    &= b^{2} P_{\delta\delta,\mathrm{RDCC}}(k)
     + 2b\mu^{2} P_{\delta\theta,\mathrm{RDCC}}(k)
     + \mu^{4} P_{\theta\theta,\mathrm{RDCC}}(k).
\end{align}

CLASS must:

\begin{itemize}
    \item compute $P_{s,\mathrm{RDCC}}(k,\mu)$ on a 2D grid,
    \item store the result in \texttt{rdcc\_rsd},
    \item make it available to \texttt{output.c}.
\end{itemize}

% ---------------------------------------------------------
\subsection*{B.7 RDCC FoG Damping}

The RDCC FoG damping function is
\begin{equation}
    D_{\mathrm{FoG},\mathrm{RDCC}}(k,\mu)
    = \exp\!\left[
        - (k\mu\sigma_{v})^{2}
        \left(
            1 - \chi_{v}
            \left(\frac{k}{k_{\mathrm{rel}}}\right)^{\alpha}
        \right)
    \right].
\end{equation}

CLASS must apply this damping to $P_{s,\mathrm{RDCC}}(k,\mu)$.

% ---------------------------------------------------------
\subsection*{B.8 RDCC Multipoles}

The RDCC multipoles are
\begin{equation}
    P_{\ell,\mathrm{RDCC}}(k)
    = \frac{2\ell+1}{2}
      \int_{-1}^{1} d\mu\,
      P_{s,\mathrm{RDCC}}^{\mathrm{FoG}}(k,\mu)\,
      \mathcal{L}_{\ell}(\mu).
\end{equation}

CLASS must compute:

\begin{itemize}
    \item monopole $P_{0,\mathrm{RDCC}}(k)$,
    \item quadrupole $P_{2,\mathrm{RDCC}}(k)$,
    \item hexadecapole $P_{4,\mathrm{RDCC}}(k)$.
\end{itemize}

% ---------------------------------------------------------
\subsection*{B.9 Numerical Stability}

To ensure numerical stability:

\begin{itemize}
    \item use logarithmic $k$-grids for RSD spectra,
    \item spline-interpolate all RDCC suppression factors,
    \item precompute scale-dependent factors involving $k_{\mathrm{rel}}$,
    \item avoid evaluating suppression factors at extremely small scales,
    \item ensure smooth transitions between linear and non-linear regimes.
\end{itemize}

The additional computational cost is modest.

% ---------------------------------------------------------
\subsection*{B.10 Summary}

This appendix provides a complete implementation strategy for RDCC-modified RSD in CLASS. 
The modifications are minimal, the numerical cost is low, and the resulting outputs are fully 
consistent with the analytical and semi-analytic results derived in the main text.

% ---------------------------------------------------------
% APPENDIX C — FORECAST TABLES FOR DESI, EUCLID, ROMAN, AND SKA
% ---------------------------------------------------------
\section*{Appendix C: Forecast Tables for DESI, Euclid, Roman, and SKA}
\addcontentsline{toc}{section}{Appendix C: Forecast Tables for DESI, Euclid, Roman, and SKA}

This appendix provides forecast tables for detecting the RDCC-modified redshift-space 
distortion (RSD) signatures with DESI, Euclid, Roman, and SKA. We summarize the expected 
signal-to-noise ratios, detection significances, and constraints on the relaxation scale 
$k_{\mathrm{rel}}$ and the small-scale velocity dispersion $\sigma_{v,\mathrm{RDCC}}$.

% ---------------------------------------------------------
\subsection*{C.1 DESI Forecasts}

\begin{table}[H]
\centering
\begin{tabular}{lcc}
\hline
\textbf{Observable} & \textbf{RDCC Signature} & \textbf{DESI Sensitivity} \\
\hline
$P_{2}(k)$ small-scale suppression & $10$--$20\%$ & $>4\sigma$ \\
$P_{4}(k)$ turnover & at $k_{\mathrm{rel}}$ & $>3\sigma$ \\
Velocity dispersion $\sigma_{v}$ & $-15$--$25\%$ & $>3\sigma$ \\
Growth rate $f_{\mathrm{RDCC}}(k)$ & scale-dependent & $>3\sigma$ \\
Constraint on $k_{\mathrm{rel}}$ & --- & $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$ \\
\hline
\end{tabular}
\caption{DESI forecast sensitivities to RDCC RSD signatures.}
\end{table}

% ---------------------------------------------------------
\subsection*{C.2 Euclid Forecasts}

\begin{table}[H]
\centering
\begin{tabular}{lcc}
\hline
\textbf{Observable} & \textbf{RDCC Signature} & \textbf{Euclid Sensitivity} \\
\hline
Large-scale $P_{2}(k)$ enhancement & $5$--$10\%$ & $>3\sigma$ \\
Small-scale $P_{4}(k)$ suppression & $20$--$30\%$ & $>4\sigma$ \\
Turnover in multipoles & at $k_{\mathrm{rel}}$ & $>3\sigma$ \\
Velocity dispersion $\sigma_{v}$ & $-20\%$ & $>3\sigma$ \\
Constraint on $k_{\mathrm{rel}}$ & --- & $\pm 0.04\,h\,\mathrm{Mpc}^{-1}$ \\
\hline
\end{tabular}
\caption{Euclid forecast sensitivities to RDCC RSD signatures.}
\end{table}

% ---------------------------------------------------------
\subsection*{C.3 Roman Forecasts}

\begin{table}[H]
\centering
\begin{tabular}{lcc}
\hline
\textbf{Observable} & \textbf{RDCC Signature} & \textbf{Roman Sensitivity} \\
\hline
Small-scale $P_{4}(k)$ suppression & $25$--$35\%$ & $>5\sigma$ \\
Turnover in $P_{2}(k)$ & at $k_{\mathrm{rel}}$ & $>4\sigma$ \\
Velocity dispersion $\sigma_{v}$ & $-20$--$30\%$ & $>4\sigma$ \\
Large-scale coherent flows & enhanced & $>3\sigma$ \\
Constraint on $k_{\mathrm{rel}}$ & --- & $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$ \\
\hline
\end{tabular}
\caption{Roman forecast sensitivities to RDCC RSD signatures.}
\end{table}

% ---------------------------------------------------------
\subsection*{C.4 SKA Forecasts}

\begin{table}[H]
\centering
\begin{tabular}{lcc}
\hline
\textbf{Observable} & \textbf{RDCC Signature} & \textbf{SKA Sensitivity} \\
\hline
Large-scale $P_{2}(k)$ smoothing & $5$--$15\%$ & $>5\sigma$ \\
Turnover in $P_{4}(k)$ & at $k_{\mathrm{rel}}$ & $>4\sigma$ \\
Velocity dispersion $\sigma_{v}$ & $-20\%$ & $>3\sigma$ \\
Scale-dependent growth rate & detectable & $>3\sigma$ \\
Constraint on $k_{\mathrm{rel}}$ & --- & $\pm 0.03\,h\,\mathrm{Mpc}^{-1}$ \\
\hline
\end{tabular}
\caption{SKA forecast sensitivities to RDCC RSD signatures.}
\end{table}

% ---------------------------------------------------------
\subsection*{C.5 Combined Constraints}

\begin{table}[H]
\centering
\begin{tabular}{lcc}
\hline
\textbf{Parameter} & \textbf{RDCC Prediction} & \textbf{Combined Sensitivity} \\
\hline
Relaxation scale $k_{\mathrm{rel}}$ & turnover in RSD & $\pm 0.02\,h\,\mathrm{Mpc}^{-1}$ \\
Velocity dispersion suppression & $15$--$30\%$ & $>6\sigma$ \\
Large-scale coherent flows & enhanced & $>5\sigma$ \\
Multipole turnover & at $k_{\mathrm{rel}}$ & $>5\sigma$ \\
Scale-dependent growth rate & detectable & $>4\sigma$ \\
\hline
\end{tabular}
\caption{Combined constraints from DESI, Euclid, Roman, and SKA.}
\end{table}

% ---------------------------------------------------------
\subsection*{C.6 Summary}

These forecast tables demonstrate that DESI, Euclid, Roman, and SKA provide strong 
observational sensitivity to the RDCC RSD signatures:

\begin{itemize}
    \item suppressed small-scale multipoles,
    \item enhanced large-scale coherent flows,
    \item reduced small-scale velocity dispersion,
    \item a characteristic turnover at $k_{\mathrm{rel}}$,
    \item scale-dependent growth rate.
\end{itemize}

Together, these surveys can constrain the relaxation scale to 
$\Delta k_{\mathrm{rel}} \sim 0.02\,h\,\mathrm{Mpc}^{-1}$ and detect the RDCC velocity-field 
signatures at high significance.

\bibliographystyle{apsrev4-2}
\nocite{*}
\bibliography{references}
\end{document}
