\documentclass[11pt,a4paper]{article}

% ---------------------------------------------------------
% PREAMBLE (stable, warning-free, Overleaf-safe)
% ---------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bm}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{authblk}
\usepackage{physics}
\usepackage{mathtools}

\geometry{margin=2.4cm}

\title{\textbf{Companion LXXXVIII: Emulator-Accelerated RDCC Inference Using Neural Transport Surrogates and Multi-Fidelity Simulation Flows}}
\author{Michael Lehmann \\ \texttt{mi.lehmann@gmx.de}}
\date{April 2026}

\begin{document}
\maketitle

\begin{abstract}
Inference in the Relaxation-Driven Cyclic Cosmology (RDCC) requires repeated 
evaluation of summary statistics across a range of $\alpha$ values. Full 
weak-lensing simulations are computationally expensive, making direct likelihood 
evaluation impractical. In this work, we introduce emulator-accelerated RDCC 
inference using neural transport surrogates and multi-fidelity simulation flows. 
Neural transport surrogates emulate the distribution of persistence diagrams, 
Minkowski functionals, and shear power spectra as a function of $\alpha$, while 
multi-fidelity flows combine low- and high-resolution simulations into a unified 
generative model. This framework reduces simulation cost by orders of magnitude 
and enables fast, accurate, and scalable RDCC inference for Euclid-like surveys.
\end{abstract}
% ---------------------------------------------------------
\section{Introduction}

Inference in RDCC requires evaluating summary statistics for many values of the 
relaxation parameter $\alpha$. Full weak-lensing simulations are expensive, and 
the summary statistics used in previous Companions—transport distances, Minkowski 
functionals, and shear power spectra—are sensitive to non-linear structure 
formation.

To make RDCC inference scalable, we introduce:

\begin{itemize}
    \item neural transport surrogates for topological statistics,
    \item multi-fidelity simulation flows combining low- and high-resolution data,
    \item emulator-based likelihood acceleration.
\end{itemize}

This reduces computational cost while preserving accuracy.
% ---------------------------------------------------------
\section{Neural Transport Surrogates}

We construct a surrogate model $S_{\theta}$ that emulates the distribution of 
persistence diagrams as a function of $\alpha$:



\[
    D \sim S_{\theta}(\alpha).
\]



The surrogate is trained using:

\begin{itemize}
    \item neural Schrödinger flows (Companion LXXX),
    \item unified diffusion--OT geometry (LXXXI),
    \item transport-based sensitivity maps (LXXXII).
\end{itemize}

The surrogate predicts:

\begin{itemize}
    \item persistence-diagram distributions,
    \item transport distances $\mathcal{T}_{\ell}(\alpha)$,
    \item their derivatives $\partial \mathcal{T}_{\ell}/\partial \alpha$.
\end{itemize}
% ---------------------------------------------------------
\section{Multi-Fidelity Simulation Flows}

We combine:

\begin{itemize}
    \item low-resolution simulations (cheap, noisy),
    \item high-resolution simulations (expensive, accurate),
\end{itemize}

into a single flow-based generative model:



\[
    f_{\mathrm{MF}} = \lambda f_{\mathrm{low}} + (1-\lambda) f_{\mathrm{high}}.
\]



Flow-matching ensures consistency across fidelities:



\[
    \mathcal{L}_{\mathrm{FM}}
    =
    \mathbb{E}
    \left[
        \| f_{\mathrm{MF}}(x) - f_{\mathrm{high}}(x) \|^{2}
    \right].
\]



This yields high-resolution accuracy at low-resolution cost.
% ---------------------------------------------------------
\section{Emulator-Accelerated Likelihood Evaluation}

Given the neural joint likelihood $\mathcal{L}_{\theta}$ from Companion LXXXVI, 
we replace expensive simulation calls with surrogate predictions:



\[
    \mathcal{L}_{\mathrm{emul}}(\mathbf{s} \mid \alpha)
    =
    \mathcal{L}_{\theta}\!\left(
        \mathbf{s} \mid S_{\theta}(\alpha)
    \right).
\]



This accelerates:

\begin{itemize}
    \item posterior sampling,
    \item Fisher forecasts,
    \item sensitivity analyses,
    \item calibration loops (Companion LXXXVII).
\end{itemize}
% ---------------------------------------------------------
\section{Accuracy Diagnostics}

We evaluate emulator accuracy using:

\begin{itemize}
    \item score discrepancies $\Delta s$,
    \item transport-based distances between surrogate and true diagrams,
    \item posterior-predictive checks,
    \item coverage tests for $\alpha$.
\end{itemize}

These diagnostics ensure that the emulator preserves RDCC signatures.
% ---------------------------------------------------------
\section{Conclusion}

Emulator-accelerated inference is essential for applying RDCC to Euclid-like 
surveys. Neural transport surrogates emulate topological statistics, 
multi-fidelity flows combine simulations of different resolutions, and 
emulator-based likelihood evaluation accelerates inference by orders of 
magnitude. This Companion establishes the computational backbone for scalable, 
accurate RDCC parameter estimation.

\bibliographystyle{apsrev4-2}
\nocite{*}
\bibliography{references}


\end{document}
